### 2019-09-20 Scripts/Denoising.all.runs.Rmd
Cleanup all done

---
title: "Denoising and decontaminating v2"
output:
  html_document:
    df_print: paged
---
After running the demultiplexer_for_dada2 (http://github.com/ramongallego/demultiplexer_for_dada2), we have to denoise the whole dataset. We will do this by using 4 different processes:

  * **Estimation of *Tag-jumping* or indices *cross-talk* **. We run multiple samples on each MiSeq run. These are identified by two sets of molecular barcodes. There is the potential of some sequences to be assigned to the wrong sample, which is a bummer. To estimate how many reads did this, on each MiSeq run we added some samples whose composition is known and extremely unlikely to be present in the enviromental samples studied. AS a result of this **Tag-jumping**, some of the positive control sequences might show in the environmental samples and viceversa. In our case, these positive controls are made of either Kangaroo or Ostrich (and Alligator). The process consists on, for each run, to model the compositon observed on the positive controls and substract it from the environmental samples from that run. The output will be a dataset with the same number of samples as before, but with fewer reads of certain sequences (ASVs)

  * **Discarding samples with extremely low number of reads**. Sometimes the number of reads sequenced from a particular replicate are really low, and hence the relative proportions of ASVs would be skewed.

  * **Full clearance from Positive control influence**. THis process also takes advantage of the known composition of the positive controls. Each ASV found in the positive controls with a higher abundace in them than in the rest of the samples will be labelled as  **Positive** and removed from the environmental dataset. The output will be a dataset with the same number of samples as before but with fewer ASVs.

  * **Occupancy modelling** . Is the presence of a ASV a reflection of a biological reality or likely a PCR artifact? This may seem trivial in extreme cases (an ASV that only appears in one PCR replicate in the whole dataset) but how to discriminate between PCR artifacts from rare but real organisms? We use Occupancy modelling to determine if the pattern of presence of a ASV in a dataset reflects that. The output of this procedure will be a datasetwith the same number of samples as before but with fewer ASVs.

  * **Dissimilarity between PCR replicates**. The workflow that leads to the sequencing of a particular sample is subject to many stochatic processes, and it is not unlikely that the composition retrieved is very different for the original community. A way to ensure that this difference is minimal is through the separate analysis of each PCR replicate. We used that approach and modeled the dissimilarity between each PCr replicate and the group centroid. This way of modeling the dissimilarity allows us to discard those PCR replicate that won't fit the normal distribution of dissimilarities. The output of this procedure will be a dataset with the same number of **Hashes** as before but with fewer **samples**.

As with everything, we will start the process by loading the required packages and datasets.

# Load the dataset and metadata

```{r load libraries, include=FALSE}
 knitr::opts_chunk$set(warning = FALSE)

 library (tidyverse)
 library (vegan)
 #library (MASS)
 library (proxy)
 library (reshape2)
 library (seqinr)
 library(patchwork)

```

We will load the ASV table and the metadata file. They are in the same folder so we use `list.files` to access them and a neat combination of `bind.rows` and `map(read_csv)`.

```{r load datasets - we will be doing that for all runs}

Local_folder <- "../Input/1.Sequencing_runs_output"

all.asvs <- list.files(path = Local_folder, pattern = "^ASV_table.csv", recursive = T, full.names = T)
all.metadata <- list.files(path = Local_folder, pattern = "^metadata.csv", recursive = T, full.names = T)
all.hashes <- list.files(path = Local_folder, pattern = "^Hash_key.csv", recursive = T, full.names = T,ignore.case = T)

ASV.table <- bind_rows(map(all.asvs, read_csv), .id = "Miseq_run")

metadata <- bind_rows(map(all.metadata, function(x) {
  read_csv(x) %>%
    dplyr::select("sample_id", "pri_index_name", "Tag")
  }),.id = "Miseq_run")

Hash.key <- bind_rows(map(all.hashes, read_csv))

Hash.key %>% 
  distinct(Hash, .keep_all = T) -> Hash.key
```

## Data Cleanup - Don't act like you don't need this

A few things we check for: That **no sample appears twice** in the metadata. That the metadata **uses Tag_01 instead of Tag_1** (so it can be sorted alphabetically). That **the structure** Site_YYYYMM[A-C].[1-3] **is the same** across the dataset.

```{r data cleaning}

# Check that no sample appears more than once in the metadata

metadata %>% 
  group_by(sample_id) %>%
  summarise(tot = n()) %>% 
  arrange(desc(tot)) # Samples only appear once

# We should change Tag_1 for Tag_01

metadata %>%
  mutate(Tag = case_when(str_detect(Tag, "\\_[0-9]{1}$")       ~     str_replace(Tag, "Tag_", "Tag_0"),
                         TRUE                                  ~     Tag  )) -> metadata

metadata %>% mutate(x= str_count(sample_id, pattern = "_")) %>% arrange(desc(x)) # Lilliwaup 201710 has 2 underscores, run 3 metadata has no underscore

ASV.table %>% distinct(sample) %>% mutate(x= str_count(sample, pattern = "_")) %>% arrange(desc(x)) # Also in the ASV table

# Fix it here

metadata %>% mutate(x= str_count(sample_id, pattern = "_")) %>%
  mutate(sample_id = case_when(x == 2 ~ str_replace(sample_id, pattern = "_B", replacement = "B"),
                               TRUE   ~ sample_id)) %>% 
  dplyr::select(-x) -> metadata

ASV.table %>% mutate(x= str_count(sample, pattern = "_")) %>%
  mutate(sample = case_when(x == 2 ~ str_replace(sample, pattern = "_B", replacement = "B"),
                               TRUE   ~ sample)) %>% 
  dplyr::select(-x) -> ASV.table

# Done

# Change Kangaroo.4 for Kangaroo.3. Also Change Ostrich1 Ostrich2 and Ostrich3 for .1, .2 .3

metadata %>% 
  mutate (sample_id = case_when (str_detect(sample_id, "Ostrich[123]") ~ str_replace(sample_id, "Ostrich", "Ostrich\\."),
                                 sample_id == "Kangaroo.4"             ~ "Kangaroo.3",
                                 TRUE                                  ~ sample_id)) -> metadata
ASV.table %>%
  mutate (sample = case_when (str_detect(sample, "Ostrich[123]") ~ str_replace(sample, "Ostrich", "Ostrich\\."),
                                 sample == "Kangaroo.4"          ~ "Kangaroo.3",
                                 TRUE                            ~ sample)) -> ASV.table

# DONE

# Add underscore to run3

ASV.table %>% 
  filter(str_detect(sample, "[A-Z][A-Z]2017")) %>% 
  distinct(Miseq_run) # it only affects one run

ASV.table %>% 
  mutate (sample = case_when(str_detect(sample, "[A-Z][A-Z]2017")    ~    str_replace(sample,"2017", "_2017" ),
                             TRUE                                    ~    sample)) -> ASV.table

# ASV table from Run 1 has the full date instead of just YYYYMM

ASV.table %>% 
  filter(str_detect(sample, "_[:digit:]{8}")) # YES

ASV.table %>% 
  mutate(sample = case_when(str_detect(sample, "_[:digit:]{8}")    ~  paste0(str_replace(sample, "_[:digit:]{8}", str_extract(sample,"_[:digit:]{6}" ))),
                            TRUE                                   ~  sample)) -> ASV.table

# in MiSeqRun1, there are duplicated entries of some Hash / sample combos, probably bc we combined the Miseq and MiSeq nano runs

ASV.table %>% 
  group_by(Miseq_run, sample, Hash) %>% 
  summarise(nReads = sum(nReads)) %>% 
  ungroup -> ASV.table

```

The output of this process are a clean ASV table and a clean metadata file.

## Cleaning Process 1: Estimation of *Tag-jumping* or sample *cross-talk*

Before we modify our datasets on any way, we can calculate how many sequences that were only supposed to be in the positives control appeared in the environmental samples, and how many did the opposite. First we divide the dataset into positive control and environmental samples. Also create an ordered list of the Hashes present in the positive controls, for ease of plotting

```{r split into two}

ASV.table %>%  mutate(source = case_when(str_detect(sample, "Kangaroo|K\\+|k\\+|Ostrich")    ~   "Positives",
                                         TRUE                                                ~   "Samples")) -> ASV.table

ASV.table %>% 
  filter (source == "Positives") %>% 
  group_by(Hash) %>% 
  summarise(tot = sum(nReads)) %>% 
  arrange(desc(tot)) %>% 
  pull(Hash) -> good.order

```

Now let's create a jumping vector. What proportion of the reads found in the positives control come from elsewhere, and what proportion of the reads in the samples come from the positives control.

### Step 1: Nest the dataset and split it in positives and samples

To streamline the process and make it easier to execute it similarly but independently on each Miseq run, we nest the dataset by run. 
So Step1 is create a nested table so we can run this analysis on each run independently.

```{r nesting the dataset}

ASV.table %>% 
  group_by(Miseq_run, source) %>% 
  nest() %>% 
  pivot_wider(names_from = source, values_from =  data) -> ASV.nested 
```

That wasn't too complicated. Let's start a summary function that keeps track of our cleaning process

```{r summary.file}

how.many <- function(ASVtable, round){
  ASVtable %>% ungroup() %>% 
    summarise(nsamples = n_distinct(sample),
              nHashes = n_distinct(Hash),
              nReads = sum(nReads), 
              Stage = paste0("Step_", round)) %>% 
    gather(starts_with("n"), value = "number", key = "Stat")
}

ASV.nested %>% 
  transmute(Miseq_run,Summary = map(Samples, ~ how.many(ASVtable = .,round = 0)))  -> ASV.summary

```

### Step 2: Model the composition of the positive controls of each run

We create a vector of the composition of each positive control and substract it from the environmental samples from their runs

```{r jumping vector}

ASV.nested %>% 
  mutate (contam.tibble = map(Positives, 
                              function(.x){
                                .x %>%
                                  group_by(sample) %>%
                                  mutate (TotalReadsperSample = sum(nReads)) %>%
                                  mutate (proportion = nReads/TotalReadsperSample) %>%
                                  group_by (Hash) %>%
                                  summarise (vector_contamination = max (proportion))
                                }) ) -> ASV.nested

ASV.nested %>% 
  group_by(Miseq_run) %>% 
  select(contam.tibble) %>% 
  unnest(cols = contam.tibble) # Check how it looks like

```

### Step 3: Substract the composition of the positive controls from the environment samples

The idea behind this procedure is that we know, for each run, how many reads from each Hash appeared in teh positive controls. These come from 2 processes: sequences we know should appear in the positive controls, and sequences that have *jumped* from the environment to the positive controls. With this procedure, we substract from every environmental sample the proportion of reads that jumped from elsewhere.

```{r cleaning step 1}
ASV.nested %>% 
  mutate(cleaned.tibble = map2(Samples, contam.tibble, function(.x,.y){ 
    .x %>%
      group_by (sample) %>%
      mutate (TotalReadsperSample = sum (nReads)) %>%
      left_join(.y, by = "Hash") %>%
      mutate (Updated_nReads = ifelse (!is.na(vector_contamination),  nReads - (ceiling(vector_contamination*TotalReadsperSample)), nReads)) %>%
      filter (Updated_nReads > 0) %>%
      ungroup() %>% 
      dplyr::select (sample, Hash, nReads = Updated_nReads)

  })) -> ASV.nested

ASV.nested %>% 
  group_by(Miseq_run) %>% 
  select(cleaned.tibble) %>% 
  unnest(cleaned.tibble) #Check how they look

```
Add this step to the summary table we were creating

```{r summary.file.2}
ASV.nested %>% 
  transmute(Miseq_run, Summary.1 = map(cleaned.tibble, ~ how.many(ASVtable = .,round = "1.Jump"))) %>% 
  left_join(ASV.summary) %>% 
  mutate(Summary   = map2(Summary, Summary.1, bind_rows)) %>%
  dplyr::select(-Summary.1) -> ASV.summary 
```

## Cleaning Process 2: **Discarding PCR replicates with low number of reads**

We will fit the number of reads assigned to each sample to a normal distribution and discard those samples with a probability of 95% of not fitting in that distribution. The output would be a dataset with less samples and potentially less number of unique Hashes.

```{r fitting nReads per sample}

ASV.nested %>% 
  select(Miseq_run,cleaned.tibble) %>% 
  unnest(cleaned.tibble) %>% 
  group_by(sample) %>%
  summarise(tot = sum(nReads)) -> all.reps

# Visualize

all.reps %>%  
  pull(tot) -> reads.per.sample

names(reads.per.sample) <- all.reps %>% pull(sample)

normparams.reads <- MASS::fitdistr(reads.per.sample, "normal")$estimate

all.reps %>%  
  mutate(prob = pnorm(tot, normparams.reads[1], normparams.reads[2])) -> all.reps

#  probs <- pnorm(all_pairwise_distances, normparams[1], normparams[2])

outliers <- 
  all.reps %>% 
  filter(prob < 0.1 & tot < normparams.reads[1])

ASV.nested %>% 
  mutate(Step.1.low.reads = map (cleaned.tibble, ~ filter(.,!sample %in% outliers$sample) %>% ungroup)) -> ASV.nested

ASV.nested %>% 
  transmute(Miseq_run, Summary.1 = map(Step.1.low.reads, ~ how.many(ASVtable = .,round = "2.Low.nReads"))) %>% 
  left_join(ASV.summary) %>% 
  mutate(Summary   = map2(Summary, Summary.1, bind_rows)) %>%
  dplyr::select(-Summary.1) -> ASV.summary

```

## Cleaning Process 3: **Full clearance from Positive control influence**

Removing the Hashes that belong to the positive controls. First, for each Hash that appeared in the positive controls, determine whether a sequence is a true positive or a true environment. For each Hash, we will calculate, maximum, mean and total number of reads in both positive and samples, and then we will use the following decission tree:

  * If all three statistics are higher in one of the groups, we will label it either of Environmental or Positive control influence.

  * If there are conflicting results, we will use the Hashes. to see if they belong to either the maximum abundance of a Hash is in a positive, then it is a positive, otherwise is a real sequence from the environment.

Now, for each Hash in each set of positives controls, calculate the proportion of reads that were missasigned - they appeared somewhere they were not expected.
We will divide that process in two: first . A second step would be to create a column named proportion switched, which states the proportion of reads from one Hash that jumped from the environment to a positive control or viceversa. The idea is that any presence below a threshold can be arguably belong to tag jumping.

```{r real or positive}

ASV.table %>% 
  filter (Hash %in% good.order) %>%
  group_by(sample) %>% 
  mutate(tot.reads = sum(nReads)) %>% 
  group_by(Hash,sample) %>% 
  mutate(prop = nReads/tot.reads) %>% 
  group_by(Hash, source) %>% 
  summarise (max.  = max(prop),
             mean. = mean(prop),
             tot.  = sum(nReads)) %>% 
  gather(contains("."), value = "number", key = "Stat") %>%
  spread(key = "source", value = "number", fill = 0) %>% 
  group_by(Hash, Stat) %>%
  mutate(origin = case_when(Positives > Samples ~ "Positive.control",
                            TRUE                ~ "Environment")) %>% 
  group_by (Hash) %>%
  mutate(tot = n_distinct(origin)) -> Hash.fate.step2

Hash.fate.step2 %>% 
  filter(tot == 1) %>% 
  group_by(Hash) %>% 
  summarise(origin = unique(origin)) %>% 
  filter(origin == "Positive.control") -> Hashes.to.remove.step2

ASV.table %>% 
  group_by(source, Hash) %>% 
  summarise(ocurrences =n()) %>% 
  spread(key = source, value = ocurrences, fill = 0) %>% 
  #left_join(Hashes.to.remove.step2) %>% 
  #mutate(origin = case_when(is.na(origin) ~ "Kept",
   #                         TRUE          ~ "Discarded")) %>% 
  mutate(second.origin = case_when(Positives >= Samples ~ "Discarded",
                                   TRUE                 ~ "Kept")) %>% 
  filter(second.origin == "Discarded") %>% 
  full_join(Hashes.to.remove.step2) -> Hashes.to.remove.step2

Hashes.to.remove.step2 %>% 
  bind_rows(tibble(Hash = c("01204d587a3c3752f426f2f6d813c0ff2b80ec8b",
                            "acebcd5c491bb273f3e4d615cafad649"))) -> Hashes.to.remove.step2

```
IN order to train DADA2 to better distinguish when positive control sequences have arrived in the environment, we will keep the sequences in a csv file

```{r ASVs from positives}

Hashes.to.remove.step2 %>% 
  left_join(Hash.key) %>% 
  select(Hash, Sequence) %>% 
  write_csv("../Output/Hashes.to.remove.csv")

```

### Remove the positive control hashes from the composition of the ASVs

```{r cleaning.Step2}

ASV.nested %>% 
  mutate(Step2.tibble = map (Step.1.low.reads, ~ filter(.,!Hash %in% Hashes.to.remove.step2$Hash) %>% ungroup)) -> ASV.nested

saveRDS(ASV.nested, file = "../Input/Cleaning.before.Occ.model")

ASV.nested <- readRDS(file ="../Input/Cleaning.before.Occ.model")

ASV.nested %>% 
  transmute(Miseq_run, Summary.1 = map(Step2.tibble, ~ how.many(ASVtable = .,round = "3.Positives"))) %>% 
  left_join(ASV.summary) %>% 
  mutate(Summary   = map2(Summary, Summary.1, bind_rows)) %>%
  dplyr::select(-Summary.1) -> ASV.summary

ASV.summary %>% 
  unnest()
```

## Cleaning Process 4: **Occupancy modelling**

What is the probabilty of a true positive presence of a Hash in a Miseq Run. We will use eDNA occupancy modeling to asses whether a hash is a rare variant that spilled out or a true presence.

The process requires to load extra packages, create some model file, and group the hashes by Run, and biological replicate, summarising the data in a presence absence format.

The occupancy model itself was performed in the Rmarkdown file `Rjags.tunning.Rmd`, so here we will upload the csv file that contains all probability of occurences of all hashes per site. Each Hash-Site combination produces a matrix of presence abascences that feeds the model - for some cases it is a 30x3 matrix, for others it is a 39x3. We summarised the number of occurences in each case and run models for each unique case (to save computing time). Each unique model was run 10 times to filter out cases in which the model converge into a local maxima.

So we will import the object `Occ.fate.csv` and reduce the dataset to those Hashes with an occ > 0.8

```{r importing Occ results}

occ.results <- read_csv("../Input/Occ.fate.csv")

occ.results %>% 
  ggplot(aes(x = real)) +
  geom_histogram(fill = "blue") +
  labs(x = "",
       y = "Number of Hashes")  -> left.plot

occ.results %>% 
  left_join(ASV.nested %>% 
              select(Step2.tibble) %>% 
              unnest(cols = Step2.tibble) %>% 
              group_by(Hash) %>% 
              summarise (tot = sum(nReads))) %>% 
  ggplot(aes(x = cut_interval(real, n = 20))) +
  geom_col(aes(y = tot), fill = "red") +
  labs(x = "",
       y = "Number of Reads")   -> right.plot

left.plot + right.plot

```

So we will throw away most of the Hashes, but will keep most of the reads - we are getting into something here

```{r actual filtering}
 occ.results %>% 
  filter(real > 0.8) %>% 
  pull (Hash) -> to.keep

ASV.nested %>% 
  mutate(Step3.tibble = map (Step2.tibble, ~ filter(.,Hash %in% to.keep))) -> ASV.nested

ASV.nested %>% 
  transmute(Miseq_run, Summary.1 = map(Step3.tibble, ~ how.many(ASVtable = .,round = "4.Occupancy"))) %>% 
  left_join(ASV.summary) %>% 
  mutate(Summary   = map2(Summary, Summary.1, bind_rows)) %>%
  dplyr::select(-Summary.1) -> ASV.summary

```

## Cleaning Process 5: **Dissimilarity between PCR replicates**

So, a second way of cleaning the dataset is to remove samples for which the dissimilarity between PCR replicates exceeds the normal distribution of dissimilarities.
Sometimes the preparation of a PCR replicate goes wrong for a number of reasons - that leads to a particular PCR replicate to be substantially different to the other 2. In that case, we will remove the PCR replicate that has higher dissimilarity with the other two.

The process starts by adding the biological information to the ASV table, then diving the dataset by their biological replicate. This will also remove any sample that is not included in the metadata, eg coming from a different project.

```{r dissimilarity between PCR replicates}

ASV.nested %>% 
  select(Miseq_run, Step3.tibble) %>% 
  unnest(Step3.tibble) %>%
  separate(sample, into = "original_sample", sep = "\\.", remove = F) -> cleaned.tibble
```

```{r quick check}
# do all samples have a name
cleaned.tibble %>% 
  filter (sample == "")
# do all of them have an original sample
cleaned.tibble %>% 
  filter(original_sample == "")
# do all of them have a Hash
cleaned.tibble %>% 
  filter(is.na(Hash))
# How many samples, how many Hashes
cleaned.tibble %>% 
  summarise(n_distinct(sample), # 770
            n_distinct(Hash))   # 4571

# Let's check the levels of replication

cleaned.tibble %>% 
  group_by(original_sample) %>% 
  summarise(nrep = n_distinct(sample)) %>% 
  #filter (nrep == 2) # 13
  filter (nrep == 1) # 0 
```
Ok, so there are 13 samples for which we only have 2 PCR replicates1.   We will get rid of those with only 1, as we can't estimate the PCR bias there.

```{r remove single replicates}
discard.1 <- cleaned.tibble %>% 
  group_by(original_sample) %>% 
  mutate(nrep = n_distinct(sample)) %>% 
  #filter (nrep == 2) # 25
  filter (nrep == 1) %>% 
  distinct(sample) %>% pull(sample)

cleaned.tibble %>% 
  filter(!sample %in% discard.1) -> cleaned.tibble
```

Anyway, let's have a visual representation of the dissimilarities between PCR replicates, biological replicates and everything else.

```{r lets do the PCR replication}
cleaned.tibble %>%
  group_by (sample) %>%
  mutate (Tot = sum(nReads),
          Row.sums = nReads / Tot) %>% 
  group_by (Hash) %>%
  mutate (Colmax = max (Row.sums),
          Normalized.reads = Row.sums / Colmax) -> cleaned.tibble
tibble_to_matrix <- function (tb) {

  tb %>% 
    group_by(sample, Hash) %>% 
    summarise(nReads = sum(Normalized.reads)) %>% 
    spread ( key = "Hash", value = "nReads", fill = 0) -> matrix_1
    samples <- pull (matrix_1, sample)
    matrix_1 %>% 
      ungroup() %>% 
    dplyr::select ( - sample) -> matrix_1
    data.matrix(matrix_1) -> matrix_1
    dimnames(matrix_1)[[1]] <- samples
    vegdist(matrix_1) -> matrix_1
}

tibble_to_matrix (cleaned.tibble) -> all.distances.full

#names(all.distances.full)

summary(is.na(names(all.distances.full)))
```

Let's make the pairwaise distances a long table
```{r}

as.tibble(subset(melt(as.matrix(all.distances.full)))) -> all.distances.melted

summary(is.na(all.distances.melted$value))

# Now, create a three variables for all distances, they could be PCR replicates, BIOL replicates, or from the same site

all.distances.melted %>%
  separate (Var1, into = "Bottle1", sep = "\\.", remove = FALSE) %>%
  separate (Bottle1, into = "Site1", remove = FALSE) %>%
  separate (Var2, into ="Bottle2", sep = "\\.", remove = FALSE) %>%
  separate (Bottle2, into = "Site2", remove = FALSE) %>%
  mutate ( Day.site1 = str_sub(Bottle1, start = 1, end = -2),
           Day.site2 = str_sub(Bottle2, start = 1, end = -2),
           Distance.type = case_when( Bottle1 == Bottle2 ~ "PCR.replicates",
                                      Day.site1 == Day.site2 ~ "Biol.replicates",
                                      Site1 == Site2 ~ "Same Site",
                                      TRUE ~ "Different Site"
                                     )) %>%
  dplyr::select(Sample1 = Var1, Sample2 = Var2 , value , Distance.type) %>%
  filter (Sample1 != Sample2) -> all.distances.to.plot

# Checking all went well

sapply(all.distances.to.plot, function(x) summary(is.na(x)))

all.distances.to.plot$Distance.type <- all.distances.to.plot$Distance.type  %>% fct_relevel( "PCR.replicates", "Biol.replicates", "Same Site")

  ggplot (all.distances.to.plot , aes (fill = Distance.type, x = value)) +
  geom_histogram (position = "dodge", stat = 'density', alpha = 0.9) +
 # facet_wrap( ~ Distance.type) +
  labs (x = "Pairwise dissimilarity", y = "density" ,
        Distance.type = "Distance")

```

```{r}
# Instead of chosing based on the pw distances, we can do a similar thing using the distance to centroid

# Find out which samples have only two pcr replicates
cleaned.tibble %>% dplyr::select(-Miseq_run) %>% group_by(original_sample) %>% nest() -> nested.cleaning

nested.cleaning %>% 
  mutate(matrix = map(data, tibble_to_matrix)) -> nested.cleaning
nested.cleaning %>% mutate(ncomparisons = map(matrix, length)) -> nested.cleaning

dist_to_centroid <- function (x,y) {
  biol <- rep(y, length(x))

  if (length(biol) == 1) {
    output = rep(x[1]/2,2)
    names(output) <- attr(x, "Labels")
  }else{

  dispersion <- betadisper(x, group = biol)
  output = dispersion$distances
  }
  output
    }

nested.cleaning <- nested.cleaning %>% mutate (distances = map2(matrix, original_sample, dist_to_centroid))

unlist (nested.cleaning$distances) -> all_distances

```

```{r}
#normparams <- fitdistr(all_pairwise_distances, "normal")$estimate
normparams <- MASS::fitdistr(all_distances, "normal")$estimate
#  probs <- pnorm(all_pairwise_distances, normparams[1], normparams[2])
probs <- pnorm(all_distances, normparams[1], normparams[2])
outliers <- which(probs>0.95)

discard <-names (all_distances[outliers])

to_write_discarded <- as.tibble(all_distances[outliers]) %>% rownames_to_column("sample") %>% dplyr::select(sample, 
                                                                                                     distance_to_centroid = value)
to_write_discarded <- to_write_discarded %>% bind_rows(tibble(sample = discard.1,
                                                              distance_to_centroid = NA))
write_csv(to_write_discarded ,"../Output/discared_samples.csv")

# Who passes this filter

all_distances %>%
  as.tibble() %>% 
  mutate(sample = names(all_distances)) %>% 
  filter(!sample %in% to_write_discarded$sample) %>% 
  separate(sample, into = "event", sep = -3) %>% 
  group_by(event) %>% 
  summarise(cases = n()) %>% 
  separate(event, into = c("Site", "Date"), remove = F) %>% 
  filter(Date != "201703") %>% 
  ggplot()+
  geom_raster(aes(x= Date, y = Site, fill = cases))+
  geom_text(aes(x= Date, y = Site, label = cases))

```

Finally, let's remove these samples from the dataset

```{r actual cleaning}

ASV.nested %>% 
  mutate(Step4.tibble = map (Step3.tibble,  ~ filter(.,! sample %in% to_write_discarded$sample))) -> ASV.nested

ASV.nested %>% 
  transmute(Miseq_run, Summary.1 = map(Step4.tibble, ~ how.many(ASVtable = .,round = "5.PCR.dissimilarity"))) %>% 
  left_join(ASV.summary) %>% 
  mutate(Summary   = map2(Summary, Summary.1, bind_rows)) %>%
  dplyr::select(-Summary.1) -> ASV.summary 
```

## Exporting the output

We will export the final cleaned table with four columns (Miseq_run, sample, Hash, nReads)

```{r}

ASV.nested %>% 
  select(Miseq_run, Step4.tibble) %>% 
  unnest(Step4.tibble) %>% 
  mutate(nReads = as.integer(nReads)) %>% 
  write_csv("../Output/ASV_table_all_together.csv")

ASV.nested %>% 
  select(Step4.tibble) %>% 
  unnest(Step4.tibble) %>% 
  distinct(Hash) %>% 
  left_join(Hash.key) %>% 
  write_csv("../Output/Hash_Key_all_together.csv")

input <- read_csv("../Output/Hash_Key_all_together.csv")
output <- "../Output/Hash_Key_all_together.csv"

write.fasta (sequences = as.list(input$Sequence),
             names = as.list(input$Hash),
             file.out = output)

```

## Summary of the cleanup process

```{r last graph}

ASV.summary %>% 
  unnest() %>% 
  ggplot(aes(x=Stage, y=number, fill = Stat))+
    geom_line(aes(group = Miseq_run, color = Miseq_run))+
  facet_grid(Stat~., scales = "free")+
  theme(axis.text.x = element_text(angle = 45, hjust =1))#,

```

### 2019-09-20 Input/1.Sequencing_runs_output/20171122.Jul_Aug/banzai_params_for_DADA2_Jul_Aug.sh

@@ -1,0 +1,266 @@
+#!/usr/bin/env bash
+
+
+################################################################################
+# INPUT
+################################################################################
+# What is the file path to the directory containing all of the libraries/reads?
+PARENT_DIR="/Users/Moncho/Google_Drive/Run_Nov17/OA_COI/171122"
+
+# Where is the sequencing metadata file? (SEE FORMATTING GUIDELINES IN README!)
+SEQUENCING_METADATA="${PARENT_DIR}"/metadata.csv
+
+
+################################################################################
+# OUTPUT
+################################################################################
+# This script will generate a directory (folder) containing the output of the script.
+# Where do you want this new folder to go?
+OUTPUT_DIRECTORY="${HOME}"/fastqs_demultiplexed_for_DADA2 #"${PARENT_DIR%/*}"
+
+# You can optionally specify a folder into which the script copies a PDF containing some results.
+# The pdf is created by default in the analysis folder specified above, but
+# if you set this to your DropBox or Google Drive Folder, you can check it out from anywhere.
+OUTPUT_PDF_DIR=""
+
+
+################################################################################
+# METADATA DETAILS
+################################################################################
+# Specify columns for raw sequencing files:
+COLNAME_FILE1="file1"
+COLNAME_FILE2="file2"
+
+# MUST be unique for each row!
+COLNAME_SAMPLE_ID="sample_id"
+
+COLNAME_DNA_ID="dna_id"
+COLNAME_SAMPLE_TYPE="dna_source"
+
+# Your metadata must have a column corresponding to the subfolders containing the raw reads.
+# In order to make this flexible across both multiple and single library preps, you must include this even if you only sequenced one library (sorry!).
+COLNAME_ID1_NAME="pri_index_name"
+COLNAME_ID1_SEQ="pri_index_seq"
+
+################################################################################
+# MERGE PAIRED READS
+################################################################################
+USE_PEAR_DEFAULTS="NO"
+# For more information on these parameters, type into a terminal window: pear -help
+# Bokulich recommends:
+# Quality_Threshold=3, r=3 (PEAR only considers r=2), UNCALLEDMAX=0
+# TRIMMIN= 0.75 * LENGTH_READ # this is hard-coded in the script banzai.sh
+
+# TODO grab this from a fragment_size column in the sequencing metadata file
+### ***** REMEMBER TO WATCH FOR ZEROS WHEN IMPLEMENTING THIS!
+# Is there a column in the metadata file for fragment size?
+insert_size_in_metadata="NO"
[…]

### 2019-11-07 Scripts/All.figures.that.go.in.the.MS.Rmd
Script for plots

---
title: "All_Figures_that_go_on_MS"
author: "Ramon Gallego"
date: "10/30/2019"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro

This Markdown will make the part of the analysis that goes in the Manuscript

```{r libraries and custom functions, echo = F, message=F}
library (RColorBrewer)
library (rstanarm)
library (fitdistrplus)
library (ggridges)
library (ggforce)
library (tidyverse)
library (cowplot)
library (concaveman)
library (vegan)
library (lubridate)
library (ggrepel)
library (factoextra)
library(nnet)
#library (OneR)
#functions for future Salish Sea
models <- readRDS("../../Analysis/Splitting.data.by.lifestyle/ModelOutputs/full_model_output_rpk_TpH.standardized_2019-08-13_1567.RDS")
source("RPK_climate_model_functions.R")

source("../../tibble.to.matrix.r")
```

## Datasets to use

To begin with: the ASV.by.taxa dataset and the environmental data (Updated salinity, DIC and pH)
### Present day datasets
```{r }
data.plankton <- read_csv("../../Data/Combined_Biol_Env_Plankton.csv")

data.plankton %>% filter(!str_detect(taxa, "Phasianidae")) %>% 
  filter (!str_detect(taxa, "Rhodomelaceae")) -> data.plankton

env.data <- read_csv("../../Analysis/Splitting.data.by.lifestyle/env.data.for.log.regression.csv")

#quality-control data
    env.data %>%
      filter(pH > 7) %>%
      mutate(
        Season = case_when(
          str_detect(sample, "1703|1710|1711|1801|1803") ~ "Winter",
          TRUE ~ "Summer"
        ),
        Area = case_when(
          str_detect(sample, "CP|LK|FH") ~ "SJI",
          TRUE ~ "Hood Canal"
        )
      ) -> now.env.data

now.env.data %>%
  gather(-Season, -sample, -Area, key = "Predictor", value = "value") %>%
  group_by(Predictor) %>%   #RPK: this was missing, which messed up the std calculations 
  mutate(std = vegan::decostand(value, method = "standardize")) -> now.env.data.with.std

# Dataset to go between standarized and real values
now.env.data.with.std %>% 
  group_by(Predictor) %>% 
  summarise(mean = mean(value),
            sd = sd(value)) %>% 
  group_by(Predictor) %>% 
  nest(params = c(mean, sd) )-> transforming.df

phylum.info <- read_csv("../../Supp.Material.OA.eDNA/Input/higher_taxonomy.csv") %>% 
  unite(family, genus, species, sep = "|", col = "taxa")

```
### Future scenarios for the Puget Sound

```{r}
#read in observed data, which will be the starting point for adjusting in future projections          
obsData <- models$data[[1]] %>% 
  dplyr::select(Temperature, pH, Area, Season)

YearSpan <- c(2017, 2095)#seq.int(2015, 2095, 20)  #taking these individual years to model
Nyears <- length(YearSpan)
NiterationsPerYear <- 50  #number of realizations of the future environmental conditions
Nreps <- Nyears*NiterationsPerYear

#here, replicate the observed dataframe Nreps times, where Nreps = Nyears * NiterationsPerYear
simData <- obsData[rep(row.names(obsData), Nreps), 1:4] %>% 
  mutate(Year = rep(YearSpan, 
                    each = nrow(obsData)*NiterationsPerYear),
         Iteration = rep(1:NiterationsPerYear, times = nrow(obsData)*Nyears))

```

function to generate future climate distributions isn't vectorized, so now adjust Temperature and pH for each year in a loop:
these use stochastic sampling functions r_T_year_STD() and r_pH_year_STD()

```{r}
set.seed(100)
      for (i in unique(simData$Year)){
        for (j in 1:NiterationsPerYear) {
          for (k in unique(simData$Season)) {
            for (m in unique(simData$Area)) {

              dataSubset <- simData$Temperature[simData$Year == i & 
                                                  simData$Iteration == j &
                                                  simData$Season == k &
                                                  simData$Area == m]

        simData$Temperature[simData$Year == i & 
                              simData$Iteration == j &
                                simData$Season == k &
                                  simData$Area == m] <-     #future condition is ....

                                             dataSubset +    #the current condition, plus...

                      r_T_year_STD_subset(i,  #the difference seen in the specified year
                                          length(dataSubset),  
                                          2017, # standardized to reference year
                                          season = k,
                                          area = m)

        simData$pH[simData$Year == i & 
                              simData$Iteration == j &
                              simData$Season == k &
                              simData$Area == m] <-     #future condition is ....
          simData$pH[simData$Year == i & 
                                simData$Iteration == j &
                                simData$Season == k &
                                simData$Area == m] +    #current condition, plus...

          r_pH_year_STD_subset(i,  #the difference seen in the specified year
                              10,  
                              2017, # standardized to reference year
                              season = k,
                              area = m)[1] #one draw per iteration for the simulated set of temperatures
          } 
         }
        }
      }

```

### Convert standardized units to natural units using reference year

```{r}
#observed = (std * SD) + mean

#useful later:
reference_Temperature_SD <- sd(now.env.data$Temperature)
reference_Temperature_mean <- mean(now.env.data$Temperature)
reference_pH_SD <- sd(now.env.data$pH)
reference_pH_mean <- mean(now.env.data$pH)

simData <- simData %>% 
  rename(TemperatureSTD = Temperature,
        pH_STD = pH) %>% 
  mutate(TemperatureCelsius = (TemperatureSTD * reference_Temperature_SD) + reference_Temperature_mean,
         pH_units = (pH_STD * reference_pH_SD) + reference_pH_mean)

#check that this is perfectly linear
# simData %>%
#   ggplot(aes(x = TemperatureSTD, y = TemperatureCelsius)) +
#   geom_point()

#check for trends
# simData %>% 
#   select(Year, Season, Area, TemperatureCelsius) %>% 
#   group_by(Year, Season, Area) %>% 
#   summarize(mean(TemperatureCelsius))
```
If we look at this data, does it have crazy points
```{r limits}
simData %>% 
  group_by(Year, Season, Area) %>% 
  summarise_at(c("TemperatureCelsius", "pH_units"), .funs = list(min = min, max = max))

# Yep, a bit crazy. Apply 90% interval

simData %>%
  group_by(Year, Area) %>% 
  nest() %>% 
  mutate(
    values = map (data, function (.x) {
      list(temp = fitdistr(.x$TemperatureCelsius ,
             densfun = "normal"),
           ph = fitdistr(.x$pH_units ,
             densfun = "normal"))
            }
          ),

    data = map2(data,values,  function (.x, .y){
      .x %>%
        mutate(pT = pnorm(TemperatureCelsius, mean  = .y[["temp"]]$estimate[1], sd = .y[["temp"]]$estimate[2]),
               ppH =pnorm(pH_units, mean  = .y[["ph"]]$estimate[1], sd = .y[["ph"]]$estimate[2]) )

    })
    ) %>% 
  unnest(data) %>% 
  mutate(keep = case_when(pT < 0.1 | ppH < 0.05 | pT > 0.9 | ppH > 0.9 ~ "Discard",
                          TRUE                   ~ "keep")) %>% 
  filter (keep == "keep")-> simData.nocrazys
simData.nocrazys %>%
  rename(Temperature = TemperatureSTD,
         pH = pH_STD) %>% 
saveRDS( file = "../../Analysis/Splitting.data.by.lifestyle/all.sim.data_climateScenarios_no.outliers.RDS")
simData.nocrazys %>% 
  group_by( Area, Season) %>% 
  summarise_at(c("TemperatureCelsius", "pH_units"), .funs = list(min = min, max = max))

```

## Figure 1

We have the datasets and some plots already made -

### Panel A Sample location
```{r Base map}
map <- readRDS("map.plot.rds")

map + ggtitle(label = element_blank()) + labs (x = "Longitude", y = "Latitude") -> map

```
### Panel B Environmental variation

Make the ridges plot

```{r}
now.env.data %>% 
  select(TA = mean.TA, DIC = mean.DIC, everything(), -contains("%")) %>% 
  filter(!is.na(TA)) %>% 
  separate(sample, into= c("Site","Date"), remove = F )-> env.data.no.outliers

env.data.no.outliers %>% 
  pivot_longer(cols = c(-Area, -Site, -Date, -Season, -sample), names_to =   "Predictor", values_to  = "value") -> env.data.to.keep

env.data.to.keep %>% 
  mutate(Site = fct_relevel(Site, "SA", "TR", "LL","PO", "TW"),
         Predictor = fct_relevel(Predictor, "DIC", "TA", "pH")) -> env.data.to.keep

env.data.to.keep %>% 
  mutate(Area = case_when (Area == "SJI" ~ "San Juan Island",
                           TRUE          ~ Area)) %>% 
  mutate(Month = month(ymd(paste0(Date, "01")), label = T, abbr = F)) %>% 
  filter (Predictor %in% c("DIC", "TA", "pH", "Temperature", "Salinity")) %>% 
  ggplot(aes(x = `value`, y = `Month`, fill = Area)) +
  geom_density_ridges_gradient() +
 # scale_fill_viridis(name = "Temp. [F]", option = "C") +
  facet_wrap(~fct_relevel(Predictor, "DIC", "TA", "pH", "Temperature", "Salinity"), scales = "free", nrow = 1) +
  theme(legend.direction ="horizontal") -> ridges

legend <- get_legend(
  # create some space to the left of the legend
  ridges + theme(legend.box.margin = margin(0, 0, 0, 12))
)
 legend
ridges + guides(fill = "none") -> ridges
```

### Panel c: Richness variation

```{r}
boxplot.richness <- readRDS("richness.rds")

boxplot.richness +
  ylim (0,60) -> boxplot.richness
```
### All together

THe top will be the environmental gradients, the lower panels will have the map and the boxplot
```{r}
plot_grid(plotlist = list(map, legend, boxplot.richness),
                     align = "v",
                     ncol = 3, 
          axis = "t",
          rel_widths = c(1,0.5,0.6))   -> bottom.half.plot1

plot_grid(plotlist = list(ridges, bottom.half.plot1),
          align = "hv",
          ncol = 1, 
          nrow = 2,
          rel_heights = c(1,0.8),
          rel_widths = 1)

plot_grid(plotlist = list(legend, boxplot.richness),
                     align = "hv",
                     ncol = 1, 
          axis = "t",
          rel_heights  = c(.5,1),
          labels = c(NA,"B" )) -> bottom.right

plot_grid (plotlist = list(map, bottom.right),
           align = "hv",
           nrow = 1,
           rel_heights = 1,
           rel_widths = c(.6,0.4),
           labels = c("A", NA)) -> bottom.plot

plot_grid(bottom.plot, ridges,
          align = "h",
          nrow = 2,
          rel_heights = c(0.6, 0.4),
          labels = c(NA,"C")) +
  ggsave(filename = "Fig1.png", width = 14, height = 14, dpi = "retina")

```

## Figure 2

```{r functions for CAPS}
cap.functions.only3things <- function(tibble, ...){ # ..  are the grouping factors

  tibble %>%

    group_by(...) %>%

    nest() %>%

    mutate(comm.matrix = map (data, ~ tibble.to.table(.x, taxon = taxa, Abundance = Normalized.reads, sample.name = event)),

           ja = map(data, ~tibble_to_matrix(.x, taxon = taxa,Abundance = Normalized.reads, sample.name = event, distance = "jaccard",transformation = "binary" )),
          bc = map(data, ~tibble_to_matrix(.x, taxon = taxa,Abundance = Normalized.reads, sample.name = event, distance = "bray",transformation = "" )), # the bc distamce
          env = map (data, ~tibble_to_env(.x, taxon = taxa, Abundance = Normalized.reads, sample.name = event, everything()) %>% 
                       select(-contains("%")) %>% 
                       separate(event, into = c("Site", "Date"), sep = "_")))
}

raw.PCAs <- function(nested.tibble){

  nested.tibble %>% 
    mutate(

  CAP.analysys.raw = map2(comm.matrix, env,

                          function(.x,.y) {
                            .y %>% select (Temperature, Salinity = Salinity.new, pH = pH_new) 
                            caps <-  capscale(formula = .x ~ Temperature + Salinity + pH ,
                                                 data = .y %>%
                                                   select(DIC = mean.DIC.new.sal, Salinity = Salinity.new, Temperature, pH = pH_new) %>% 
                                                   select_if(is.numeric) %>%
                                                   decostand(method = "standardize", na.action = na.omit),
                                                 distance = "jaccard", 
                                                binary = T)

                            sppscores(caps) <- sqrt(decostand(.x, "total"))

                            return(caps) }),

  CAP.plot = map (CAP.analysys.raw, ~ plot(.x)),

  CAP.signficance = map(CAP.analysys.raw, ~ permutest(.x, permutations = 999 )),

  gp.plot = map2 (CAP.plot, env, function (.x, .y){

                         .x$sites %>%
                          as.data.frame() %>%
                          rownames_to_column("sample") %>%

                          separate(sample, into = c("Site", "Date"), sep = "_",remove = F) %>%
                          mutate(Date = ymd(Date),
                                 Date2 = paste(month(Date, label = T, abbr = T), year(Date), sep = "' "),
                                 Month = month(Date, label = T)) -> site.scores # pull the  site scores

                        site.scores %>% 
                          mutate(Date = as.character(Date)) %>% 
                          left_join(.y) -> site.scores.plot

                        scalate.corrs <- c(CAP1max =  max(abs(site.scores$CAP1)), CAP2max =  max(abs(site.scores$CAP2)))

                        .x$species %>%
                        as.data.frame() %>%
                        rownames_to_column("taxa") %>%
                          separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
                          mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
                                                   species == "NA" ~  paste0(genus, " sp."),
                                                   TRUE            ~ species),
                                 scaledCAP1 = CAP1 * scalate.corrs["CAP1max"],
                                 scaledMDS1 = CAP2 * scalate.corrs["CAP2max"]) %>%
                          mutate (rate = rank(abs(CAP1)),
                                  corr = sqrt(CAP1^2 + CAP2^2)) %>%
                         arrange (desc(rate)) %>%
                          slice (1:10) -> species.scores # Keep only the top 20 matches

                       .x$biplot %>%
                         as.data.frame() %>%
                        rownames_to_column("env.variable") -> var.scores

                        ggplot (aes (x= CAP1, y = CAP2), data = site.scores.plot) +
                          geom_point(aes(
                            color = pH < 8 )) +
                          geom_segment(aes(x = 0, y = 0,
                                           xend = CAP1,
                                           yend = CAP2), data = var.scores, color = "blue", arrow = arrow(length = unit(0.1,"cm"))) +
                          geom_label_repel(aes(x= CAP1  ,
                                         y= CAP2 ,
                                         label = env.variable), data = var.scores, fill = "pink", alpha = 0.75) +

                          guides(color = guide_legend(title = "pH < 8")) -> p # The basic plot

                        p +

                          geom_segment(aes(x = 0, y = 0,
                                           xend = CAP1 * scalate.corrs["CAP1max"] ,
                                           #yend = 0),
                                           yend = CAP2 * scalate.corrs["CAP2max"]), 
                                       data = species.scores, color = "red", arrow = arrow(length = unit(0.1,"cm"))) +

                          geom_label_repel(aes(x= CAP1 * scalate.corrs["CAP1max"] ,
                                         #      y = 0,
                                         y= CAP2 * scalate.corrs["CAP2max"],
                                         label = label), data = species.scores, fill = "lightblue", alpha = 0.75, size = 1.5) -> p.with.spp # The species corr
                          # ggsave("CAP_with_species.png",
                          #        dpi = "retina",
                          #        width = 7) +
                          #ggedit::remove_geom('label',) +

                        return (list (basic = p, spp = p.with.spp))
                       #return(bind_rows(site.scores, var.scores, species.scores ))
                        } ) ,

  nclusters = map (CAP.plot, ~  fviz_nbclust( .x$sites, kmeans, k.max = 5, method = "silhouette")),

  n_clust = c(4,3,3,3),

  CAP.clusters = map2 (CAP.plot,n_clust,  function (.x, .y) {

    final <- kmeans(.x$sites, .y, nstart = 15)

    fviz_cluster(final, data = .x$sites, geom = "point", palette = "Set2")

  } ),

  Cluster.df = map2 (CAP.plot,n_clust,  function (.x, .y) {

    final <- kmeans(.x$sites, .y, nstart = 15)

    tibble(event = names(final$cluster),
           cluster = final$cluster) }

    ),

  Final.plot = map2(CAP.plot, CAP.clusters, function (.x, .y) {

                        .x$biplot %>%
                         as.data.frame() %>%
                        rownames_to_column("env.variable") -> var.scores

                          .x$sites %>%
                          as.data.frame() %>%
                          rownames_to_column("sample") %>%
                            separate(sample, into = c("Site", "Date"), sep = "_",remove = F) %>%
                            mutate(Date = ymd(Date),
                                 Date2 = paste(month(Date, label = T, abbr = T), year(Date), sep = "' "),
                                 Month = month(Date, label = T)) -> site.scores

    .y +
      #geom_label(aes(x= CAP1, y = CAP2, label = Month), data = site.scores) +
      geom_segment(aes(x = 0, y = 0,
                       xend = CAP1,
                       yend = CAP2), data = var.scores, color = "blue", arrow = arrow(length = unit(0.1,"cm"))) +
      geom_label_repel(aes(x= CAP1  ,
                           y= CAP2 ,
                           label = env.variable), data = var.scores, fill = "pink", alpha = 0.75)

  })

    )}

```

We will do the Full CAP analysis - but will only keep Hood Canal in summer for the MS - the rest will have to go to the Supplementary information

The figure will have three panels: the CAP for HC-summer , the probability of presence of each cluster in the pH/Temperature space

```{r CAP analysis FULL}

data.plankton %>% 
  separate (event, into = c("Site", "Date"), sep = "_") %>% 
  mutate(Season = case_when(str_detect(Date,"17-03|17-10|17-11|18-01|18-03") ~ "Winter",
                             TRUE ~                                         "Summer"),
         Area = case_when(Site %in% c("CP","LK", "FH") ~ "San Juan Island",
                          TRUE                         ~ "Hood Canal")) %>% 
   unite (Site, Date, col = "event", sep = "_") %>% 
  cap.functions.only3things(Area, Season) -> joined.dataset.for.CAP
```
Use the functions
```{r lets CAP this}
joined.dataset.for.CAP %>% 
 # select(DIC = mean.DIC.new.sal, Salinity = Salinity.new, Temperature, pH = pH_new, everything())
  raw.PCAs  %>% 
  mutate(Titles = paste(Area,Season, sep = " - " )) -> Clusters.analysis.and.cap
```

```{r get the plots}

Clusters.analysis.and.cap %>% 
  select(Titles, Final.plot) %>% 
  mutate(Final.plot = map2(Titles, Final.plot, ~ .y + ggtitle(.x) )) %>% 
  pull(Final.plot) -> list.of.cluster.plots

names(list.of.cluster.plots) <- Clusters.analysis.and.cap$Titles

# We keep only the Hood Canal in Summer - we are using _scale_color_brewer set 2

list.of.cluster.plots[[ "Hood Canal - Summer"]] -> left.most.plot

left.most.plot + ggtitle ("") +coord_equal() + scale_color_brewer(palette = "Set2")-> left.most.plot
#+ theme_cowplot() + 
left.most.plot + theme(legend.position = "bottom")  -> left.most.plot

legend <- get_legend(left.most.plot)

left.most.plot + theme(legend.position = "none") -> left.most.plot

Clusters.analysis.and.cap %>% 
  mutate(full.cluster.info = map2(env, Cluster.df, ~ .x %>% unite(Site, Date, col = "event") %>%right_join(.y) ),
         full.cluster.info = map2(full.cluster.info,CAP.plot, ~ .y$sites %>% 
                                    as.data.frame() %>% 
                                    rownames_to_column("event") %>% 
                                    right_join(.x) %>% 
                                    mutate(pH = round(pH,1))),
         Final.plot2 = map2(Final.plot,full.cluster.info, ~ .x +
                             geom_label_repel(data = .y, 
                                              aes(x = CAP1, y = CAP2, label = cluster)))) -> Clusters.analysis.and.cap

Clusters.analysis.and.cap %>% 
  select(Area, Season, full.cluster.info) %>% 
  unnest() %>% #filter (Area == "Hood Canal", Season == "Summer") %>% 
  write_csv("../../Analysis/Splitting.data.by.lifestyle/clusters.of.events.csv")
Clusters.analysis.and.cap %>% pull(Final.plot2) 
```

```{r capture legend}

Clusters.analysis.and.cap %>% 
  slice(3) %>% 
  mutate(Hand.made.CAP = map2(CAP.plot, full.cluster.info, function (.x, .y){

    .x$biplot %>%
                         as.data.frame() %>%
                        rownames_to_column("env.variable") -> var.scores

    # return(.y %>% 
    #   select(event, CAP1, CAP2, cluster))
    .y %>% 
      select(event, CAP1, CAP2, cluster) %>% 
      ggplot(aes(x = CAP1,
                 y = CAP2)) +
      ggforce::geom_mark_hull(aes( group= cluster, color = as.factor(cluster),label = as.factor(cluster), fill = as.factor(cluster)),
                          expand = unit(1, "mm"),
                 label.margin = margin(1, 1, 1, 1, "mm"),
                 label.buffer = unit(0, "mm"),
                 con.cap = 1 ,
                 alpha = 0.5) +

      geom_point(size = 1.5) +
       geom_point(aes(color = as.factor(cluster)), size = 1) +
    #  geom_label(aes(x= CAP1, y = CAP2, label = Month), data = site.scores) +
      geom_segment(aes(x = 0, y = 0,
                       xend = CAP1,
                       yend = CAP2), data = var.scores, color = "blue", arrow = arrow(length = unit(0.1,"cm"))) +
      geom_label_repel(aes(x= CAP1  ,
                           y= CAP2 ,
                           label = env.variable), data = var.scores, fill = "pink", alpha = 0.75) +
      ggtitle ("") + coord_equal() + scale_color_brewer(name = "Cluster", palette = "Set2") + scale_fill_brewer(name = "Cluster",palette = "Set2") +
     # guides(color ="Cluster" ) +
      theme(legend.position = "bottom")

  })) %>% pull(Hand.made.CAP) -> leftplot.Fig2

legend <- get_legend(leftplot.Fig2[[1]])
leftplot.Fig2 [[1]]+  theme(legend.position = "none") -> leftplot.Fig2

```

###Expand the plots to all seasons and areas for supplementary material
```{r}
Clusters.analysis.and.cap %>%

  mutate(Hand.made.CAP = map2(CAP.plot, full.cluster.info, function (.x, .y){

    .x$biplot %>%
                         as.data.frame() %>%
                        rownames_to_column("env.variable") -> var.scores

    # return(.y %>% 
    #   select(event, CAP1, CAP2, cluster))
    .y %>% 
      select(event, CAP1, CAP2, cluster) %>% 
      ggplot(aes(x = CAP1,
                 y = CAP2)) +
       geom_mark_ellipse(aes(fill = as.factor(cluster))) +
      # ggforce::geom_mark_hull(aes( group= cluster, color = as.factor(cluster),label = as.factor(cluster), fill = as.factor(cluster)),
      #                     expand = unit(1, "mm"),
      #            label.margin = margin(1, 1, 1, 1, "mm"),
      #            label.buffer = unit(0, "mm"),
      #            con.cap = 1 ,
      #            alpha = 0.5) +

      geom_point(size = 1.5) +
       geom_point(aes(color = as.factor(cluster)), size = 1) +
    #  geom_label(aes(x= CAP1, y = CAP2, label = Month), data = site.scores) +
      geom_segment(aes(x = 0, y = 0,
                       xend = CAP1,
                       yend = CAP2), data = var.scores, color = "blue", arrow = arrow(length = unit(0.1,"cm"))) +
      geom_label_repel(aes(x= CAP1  ,
                           y= CAP2 ,
                           label = env.variable), data = var.scores, fill = "pink", alpha = 0.75) +
      ggtitle ("") + coord_equal() + scale_color_brewer(name = "Cluster", palette = "Set2") + scale_fill_brewer(name = "Cluster",palette = "Set2") +
     # guides(color ="Cluster" ) +
      theme(legend.position = "bottom")

  })) %>% pull(Hand.made.CAP) -> list.of.left.plots
```

### Panel B: Get bits of script from RPK

```{r}

  TpH_likelihood <- function(focalArea, focalYear, focalSeason, minT, pHmax) {
    require(tidyverse)

    all.sim.data %>% 
      filter(Area == focalArea,
             Year == focalYear,
             Season == focalSeason) %>% 
      group_by(TemperatureCelsius > minT & pH_units < pHmax) %>% 
      summarize(NumberObs = n()) %>% 
      summarize(NumberObs[2]/(NumberObs[1] + NumberObs[2])) %>% 
      as.numeric()
  }
```
Load the objects
```{r}
all.sim.data <- simData.nocrazys  #to see simulation data

clusters <- read.csv("../../Analysis/Splitting.data.by.lifestyle/clusters.of.events.csv") %>% 
  filter(Season == "Summer",
         Area == "Hood Canal")
  clusters$cluster <- as.factor(clusters$cluster)
```

```{r models and so for getting the communities}
multi.mod <- multinom(cluster ~ Temperature + pH_new, data = clusters)
  z <- summary(multi.mod)$coefficients/summary(multi.mod)$standard.errors
  p <- (1 - pnorm(abs(z), 0, 1)) * 2
  p  #T and pH signif predictors for all conditions

  #predict(multi.mod, type = "probs") #get probs for each cluster at each set of enviro conditions

  #at median conditions for HC Summer 2015
  predict(multi.mod, 
          type = "probs", 
          newdata = data.frame(Temperature = median(clusters$Temperature),
                       pH_new = median(clusters$pH_new))
  )

  #in scenario w Temp = 19, pH = 7.9, the low-pH community is nearly certain under our model
  predict(multi.mod, 
          type = "probs", 
          newdata = data.frame(Temperature = 19,
                               pH_new = 7.9)
  )

#build a heatmap; likelihood of low-pH / high-T community under different parameter sets

pH.range <- seq(7.5, 8.5, length.out = 12)
T.range <- seq(12, 27, length.out = 22)

#build heatmap of frequency of these climate thresholds
        threshold.res <- as.data.frame(matrix(NA, nrow = length(pH.range), ncol = length(T.range)))
        for (i in 1:length(pH.range)){
          for (j in 1:length(T.range)) {

            threshold.res[i,j] <-    predict(multi.mod, 
                    type = "probs", 
                    newdata = data.frame(Temperature = T.range[j],
                                         pH_new = pH.range[i])
            )[2]  #second element is the low-pH/high-T community
          }
        }
        threshold.res <- round(threshold.res, 3)

        row.names(threshold.res) <- paste0("pH_",pH.range)
        colnames(threshold.res) <- paste0("T_",T.range)

whichComm.res <- as.data.frame(matrix(NA, nrow = length(pH.range), ncol = length(T.range)))
          for (i in 1:length(pH.range)) {
            for (j in 1:length(T.range)) {
              whichComm.res[i, j] <- as.numeric(
                which.max(
                  predict(multi.mod,
                    type = "probs",
                    newdata = data.frame(
                      Temperature = T.range[j],
                      pH_new = pH.range[i]
                    )
                  )
                )
              )
            }
          }

        row.names(whichComm.res) <- paste0("pH_",pH.range)
        colnames(whichComm.res) <- paste0("T_",T.range)

```

And the plot

```{r}

 whichComm.res %>% 
          rownames_to_column("pH") %>% 
          gather(key = "Temperature", value = "Community", -pH) %>% 
          separate(pH, into = c("p", "pHvalue"), sep = "_") %>% 
          separate(Temperature, into = c("Td", "Tvalue")) %>% 
          dplyr::select(-p, -Td) %>% 
          mutate(pHvalue = as.numeric(pHvalue),
                 Tvalue = as.numeric(Tvalue)) %>% 
  group_by(Community) %>% 
  summarise_all(mean) -> centroids.of.comms

 whichComm.res %>% 
          rownames_to_column("pH") %>% 
          gather(key = "Temperature", value = "Community", -pH) %>% 
          separate(pH, into = c("p", "pHvalue"), sep = "_") %>% 
          separate(Temperature, into = c("Td", "Tvalue")) %>% 
          dplyr::select(-p, -Td) %>% 
          mutate(pHvalue = as.numeric(pHvalue),
                 Tvalue = as.numeric(Tvalue)) %>% 
          ggplot(aes(y = pHvalue, x = Tvalue, fill = as.factor(Community))) + 
          geom_tile() +
          geom_label(data = centroids.of.comms,
                     aes (x = Tvalue,
                          y = pHvalue,
                          label = Community), color = "white") +

          scale_fill_brewer(palette = "Set2") +
          xlab("Temperature") + ylab("pH") +
          guides(fill = "none") -> right_top
          #ggtitle("Most-Likely Community")+ 
          #theme_cowplot()

# right_top + coord_equal () -> right_top 
 ggsave("Figure2.topright.png", width = 7)
```

### Repeat the above for all seasons and areas

```{r}
all.sim.data <- simData.nocrazys  #to see simulation data

clusters.all <- read.csv("../../Analysis/Splitting.data.by.lifestyle/clusters.of.events.csv")

  clusters.all$cluster <- as.factor(clusters.all$cluster)
```

Models

```{r}
simData.nocrazys %>% 
  mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
                          TRUE          ~ Area)) %>% 
  group_by( Area, Season) %>% 
  summarise_at(c("TemperatureCelsius", "pH_units"), .funs = list(min = ~ min(round(.x, 1)), max = ~max(round(.x, 1)))) %>% 
  group_by (Season, Area) %>% 
  nest() %>% 
  mutate(data = map (data, ~ expand_grid(Temperature = seq(round(.x$TemperatureCelsius_min,1), round(.x$TemperatureCelsius_max, 1),by = 0.1),
                                         pH_new = seq(round(.x$pH_units_min,1), round(.x$pH_units_max,1), by = 0.1)))) -> limits.for.planb

Clusters.analysis.and.cap %>% 
  select(Area, Season, full.cluster.info) %>% 
  left_join(limits.for.planb) %>% 
  mutate(multimod = map (full.cluster.info, ~ .x %>% 
                           mutate(cluster = as.factor(cluster)) %>% 
                           multinom(cluster ~ Temperature + pH_new, data = . )),
         z = map (multimod, ~ summary(.x)$coefficients / summary(.x)$standard.errors),
         p = map (z, ~ (1 - pnorm(abs(.x), 0, 1)) * 2),
         threshold.res = map2(multimod, data, ~ predict (.x, 
                                                         type = "probs",
                                                         newdata = .y) %>% 
                                as.data.frame() %>% 
                                bind_cols(.y) %>% 
                                pivot_longer(cols = c(-Temperature, -pH_new),
                                             names_to = "Cluster",
                                             values_to = "p") %>% 
                                group_by(Temperature, pH_new) %>% 
                                arrange(desc(p)) %>% 
                                slice(1)),
         plot = map (threshold.res, ~.x %>% 
                       ggplot(aes(y = pH_new, x = Temperature, fill = as.factor(Cluster))) + 
          geom_tile() +
          scale_fill_brewer(name = "Cluster",palette = "Set2") +
          geom_label(data = .x %>% group_by(Cluster) %>% summarise_all(mean),
                     aes (x = Temperature,
                          y = pH_new,
                          label = Cluster), color = "white"))) %>% pull(plot) -> list.of.mid.plots

```

### Panel C: For each cluster, follow the relative abundance of the taxa driving the differences

Generate the SIMPER clusters
```{r}
Clusters.analysis.and.cap %>% 
  mutate(SIMPER.clusters = map2(comm.matrix, Cluster.df, function(.x, .y) {

    simper(comm = .x,
          group = as.factor(.y$cluster),
          permutations = 999)

  })) %>% 
  select(Area, Season, SIMPER.clusters) -> SIMPER.clusters

SIMPER.clusters %>% 
  mutate(summary.simper = map(SIMPER.clusters, ~ summary(.x, ordered  = TRUE) )) %>% 
  mutate(taxa.simper = map(summary.simper, ~ map(.,function(x) rownames_to_column(x, var = "taxa") %>% pull(var = taxa)   )[[1]])) -> SIMPER.clusters

SIMPER.clusters %>% 
  filter (Area == "Hood Canal", Season == "Summer") %>% 
  pull(summary.simper) -> list.simpr 
list.simpr %>% flatten_df(.id = "comparison") ->simper2

  simper2 %>% 
    group_by(comparison) %>% 
    nest() %>% 
    mutate(data = map (data, ~ .x %>% bind_cols(SIMPER.clusters %>% 
                                                filter (Area == "Hood Canal", Season == "Summer") %>%
                                                ungroup() %>% 
                                                select(taxa.simper) %>% 
                                                unnest()))) %>% 
    unnest(data) %>% 
  filter (p < 0.05) %>% 
  write_csv("../../Analysis/Splitting.data.by.lifestyle/simper.HoodCanal.summer.csv")

## Get the data for all seasons and areas
 SIMPER.clusters %>%
   pull(summary.simper)-> list.of.simpers

 list.of.simpers %>% 
   map_df(~bind_rows(.x, .id = 'Comparison'), .id = 'var1') %>%
   separate(var1, into = c("Area", "Season"), sep = "_") %>%
   group_by(Area, Season, Comparison) %>% 
   nest() %>% 
   left_join(SIMPER.clusters) %>%
   mutate(data = map2(data, taxa.simper, ~bind_cols(.x, as_tibble(.y)))) %>% 
   select(Area, Season, Comparison, data) %>% 
   unnest(data) %>% 
  # rename(taxa = value) %>% 
  write_csv("../../Analysis/Splitting.data.by.lifestyle/simper.all.csv")

```
#### Now do the dreamed panel
```{r}
taxa.to.follow <- read_csv("../../Analysis/Splitting.data.by.lifestyle/simper.HoodCanal.summer.csv")

taxa.to.follow %>% 
  filter (p<0.01, cumsum < 0.7) %>%  
  transmute(comparison, taxa = taxa.simper, diff = avb - ava) %>% 
  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
                           species == "NA" ~  paste0(genus, " sp."),
                           TRUE            ~ species)) %>%
  mutate(comparison = fct_relabel(comparison, ~ str_replace(.,"_", " vs "))) %>% 
  separate(comparison, into = c("a", "b"), remove = F) %>% 
  ggplot(aes(y = diff,
             x = fct_reorder(label, diff)))+
  geom_col(aes(fill = diff>0))+
  geom_hline(yintercept = 0, color = "black")+
  labs (x = "taxa",
        y = "Mean difference")+
  coord_flip()+
  facet_wrap(~comparison)+
  guides(fill = "none") +
  theme(axis.text.y = element_text(face = "italic"))  -> right_most

## And now do the same thing for all seasons and areas

all.taxa.to.follow <- read_csv("../../Analysis/Splitting.data.by.lifestyle/simper.all.csv")

all.taxa.to.follow %>% 
  group_by(Area, Season) %>% 
  nest() %>% 
  mutate(plot = map (data, ~ .x %>% 
                       filter (p<0.01, cumsum < 0.7) %>%  
                       transmute(comparison = Comparison, taxa = value, diff = avb - ava) %>% 
                       separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
                       mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
                           species == "NA" ~  paste0(genus, " sp."),
                           TRUE            ~ species)) %>%
                       mutate(comparison = fct_relabel(comparison, ~ str_replace(.,"_", " vs "))) %>% 
                       separate(comparison, into = c("a", "b"), remove = F) %>% 
                       ggplot(aes(y = diff,
                                  x = fct_reorder(label, diff))) +
                       geom_col(aes(fill = diff>0)) +
                       geom_hline(yintercept = 0, color = "black") +
                       labs (x = "taxa",
                             y = "Mean difference") +
                       coord_flip() +
                       facet_wrap(~comparison) +
                       guides(fill = "none") +
                       theme(axis.text.y = element_text(face = "italic")) )) %>% pull(plot) -> list.of.right.plots

```
### Final Figure 2

```{r assembling Figure 2}

plot_grid(leftplot.Fig2, right_top, right_most,align = "hv",nrow = 1, labels = "AUTO") -> top.plots

plot_grid(top.plots , legend, align = "v", nrow = 2, rel_heights = c(1, 0.1))

ggsave(filename = paste0("Figure2.", Sys.Date(), ".png"),
         width = 14,height = 5, dpi = "retina")

```

```{r assembling Figure for Supplementary material}

list.of.left.plots %>% map (get_legend) -> list.of.legends
list.of.left.plots %>% map(~.x + theme(legend.position = "none")) -> list.of.left.plots

list.of.mid.plots  %>% map(~.x + theme(legend.position = "none")) -> list.of.mid.plots

list.of.right.plots

pmap(.l = list(list.of.left.plots, list.of.mid.plots, list.of.right.plots),
      .f = function(a,b,c) plot_grid(a,b,c,align = "hv",nrow = 1, labels = "auto")) -> top.plots.in.a.list

pmap(.l = list(top.plots.in.a.list, list.of.legends, list.of.labels),
     .f = function(a,b,c) plot_grid(a,b, align = "v", nrow = 2, rel_heights = c(1, 0.1), labels = c) +
       ggsave(filename = paste0(c,"CAP.plot.png"),  width = 14,height = 5, dpi = "retina")) -> plots.ready.to.go

plots.ready.to.go %>% saveRDS("All_Cap_for_suppplemental.rds")
```

## Figure 3: a quick glance at spp responses
Get the models output and calculate the posterior across the simulated scenario

Now, calculate the posterior of each taxa across the 7600 scenarios simulated

### Posterior per taxa per point in the future

```{r}
## Rename variables so the model can be projected

all.sims.for.projection <- simData.nocrazys %>% 
  rename(Temperature = TemperatureSTD,
         pH = pH_STD)

## Get 
models %>% 
  transmute(taxa, 
            post.draws = map2 (model, taxa, ~ posterior_predict(object = .x,
                                                                newdata = all.sims.for.projection,
                                                                draws = 100) %>% 
                                 as_tibble() %>% 
                                 t(.) %>% 
                                 as_tibble() %>% 
                                 rownames_to_column("sim.row") )) -> posterior.draws.all

# This is an object with as many rows as environmental points there are. The first two columns are taxa and nrow. The rest are the binary output of the posterior
# predict

posterior.draws.all %>% 
 left_join(phylum.info %>% dplyr::select(phylum, taxa)) %>% 
  group_by(phylum) %>% # Add phylum information
  nest() %>% 
  mutate (data = map (data, ~ .x %>% 
                        unnest() %>% 
                        pivot_longer(cols = c(-taxa, -sim.row),
                                    names_to = "draw.row", values_to = "presence") %>% 
  group_by(taxa, sim.row) %>% 
  summarise(prob.of.presence = mean(presence)) %>% 
  left_join(all.sims.for.projection %>%  
              rownames_to_column("sim.row")) 
  #filter (Year %in% c(2015, 2095))
  )) -> posterior.probabilities.by.taxa

posterior.probabilities.by.taxa %>% mutate(suma = map (data,~.x %>% group_by(taxa, Year) %>% summarise(tot = sum(prob.of.presence)) %>% ungroup() %>% 
                                                         pivot_wider(names_from = Year, values_from = tot) %>% 
                                                         group_by(taxa) %>% mutate(diff = `2017` - `2095`))) %>% select(phylum, suma) %>% unnest(suma) %>% arrange(desc(abs(diff))) %>% 
  filter (phylum %in% c("Bacillariophyta", "Dinophyceae", "Haptophyceae", "Arthropoda"))

posterior.probabilities.by.taxa %>% 
  mutate( plot.by.taxa = map2 (data, phylum, ~ .x %>% 
      separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
                           species == "NA" ~  paste0(genus, " sp."),
                           TRUE            ~ species))    %>% 
    mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
                            TRUE          ~ Area)) %>%

  ggplot(aes(x= prob.of.presence,
             y = label)) +
  #geom_density_ridges(fill = as.factor(Year))
  stat_density_ridges(aes(fill = as.factor(Year)),
    # geom = "density_ridges_gradient", calc_ecdf = TRUE,
    quantiles =2, quantile_lines = TRUE, alpha = 0.5
  ) + 
  scale_fill_brewer(name = "Year",  type = "qual", palette = 2) + 
  labs (x = "Probability of presence",
        y = "") +
    facet_wrap(~Area) +

 # facet_grid(phylum ~ Area) + coord_flip(xlim = c(0,1)) +
 theme(strip.text.y = element_text(angle = 0),
       axis.text.y = element_text(face = "italic")) +
   ggtitle(.y) +
  theme(legend.direction = "horizontal",
        legend.position  = "bottom") ))   -> list.of.plots

# posterior.draws.all %>% 
#  left_join(phylum.info %>% dplyr::select(phylum, taxa)) %>% 
#   group_by(phylum) %>% 
#   nest() %>% pull(phylum) -> list.of.names

# map(list.of.plots,  ~ .x + facet_wrap(~Area)  ) -> list.of.plots
# 
#pdf(file = "taxadensityplots.by.phylum/all.by.area.pdf")
png(file = "taxadensityplots.by.phylum/all.by.area.png", width = 14, units = "in", res = 300)
list.of.plots %>% pull(plot.by.taxa)
dev.off()
#
list.of.plots %>% slice(1) %>% pull(plot.by.taxa)-> one.plot
legend.left <- get_legend(one.plot[[1]] )

list.of.plots %>% mutate(plot.by.taxa= map(plot.by.taxa, ~.x + theme(legend.position = "none") )) -> list.of.plots

```

#### Now do change in suitability by taxa by time
```{r}
selected.taxa <- models %>% select(taxa) %>% filter (str_detect(taxa, "Chaetoceros|huxleyi|Alexandrium|Pseudochatt|Thalassionema|Navicula |Dityl|Nitzschia|Karlodinium|Coscinodiscus|Attheya|Phaeocystis globosa|Chattonella|Wolosz|Dictyocha|Emili")) %>% pull(taxa)

posterior.probabilities.by.taxa %>% 
  unnest(data) ->  posterior.probabilities.by.taxa
```

Now something for the supporting information

```{r}
posterior.probabilities.by.taxa %>% 
  filter (taxa %in% selected.taxa) %>% 
  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
                           species == "NA" ~  paste0(genus, " sp."),
                           TRUE            ~ species)) %>% 
   ggplot(aes(x= prob.of.presence,
             y = label)) +
  #geom_density_ridges(fill = as.factor(Year))
  stat_density_ridges(aes(fill = as.factor(Year)),
    # geom = "density_ridges_gradient", calc_ecdf = TRUE,
    quantiles =2, quantile_lines = TRUE, alpha = 0.5
  ) + 
  scale_fill_brewer(name = "Year",  type = "qual", palette = 2) + 
  labs (x = "Probability of presence",
        y = "") +
    facet_wrap(~Area) +

 # facet_grid(phylum ~ Area) + coord_flip(xlim = c(0,1)) +
 theme(strip.text.y = element_text(angle = 0),
       axis.text.y = element_text(face = "italic")) +
  # ggtitle(.y) +
  theme(legend.direction = "horizontal",
        legend.position  = "bottom") +
  ggsave("Most.varied.taxa.png", width = 14, dpi = "retina")
```

```{r}

posterior.probabilities.by.taxa %>% 
  group_by(taxa, Area) %>% 
  filter (Year == 2017) %>% 
  summarise(median2017 = median (prob.of.presence)) %>% 
  right_join(posterior.probabilities.by.taxa %>% 
               filter (taxa %in% selected.taxa)) %>% 
  mutate(Corrected.presence = prob.of.presence - median2017) -> almost.there

almost.there %>% 
  group_by(taxa, Area, Year, phylum) %>% 
  summarise_at ("Corrected.presence", .funs = list (median = median, sd = sd)) %>% 
  filter (Year == 2095) %>% 
  ggplot(aes (x = median)) +
  geom_histogram() # establish a cutoff of +- 0.2

almost.there %>% 
  group_by(taxa, Area, Year, phylum) %>% 
  summarise_at ("Corrected.presence", .funs = list (mean = median, sd = sd)) %>% 
 filter (Year == 2095, abs(mean) > 0.10 ) %>%  ungroup %>% select(taxa, Area) -> subset.for.the.plot

almost.there %>% 
  group_by(taxa, Area, Year, phylum) %>% 
  summarise_at ("Corrected.presence", .funs = list (mean = median, sd = sd)) %>% 
  right_join(subset.for.the.plot) %>% 
  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
                           species == "NA" ~  paste0(genus, " sp."),
                           TRUE            ~ species)) %>% 
  ungroup() %>% 
  mutate(Area = case_when(Area == "SJI"    ~ "San Juan Island",
                          TRUE             ~ Area)) %>%

ggplot(aes(x = Year,
             y = mean))+
  geom_pointrange(aes (color = phylum,
                       ymin = mean - sd,
                       ymax = mean + sd))+
  geom_line(aes(group = taxa)) +
  geom_label_repel(data = . %>% filter (Year == 2095), aes(label = label,
                                                     fill = phylum))+
  facet_wrap(~Area) +
  theme(legend.position ="bottom") +
  ggsave("Suitability change.png", width = 14) -> bottom.plot.figure3

# almost.there %>% 
# ggplot(aes (x = as.factor(Year),
#               y = Corrected.presence, 
#               fill = taxa)) +
#   geom_boxplot()+
#   facet_wrap(~Area) +
#   guides(fill = "none")
```

### Richness by phyla with time
 First create the square matrix of points
```{r}
all.sims.for.projection  %>% 
  group_by( Season, Area) %>% 
  filter (keep == "keep") %>%   
  summarise_at(c("TemperatureCelsius", "pH_units"), list(max = max, min = min)) %>%

  group_by (Season, Area) %>% 
  nest() %>% 
  mutate(data = map (data, ~ expand_grid(TemperatureCelsius = seq(round(.x$TemperatureCelsius_min,1), round(.x$TemperatureCelsius_max, 1),by = 0.1),
                                         pH_units = seq(round(.x$pH_units_min,1), round(.x$pH_units_max,1), by = 0.1)))) %>% 
  unnest(data) -> full.grid
```

```{r}
full.grid %>% 
  ungroup() %>% 
  rownames_to_column("Case") %>% 
  rename(Temperature = TemperatureCelsius,
         pH = pH_units) %>% 
  pivot_longer(cols = c(Temperature, pH),
               names_to  = "Predictor",
               values_to = "value") %>% 
  group_by(Predictor) %>% 
  nest() %>% 
  left_join(transforming.df) %>% 
  mutate(data = map2(data, params, ~.x %>% 
                       mutate(std = (value - .y$mean)/(.y$sd) ))) %>% 
  select(-params) %>% 
  unnest(data) %>% 
  pivot_wider(names_from = Predictor,
              values_from = c(std,value)) %>% 
  rename(Temperature = std_Temperature,
         pH = std_pH,
         TemperatureCelsius = value_Temperature,
         pH_units           = value_pH) -> full.grid

models %>% 
  ungroup() %>%  #rpk added
  transmute(taxa, 
            post.draws = map2 (model, taxa, ~ posterior_predict(object = .x,
                                                                newdata = full.grid,
                                                                draws = 100) %>% 
                                 as_tibble() %>% 
                                 t(.) %>% 
                                 as_tibble() %>% 
                                 rownames_to_column("sim.row") )) -> posterior.draws.raster

posterior.draws.raster %>% 
 left_join(phylum.info %>% dplyr::select(phylum, taxa)) %>% 
  group_by(phylum) %>% 
  nest() %>% 
  mutate(data = map (data, ~ .x %>% unnest(cols= post.draws) %>% 
                       pivot_longer(cols = c(-taxa, -sim.row),
                                    names_to = "draw.row", values_to = "presence") %>% 
                       group_by(draw.row, sim.row) %>% 
                       summarise(tot = sum(presence)) %>% 
                       group_by(sim.row) %>% 
                       summarise_at(.vars = "tot", .funs = list(Richness = mean,sd = sd)))) %>% 
  unnest(data) -> ready_plot

ready_plot %>% 
  left_join(full.grid %>%  
              rownames_to_column("sim.row")) -> ready_plot

# Normailzed the data for each phyla

ready_plot %>% 
  group_by(phylum) %>% 
  mutate(st.richness = Richness / max(Richness)) -> ready_plot

```

```{r}
plot3 <- function(tibble,string , response) {

  response = enquo(response)

  tibble %>% 
    mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
                            TRUE          ~ Area)) %>% 
    group_by(TemperatureCelsius, pH_units, Area) %>%

    summarise (!!response := mean (!!response))  %>% 
   #  filter(str_detect(phylum, string))   %>% 
 # full_join(full.grid %>% unnest(data)) %>% 
  ggplot(aes(x = round(TemperatureCelsius,10),
             y = round(pH_units, 10))) +

  geom_raster( interpolate = T,aes( fill = !!response)) +

  scale_fill_distiller(name = "Richness", type = "seq", palette = "Spectral") +

  # geom_density2d(data = simData %>%
  #                   filter (Year %in% c(2015, 2095)),
  #                 aes(x= TemperatureCelsius, y = pH_units, group= Year, color = as.factor(Year)), bins = 5, alpha = 0.5) +

  scale_color_brewer(name = "Year",  type = "qual", palette = 2) +
  ggforce::geom_mark_hull(data = all.sims.for.projection %>%
                            mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
                            TRUE          ~ Area)) %>%
                            filter (Year %in% c(2017, 2095), keep == "keep") ,
                          aes(x= TemperatureCelsius, y = pH_units, group= Year, color = as.factor(Year),label = as.factor(Year)),
                          expand = unit(0.5, "mm"),
                 label.margin = margin(1, 1, 1, 1, "mm"),
                 label.buffer = unit(0, "mm"),
                 con.cap = 0 ,
                 alpha = 0.5) +
     facet_grid(.~Area) +
  guides(color = "none") +
    labs (x = "Temperature",
          y= "pH") +
    #ggtitle(label = string) +
    theme(legend.position = "bottom",
          legend.key.width  = unit(2 ,"cm"))

}
```

```{r}
ready_plot %>% 
  nest(-phylum) %>% 
  mutate(plot = map2(data, phylum, ~ plot3(.x, .y, response = st.richness)))  -> list.of.richness.plots

# Get the other legend
list.of.richness.plots %>% slice(1) %>% pull(plot) -> one.plot
legend.right <- get_legend(one.plot[[1]]  )
list.of.richness.plots %>% 
  mutate(plot = map (plot, ~.x + theme(legend.position = "none"))) -> list.of.richness.plots
legend.left

```

```{r}
posterior.probabilities.by.taxa %>% 
  group_by(phylum) %>% 
  summarise(n_distinct(taxa)) %>% 
  pull -> list.of.number.of.taxa

list.of.plots %>%  pull(phylum) -> list.of.phyla

list.of.plots %>% pull(plot.by.taxa) -> list.of.ridge.plots

list.of.ridge.plots <- set_names(list.of.ridge.plots, nm = list.of.phyla)
#list.of.ridge.plots %>% map(~.x + )
list.of.richness.plots %>% pull(plot) -> list.of.raster.plots

list.of.raster.plots <- set_names(list.of.raster.plots, nm = list.of.phyla)

map2(list.of.ridge.plots, list.of.raster.plots, function(.x, .y) plot_grid (.x, .y, nrow = 2, align = "hv", axis = "right")) -> combined.plots

pmap(.l = list(list.of.ridge.plots,list.of.phyla,list.of.number.of.taxa),
     .f = function (a,b) ggsave(a, filename = paste0(b, "ridges.png"), width = 14, 
                                  #height = c*20,
                                  dpi = "retina"))
map2(list.of.ridge.plots,list.of.phyla, 
     ~ ggsave(.x, filename = paste0(.y, "ridges.png"), width = 14, 
                                  #height = c*20,
                                  dpi = "retina"))
ggsave(list.of.ridge.plots[["Bacillariophyta"]], filename = "Bacillariophytaridges.png", width = 14, height = 14,                                  dpi = "retina")

map2(list.of.raster.plots,list.of.phyla, 
     ~ ggsave(.x, filename = paste0(.y, "raster.png"), width = 7, 
                                  #height = c*20,
                                  dpi = "retina"))
map2(list.of.raster.plots,list.of.phyla, 
     ~ .x + 
       ggtitle(.y) + ggsave(filename = paste0(.y, "rastertitle.png"), width = 7, 
                                  #height = c*20,
                                  dpi = "retina"))
```

```{r}
combined.plots <- set_names(combined.plots, list.of.phyla)

top.line <- combined.plots[c("Arthropoda", "Dinophyceae", "Bacillariophyta", "Haptophyceae")]

plot_grid(plotlist = top.line,
          nrow = 1,
          align = "hv" ) -> top.all
plot_grid(legend.left, legend.right, nrow = 2, align = "hv" ) -> midlayer

plot_grid(bottom.plot.figure3, midlayer,
          nrow = 1,
          rel_widths = c(1,0.2),
          align = "hv") -> bottom.plots

# bottom.line <- list( combined.plots[c("Haptophyceae")][[1]], bottom.plot.figure3)
# plot_grid(plotlist = bottom.line,
#           nrow = 1,
#           align = "hv") -> bottomline

top.all + ggsave("Top_fig3.png", dpi = "retina", width = 28, height = 10)

bottom.plot.figure3 + ggsave("bottom.fig.3.png", dpi ="retina", width = 14)

midlayer + ggsave("legends.fig3.png", dpi = "retina", height = 2)

```

### Besides Figures: other stats reported in the MS
```{r}
ASV.table  <- read_csv("../../Data/3.Growing_database/ASV_table_all_together.csv")

Annotation <- read_csv("../../Analysis/Annotation/Annotated.hashes.csv")

## Remove HUmans and chickenfilter(!str_detect(taxa, "Phasianidae"))

ASV.table %>% 
  filter(!Hash %in%  (Annotation %>% 
           filter (str_detect(species, "Homo")) %>% 
           pull(Hash))) -> ASV.table

ASV.table %>% 
  summarise(runs = n_distinct(Miseq_run),
            reads = sum(nReads),
            asvs = n_distinct(Hash)) -> totalreads

## collapse the dataset by taxonomy

left_join(ASV.table, Annotation) %>% 
  unite(family, genus, species, col = "taxa", sep = "|") %>% 
  filter(taxa != "NA|NA|NA") %>% 
  select(taxa, sample, nReads, benthos) %>% 
  group_by(taxa, benthos, sample) %>% 
  summarise(nReads = sum(nReads)) -> ASV.by.taxa

ASV.by.taxa %>% 
  group_by(benthos) %>% 
  summarise(reads = sum(nReads),
            taxa = n_distinct(taxa)) -> benthostaxareads

ASV.by.taxa %>% 
  ungroup() %>% 
  summarise(reads = sum(nReads),
            taxa = n_distinct(taxa)) -> taxareads

ASV.by.taxa %>% 
  separate(sample, into = c("Site", "biol"), sep = "_", remove = F) %>% 
  separate(biol, into = c("biol", "tech.rep"), sep = "\\.") %>% 
  separate(biol, into = c("date", "biol.rep"), sep = -1) %>% 
  unite (Site, date, col = "event") -> ASV.by.taxa

ASV.by.taxa %>% 
  filter (benthos != "BEN") %>% 
      mutate(
        Season = case_when(
          str_detect(sample, "1703|1710|1711|1801|1803") ~ "Winter",
          TRUE ~ "Summer"
        ),
        Area = case_when(
          str_detect(sample, "CP|LK|FH") ~ "SJI",
          TRUE ~ "Hood Canal"
        )
      ) %>% 
 group_by(Area) %>%
  nest() %>% 
  mutate(data = map (data, ~.x %>% group_by(taxa) %>% summarise (tot = sum(nReads)))) %>% unnest() %>% 
  pivot_wider(names_from = Area, values_from = tot, values_fill = list(tot = 0)) %>% 
  mutate(shared = case_when(SJI > 0 & `Hood Canal` > 0 ~ "Shared",
                            SJI != 0 & `Hood Canal` == 0 ~ "Only SJI",
                            SJI == 0 & `Hood Canal` != 0 ~ "Only HC")) %>% 
  group_by(shared) %>% #tally
summarise (tot = sum(SJI + `Hood Canal`),
                                 ncases = n())
```
#### Does richness change within phylum between Years

```{r}
  posterior.draws.all %>% 
    left_join(phylum.info %>% select(phylum, taxa)) %>% 
    group_by(phylum) %>% 
    nest() %>% 
  #  slice(1) %>%
    mutate(data = map (data, ~ .x %>%
                         unnest() %>% 
                         pivot_longer(cols = c(-taxa, -sim.row),
                                      names_to = "draw.row",
                                      values_to = "presence")  %>% 
                         group_by(draw.row, sim.row) %>% 
                         summarise(tot = sum(presence)) %>% 
                         group_by(sim.row) %>% 
                         summarise_at(.vars = "tot", .funs = list(Richness = mean,sd = sd)) %>% 
  left_join(all.sims.for.projection %>%  
              rownames_to_column("sim.row")) ))%>%

   mutate(lm = map(data, ~ lm(Richness ~ Year, data = .x)),
         lm.results = map (lm, broom::tidy))  -> results.lm

### Using a wilcox test to see if there is a difference between Years

results.lm %>% 
  mutate(wilcoxon = map (data, ~ wilcox.test(Richness ~ Year, data = .x) ),
         tidy.w   = map (wilcoxon, broom::tidy)) -> results.lm

results.lm %>% select(phylum, wilcoxon) %>% filter (phylum == "Bacillariophyta") %>% pull(wilcoxon) 
```

### 2020-08-03 Input/1.Sequencing_runs_output/20171122.Jul_Aug/banzai_params_for_DADA2_Jul_Aug.sh
INput files

@@ -1,266 +1,0 @@
-#!/usr/bin/env bash
-
-
-################################################################################
-# INPUT
-################################################################################
-# What is the file path to the directory containing all of the libraries/reads?
-PARENT_DIR="/Users/Moncho/Google_Drive/Run_Nov17/OA_COI/171122"
-
-# Where is the sequencing metadata file? (SEE FORMATTING GUIDELINES IN README!)
-SEQUENCING_METADATA="${PARENT_DIR}"/metadata.csv
-
-
-################################################################################
-# OUTPUT
-################################################################################
-# This script will generate a directory (folder) containing the output of the script.
-# Where do you want this new folder to go?
-OUTPUT_DIRECTORY="${HOME}"/fastqs_demultiplexed_for_DADA2 #"${PARENT_DIR%/*}"
-
-# You can optionally specify a folder into which the script copies a PDF containing some results.
-# The pdf is created by default in the analysis folder specified above, but
-# if you set this to your DropBox or Google Drive Folder, you can check it out from anywhere.
-OUTPUT_PDF_DIR=""
-
-
-################################################################################
-# METADATA DETAILS
-################################################################################
-# Specify columns for raw sequencing files:
-COLNAME_FILE1="file1"
-COLNAME_FILE2="file2"
-
-# MUST be unique for each row!
-COLNAME_SAMPLE_ID="sample_id"
-
-COLNAME_DNA_ID="dna_id"
-COLNAME_SAMPLE_TYPE="dna_source"
-
-# Your metadata must have a column corresponding to the subfolders containing the raw reads.
-# In order to make this flexible across both multiple and single library preps, you must include this even if you only sequenced one library (sorry!).
-COLNAME_ID1_NAME="pri_index_name"
-COLNAME_ID1_SEQ="pri_index_seq"
-
-################################################################################
-# MERGE PAIRED READS
-################################################################################
-USE_PEAR_DEFAULTS="NO"
-# For more information on these parameters, type into a terminal window: pear -help
-# Bokulich recommends:
-# Quality_Threshold=3, r=3 (PEAR only considers r=2), UNCALLEDMAX=0
-# TRIMMIN= 0.75 * LENGTH_READ # this is hard-coded in the script banzai.sh
-
-# TODO grab this from a fragment_size column in the sequencing metadata file
-### ***** REMEMBER TO WATCH FOR ZEROS WHEN IMPLEMENTING THIS!
-# Is there a column in the metadata file for fragment size?
-insert_size_in_metadata="NO"
[…]

### 2020-08-03 Scripts/All.figures.that.go.in.the.MS.no.seasons.Rmd
Updated scripts

---
title: "Analysis and Figures for Gallego et al, 2020"
author: "Ramon Gallego"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

## Intro

This Markdown will make the part of the analysis that goes in the Manuscript:

```{r libraries and custom functions, echo = F, message=F}
library (RColorBrewer)
library (patchwork)
library (rstanarm)
library (fitdistrplus)
library (ggridges)
library (ggforce)
library (tidyverse)
library (cowplot)
library (concaveman)
library (vegan)
library (lubridate)
library (ggrepel)
library (factoextra)
library (nnet)
library (here)
library (conflicted)
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
conflict_prefer("area", "patchwork")
#functions for future Salish Sea
models <- readRDS(here("Input","full_model_output_rpk_TpH.standardized_nonSeasonal2020-06-24_3262.RDS"))

source(here("Scripts","Future_climate_conditions.R"))

source(here("Scripts","tibble.to.matrix.r"))

stds_to_value <- read_rds(here("Input", "std_to_real.rds"))

```

## Datasets to use

To begin with: the ASV.by.taxa dataset and the environmental data (Updated salinity, DIC and pH).

### Present day datasets

```{r }
data.plankton <- read_csv(here("Input","Combined_Biol_Env_Plankton.csv"))

data.plankton %>% filter(!str_detect(taxa, "Phasianidae|Laridae")) %>% 
  filter (!str_detect(taxa, "Rhodomelaceae")) -> data.plankton

env.data <- read_csv(here("Input","env.data.for.log.regression.csv"))

#quality-control data
    env.data %>%
      filter(pH > 7) %>%
      mutate(
        Season = case_when(
          str_detect(sample, "1703|1710|1711|1801|1803") ~ "Winter",
          TRUE ~ "Summer"
        ),
        Area = case_when(
          str_detect(sample, "CP|LK|FH") ~ "SJI",
          TRUE ~ "Hood Canal"
        )
      ) -> now.env.data

now.env.data %>%
  gather(-Season, -sample, -Area,  key = "Predictor", value = "value") %>%
  group_by(Area, Predictor) %>%   #RPK: this was missing, which messed up the std calculations 
  mutate(std = vegan::decostand(value, method = "standardize")) -> now.env.data.with.std

#### TODO: remove

# Dataset to go between standarized and real values
now.env.data.with.std %>%
  group_by(Area, Predictor) %>%
  # summarise(mean = mean(value),
  #           sd = sd(value)) %>%
  # group_by(Area, Predictor) %>%
  nest() %>% 
  summarise(tovalue = map(data, ~ lm(value ~ std, data = .x) ),
            tostd   = map(data, ~ lm(std ~ value, data = .x)))-> transforming.df

transforming.df[1,4] %>% pull -> test
broom::tidy(test[[1]])
extract <- function (lm) {broom::tidy(lm)$estimate}

extract(test[[1]])

transforming.df %>% 
  mutate(to.values = map(tovalue, extract),
         to.std= map(tostd, extract)) ->transforming.df

phylum.info <- read_csv(here("Input","higher_taxonomy.csv")) %>% 
  unite(family, genus, species, sep = "|", col = "taxa")

```
### Future scenarios for the Puget Sound

In order to project the suitability of the climate conditions in the Puget Sound and Salish Sea, we use the work from based upon Khangaonkar, T., Nugraha, A., Xu, W., & Balaguru, K. (2019). Salish Sea response to global climate change, sea level rise, and future nutrient loads. Journal of Geophysical Research: Oceans, 124. https://doi.org/10.1029/2018JC014670 and in particular, upon Figure 14 in that paper. Check the script Future_climate_conditions.R for further details.

```{r Future scenarios}
Year.area.dataset <- function(Area, Year, n){

    Temp.conv <- stds_to_value[which (stds_to_value$Area == Area & stds_to_value$key == "Temperature"),"toValue"] %>% pull %>% .[[1]]
    pH.conv   <- stds_to_value[which (stds_to_value$Area == Area & stds_to_value$key == "pH")         ,"toValue"] %>% pull %>% .[[1]]

    one<-  data.frame(
              Area = Area,
              Year = Year,
              Temperature = r_Temperature_year_STD(Year, Area, n, 2017),
              pH          = r_pH_year_STD(Year, Area, n, 2017)
              )

    one$TemperatureCelsius = sapply(one$Temperature,function(x){Temp.conv$coefficients[1]+Temp.conv$coefficients[2]*x})
    one$pH_units            = sapply(one$pH,function(x){pH.conv$coefficients[1]+pH.conv$coefficients[2]*x})

    return(one)

}

j <- Year.area.dataset("Hood Canal", 2095, 5000)
k <- Year.area.dataset("Hood Canal", 2017, 5000)
l <- Year.area.dataset("SJI", 2095, 5000)
m <- Year.area.dataset("SJI", 2017, 5000)

simulated.scenario <- rbind(j,k,l,m)

rm(j,k,l,m)

simulated.scenario %>%
  group_by(Year, Area) %>% 
  nest() %>% 
  mutate(
    values = map (data, function (.x) {
      list(temp = fitdistr(.x$TemperatureCelsius ,
             densfun = "normal"),
           ph = fitdistr(.x$pH_units ,
             densfun = "normal"))
            }
          ),

    data = map2(data,values,  function (.x, .y){
      .x %>%
        mutate(pT = pnorm(TemperatureCelsius, mean  = .y[["temp"]]$estimate[1], sd = .y[["temp"]]$estimate[2]),
               ppH =pnorm(pH_units, mean  = .y[["ph"]]$estimate[1], sd = .y[["ph"]]$estimate[2]) )
      })
    ) %>% 
  unnest(data) %>% 
  mutate(keep = case_when(pT < 0.05 | ppH < 0.05 | pT > 0.95 | ppH > 0.95 ~ "Discard",
                          TRUE                   ~ "keep")) %>% 
  filter (keep == "keep")-> simData.nocrazys

simData.nocrazys %>% 
  select(-values) %>% write_csv(here("Output","simulation.from.std.functions.cleaned.csv"))
```

### Save future scenarios

```{r}

simData.nocrazys %>%

saveRDS( file = here("Output","all.sim.data_climateScenarios_no.outliers.RDS"))

simData.nocrazys %>% 
  group_by( Area, Year) %>% 
  summarise_at(c("TemperatureCelsius", "pH_units"), .funs = list(min = min, max = max))

```

## Figure 1

We have the datasets and some plots already made -

### Panel A Sample location
```{r Base map}
map <- readRDS(here("Input","map.plot.rds"))

map + ggtitle(label = element_blank()) + labs (x = "Longitude", y = "Latitude") -> map

```
### Panel B Environmental variation

Make the ridges plot

```{r}
now.env.data %>% 
  select(TA = mean.TA, DIC = mean.DIC, everything(), -contains("%")) %>% 
  filter(!is.na(TA)) %>% 
  separate(sample, into= c("Site","Date"), remove = F )-> env.data.no.outliers

env.data.no.outliers %>% 
  pivot_longer(cols = c(-Area, -Site, -Date, -Season, -sample), names_to =   "Predictor", values_to  = "value") -> env.data.to.keep

env.data.to.keep %>% 
  mutate(Site = fct_relevel(Site, "SA", "TR", "LL","PO", "TW"),
         Predictor = fct_relevel(Predictor, "DIC", "TA", "pH")) -> env.data.to.keep

env.data.to.keep %>% 
  mutate(Area = case_when (Area == "SJI" ~ "San Juan Island",
                           TRUE          ~ Area)) %>% 
  mutate(Month = month(ymd(paste0(Date, "01")), label = T, abbr = F)) %>% 
  filter (Predictor %in% c("DIC", "TA", "pH", "Temperature", "Salinity")) %>% 
  ggplot(aes(x = `value`, y = `Month`, fill = Area)) +
  geom_density_ridges_gradient() +
 # scale_fill_viridis(name = "Temp. [F]", option = "C") +
  facet_wrap(~fct_relevel(Predictor, "DIC", "TA", "pH", "Temperature", "Salinity"), scales = "free", nrow = 1) +
  theme(legend.direction ="horizontal") -> ridges

legend <- get_legend(
  # create some space to the left of the legend
  ridges + theme(legend.box.margin = margin(0, 0, 0, 12))
)
 legend
ridges + guides(fill = "none") -> ridges
```

### Panel c: Richness variation

```{r}
boxplot.richness <- readRDS(here("Input","richness.rds"))

boxplot.richness +
  ylim (0,60) -> boxplot.richness
```
### All together

THe top will be the environmental gradients, the lower panels will have the map and the boxplot
```{r}
plot_grid(plotlist = list(map, legend, boxplot.richness),
                     align = "v",
                     ncol = 3, 
          axis = "t",
          rel_widths = c(1,0.5,0.6))   -> bottom.half.plot1

plot_grid(plotlist = list(ridges, bottom.half.plot1),
          align = "hv",
          ncol = 1, 
          nrow = 2,
          rel_heights = c(1,0.8),
          rel_widths = 1)

plot_grid(plotlist = list(legend, boxplot.richness),
                     align = "hv",
                     ncol = 1, 
          axis = "t",
          rel_heights  = c(.5,1),
          labels = c(NA,"B" )) -> bottom.right

plot_grid (plotlist = list(map, bottom.right),
           align = "hv",
           nrow = 1,
           rel_heights = 1,
           rel_widths = c(.6,0.4),
           labels = c("A", NA)) -> bottom.plot

plot_grid(bottom.plot, ridges,
          align = "h",
          nrow = 2,
          rel_heights = c(0.6, 0.4),
          labels = c(NA,"C")) +
  ggsave(filename = here("Figures","Manuscript","Fig1.png"), width = 14, height = 14, dpi = "retina")

```

## Figure 2

To reuse for future analysis, create a function that does the fist step on a multivariate analysis: Create a horizontal matrix, a compatible matrix with the environmental data, and a dissimilarity object.

```{r functions for CAPS now without season}
cap.functions.only3things <- function(tibble, ...){ # ..  are the grouping factors

  tibble %>%

    group_by(...) %>%

    nest() %>%

    mutate(comm.matrix = map (data, ~ tibble.to.table(.x, taxon = taxa, Abundance = Normalized.reads, sample.name = event)),

           ja = map(data, ~tibble_to_matrix(.x, taxon = taxa,Abundance = Normalized.reads, sample.name = event, distance = "jaccard",transformation = "binary" )),
          bc = map(data, ~tibble_to_matrix(.x, taxon = taxa,Abundance = Normalized.reads, sample.name = event, distance = "bray",transformation = "" )), # the bc distamce
          env = map (data, ~tibble_to_env(.x, taxon = taxa, Abundance = Normalized.reads, sample.name = event, everything()) %>% 
                       select(-contains("%")) %>% 
                       separate(event, into = c("Site", "Date"), sep = "_")))
}

```

We will do the Full CAP analysis - but will only keep Hood Canal for the MS - the rest will have to go to the Supplementary information

The figure will have three panels: the CAP for Hood Canal, the probability of presence of each cluster in the pH/Temperature space and the main contributors to the dissimilarity betweeen clusters.

```{r CAP analysis FULL}

data.plankton %>% 
  separate (event, into = c("Site", "Date"), sep = "_") %>% 
  mutate(Season = case_when(str_detect(Date,"17-03|17-10|17-11|18-01|18-03") ~ "Winter",
                             TRUE ~                                         "Summer"),
         Area = case_when(Site %in% c("CP","LK", "FH") ~ "San Juan Island",
                          TRUE                         ~ "Hood Canal")) %>% 
   unite (Site, Date, col = "event", sep = "_") %>% 
  cap.functions.only3things(Area) -> joined.dataset.for.CAP
```

Step by step
```{r CAP1}
joined.dataset.for.CAP %>% 
  mutate(
    CAP.analysys.raw = map2(comm.matrix, env,

                          function(.x,.y) {
                            .y %>% select (Temperature, Salinity = Salinity.new, pH = pH_new) 
                            caps <-  capscale(formula = .x ~ Temperature + Salinity + pH ,
                                                 data = .y %>%
                                                   select(DIC = mean.DIC.new.sal, Salinity = Salinity.new, Temperature, pH = pH_new) %>% 
                                                   select_if(is.numeric) %>%
                                                   decostand(method = "standardize", na.action = na.omit),
                                                 distance = "jaccard", 
                                                binary = T)

                            sppscores(caps) <- sqrt(decostand(.x, "total"))

                            return(caps) }),
         CAP.plot = map (CAP.analysys.raw, ~ plot(.x)),

         CAP.significance = map(CAP.analysys.raw, ~ permutest(.x, permutations = 999 )) ) -> joined.dataset.for.CAP

joined.dataset.for.CAP %>% pull(CAP.significance)
```

So the best numbers of clusters are

```{r CAP2}
joined.dataset.for.CAP %>% mutate(plot3 =  map2 ( CAP.plot, Area, function(.x, .y) {fviz_nbclust(.x$sites,
                                                                             kmeans,
                                                                             k.max = 5,
                                                                             method = "wss") + ggtitle(paste0 (.y, "-WSS"))})) %>% pull(plot3)

joined.dataset.for.CAP %>% mutate(plot3 =  map2 ( CAP.plot, Area, function(.x, .y) {fviz_nbclust(.x$sites,
                                                                             kmeans,
                                                                             k.max = 5,
                                                                             method = "gap_stat") + ggtitle(paste0 (.y, " gap_stat"))})) %>% pull(plot3)

joined.dataset.for.CAP %>% mutate(plot3 =  map2 ( CAP.plot, Area, function(.x, .y) {fviz_nbclust(.x$sites,
                                                                             kmeans,
                                                                             k.max = 5,
                                                                             method = "silhouette") + ggtitle(paste0 (.y, "-Silhouette"))})) %>% pull(plot3)
```

These three methods would not agree on a "perfect" number of clusters - but using 3 clusters for both areas returned good scores in all three criteria.

So using 3 clusters - generate these clusters and join the datasets

```{r CAP2b}
joined.dataset.for.CAP %>%
  ungroup %>% 
  mutate(n_clust = c(3,3),

  CAP.clusters = map2 (CAP.plot,n_clust,  function (.x, .y) {

    final <- kmeans(.x$sites, .y, nstart = 15)

    fviz_cluster(final, data = .x$sites, geom = "point", palette = "Set2")

  } ),

  Cluster.df = map2 (CAP.plot,n_clust,  function (.x, .y) {

    final <- kmeans(.x$sites, .y, nstart = 15)

    tibble(event = names(final$cluster),
           cluster = final$cluster) }

    )
  ) -> Clusters.analysis.and.cap

Clusters.analysis.and.cap %>% 
  mutate(full.cluster.info = map2(env, Cluster.df, ~ .x %>% unite(Site, Date, col = "event") %>%right_join(.y) ),
         full.cluster.info = map2(full.cluster.info,CAP.plot, ~ .y$sites %>% 
                                    as.data.frame() %>% 
                                    rownames_to_column("event") %>% 
                                    right_join(.x) %>% 
                                    mutate(pH = round(pH,1)))

         ) -> Clusters.analysis.and.cap

# Write out clusers and CAP points, and raw env data

Clusters.analysis.and.cap %>% 
  select(Area,  full.cluster.info) %>% 
  unnest() %>%
  write_csv(here("Output","clusters.of.events.csv"))

```

Now create the plot that combines Constrained Analysis of Principal Coordinates and clusterings
### Panel A: CAP analysis with environmnental correlation and cluster info
```{r Hand made CAP}

Clusters.analysis.and.cap %>%

  mutate(Hand.made.CAP = map2(CAP.plot, full.cluster.info, function (.x, .y){

    .x$biplot %>%
                         as.data.frame() %>%
                        rownames_to_column("env.variable") -> var.scores

    .y %>% 
      select(event, CAP1, CAP2, cluster) %>% 
      ggplot(aes(x = CAP1,
                 y = CAP2)) +
      ggforce::geom_mark_hull(aes( group= cluster, color = as.factor(cluster),label = as.factor(cluster), fill = as.factor(cluster)),
                          expand = unit(2, "mm"),
                 label.margin = margin(1, 1, 1, 1, "mm"),
                 label.buffer = unit(0, "mm"),
                 con.cap = 1 ,
                 alpha = 0.5) +

      geom_point(size = 1.5) +
       geom_point(aes(color = as.factor(cluster)), size = 1) +
      geom_segment(aes(x = 0, y = 0,
                       xend = CAP1,
                       yend = CAP2), data = var.scores, color = "blue", arrow = arrow(length = unit(0.1,"cm"))) +
      geom_label_repel(aes(x= CAP1  ,
                           y= CAP2 ,
                           label = env.variable), data = var.scores, fill = "pink", alpha = 0.75) +
      ggtitle ("") + coord_equal() + scale_color_brewer(name = "Cluster", palette = "Set2") + scale_fill_brewer(name = "Cluster",palette = "Set2") +

      theme(legend.position = "bottom")

  })) %>% pull(Hand.made.CAP) -> leftplot.Fig2

 leftplot.Fig2 %>% set_names(Clusters.analysis.and.cap$Area) %>% map(~ .x +  theme_light() + theme(legend.position = "none") ) -> left_top

```

### Panel B: Cluster prevalence in the Temperature / pH space

We need the simulated data of the future conditions, and the cluster information.

```{r}
all.sim.data <- simData.nocrazys  #to see simulation data

clusters <- read.csv(here("Output","clusters.of.events.csv")) %>% 
  mutate(cluster  = as.factor(cluster)) %>% 
  nest(clusterdata = -Area)
```

Use a multinomial model to predict the cluster identity as a function of Temperature and pH.

```{r models and so for getting the communities}
clusters %>% 
  mutate (multi.mod = map (clusterdata, ~ multinom(cluster ~ Temperature + pH_new, data = .x))) -> clusters

```

Now create the prediction of the most likely community under each Temp-pH combination.

```{r models and so for getting the communities2}

  all.sim.data %>% 
    mutate (Area = fct_recode(Area, "San Juan Island" = 
                                "SJI")) %>% 
    group_by(Area) %>% 
    nest() %>% 
    mutate(pH.range = map(data, ~ seq(round(min(.x$pH_units),1), 
                                      round(max(.x$pH_units),1),
                                      length.out = 12)),
           T.range =  map(data, ~ seq(round(min(.x$TemperatureCelsius),1), 
                                      round(max(.x$TemperatureCelsius),1),
                                      length.out = 22)),
           output.range = map2(pH.range, T.range, ~ expand_grid(pH_new = .x, Temperature = .y))) %>%

    left_join(clusters) %>% 
    mutate(whichComm.res = map2(output.range,multi.mod,~ .x %>% 
                                  mutate(comm = as.numeric (apply(predict(.y, type = "probs", newdata = .x),1, which.max)) )
              ) ,
           plots = map(whichComm.res, ~ 
                         ggplot(.x, aes(y = pH_new,
                                        x = Temperature,
                                        fill = as.factor(comm),
                                        )) +
                         geom_tile()+
                         geom_label(data = . %>% group_by(comm) %>% 
                                      summarise_all( mean),
                                    aes(x= Temperature,
                                        y= pH_new,
                                        label = comm),
                                    fill = "white")+
                         scale_fill_brewer(name = "Cluster", palette = "Set2") +
                         xlab("Temperature") + ylab("pH")+
                         theme_light())) -> testquick

testquick %>% pull(plots) %>% set_names(testquick$Area) %>% map(~ .x +  theme(axis.title.y = element_text(angle = 0),legend.position = "none") )-> center_top
```

### Panel C: For each cluster, follow the relative abundance of the taxa driving the differences

#### Generate the SIMPER clusters
```{r}
Clusters.analysis.and.cap %>% 
  mutate(SIMPER.clusters = map2(comm.matrix, Cluster.df, function(.x, .y) {

    simper(comm = .x,
          group = as.factor(.y$cluster),
          permutations = 999)

  })) %>% 
  select(Area, SIMPER.clusters) -> SIMPER.clusters

SIMPER.clusters %>% 
  mutate(summary.simper = map(SIMPER.clusters, ~ summary(.x, ordered  = TRUE) )) %>% 
  mutate(taxa.simper = map(summary.simper, ~ map(.,function(x) rownames_to_column(x, var = "taxa") %>% pull(var = taxa)   )[[1]])) -> SIMPER.clusters

SIMPER.clusters %>% 
  mutate(summary.simper = map(summary.simper,function(.x) bind_rows(.x[1:length(.x)], .id = "comparison") %>% rownames_to_column("taxa"))) %>% pull(summary.simper) %>% set_names(SIMPER.clusters$Area) %>% bind_rows(.id = "Area") %>% separate(taxa, into = c("taxa",NA), sep = "\\...") %>% 
  filter (p < 0.05) %>% 
  write_csv(here("Output","simper.all.csv"))

## Get the data for all seasons and areas
 SIMPER.clusters %>%
   pull(summary.simper)-> list.of.simpers
names(list.of.simpers) <- paste(SIMPER.clusters$Area, SIMPER.clusters$Season, sep = "_")

 list.of.simpers %>%
  map(~ .x %>% map(~ bind_rows(.)) %>% 
                          bind_rows(., .id = "Comparison")) %>% 
  bind_rows(., .id = "var1") %>%
   separate(var1, into = c("Area", "Season"), sep = "_") %>%
   select(-Season) %>% 
   group_by(Area,  Comparison) %>% 
   nest() %>% 
   left_join(SIMPER.clusters) %>%
   mutate(data = map2(data, taxa.simper, ~bind_cols(.x, as_tibble(.y)))) %>% 
   select(Area,  Comparison, data) %>% 
   unnest(data) %>% 
  # rename(taxa = value) %>% 
  write_csv(here("Output","simper.all.csv") )

```
#### Now do the topright
 get the data and plt a distribution of the average contributions of each taxa for each 
```{r}
taxa.to.follow <- read_csv(here("Output","simper.all.csv"))

taxa.to.follow %>% 
  ggplot(aes(x = average)) + 
  geom_histogram() +
  facet_grid(Area ~ Comparison)
```

### Choose a threshold of 0.01 for average contribution

```{r}
taxa.to.follow %>% 
  filter (p<0.01, average > 0.01) %>%  
  arrange(desc(average)) %>% 
  group_by(Area, Comparison) %>% 
    slice(1:5) %>%

  transmute(Area, Comparison, taxa = value, diff = avb - ava, average = case_when(ava > avb ~ average, TRUE ~ -average), sd) %>% 
  left_join(phylum.info %>% select(phylum,taxa)) %>% 
  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
                           species == "NA" ~  paste0(genus, " sp."),
                           TRUE            ~ species)) %>%
  mutate(label2 = paste(phylum, label, sep = " | ")) %>% 
  mutate(Comparison = fct_relabel(Comparison, ~ str_replace(.,"_", " vs "))) %>% 
  nest(-Area) %>% 
  mutate(plot = map(data, function(.x){
    lims = c(-max(abs(.x$average) + (.x$sd)), max(abs(.x$average) + (.x$sd)))
    ggplot(.x, aes(y = average,
             x = fct_reorder(label2, average))) +
  geom_col(aes(fill = diff>0), size = 0.5, color = "black") +
      geom_errorbar(aes(ymin = average -sd, ymax = average + sd ), width = 0.25) +
  geom_hline(yintercept = 0, color = "black")+
  labs (x = "taxa",
        y = "Contribution to dissimilarity")+
  coord_flip(ylim=lims) +
    scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
  facet_wrap(~Comparison) +
  guides(fill = "none") +

    theme_light() + theme(axis.text.y = element_text(face = "bold.italic", colour ="#3B4252", hjust = 0 ),
                          strip.background = element_rect(fill = "#3B4252"),
                          strip.text = element_text(face = "bold", color = "#EBCB8B"))} 
  ))   -> right_most

right_most %>% pull(plot) %>% set_names(right_most$Area) -> right_top
right_top
```

### Final Figure 2

Here is the code for the figure that goes in the MS - the San Juan Island's is on the next chunk of code.

```{r assembling Figure 2}

left_top[["Hood Canal"]] + theme(legend.position  = "right", 
                                 axis.title.y = element_text(vjust = 0.5, angle = 0))  -> p1
 center_top[["Hood Canal"]] + theme(axis.title.y = element_text(vjust = 0.5, angle = 0)) -> p2
 right_top[["Hood Canal"]] +  theme(axis.title.y = element_blank())-> p3

layout5 <-c(
  patchwork::area(t= 1, l=1 , b=6 , r=4),
  patchwork::area(t= 7, l=1 , b=11 , r=4),
 # area(t= 5, b=5, l=4),
  patchwork::area(t= 1, l= 6, b= 11, r=9)
)

plot(layout5)
p1 + 
  p2 +
  # guide_area()+ 
  p3 + plot_layout(design = layout5) + plot_annotation(tag_levels = "A")  -> plot

ggsave(plot,filename = here("Figures","Manuscript",paste0("Figure2.", Sys.Date(), ".png")),
         width = 14,height = 11, dpi = "retina")

```

```{r assembling Figure for Supplementary material}

left_top[["San Juan Island"]] + theme(legend.position  = "right")  -> p1
 center_top[["San Juan Island"]] -> p2
 right_top[["San Juan Island"]] -> p3
p1 + 
  p2 +

  p3 + plot_layout(design = layout5) + plot_annotation(tag_levels = "A")  -> plot

ggsave(plot,filename = here("Figures","Supplementary",paste0("Supp.info.San.Juan.Island.Fig2", Sys.Date(), ".png")),
         width = 14,height = 11, dpi = "retina")

```

## Figure 3: a quick glance at spp responses
Get the models output and calculate the posterior across the simulated scenario

Now, calculate the posterior of each taxa across the 16,276 scenarios simulated

### Posterior per taxa per point in the future

We use the climate projection to predict changes in suitability

```{r}

all.sims.for.projection <- simData.nocrazys %>% 
  select(Area, Year, Temperature, TemperatureCelsius, pH, pH_units)

## Get 
models %>% 
  transmute(taxa, 
            post.draws = map2 (model, taxa, ~ posterior_predict(object = .x,
                                                                newdata = all.sims.for.projection,
                                                                draws = 100) %>% 
                                 as_tibble() %>% 
                                 t(.) %>% 
                                 as_tibble() %>% 
                                 rownames_to_column("sim.row") )) -> posterior.draws.all

# This is an object with as many rows as environmental points there are. The first two columns are taxa and nrow. The rest are the binary output of the posterior
# predict

posterior.draws.all %>% 
 left_join(phylum.info %>% dplyr::select(phylum, taxa)) %>% 
  group_by(phylum) %>% # Add phylum information
  nest() %>% 
   mutate (data = map (data, ~ .x %>% 
                        unnest() %>% 
                        pivot_longer(cols = c(-taxa, -sim.row),
                                    names_to = "draw.row", values_to = "presence") %>% 
  group_by(taxa, sim.row) %>% 
  summarise(prob.of.presence = mean(presence)) %>% 
  left_join(all.sims.for.projection %>%  
              rownames_to_column("sim.row"))

  )) -> posterior.probabilities.by.taxa
```
Now we have a nested dataframe per each phylum. Not that we need it. But the nested part is a dataframe with the mean result of the 100 draws for each environmental point.

### Richness by phyla with time
 First create the square matrix of points - the ggplot geometries requires one entry per grid - so we have to comply
```{r create full grid}

all.sims.for.projection  %>% 
  group_by(  Area) %>% 
  # filter (keep == "keep") %>%   
  summarise_at(c("TemperatureCelsius", "pH_units"), list(max = max, min = min)) %>%

  group_by ( Area) %>% 
  nest() %>% 
  mutate(data = map (data, ~ expand_grid(TemperatureCelsius = seq(round(.x$TemperatureCelsius_min,1), round(.x$TemperatureCelsius_max, 1),by = 0.1),
                                         pH_units = seq(round(.x$pH_units_min,1), round(.x$pH_units_max,1), by = 0.1)))) %>% 
  unnest(data) -> full.grid
```

```{r add standarized temperatures to the grid}

full.grid %>% 
  ungroup() %>% 
  rownames_to_column("Case") %>% 
  rename(Temperature = TemperatureCelsius,
         pH = pH_units) %>%

  pivot_longer(cols = c(Temperature, pH),
               names_to  = "Predictor",
               values_to = "value") %>% 
  group_by(Area, Predictor) %>% 
  nest() %>% 
  left_join(transforming.df) %>% 
  mutate(data = map2(data, to.std, ~.x %>% 
                       mutate(std = .y[1]+.y[2]*value ))) %>% 
  select(Area, Predictor, data) %>% 
  unnest(data) %>% 
  pivot_wider(names_from = Predictor,
              values_from = c(std,value)) %>% 
  rename(Temperature = std_Temperature,
         pH = std_pH,
         TemperatureCelsius = value_Temperature,
         pH_units           = value_pH) -> full.grid
```

`full.grid` has the temperature and pH values we want to project our models to - so now we use it with the `models` output.

```{r calculate the 100 draws from the posterior}
models %>% 
  ungroup() %>%  #rpk added
  transmute(taxa, 
            post.draws = map2 (model, taxa, ~ posterior_predict(object = .x,
                                                                newdata = full.grid,
                                                                draws = 100) %>% 
                                 as_tibble() %>% 
                                 t(.) %>% 
                                 as_tibble() %>% 
                                 rownames_to_column("sim.row") )) -> posterior.draws.raster
```

```{r calculate richness}
posterior.draws.raster %>% 
 left_join(phylum.info %>% dplyr::select(phylum, taxa)) %>% 
  group_by(phylum) %>% 
  nest() %>% 
  mutate(data = map (data, ~ .x %>% unnest(cols= post.draws) %>% 
                       pivot_longer(cols = c(-taxa, -sim.row),
                                    names_to = "draw.row", values_to = "presence") %>% 
                       group_by(draw.row, sim.row) %>% 
                       summarise(tot = sum(presence)) %>% 
                       group_by(sim.row) %>% 
                       summarise_at(.vars = "tot", .funs = list(Richness = mean,sd = sd)))) %>% 
  unnest(data) -> ready_plot

ready_plot %>% 
  left_join(full.grid %>%  
              rownames_to_column("sim.row")) -> ready_plot

# Normailzed the data for each phyla

ready_plot %>% 
  group_by(phylum) %>% 
  mutate(st.richness = Richness / max(Richness)) -> ready_plot

```

```{r plot3 function}
plot3 <- function(tibble,string , response) {

  response = enquo(response)

  tibble %>% 
    ungroup() %>% 
    mutate(Area = as.character(Area)) %>% 
    mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
                            TRUE          ~ Area)) %>% 
    group_by(TemperatureCelsius, pH_units, Area) %>%

    summarise (!!response := mean (!!response))  %>%

    ggplot(aes(x = round(TemperatureCelsius,10),
             y = round(pH_units, 10))) +

  geom_raster( interpolate = T,aes( fill = !!response)) +

  scale_fill_distiller(name = "Standarized Richness", type = "seq", palette = "Spectral",limits=c(0,1) ) +

  # geom_density2d(data = simData %>%
  #                   filter (Year %in% c(2015, 2095)),
  #                 aes(x= TemperatureCelsius, y = pH_units, group= Year, color = as.factor(Year)), bins = 5, alpha = 0.5) +

  scale_color_brewer(name = "Year",  type = "qual", palette = 4, direction = -1, breaks=c("2017","2095")) +
  ggforce::geom_mark_hull(data = all.sims.for.projection %>%
                            ungroup() %>%
                            mutate(Area = as.character(Area)) %>% 
                            mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
                            TRUE          ~ Area)) %>%
                            filter (Year %in% c(2017, 2095)) ,
                          aes(x= TemperatureCelsius, y = pH_units, group= Year, color = as.factor(Year),label = as.factor(Year)),
                          expand = unit(0.5, "mm"),
                 label.margin = margin(1, 1, 1, 1, "mm"),
                 label.buffer = unit(0, "mm"),
                 con.cap = 0 ,
                 alpha = 0.5) +
     facet_grid(.~Area) +
  guides(color = "none") +
    labs (x = "Temperature",
          y= "pH") +
    theme_minimal()+
    theme(legend.position = "bottom",
          legend.key.width  = unit(2 ,"cm"),
          axis.title.x = element_text(face = "bold"),
          legend.direction = "horizontal") +
    scale_x_continuous(position = "top")

}
```

```{r}
ready_plot %>% 
  nest(-phylum) %>% 
  mutate(plot = map2(data, phylum, ~ plot3(.x, .y, response = st.richness)))  -> list.of.richness.plots

```

```{r}

list.of.richness.plots %>%  pull(phylum) -> list.of.phyla

list.of.richness.plots %>% pull(plot) -> list.of.raster.plots

list.of.raster.plots <- set_names(list.of.raster.plots, nm = list.of.phyla)

```

# Assembling Figure 3

Following reviewer's suggestions, we are limiting the information displayed in the main Figures to two of most important phyla for coastal ecosystems dinamics: Diatoms and Dinoflagellates.

```{r select some }
posterior.probabilities.by.taxa %>% 
  ungroup() %>% 
  filter (phylum %in%c("Dinophyceae", "Bacillariophyta")) %>% 
  mutate(suma = map (data,~.x %>% group_by(taxa,Area, Year) %>%
                       summarise(tot = sum(prob.of.presence)) %>%
                       ungroup() %>% 
                       pivot_wider(names_from = Year, values_from = tot) %>% 
                        group_by(taxa) %>% mutate(diff = `2017` - `2095`)
                     )) %>% 
  select(phylum, suma) %>% 
  unnest(suma) %>%
  group_by(phylum) %>% arrange(desc(abs(diff))) %>% 
  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
                           species == "NA" ~  paste0(genus, " sp."),
                           TRUE            ~ species)) -> taxa.and.diff

taxa.and.diff %>%
  slice(1:5) %>%
  arrange(diff) %>% pull(taxa) -> ahorasi

taxa.and.diff %>% 
  slice(1:5) %>%
  arrange(diff) %>% pull(label) -> reorder.labels

# Write to file the posterior probabilietes by taxa so Ryan can play with it
# posterior.probabilities.by.taxa %>% 
#   unnest(data) %>% 
#   write_csv(here("Output", "posterior.probability.by.taxa.csv"))
```

```{r the final one}

posterior.probabilities.by.taxa %>%

ungroup() %>%
filter (phylum %in%c("Dinophyceae", "Bacillariophyta")) %>%
mutate( plot.by.taxa = map2 (data, phylum, ~ .x %>%
filter(taxa %in% ahorasi) %>% 
separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
species == "NA" ~  paste0(genus, " sp."),
TRUE            ~ species))    %>%
mutate(Area = as.character(Area)) %>%
mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
TRUE          ~ Area)) %>%
mutate(Year = fct_rev(as.factor(Year) )) %>%
ggplot(aes(x= prob.of.presence,
y = fct_relevel(label,reorder.labels))) +
#   geom_violin(aes(fill= (Year)), weight = 2, adjust = 1, draw_quantiles =  0.5, scale = "count")+
scale_fill_brewer(name = "Year",  type = "qual", palette = 4, direction = 1, breaks=c("2017","2095")) +
labs (x = "Probability of presence",
y = "") +
facet_wrap(~Area, ) +
theme_minimal() +
theme(strip.text.y = element_text(angle = 0),
axis.text.y = element_text(face = "italic"),
legend.direction = "horizontal",
legend.position  = "bottom",
plot.title.position = "plot",
plot.title = element_text(face = "bold",vjust =  -5 ),
axis.title.x.top  = element_text(face = "bold",vjust = 3) ) +
ggtitle(.y)  +
scale_x_continuous(limits = c(0,1),position = "top") )) %>%
pull(plot.by.taxa) %>% set_names(.$phylum) -> fig3.left

map(fig3.left, function(x) x+ geom_boxplot(aes(fill = as.factor(Year)))) -> fig3.left

posterior.probabilities.by.taxa %>%
ungroup() %>%
filter (phylum %in%c("Dinophyceae", "Bacillariophyta")) %>%
  pull(phylum) %>%
  set_names(fig3.left, .) -> fig3.left

list.of.raster.plots[c("Dinophyceae", "Bacillariophyta")] -> fig3.right

layout <- "
AA
AA
FF
EE
EE
GG
"

wrap_plots(A = fig3.left[["Dinophyceae"]] +   labs(tag = "A"),
           F = plot_spacer(),
           E = fig3.left[["Bacillariophyta"]] ,
           G = guide_area(),
            design = layout, heights = c(5,5,1,5,5,5), guides = "collect") -> left

layout2 <- "
#BB
#BB
JJJ
#HH
#HH
III
"
wrap_plots(B = fig3.right[["Dinophyceae"]] +   labs(tag = "B"),
           J = plot_spacer(),
           H = fig3.right[["Bacillariophyta"]] ,
           I = guide_area(),
            design = layout2,heights = c(5,5,1,5,5,5),
           widths = c(1,2,2), guides = "collect") -> right

(left|right)+plot_layout(widths = c(2,3))  -> plot 
   ggsave(plot , filename = here("Figures","Manuscript", paste0("Top_fig3_",Sys.Date(),".Other.sel.png")), units = "in", dpi = "retina", width = 20, height = 12)

```

#### Bonus track:Create phylum-specfic plots for Supplementary Information

```{r}
posterior.probabilities.by.taxa %>% 
  ungroup() %>% 
  # filter (phylum %in%c("Dinophyceae", "Bacillariophyta")) %>% 
  mutate(suma = map (data,~.x %>% group_by(taxa,Area, Year) %>%
                       summarise(tot = sum(prob.of.presence)) %>%
                       ungroup() %>% 
                       pivot_wider(names_from = Year, values_from = tot) %>% 
                        group_by(taxa) %>% mutate(diff = `2017` - `2095`)
                     )) %>% 
  select(phylum, suma) %>% 
  unnest(suma) %>%
  group_by(phylum) %>% arrange(desc(abs(diff))) %>% 
  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
                           species == "NA" ~  paste0(genus, " sp."),
                           TRUE            ~ species)) -> taxa.and.diff

taxa.and.diff %>%
  slice(1:5) %>%
  arrange(diff) %>% pull(taxa) -> ahorasi

taxa.and.diff %>% 
  slice(1:5) %>%
  arrange(diff) %>% pull(label) -> reorder.labels
```

```{r}
posterior.probabilities.by.taxa %>%

ungroup() %>%
mutate( plot.by.taxa = map2 (data, phylum, ~ .x %>%
#filter(taxa %in% ahorasi) %>% 
separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
species == "NA" ~  paste0(genus, " sp."),
TRUE            ~ species))    %>%
mutate(Area = as.character(Area)) %>%
mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
TRUE          ~ Area)) %>%
mutate(Year = fct_rev(as.factor(Year) )) %>%
ggplot(aes(x= prob.of.presence,
# y = fct_relevel(label,reorder.labels))) +
 y = label)) + 
#   geom_violin(aes(fill= (Year)), weight = 2, adjust = 1, draw_quantiles =  0.5, scale = "count")+
scale_fill_brewer(name = "Year",  type = "qual", palette = 4, direction = 1, breaks=c("2017","2095")) +
labs (x = "Probability of presence",
y = "") +
facet_wrap(~Area, ) +
theme_minimal() +
theme(strip.text.y = element_text(angle = 0),
axis.text.y = element_text(face = "italic"),
legend.direction = "horizontal",
legend.position  = "bottom",
plot.title.position = "plot",
plot.title = element_text(face = "bold",vjust =  -5 ),
axis.title.x.top  = element_text(face = "bold",vjust = 3) ) +
ggtitle(.y)  +
scale_x_continuous(limits = c(0,1),position = "top") )) %>%

pull(plot.by.taxa)   -> Suppl.Material.plots

map(Suppl.Material.plots, function(x) x+ geom_boxplot(aes(fill = as.factor(Year)))) -> Suppl.Material.plots

posterior.probabilities.by.taxa %>%
ungroup() %>%
  pull(phylum) %>%
  set_names(Suppl.Material.plots, .) -> Suppl.Material.plots

list.of.raster.plots[[1]]

names(Suppl.Material.plots) == names(list.of.raster.plots)

map2(Suppl.Material.plots, list.of.raster.plots, 
     function(.x,.y){

       .x  +plot_spacer()+ .y + plot_layout(widths = c(4,1,4)) -> plot

        return(plot)

     }) -> Suppl.Material.plots.final

names (Suppl.Material.plots.final) <- names(Suppl.Material.plots)
map2 (Suppl.Material.plots.final,names(Suppl.Material.plots.final),   ~ ggsave(plot = .x, filename = here("Figures","Supplementary",paste0("Supp.fig",.y,".png")),device = "png",
                units = "in", dpi = "retina", width = 20, height = 12)
                                                                               )
```

### 2020-08-03 Scripts/All.figures.that.go.in.the.MS.reviewed.Rmd
Updated scripts

--- 2019-11-07 Scripts/All.figures.that.go.in.the.MS.Rmd
@@
 editor_options: 
-  chunk_output_type: inline
+  chunk_output_type: console
 ---
@@
 library (factoextra)
-library(nnet)
-#library (OneR)
+library (nnet)
+library (here)
+library (conflicted)
+conflict_prefer("filter", "dplyr")
+conflict_prefer("select", "dplyr")
 #functions for future Salish Sea
-models <- readRDS("../../Analysis/Splitting.data.by.lifestyle/ModelOutputs/full_model_output_rpk_TpH.standardized_2019-08-13_1567.RDS")
-source("RPK_climate_model_functions.R")
-
-source("../../tibble.to.matrix.r")
+models <- readRDS(here("Input","full_model_output_rpk_TpH.standardized_nonSeasonal2020-06-24_3262.RDS"))
+source(here("Scripts","RPK_climate_model_functions_REVISIONS.R"))
+
+source(here("Scripts","tibble.to.matrix.r"))
+
+stds_to_value <- read_rds(here("Input", "std_to_real.rds"))
+
 ```
@@
 ```{r }
-data.plankton <- read_csv("../../Data/Combined_Biol_Env_Plankton.csv")
+data.plankton <- read_csv(here("Input","Combined_Biol_Env_Plankton.csv"))
 
@@
 
-env.data <- read_csv("../../Analysis/Splitting.data.by.lifestyle/env.data.for.log.regression.csv")
+env.data <- read_csv(here("Input","env.data.for.log.regression.csv"))
 
@@
   gather(-Season, -sample, -Area,  key = "Predictor", value = "value") %>%
-  group_by(Predictor) %>%   #RPK: this was missing, which messed up the std calculations 
+  group_by(Area, Predictor) %>%   #RPK: this was missing, which messed up the std calculations 
   mutate(std = vegan::decostand(value, method = "standardize")) -> now.env.data.with.std
+
+#### TODO: remove
 
@@
 now.env.data.with.std %>%
-  group_by(Predictor) %>% 
-  summarise(mean = mean(value),
-            sd = sd(value)) %>% 
-  group_by(Predictor) %>% 
-  nest(params = c(mean, sd) )-> transforming.df
-
-phylum.info <- read_csv("../../Supp.Material.OA.eDNA/Input/higher_taxonomy.csv") %>% 
+  group_by(Area, Predictor) %>%
+  # summarise(mean = mean(value),
+  #           sd = sd(value)) %>%
+  # group_by(Area, Predictor) %>%
+  nest() %>% 
+  summarise(tovalue = map(data, ~ lm(value ~ std, data = .x) ),
+            tostd   = map(data, ~ lm(std ~ value, data = .x)))-> transforming.df
+
+transforming.df[1,4] %>% pull -> test
+broom::tidy(test[[1]])
+extract <- function (lm) {broom::tidy(lm)$estimate}
+
+extract(test[[1]])
+
+transforming.df %>% 
+  mutate(to.values = map(tovalue, extract),
+         to.std= map(tostd, extract)) ->transforming.df
+
+phylum.info <- read_csv(here("Input","higher_taxonomy.csv")) %>% 
   unite(family, genus, species, sep = "|", col = "taxa")
@@
 ```{r}
-#read in observed data, which will be the starting point for adjusting in future projections          
-obsData <- models$data[[1]] %>% 
-  dplyr::select(Temperature, pH, Area, Season)
-
-YearSpan <- c(2017, 2095)#seq.int(2015, 2095, 20)  #taking these individual years to model
-Nyears <- length(YearSpan)
-NiterationsPerYear <- 50  #number of realizations of the future environmental conditions
-Nreps <- Nyears*NiterationsPerYear
-
-#here, replicate the observed dataframe Nreps times, where Nreps = Nyears * NiterationsPerYear
-simData <- obsData[rep(row.names(obsData), Nreps), 1:4] %>% 
-  mutate(Year = rep(YearSpan, 
-                    each = nrow(obsData)*NiterationsPerYear),
-         Iteration = rep(1:NiterationsPerYear, times = nrow(obsData)*Nyears))
-
-```
-
-function to generate future climate distributions isn't vectorized, so now adjust Temperature and pH for each year in a loop:
-these use stochastic sampling functions r_T_year_STD() and r_pH_year_STD()
-
-```{r}
-set.seed(100)
-      for (i in unique(simData$Year)){
-        for (j in 1:NiterationsPerYear) {
-          for (k in unique(simData$Season)) {
-            for (m in unique(simData$Area)) {
-
-              dataSubset <- simData$Temperature[simData$Year == i & 
-                                                  simData$Iteration == j &
-                                                  simData$Season == k &
-                                                  simData$Area == m]
-
-        simData$Temperature[simData$Year == i & 
-                              simData$Iteration == j &
-                                simData$Season == k &
-                                  simData$Area == m] <-     #future condition is ....
-
-                                             dataSubset +    #the current condition, plus...
-
-                      r_T_year_STD_subset(i,  #the difference seen in the specified year
-                                          length(dataSubset),  
-                                          2017, # standardized to reference year
-                                          season = k,
-                                          area = m)
-
-        simData$pH[simData$Year == i & 
-                              simData$Iteration == j &
-                              simData$Season == k &
-                              simData$Area == m] <-     #future condition is ....
-          simData$pH[simData$Year == i & 
-                                simData$Iteration == j &
-                                simData$Season == k &
-                                simData$Area == m] +    #current condition, plus...
-
-          r_pH_year_STD_subset(i,  #the difference seen in the specified year
-                              10,  
-                              2017, # standardized to reference year
-                              season = k,
-                              area = m)[1] #one draw per iteration for the simulated set of temperatures
+Year.area.dataset <- function(Area, Year, n){
+
+    Temp.conv <- stds_to_value[which (stds_to_value$Area == Area & stds_to_value$key == "Temperature"),"toValue"] %>% pull %>% .[[1]]
+    pH.conv   <- stds_to_value[which (stds_to_value$Area == Area & stds_to_value$key == "pH")         ,"toValue"] %>% pull %>% .[[1]]
+
+    one<-  data.frame(
+              Area = Area,
+              Year = Year,
+              Temperature = r_Temperature_year_STD(Year, Area, n, 2017),
+              pH          = r_pH_year_STD(Year, Area, n, 2017)
+              )
+
+    one$TemperatureCelsius = sapply(one$Temperature,function(x){Temp.conv$coefficients[1]+Temp.conv$coefficients[2]*x})
+    one$pH_units            = sapply(one$pH,function(x){pH.conv$coefficients[1]+pH.conv$coefficients[2]*x})
+
+    return(one)
+
 }
-         }
-        }
-      }
-
-```
-
-### Convert standardized units to natural units using reference year
-
-```{r}
-#observed = (std * SD) + mean
-
-#useful later:
-reference_Temperature_SD <- sd(now.env.data$Temperature)
-reference_Temperature_mean <- mean(now.env.data$Temperature)
-reference_pH_SD <- sd(now.env.data$pH)
-reference_pH_mean <- mean(now.env.data$pH)
-
-simData <- simData %>% 
-  rename(TemperatureSTD = Temperature,
-        pH_STD = pH) %>% 
-  mutate(TemperatureCelsius = (TemperatureSTD * reference_Temperature_SD) + reference_Temperature_mean,
-         pH_units = (pH_STD * reference_pH_SD) + reference_pH_mean)
-
-#check that this is perfectly linear
-# simData %>%
-#   ggplot(aes(x = TemperatureSTD, y = TemperatureCelsius)) +
-#   geom_point()
-
-#check for trends
-# simData %>% 
-#   select(Year, Season, Area, TemperatureCelsius) %>% 
-#   group_by(Year, Season, Area) %>% 
-#   summarize(mean(TemperatureCelsius))
-```
-If we look at this data, does it have crazy points
-```{r limits}
-simData %>% 
-  group_by(Year, Season, Area) %>% 
-  summarise_at(c("TemperatureCelsius", "pH_units"), .funs = list(min = min, max = max))
-
-# Yep, a bit crazy. Apply 90% interval
-
-simData %>%
+
+j <- Year.area.dataset("Hood Canal", 2095, 5000)
+k <- Year.area.dataset("Hood Canal", 2017, 5000)
+l <- Year.area.dataset("SJI", 2095, 5000)
+m <- Year.area.dataset("SJI", 2017, 5000)
+
+simulated.scenario <- rbind(j,k,l,m)
+
+rm(j,k,l,m)
+
+simulated.scenario %>%
   group_by(Year, Area) %>% 
@@
                ppH =pnorm(pH_units, mean  = .y[["ph"]]$estimate[1], sd = .y[["ph"]]$estimate[2]) )
-
       })
@@
   unnest(data) %>% 
-  mutate(keep = case_when(pT < 0.1 | ppH < 0.05 | pT > 0.9 | ppH > 0.9 ~ "Discard",
+  mutate(keep = case_when(pT < 0.05 | ppH < 0.05 | pT > 0.95 | ppH > 0.95 ~ "Discard",
                           TRUE                   ~ "keep")) %>% 
   filter (keep == "keep")-> simData.nocrazys
+
 simData.nocrazys %>% 
-  rename(Temperature = TemperatureSTD,
-         pH = pH_STD) %>% 
-saveRDS( file = "../../Analysis/Splitting.data.by.lifestyle/all.sim.data_climateScenarios_no.outliers.RDS")
+  select(-values) %>% write_csv("Output/simulation.from.std.functions.cleaned.csv")
+```
+
+### Save future scenarios
+
+```{r}
+
 simData.nocrazys %>%
-  group_by( Area, Season) %>% 
+
+saveRDS( file = here("Output","all.sim.data_climateScenarios_no.outliers.RDS"))
+
+simData.nocrazys %>% 
+  group_by( Area, Year) %>% 
   summarise_at(c("TemperatureCelsius", "pH_units"), .funs = list(min = min, max = max))
 
+```
+
+### Seasonality
+
+To proyect the different community clusters fate into the pH - Temperature continuum, we need to include the seasonality or ditch it from the CAP Analysis. Here I am 
+Now create a seasonal data, we can sample each year's data and select them with the profile for the seasonal data in the real dataset.
+By calculating the quantiles of the Temperature distributions, we see that both in the San Juan Islands and the Hood Canal, the Winter months occupy the lower 30% quantile of the data.
+In order to create a matrix of possible ranges of pH and Temperature for each season in the future scenario, we will apply these quantiles to the simulated data
+
+```{r}
+simData.nocrazys %>% 
+  mutate(Season = case_when(pT < 0.3 ~ "Winter",
+                            TRUE     ~ "Summer")) %>% 
+  ggplot(aes(x = TemperatureCelsius, fill = Season)) +
+  geom_density_ridges_gradient(aes(y = as.factor(Year))) +
+  facet_wrap(~Area, nrow = 2)
+```
+
+Now get the ranges and create the grid for each season / Area
+
+```{r get the ranges}
+
+simData.nocrazys %>% 
+  mutate(Season = case_when(pT < 0.3 ~ "Winter",
+                            TRUE     ~ "Summer")) %>% 
+  group_by(Area, Season) %>% 
+  summarise_at (c("TemperatureCelsius", "pH_units"), .funs = list(min = min, max = max)) -> seasonal.ranges
+```
+
+```{r}
+simData.nocrazys %>% 
+  select(Area, Temperature, pH, Year) %>% 
+  pivot_longer(cols = c(pH, Temperature), names_to = "Predictor", values_to = "value") -> simData.nocrazys.long
+
+ simTemp <- r_T_year_STD(2017, 1000, 2017) %>% 
+    sort()
+  #what quantile of the simulated data is each of the observed points in?
+  simECDF <- ecdf(simTemp) #create ecdf function from sim distribution, then find quantile for each observed point
+quantile_df <- function(x, probs, na.rm =F, names = F, type = 7, ...){
+  z <- quantile(x, probs, na.rm, names, type)
+  output <- tibble(id = probs,
+                       values = z)
+
+      return(output)   
+}
+
+ models$data[[1]] %>% 
+    select(-sample, -Noccur, -phylum, -binary) %>% 
+    pivot_longer(cols = c(-Season, -Area),names_to = "Predictor", values_to = "value") %>% 
+    filter (Predictor == "Temperature") %>% 
+   group_by(Season, Area, Predictor) %>%
+    nest()-> testdata
+
+testdata %>% 
+  unnest(data) %>% ungroup() %>% 
+  select(-Season) %>% 
+  group_by(Area, Predictor) %>% 
+  nest(value) %>% 
+  mutate(Season = "All") %>% 
+  bind_rows(testdata, .) %>%
+
+  mutate(
+    betas = map(data, function(.x){
+      pull(.x) %>% 
+      sort() %>% 
+        simECDF() %>% 
+        fitdist("beta")}),
+    quantiles = map (data,  ~ quantile_df(.x$value, probs = c(0, 0.25, 0.5, 0.75, 1)))) -> testdata
+
+testdata %>% 
+  left_join(transforming.df %>% select(Area, Predictor, to.values)) %>% 
+  mutate(quantiles = map2(quantiles, to.values, ~.x %>% 
+                       mutate(real = .y[1]+.y[2]*values ))) %>% 
+# unnest(quantiles) %>% 
+#   select(-values) %>% 
+#   pivot_wider(names_from = id, values_from = real) %>% 
+#   filter(Predictor == "Temperature") %>% 
+#   arrange(Area)
+  nest(-Area) %>% 
+  expand_grid(., Year = c(2017, 2095)) %>% 
+  unnest(data) %>% 
+  nest_join(simData.nocrazys.long, by = c("Area","Predictor", "Year"), name = "new") %>% 
+  mutate(lims = map2(new, betas,
+                     function(.x, .y){
+                       .x %>%
+                         pull(value) %>% 
+                         sort() -> temps 
+          base::sample(temps, 
+                       size = 500, 
+                       replace = TRUE, 
+                       prob = rbeta(length(temps), shape1 = .y$estimate[1], shape2 =.y$estimate[2])) -> temps
+
+          quantile_df(temps, probs = c(0,0.2, 0.5, 0.7, 1))
+          } )) %>%
+
+  mutate(lims.real = map2(lims, to.values, ~.x %>% 
+                       mutate(real = .y[1]+.y[2]*values ))) %>% 
+# unnest(lims.real) %>% 
+  unnest(quantiles) %>% 
+  select(-values) %>% 
+  pivot_wider(names_from = id, values_from = real)
 ```
@@
 ```{r Base map}
-map <- readRDS("map.plot.rds")
+map <- readRDS(here("Input","map.plot.rds"))
 
@@
 ```{r}
-boxplot.richness <- readRDS("richness.rds")
+boxplot.richness <- readRDS(here("Input","richness.rds"))
 
@@
           labels = c(NA,"C")) +
-  ggsave(filename = "Fig1.png", width = 14, height = 14, dpi = "retina")
+  ggsave(filename = here("Figures","Fig1.png"), width = 14, height = 14, dpi = "retina")
 
@@
 joined.dataset.for.CAP %>% 
+  ungroup %>% 
  # select(DIC = mean.DIC.new.sal, Salinity = Salinity.new, Temperature, pH = pH_new, everything())
@@
   unnest() %>% #filter (Area == "Hood Canal", Season == "Summer") %>% 
-  write_csv("../../Analysis/Splitting.data.by.lifestyle/clusters.of.events.csv")
+  write_csv(here("Output","clusters.of.events.csv"))
 Clusters.analysis.and.cap %>% pull(Final.plot2) 
@@
 
-clusters <- read.csv("../../Analysis/Splitting.data.by.lifestyle/clusters.of.events.csv") %>% 
+clusters <- read.csv(here("Input","clusters.of.events.csv")) %>% 
   filter(Season == "Summer",
@@
 # right_top + coord_equal () -> right_top 
- ggsave("Figure2.topright.png", width = 7)
+ ggsave(here("Figures","Figure2.topright.png"), width = 7)
 ```
@@
 
-clusters.all <- read.csv("../../Analysis/Splitting.data.by.lifestyle/clusters.of.events.csv")
+clusters.all <- read.csv(here("Input","clusters.of.events.csv") )
 
@@
 simData.nocrazys %>% 
-  mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
+  ungroup %>% 
+  mutate(Area = as.character(Area),
+         Area = case_when(Area == "SJI" ~ "San Juan Island",
                           TRUE          ~ Area)) %>% 
-  group_by( Area, Season) %>% 
+  group_by( Area) %>% 
   summarise_at(c("TemperatureCelsius", "pH_units"), .funs = list(min = ~ min(round(.x, 1)), max = ~max(round(.x, 1)))) %>% 
@@
   filter (p < 0.05) %>% 
-  write_csv("../../Analysis/Splitting.data.by.lifestyle/simper.HoodCanal.summer.csv")
+  write_csv(here("Output","simper.HoodCanal.summer.csv"))
 
@@
    pull(summary.simper)-> list.of.simpers
+names(list.of.simpers) <- paste(SIMPER.clusters$Area, SIMPER.clusters$Season, sep = "_")
 
  list.of.simpers %>%
-   map_df(~bind_rows(.x, .id = 'Comparison'), .id = 'var1') %>%
+  map(~ .x %>% map(~ bind_rows(.)) %>% 
+                          bind_rows(., .id = "Comparison")) %>% 
+  bind_rows(., .id = "var1") %>%
    separate(var1, into = c("Area", "Season"), sep = "_") %>%
@@
   # rename(taxa = value) %>% 
-  write_csv("../../Analysis/Splitting.data.by.lifestyle/simper.all.csv")
+  write_csv(here("Output","simper.all.csv") )
 
@@
 ```{r}
-taxa.to.follow <- read_csv("../../Analysis/Splitting.data.by.lifestyle/simper.HoodCanal.summer.csv")
+taxa.to.follow <- read_csv(here("Output","simper.HoodCanal.summer.csv"))
 
@@
 
-all.taxa.to.follow <- read_csv("../../Analysis/Splitting.data.by.lifestyle/simper.all.csv")
+all.taxa.to.follow <- read_csv(here("Output","simper.all.csv"))
 
@@
 
-ggsave(filename = paste0("Figure2.", Sys.Date(), ".png"),
+ggsave(filename = here("Figures",paste0("Figure2.", Sys.Date(), ".png")),
          width = 14,height = 5, dpi = "retina")
@@
 list.of.right.plots
+Clusters.analysis.and.cap %>% 
+  mutate (label = paste (Area, Season, sep = " ")) %>% 
+  pull(label) -> list.of.labels
 
@@
      .f = function(a,b,c) plot_grid(a,b, align = "v", nrow = 2, rel_heights = c(1, 0.1), labels = c) +
-       ggsave(filename = paste0(c,"CAP.plot.png"),  width = 14,height = 5, dpi = "retina")) -> plots.ready.to.go
-
-plots.ready.to.go %>% saveRDS("All_Cap_for_suppplemental.rds")
+       ggsave(filename = here("Figures",paste0(c,"CAP.plot.png")),  width = 14,height = 5, dpi = "retina")) -> plots.ready.to.go
+
+plots.ready.to.go %>% saveRDS(here("Output","All_Cap_for_suppplemental.rds"))
 ```
@@
 
-```{r}
-## Rename variables so the model can be projected
+We use the climate projection to predict changes in suitability
+
+```{r}
 
 all.sims.for.projection <- simData.nocrazys %>% 
-  rename(Temperature = TemperatureSTD,
-         pH = pH_STD)
+  select(Area, Year, Temperature, TemperatureCelsius, pH, pH_units)
 
@@
   )) -> posterior.probabilities.by.taxa
-
-posterior.probabilities.by.taxa %>% mutate(suma = map (data,~.x %>% group_by(taxa, Year) %>% summarise(tot = sum(prob.of.presence)) %>% ungroup() %>% 
+```
+Now we have a nested dataframe per each phylum. Not that we need it. But the nested part is a dataframe with the mean result of the 100 draws for each environmental point.
+
+```{r subset of taxa for manuscript}
+## Now create a set of phyla and taxa to show on figure 3
+
+posterior.probabilities.by.taxa %>%
+  mutate(suma = map (data,~.x %>% group_by(taxa, Year) %>%
+                       summarise(tot = sum(prob.of.presence)) %>%
+                       ungroup() %>% 
                        pivot_wider(names_from = Year, values_from = tot) %>% 
-                                                         group_by(taxa) %>% mutate(diff = `2017` - `2095`))) %>% select(phylum, suma) %>% unnest(suma) %>% arrange(desc(abs(diff))) %>% 
-  filter (phylum %in% c("Bacillariophyta", "Dinophyceae", "Haptophyceae", "Arthropoda"))
-
+                       group_by(taxa) %>% mutate(diff = `2017` - `2095`))) %>% 
+  select(phylum, suma) %>% 
+  unnest(suma) %>%
+  arrange(desc(abs(diff))) %>% 
+  filter (phylum %in% c("Bacillariophyta", "Dinophyceae", "Haptophyceae", "Arthropoda")) %>% 
+  group_by(phylum) %>% 
+  slice(1:5) %>% 
+  pull(taxa)-> taxa.for.Fig3
+
+posterior.probabilities.by.taxa %>%
+  mutate(suma = map (data,~.x %>% group_by(taxa,Area, Year) %>%
+                       summarise(tot = sum(prob.of.presence)) %>%
+                       ungroup() %>% 
+                       pivot_wider(names_from = Year, values_from = tot) %>% 
+                        group_by(taxa) %>% mutate(diff = `2017` - `2095`)
+                     )) %>% 
+  select(phylum, suma) %>% 
+  unnest(suma) %>%
+ # arrange(desc(abs(diff))) %>% 
+  group_by(phylum) %>% 
+  arrange(desc(abs(diff))) %>%
+  slice(1:5) %>% 
+  ggplot(aes(x = fct_reorder(phylum,abs(diff)), y = abs(diff))) +
+  geom_boxplot()+
+  geom_point(aes(color = diff>0)) +
+  coord_flip()
+
+posterior.probabilities.by.taxa %>%
+  mutate(suma = map (data,~.x %>% group_by(taxa,Area, Year) %>%
+                       summarise(tot = sum(prob.of.presence)) %>%
+                       ungroup() %>% 
+                       pivot_wider(names_from = Year, values_from = tot) %>% 
+                        group_by(taxa) %>% mutate(diff = `2017` - `2095`)
+                     )) %>% 
+  select(phylum, suma) %>% 
+  unnest(suma) %>%
+ # arrange(desc(abs(diff))) %>% 
+  group_by(phylum) %>% 
+  arrange(desc(abs(diff))) %>%
+  slice(1:5) %>% 
+  ungroup() %>% 
+  filter(phylum %in% c("Bacillariophyta", "Dinophyceae", "Haptophyceae", "Chlorophyta")) %>% 
+  distinct(taxa) %>% 
+  pull() -> taxa.for.Fig3.2
+```
+
+```{r plotting them}
+### some for fig3 ### 
+posterior.probabilities.by.taxa %>% 
+  filter (phylum %in% c("Bacillariophyta", "Dinophyceae", "Haptophyceae", "Arthropoda")) %>% 
+  mutate( plot.by.taxa = map2 (data, phylum, ~ .x %>% 
+                                 filter(taxa %in% taxa.for.Fig3) %>% 
+      separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
+  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
+                           species == "NA" ~  paste0(genus, " sp."),
+                           TRUE            ~ species))    %>% 
+    mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
+                            TRUE          ~ Area)) %>% 
+    mutate(Year = fct_rev(as.factor(Year) )) %>%
+
+  ggplot(aes(x= prob.of.presence,
+             y = label)) +
+  #geom_density_ridges(fill = as.factor(Year))
+  # stat_density_ridges(aes(fill = as.factor(Year)),
+  #    geom = "density_ridges_gradient", calc_ecdf = TRUE,
+  #   quantiles =2, quantile_lines = TRUE, alpha = 0.5
+  # ) +
+    geom_violin(aes(fill= (Year)), weight = 2, adjust = .5, draw_quantiles =  0.5)+
+# stat_summary(fun.y=mean, geom="line", aes(group=Year)) +
+  scale_fill_brewer(name = "Year",  type = "qual", palette = 4, direction = 1, breaks=c("2017","2095")) + 
+  labs (x = "Probability of presence",
+        y = "") +
+    facet_wrap(~Area) +
+
+ # facet_grid(phylum ~ Area) + coord_flip(xlim = c(0,1)) +
+ theme(strip.text.y = element_text(angle = 0),
+       axis.text.y = element_text(face = "italic")) +
+   ggtitle(.y) +
+  theme(legend.direction = "horizontal",
+        legend.position  = "bottom") ))   -> list.of.plots.for.fig3
+
+#A new subset to high;ight change
+
+posterior.probabilities.by.taxa %>% 
+  filter (phylum %in% c("Bacillariophyta", "Dinophyceae", "Haptophyceae", "Chlorophyta")) %>% 
+  mutate( plot.by.taxa = map2 (data, phylum, ~ .x %>% 
+                                 filter(taxa %in% taxa.for.Fig3.2) %>% 
+      separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
+  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
+                           species == "NA" ~  paste0(genus, " sp."),
+                           TRUE            ~ species))    %>% 
+    mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
+                            TRUE          ~ Area)) %>% 
+    mutate(Year = fct_rev(as.factor(Year) )) %>%
+
+  ggplot(aes(x= prob.of.presence,
+             y = label)) +
+  #geom_density_ridges(fill = as.factor(Year))
+  # stat_density_ridges(aes(fill = as.factor(Year)),
+  #    geom = "density_ridges_gradient", calc_ecdf = TRUE,
+  #   quantiles =2, quantile_lines = TRUE, alpha = 0.5
+  # ) +
+    geom_violin(aes(fill= (Year)), weight = 2, adjust = 1, draw_quantiles =  0.5, scale = "count")+
+# stat_summary(fun.y=mean, geom="line", aes(group=Year)) +
+  scale_fill_brewer(name = "Year",  type = "qual", palette = 4, direction = 1, breaks=c("2017","2095")) + 
+  labs (x = "Probability of presence",
+        y = "") +
+    facet_wrap(~Area) +
+
+  theme_minimal() +
+
+ theme(strip.text.y = element_text(angle = 0),
+       axis.text.y = element_text(face = "italic"),
+       legend.direction = "horizontal",
+       legend.position  = "bottom",
+       plot.title.position = "plot",
+       plot.title = element_text(face = "bold",vjust =  -5 ),
+       axis.title.x.top  = element_text(face = "bold",vjust = 3) ) +
+   ggtitle(.y)  + 
+  scale_x_continuous(limits = c(0,1),position = "top")))  -> list.of.plots.for.fig3.2
+
+list.of.plots.for.fig3.2$plot.by.taxa[[1]]
+```
+
+```{r plots for supp}
+## some for suppmat
 posterior.probabilities.by.taxa %>% 
@@
 
+  ggplot(aes(x= prob.of.presence,
+             y = label)) +
+  #geom_density_ridges(fill = as.factor(Year))
+ #stat_density_ridges(aes(fill = as.factor(Year)),
+    # geom = "density_ridges_gradient", calc_ecdf = TRUE,
+  #  quantiles =2, quantile_lines = TRUE, alpha = 0.5
+  #) +
+   geom_violin(aes(group = ))
+  scale_fill_brewer(name = "Year",  type = "qual", palette = 2) + 
+  labs (x = "Probability of presence",
+        y = "") +
+    facet_wrap(~Area) +
+
+ # facet_grid(phylum ~ Area) + coord_flip(xlim = c(0,1)) +
+ theme(strip.text.y = element_text(angle = 0),
+       axis.text.y = element_text(face = "italic")) +
+   ggtitle(.y) +
+  theme(legend.direction = "horizontal",
+        legend.position  = "bottom") ))   -> list.of.plots
+
+png(file = here("Figures","taxadensityplots.by.phylum","all.by.area.png"), width = 14, units = "in", res = 300)
+list.of.plots %>% pull(plot.by.taxa)
+dev.off()
+#
+# list.of.plots %>% slice(1) %>% pull(plot.by.taxa)-> one.plot
+# legend.left <- get_legend(one.plot[[1]] )
+# 
+# list.of.plots %>% mutate(plot.by.taxa= map(plot.by.taxa, ~.x + theme(legend.position = "none") )) -> list.of.plots
+
+```
+
+#### Now do change in suitability by taxa by time - for the supp
+
+```{r}
+selected.taxa <- models %>% select(taxa) %>% filter (str_detect(taxa, "Chaetoceros|huxleyi|Alexandrium|Nitzschia|Karlodinium")) %>% pull(taxa)
+```
+
+Here I will unnest the whole thing - maybe better to always keep it unnested
+
+```{r}
+posterior.probabilities.by.taxa %>% 
+  unnest(data) ->  posterior.probabilities.by.taxa
+```
+
+Now something for the supporting information
+
+```{r}
+posterior.probabilities.by.taxa %>% 
+  filter (taxa %in% selected.taxa) %>% 
+  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
+  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
+                           species == "NA" ~  paste0(genus, " sp."),
+                           TRUE            ~ species)) %>% 
+  group_by(phylum) %>% 
    ggplot(aes(x= prob.of.presence,
@@
         y = "") +
-    facet_wrap(~Area) +
-
- # facet_grid(phylum ~ Area) + coord_flip(xlim = c(0,1)) +
- theme(strip.text.y = element_text(angle = 0),
-       axis.text.y = element_text(face = "italic")) +
-   ggtitle(.y) +
-  theme(legend.direction = "horizontal",
-        legend.position  = "bottom") ))   -> list.of.plots
-
-# posterior.draws.all %>% 
-#  left_join(phylum.info %>% dplyr::select(phylum, taxa)) %>% 
-#   group_by(phylum) %>% 
-#   nest() %>% pull(phylum) -> list.of.names
-
-# map(list.of.plots,  ~ .x + facet_wrap(~Area)  ) -> list.of.plots
-# 
-#pdf(file = "taxadensityplots.by.phylum/all.by.area.pdf")
-png(file = "taxadensityplots.by.phylum/all.by.area.png", width = 14, units = "in", res = 300)
-list.of.plots %>% pull(plot.by.taxa)
-dev.off()
-#
-list.of.plots %>% slice(1) %>% pull(plot.by.taxa)-> one.plot
-legend.left <- get_legend(one.plot[[1]] )
-
-list.of.plots %>% mutate(plot.by.taxa= map(plot.by.taxa, ~.x + theme(legend.position = "none") )) -> list.of.plots
-
-```
-
-#### Now do change in suitability by taxa by time
-```{r}
-selected.taxa <- models %>% select(taxa) %>% filter (str_detect(taxa, "Chaetoceros|huxleyi|Alexandrium|Pseudochatt|Thalassionema|Navicula |Dityl|Nitzschia|Karlodinium|Coscinodiscus|Attheya|Phaeocystis globosa|Chattonella|Wolosz|Dictyocha|Emili")) %>% pull(taxa)
-
-posterior.probabilities.by.taxa %>% 
-  unnest(data) ->  posterior.probabilities.by.taxa
-```
-
-Now something for the supporting information
-
-```{r}
-posterior.probabilities.by.taxa %>% 
-  filter (taxa %in% selected.taxa) %>% 
-  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
-  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
-                           species == "NA" ~  paste0(genus, " sp."),
-                           TRUE            ~ species)) %>% 
-   ggplot(aes(x= prob.of.presence,
-             y = label)) +
-  #geom_density_ridges(fill = as.factor(Year))
-  stat_density_ridges(aes(fill = as.factor(Year)),
-    # geom = "density_ridges_gradient", calc_ecdf = TRUE,
-    quantiles =2, quantile_lines = TRUE, alpha = 0.5
-  ) + 
-  scale_fill_brewer(name = "Year",  type = "qual", palette = 2) + 
-  labs (x = "Probability of presence",
-        y = "") +
-    facet_wrap(~Area) +
+  facet_grid(phylum ~ Area, scales = "free_y") +
+   # facet_wrap(~Area) +
 
@@
         legend.position  = "bottom") +
-  ggsave("Most.varied.taxa.png", width = 14, dpi = "retina")
-```
-
-```{r}
-
+  ggsave(here("Figures","Most.varied.taxa.png"), width = 14, dpi = "retina")
+```
+
+```{r change over time only for supp}
 posterior.probabilities.by.taxa %>% 
@@
   theme(legend.position ="bottom") +
-  ggsave("Suitability change.png", width = 14) -> bottom.plot.figure3
-
-# almost.there %>% 
-# ggplot(aes (x = as.factor(Year),
-#               y = Corrected.presence, 
-#               fill = taxa)) +
-#   geom_boxplot()+
-#   facet_wrap(~Area) +
-#   guides(fill = "none")
+  ggsave(here("Figures","Suitability change.png"), width = 14) -> bottom.plot.figure3
+
 ```
@@
  First create the square matrix of points
-```{r}
+```{r create full grid}
+
 all.sims.for.projection  %>% 
-  group_by( Season, Area) %>% 
-  filter (keep == "keep") %>%   
+  group_by(  Area) %>% 
+  # filter (keep == "keep") %>%   
   summarise_at(c("TemperatureCelsius", "pH_units"), list(max = max, min = min)) %>%
 
-  group_by (Season, Area) %>% 
+  group_by ( Area) %>% 
   nest() %>% 
@@
 
-```{r}
+```{r add real temperatures to the grid}
+
 full.grid %>% 
@@
          pH = pH_units) %>%
+
   pivot_longer(cols = c(Temperature, pH),
@@
                values_to = "value") %>% 
-  group_by(Predictor) %>% 
+  group_by(Area, Predictor) %>% 
   nest() %>% 
   left_join(transforming.df) %>% 
-  mutate(data = map2(data, params, ~.x %>% 
-                       mutate(std = (value - .y$mean)/(.y$sd) ))) %>% 
-  select(-params) %>% 
+  mutate(data = map2(data, to.std, ~.x %>% 
+                       mutate(std = .y[1]+.y[2]*value ))) %>% 
+  select(Area, Predictor, data) %>% 
   unnest(data) %>% 
@@
          pH_units           = value_pH) -> full.grid
-
+```
+
+`full.grid` has the temperature and pH values we want to project our models to - so now we use it with the `models` output.
+
+```{r calculate the 100 draws from the posterior}
 models %>% 
@@
                                  rownames_to_column("sim.row") )) -> posterior.draws.raster
-
+```
+
+```{r calculate richness}
 posterior.draws.raster %>% 
@@
 
-```{r}
+```{r plot3 function}
 plot3 <- function(tibble,string , response) {
@@
   tibble %>% 
+    ungroup() %>% 
     mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
@@
     summarise (!!response := mean (!!response))  %>%
-   #  filter(str_detect(phylum, string))   %>% 
- # full_join(full.grid %>% unnest(data)) %>% 
+
     ggplot(aes(x = round(TemperatureCelsius,10),
@@
 
-  scale_fill_distiller(name = "Richness", type = "seq", palette = "Spectral") +
+  scale_fill_distiller(name = "Standarized Richness", type = "seq", palette = "Spectral",limits=c(0,1) ) +
 
@@
 
-  scale_color_brewer(name = "Year",  type = "qual", palette = 2) +
+  scale_color_brewer(name = "Year",  type = "qual", palette = 4, direction = -1, breaks=c("2017","2095")) +
+  # ggforce::geom_mark_hull(data = all.sims.for.projection %>%
+  #                           ungroup() %>% 
+  #                           mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
+  #                           TRUE          ~ Area)) %>%
+  #                           filter (Year %in% c(2017, 2095), keep == "keep") ,
+  #                         aes(x= TemperatureCelsius, y = pH_units, group= Year, color = as.factor(Year),label = as.factor(Year)),
+  #                         expand = unit(0.5, "mm"),
+  #                label.margin = margin(1, 1, 1, 1, "mm"),
+  #                label.buffer = unit(0, "mm"),
+  #                con.cap = 0 ,
+  #                alpha = 0.5) +
+     facet_grid(.~Area) +
+  guides(color = "none") +
+    labs (x = "Temperature",
+          y= "pH") +
+    theme_minimal()+
+    theme(legend.position = "bottom",
+          legend.key.width  = unit(2 ,"cm"),
+          axis.title.x = element_text(face = "bold"),
+          legend.direction = "horizontal") +
+    scale_x_continuous(position = "top")
+
+}
+```
+
+```{r}
+ready_plot %>% 
+  nest(-phylum) %>% 
+  mutate(plot = map2(data, phylum, ~ plot3(.x, .y, response = st.richness)))  -> list.of.richness.plots
+
+# # Get the other legend
+# list.of.richness.plots %>% slice(1) %>% pull(plot) -> one.plot
+# legend.right <- get_legend(one.plot[[1]]  )
+# list.of.richness.plots %>% 
+#   mutate(plot = map (plot, ~.x + theme(legend.position = "none"))) -> list.of.richness.plots
+# legend.left
+
+```
+
+```{r}
+posterior.probabilities.by.taxa %>% 
+  group_by(phylum) %>% 
+  summarise(n_distinct(taxa)) %>% 
+  pull -> list.of.number.of.taxa
+
+list.of.plots %>%  pull(phylum) -> list.of.phyla
+
+list.of.plots %>% pull(plot.by.taxa) -> list.of.ridge.plots
+
+list.of.ridge.plots <- set_names(list.of.ridge.plots, nm = list.of.phyla)
+
+list.of.richness.plots %>% pull(plot) -> list.of.raster.plots
+
+list.of.raster.plots <- set_names(list.of.raster.plots, nm = list.of.phyla)
+
+# # Use the fig3 ridges
+# list.of.plots.for.fig3 %>% pull(plot.by.taxa) %>% 
+#   map(function(.x) .x + theme(legend.position = "none")) %>% 
+#   set_names(nm = list.of.plots.for.fig3$phylum) -> list.of.ridge.plots.fig3
+# 
+# 
+# map2(list.of.ridge.plots.fig3, list.of.raster.plots[list.of.plots.for.fig3$phylum],
+#      function(.x, .y) plot_grid (.x, .y, nrow = 1, align = "hv", axis = "right")) %>% 
+#   set_names(list.of.plots.for.fig3$phylum)-> combined.plots
+# 
+# 
+# 
+# map2(list.of.ridge.plots,list.of.phyla, 
+#      ~ ggsave(.x, filename = here("Figures",paste0(.y, "ridges.png")), width = 14, 
+#                                   #height = c*20,
+#                                   dpi = "retina"))
+# ggsave(list.of.ridge.plots[["Bacillariophyta"]], filename = here("Figures","Bacillariophytaridges.png"), width = 14, height = 14,                                  dpi = "retina")
+# 
+# map2(list.of.raster.plots,list.of.phyla, 
+#      ~ ggsave(.x, filename = here("Figures",paste0(.y, "raster.png")), width = 7, 
+#                                   #height = c*20,
+#                                   dpi = "retina"))
+# map2(list.of.raster.plots,list.of.phyla, 
+#      ~ .x + 
+#        ggtitle(.y) + ggsave(filename = here("Figures",paste0(.y, "rastertitle.png")), width = 7, 
+#                                   #height = c*20,
+#                                   dpi = "retina"))
+```
+
+```{r}
+
+# plot_grid(plotlist = combined.plots,
+#           nrow = 1,axis = "l",
+#           align = "hv",
+#             ) -> top.all
+# 
+# plot_grid(legend.left, legend.right, 
+#           nrow = 2, align = "hv" ) -> midlayer
+# 
+# 
+# plot_grid(bottom.plot.figure3, midlayer,
+#           nrow = 1,
+#           rel_widths = c(1,0.2),
+#           align = "hv") -> bottom.plots
+# 
+# 
+# top.all + ggsave(here("Figures","Top_fig3.png"), dpi = "retina", width = 28, height = 10)
+# 
+# bottom.plot.figure3 + ggsave(here("Figures","bottom.fig.3.png"), dpi ="retina", width = 14)
+# 
+# midlayer + ggsave(here("Figures","legends.fig3.png"), dpi = "retina", height = 2)
+
+```
+
+# Redesign
+
+```{r}
+list.of.ridge.plots.fig3 %>% map(get_title) -> list.of.titles
+list.of.ridge.plots.fig3 %>% map(function(x) x + theme(plot.title = element_blank())) -> list.of.ridge.plots.fig3
+
+plot_grid(list.of.titles[[1]],list.of.titles[[2]], 
+          list.of.titles[[3]],list.of.titles[[4]],  nrow = 4,
+          align = "hv") -> titles
+plot_grid(plotlist = list.of.ridge.plots.fig3, axis = "r",
+          nrow = 4,
+          align = "hv" ) -> topleft
+
+plot_grid(plotlist = list.of.raster.plots[list.of.plots.for.fig3$phylum], axis = "l",
+          nrow = 4,
+          align = "hv" ) -> topright
+
+plot_grid(titles, topleft, topright,rel_widths = c(1,5,5),
+          ncol = 3,align = "v",axis = "tblr", labels = c("A",NA,"B") ) -> top.together
+
+plot_grid(legend.left, legend.right, get_legend(bottom.plot.figure3),
+          nrow = 3, align = "hv" ) -> midlayer
+bottom.plot.figure3 + theme(legend.position = "none") -> goodnow
+plot_grid(goodnow, midlayer,
+          rel_widths = c(2,1),
+          labels = c("C",NA),
+          ncol = 2,align = "v", axis = "t") -> bottom.together
+
+plot_grid(top.together, bottom.together,
+          nrow = 2, rel_heights = c(4,1),
+          align = "h", axis = "rl") + 
+  ggsave(here("Figures","Top_fig3.png"), dpi = "retina", width = 24, height = 32)
+ggsave(here("Figures","Top_fig3.18.png"), dpi = "retina", width = 18, height = 24)
+```
+
+# Patchwork
+
+```{r}
+library(patchwork)
+
+# Ridge plots
+list.of.plots.for.fig3 %>%
+  pull(plot.by.taxa) %>% set_names(list.of.plots.for.fig3$phylum) %>%
+  map(function(x) x +  theme_minimal()+ theme(plot.title.position = "plot",
+                                              plot.title = element_text(face = "bold",vjust =  -5 ),
+                                              axis.title.x.top  = element_text(face = "bold",vjust = 3),
+                                              legend.direction = "horizontal",
+                                              axis.text.y = element_text(face = "italic"))
+       + scale_x_continuous(limits = c(0,1),position = "top")
+      ) -> lefts
+lefts[[1]]
+# remove redundant axis and facet labels
+
+lefts[2:4] <- map(lefts[2:4], function(x) x + labs(x= "") + theme( strip.text.x = element_blank())  + scale_x_continuous(position = "top"))
+
+# Raster plots
+list.of.richness.plots %>%
+  pull(plot) %>%
+  set_names(list.of.richness.plots$phylum) %>%
+  .[list.of.plots.for.fig3$phylum] %>%
+  map(function(x) x + theme_minimal()+ theme(axis.title.x = element_text(face = "bold"),
+                                                                                                                                                                       legend.direction = "horizontal") + scale_x_continuous(position = "top")) -> rights
+
+# remove redundant axis and facet labels
+
+rights[2:4] <- map(rights[2:4], function(x) x +
+                     labs(x= "") +
                      ggforce::geom_mark_hull(data = all.sims.for.projection %>%
+                            ungroup() %>% 
+                            mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
+                             TRUE          ~ Area)) %>%
+                            filter (Year %in% c(2017, 2095), keep == "keep") ,
+                          aes(x= TemperatureCelsius, y = pH_units, group= Year, color = as.factor(Year)),
+                          expand = unit(0.5, "mm"),
+                 con.cap = 0 ,
+                 alpha = 0.5) +
+    theme( strip.text.x = element_blank()))
+
+# Variation plot
+bottom.plot.figure3 + theme_minimal() -> bottom.plot.figure3
+
+lefts[["Arthropoda"]] +  rights[["Arthropoda"]] +
+   lefts[["Bacillariophyta"]] +   rights[["Bacillariophyta"]] +
+   lefts[["Haptophyceae"]]  + rights[["Haptophyceae"]] +
+   lefts[["Dinophyceae"]]  + rights[["Dinophyceae"]] +
+   #bottom.plot.figure3 + guide_area() +
+
+  plot_layout(ncol = 2, widths = c(3,2),
+  guides = "collect",) -> test
+
+## Guides differently
+
+((guide_area()/
+  lefts[["Arthropoda"]]/
+  lefts[["Bacillariophyta"]] /
+   lefts[["Haptophyceae"]] /
+    lefts[["Dinophyceae"]] + 
+    plot_layout(heights = c(0.5,3,3,3,3), guides = "collect")) |
+    plot_spacer() |
+    (guide_area() /
+                                          rights[["Arthropoda"]]/
+                                          rights[["Bacillariophyta"]]/
+                                          rights[["Haptophyceae"]]/
+                                          rights[["Dinophyceae"]] + 
+                                          plot_layout(heights = c(0.5,3,3,3,3), guides = "collect"))) +
+  plot_layout(widths = c(3,0.25,2))-> test
+ggsave( filename = here("Figures","Top_fig3_20200615.18.png"),plot = test, units = "in", dpi = "retina", width = 12, height = 18)
+
+##
+list.of.plots.for.fig3.2 %>%
+  pull(plot.by.taxa) %>% set_names(list.of.plots.for.fig3.2$phylum) -> lefts
+
+lefts[[1]]
+# remove redundant axis and facet labels
+
+lefts[2:4] <- map(lefts[2:4], function(x) x + labs(x= "") + theme( strip.text.x = element_blank())  + scale_x_continuous(position = "top"))
+
+list.of.richness.plots %>%
+  pull(plot) %>%
+  set_names(list.of.richness.plots$phylum) %>%
+  .[list.of.plots.for.fig3.2$phylum] -> rights
+
+rights[[1]]<- rights[[1]]+ ggforce::geom_mark_hull(data = all.sims.for.projection %>%
+                            ungroup() %>%
                             mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
@@
                  con.cap = 0 ,
+                 alpha = 0.5)
+
+# remove redundant axis and facet labels
+
+rights[2:4] <- map(rights[2:4], function(x) x +
+                     labs(x= "") +
+                     ggforce::geom_mark_hull(data = all.sims.for.projection %>%
+                            ungroup() %>% 
+                            mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
+                            TRUE          ~ Area)) %>%
+                            filter (Year %in% c(2017, 2095), keep == "keep") ,
+                          aes(x= TemperatureCelsius, y = pH_units, group= Year, color = as.factor(Year)),
+                          expand = unit(0.5, "mm"),
+                 con.cap = 0 ,
                  alpha = 0.5) +
-     facet_grid(.~Area) +
-  guides(color = "none") +
-    labs (x = "Temperature",
-          y= "pH") +
-    #ggtitle(label = string) +
-    theme(legend.position = "bottom",
-          legend.key.width  = unit(2 ,"cm"))
-
-}
-```
-
-```{r}
-ready_plot %>% 
-  nest(-phylum) %>% 
-  mutate(plot = map2(data, phylum, ~ plot3(.x, .y, response = st.richness)))  -> list.of.richness.plots
-
-# Get the other legend
-list.of.richness.plots %>% slice(1) %>% pull(plot) -> one.plot
-legend.right <- get_legend(one.plot[[1]]  )
-list.of.richness.plots %>% 
-  mutate(plot = map (plot, ~.x + theme(legend.position = "none"))) -> list.of.richness.plots
-legend.left
-
-```
-
+    theme( strip.text.x = element_blank()))
+
+((guide_area()/
+  lefts[["Bacillariophyta"]]/
+  lefts[["Chlorophyta"]] /
+   lefts[["Haptophyceae"]] /
+    lefts[["Dinophyceae"]] + 
+    plot_layout(heights = c(0.5,3,3,3,3), guides = "collect")) |
+    plot_spacer() |
+    (guide_area() /
+                                          rights[["Bacillariophyta"]]/
+                                          rights[["Chlorophyta"]]/
+                                          rights[["Haptophyceae"]]/
+                                          rights[["Dinophyceae"]] + 
+                                          plot_layout(heights = c(0.5,3,3,3,3), guides = "collect"))) +
+  plot_layout(widths = c(2,0.25,3))-> test
+
+ggsave( filename = here("Figures","Top_fig3_20200615.18.Other.sel.png"),plot = test, units = "in", dpi = "retina", width = 12, height = 18)
+```
+
+? Emiliainia huxleyi
 ```{r}
 posterior.probabilities.by.taxa %>% 
-  group_by(phylum) %>% 
-  summarise(n_distinct(taxa)) %>% 
-  pull -> list.of.number.of.taxa
-
-list.of.plots %>%  pull(phylum) -> list.of.phyla
-
-list.of.plots %>% pull(plot.by.taxa) -> list.of.ridge.plots
-
-list.of.ridge.plots <- set_names(list.of.ridge.plots, nm = list.of.phyla)
-#list.of.ridge.plots %>% map(~.x + )
-list.of.richness.plots %>% pull(plot) -> list.of.raster.plots
-
-list.of.raster.plots <- set_names(list.of.raster.plots, nm = list.of.phyla)
-
-map2(list.of.ridge.plots, list.of.raster.plots, function(.x, .y) plot_grid (.x, .y, nrow = 2, align = "hv", axis = "right")) -> combined.plots
-
-pmap(.l = list(list.of.ridge.plots,list.of.phyla,list.of.number.of.taxa),
-     .f = function (a,b) ggsave(a, filename = paste0(b, "ridges.png"), width = 14, 
-                                  #height = c*20,
-                                  dpi = "retina"))
-map2(list.of.ridge.plots,list.of.phyla, 
-     ~ ggsave(.x, filename = paste0(.y, "ridges.png"), width = 14, 
-                                  #height = c*20,
-                                  dpi = "retina"))
-ggsave(list.of.ridge.plots[["Bacillariophyta"]], filename = "Bacillariophytaridges.png", width = 14, height = 14,                                  dpi = "retina")
-
-map2(list.of.raster.plots,list.of.phyla, 
-     ~ ggsave(.x, filename = paste0(.y, "raster.png"), width = 7, 
-                                  #height = c*20,
-                                  dpi = "retina"))
-map2(list.of.raster.plots,list.of.phyla, 
-     ~ .x + 
-       ggtitle(.y) + ggsave(filename = paste0(.y, "rastertitle.png"), width = 7, 
-                                  #height = c*20,
-                                  dpi = "retina"))
-```
-
-```{r}
-combined.plots <- set_names(combined.plots, list.of.phyla)
-
-top.line <- combined.plots[c("Arthropoda", "Dinophyceae", "Bacillariophyta", "Haptophyceae")]
-
-plot_grid(plotlist = top.line,
-          nrow = 1,
-          align = "hv" ) -> top.all
-plot_grid(legend.left, legend.right, nrow = 2, align = "hv" ) -> midlayer
-
-plot_grid(bottom.plot.figure3, midlayer,
-          nrow = 1,
-          rel_widths = c(1,0.2),
-          align = "hv") -> bottom.plots
-
-# bottom.line <- list( combined.plots[c("Haptophyceae")][[1]], bottom.plot.figure3)
-# plot_grid(plotlist = bottom.line,
-#           nrow = 1,
-#           align = "hv") -> bottomline
-
-top.all + ggsave("Top_fig3.png", dpi = "retina", width = 28, height = 10)
-
-bottom.plot.figure3 + ggsave("bottom.fig.3.png", dpi ="retina", width = 14)
-
-midlayer + ggsave("legends.fig3.png", dpi = "retina", height = 2)
-
-```
-
-### Besides Figures: other stats reported in the MS
-```{r}
-ASV.table  <- read_csv("../../Data/3.Growing_database/ASV_table_all_together.csv")
-
-Annotation <- read_csv("../../Analysis/Annotation/Annotated.hashes.csv")
+  filter(phylum == "Haptophyceae") %>% 
+  unnest() %>% 
+  filter(str_detect(taxa, "huxleyi")) %>% 
+  ggplot(aes(x = prob.of.presence))+
+ # geom_violin(scale="area") +
+ # geom_density_ridges() +
+  facet_wrap(~Area) +
+  theme(legend.position = "none") -> p
+
+p +geom_violin(aes( y= as.factor(Year), fill = as.factor(Year)), scale="area")+ ggtitle("scale = 'area' ") -> left.1
+
+p +geom_violin(aes( y= as.factor(Year)),scale="count")+ ggtitle("scale = 'count' ") -> cent.1
+
+p +geom_violin(aes( y= as.factor(Year)),scale="width")+ ggtitle("scale = 'width' ") -> right.1
+
+left.1 + geom_boxplot(aes( y= as.factor(Year), fill = as.factor(Year)))
+cent.1 + facet_grid(Year~Area)
+right.1 + facet_grid(Year~Area)
+
+p + geom_violin(scale="area", )
+  cent.1 + right.1
+```
+
+# Less is more
+
+What if we reduce the figure 3 to a dinoflagellate vs diatoms and
+
+```{r which to choose}
+posterior.probabilities.by.taxa %>% 
+  ungroup() %>% 
+  filter (phylum %in%c("Dinophyceae", "Bacillariophyta")) %>% 
+  mutate(suma = map (data,~.x %>% group_by(taxa,Area, Year) %>%
+                       summarise(tot = sum(prob.of.presence)) %>%
+                       ungroup() %>% 
+                       pivot_wider(names_from = Year, values_from = tot) %>% 
+                        group_by(taxa) %>% mutate(diff = `2017` - `2095`)
+                     )) %>% 
+  select(phylum, suma) %>% 
+  unnest(suma) %>%
+  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
+  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
+                           species == "NA" ~  paste0(genus, " sp."),
+                           TRUE            ~ species))  %>%  
+  ggplot(aes(x = fct_reorder(phylum,abs(diff)), y = diff)) +
+  geom_boxplot()+
+  geom_point(aes(color = diff>0)) +
+  geom_label_repel(data = . %>% group_by(phylum) %>% arrange(desc(abs(diff))) %>% slice(1:5) , aes(label = label))+
+  coord_flip() +
+  facet_wrap(~Area)
+```
+
+```{r select some }
+posterior.probabilities.by.taxa %>% 
+  ungroup() %>% 
+  filter (phylum %in%c("Dinophyceae", "Bacillariophyta")) %>% 
+  mutate(suma = map (data,~.x %>% group_by(taxa,Area, Year) %>%
+                       summarise(tot = sum(prob.of.presence)) %>%
+                       ungroup() %>% 
+                       pivot_wider(names_from = Year, values_from = tot) %>% 
+                        group_by(taxa) %>% mutate(diff = `2017` - `2095`)
+                     )) %>% 
+  select(phylum, suma) %>% 
+  unnest(suma) %>%
+  group_by(phylum) %>% arrange(desc(abs(diff))) %>% 
+  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
+  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
+                           species == "NA" ~  paste0(genus, " sp."),
+                           TRUE            ~ species)) -> taxa.and.diff
+
+taxa.and.diff %>%
+  slice(1:5) %>%
+  arrange(diff) %>% pull(taxa) -> ahorasi
+
+taxa.and.diff %>% 
+  slice(1:5) %>%
+  arrange(diff) %>% pull(label) -> reorder.labels
+
+```
+
+# Besides Figures: other stats reported in the MS
+```{r}
+ASV.table  <- read_csv(here("Output","ASV_table_all_together.csv"))
+
+Annotation <- read_csv(here("Input","Annotated.hashes.csv"))
 

### 2020-08-03 Scripts/Comparing_climate_simulations.Rmd
Updated scripts

---
title: "A simulated scenario"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)

```

## The nested-for-loop scenario

```{r}
data_for_loop_1 <- read_csv("Output/simulation.from.real.for.loop.csv")

data_for_loop_2 <- read_csv("Output/simulation.from.std.for.loop.csv")

data_from_function <- read_csv("Output/simulation.from.std.functions.csv")

data_from_function_cleaned <- read_csv("Output/simulation.from.std.functions.cleaned.csv")

data_from_function_cleaned %>% 
  select(Area, Year, Temperature = TemperatureCelsius, pH = pH_units) -> data_from_function_cleaned
```

```{r}
now.env.data <- read_csv(here("Input","env.data.for.log.regression.csv")) %>% 
  mutate (Year = 2017) %>% 
      filter(pH > 7) %>%
      mutate(
        Season = case_when(
          str_detect(sample, "1703|1710|1711|1801|1803") ~ "Winter",
          TRUE ~ "Summer"
        ),
        Area = case_when(
          str_detect(sample, "CP|LK|FH") ~ "SJI",
          TRUE ~ "Hood Canal"
        )
      )
```

```{r}

list.of.dfs <- list(data_for_loop_1,data_for_loop_2,data_from_function, data_from_function_cleaned, now.env.data )

names(list.of.dfs) <- c("for loop from real", "for loop from std" , "function data", "function data cleaned", "real data")

all.together <- bind_rows(list.of.dfs, .id = "Source")

```

Let's see them in hex
```{r}
all.together %>% 
  filter (Area == "SJI", Source != "for loop from real") %>% 
  ggplot(aes(x = Temperature, y = pH)) +
  geom_hex(bins= 40, aes(group = as.factor(Year), fill = ..density..)) +
  facet_grid (Year~Source) +
  ggtitle("San Juan Islands") -> left

all.together %>% 
  filter (Area != "SJI", Source != "for loop from real") %>% 
  ggplot(aes(x = Temperature, y = pH)) +
  geom_hex(bins= 40, aes(group = as.factor(Year), fill = ..density..)) +
  facet_grid (Year~Source) +
  ggtitle("Hood Canal") -> right

library(patchwork)

(left + right)  +  plot_layout(guides = "collect")-> all
all & scale_fill_continuous(limits = c(0,0.1))
```

```{r Check each }

expl.plots <- function (tibble){

tibble %>% 
  separate(sample, into = c("Site", "Date")) %>% 
  mutate(Date = lubridate::ymd(paste0(Date, "01")) %>% month(label = T)) ->temp

temp %>% 
ggplot(aes(x = Date, y = Temperature, fill = as.factor(Year))) +
  #geom_point(position = "jitter")+
  geom_boxplot()+
#  geom_smooth(aes(x = as.numeric(Date)), method = "loess") +
  facet_wrap(~Area)+ scale_fill_discrete(name = "Year", direction = -1) +
  guides(fill = "none") +
  labs(x = "Month") -> l1

temp %>% 
ggplot(aes(x = Date, y = pH, fill = as.factor(Year))) +
  #geom_point(position = "jitter")+
  geom_boxplot()+
#  geom_smooth(aes(x = as.numeric(Date)), method = "loess") +
  facet_wrap(~Area)+ scale_fill_discrete(name = "Year", direction = -1) +
  guides(fill = "none") +
  labs(x = "Month") -> l2

temp %>% 
ggplot(aes(x = Date, y = Temperature, color = as.factor(Year))) +
  #geom_point(position = "jitter")+
 # geom_boxplot()+
  geom_smooth(aes(x = as.numeric(Date)), method = "loess") +
  facet_wrap(~Area)+ scale_color_discrete(name = "Year", direction = -1) +
  labs(x = "Month") -> l3

temp %>% 
ggplot(aes(x = Date, y = pH, color = as.factor(Year))) +
  #geom_point(position = "jitter")+
  #geom_boxplot()+
  geom_smooth(aes(x = as.numeric(Date)), method = "loess") +
  facet_wrap(~Area)+ scale_color_discrete(name = "Year", direction = -1) +
  labs(x = "Month") -> l4

library(patchwork)

(l1+l2)/(l3 +l4) + patchwork::plot_layout(guides = "collect") 
}

```

```{r}
expl.plots(data_for_loop_1)
```

```{r}
now.env.data %>% ggplot(aes(x = pH , y = Temperature)) + geom_smooth(method = "lm") + facet_wrap(~Area) + geom_point ()

```

### 2020-08-03 Scripts/Denoising.all.runs.Rmd
Updated scripts

@@
 ASV.nested %>% 
-  transmute(Miseq_run,Summary = map(Samples, ~ how.many(ASVtable = .,round = 0)))  -> ASV.summary
+  transmute(Summary = map(Samples, ~ how.many(ASVtable = .,round = 0)))  -> ASV.summary
 
@@
 ASV.nested %>% 
-  transmute(Miseq_run, Summary.1 = map(cleaned.tibble, ~ how.many(ASVtable = .,round = "1.Jump"))) %>% 
+  transmute( Summary.1 = map(cleaned.tibble, ~ how.many(ASVtable = .,round = "1.Jump"))) %>% 
   left_join(ASV.summary) %>% 
@@
 ASV.nested %>% 
-  transmute(Miseq_run, Summary.1 = map(Step.1.low.reads, ~ how.many(ASVtable = .,round = "2.Low.nReads"))) %>% 
+  transmute( Summary.1 = map(Step.1.low.reads, ~ how.many(ASVtable = .,round = "2.Low.nReads"))) %>% 
   left_join(ASV.summary) %>% 
@@
 ASV.nested %>% 
-  transmute(Miseq_run, Summary.1 = map(Step2.tibble, ~ how.many(ASVtable = .,round = "3.Positives"))) %>% 
+  transmute( Summary.1 = map(Step2.tibble, ~ how.many(ASVtable = .,round = "3.Positives"))) %>% 
   left_join(ASV.summary) %>% 
@@
   geom_col(aes(y = tot), fill = "red") +
+
   labs(x = "",
@@
 ASV.nested %>% 
-  transmute(Miseq_run, Summary.1 = map(Step3.tibble, ~ how.many(ASVtable = .,round = "4.Occupancy"))) %>% 
+  transmute(Summary.1 = map(Step3.tibble, ~ how.many(ASVtable = .,round = "4.Occupancy"))) %>% 
   left_join(ASV.summary) %>% 
@@
 
-Let's make the pairwaise distances a long table
+Let's make the pairwise distances a long table
 ```{r}
@@
   geom_histogram (position = "dodge", stat = 'density', alpha = 0.9) +
- # facet_wrap( ~ Distance.type) +
+  facet_wrap( ~ Distance.type) +
   labs (x = "Pairwise dissimilarity", y = "density" ,
         Distance.type = "Distance")
-
+ggsave("visual.anova.png", dpi = "retina")
 ```
@@
   filter(Date != "201703") %>% 
+  mutate(Date = lubridate::ymd(paste0(Date, "01"))) %>% 
+  mutate(Area = case_when(Site %in% c("CP", "LK", "FH") ~ "San Juan Island",
+                          TRUE                          ~ "Hood Canal"),
+         Site = fct_recode(Site, "Cattle Point" = "CP",
+                           "Lime Kiln" = "LK",
+                           "Friday Harbor" = "FH",
+                           "Twanoh" = "TW",
+                           "Potlatch"= "PO",
+                           "Lilliwaup" = "LL",
+                           "Triton Cove"= "TR",
+                           "Salsbury Park"= "SA")) -> Coverage.dataset
+
+Coverage.dataset %>% 
   ggplot()+
@@
 ASV.nested %>% 
-  transmute(Miseq_run, Summary.1 = map(Step4.tibble, ~ how.many(ASVtable = .,round = "5.PCR.dissimilarity"))) %>% 
+  transmute(Summary.1 = map(Step4.tibble, ~ how.many(ASVtable = .,round = "5.PCR.dissimilarity"))) %>% 
   left_join(ASV.summary) %>% 
@@
 ```
+
+# Coverage Table with nReads
+
+```{r}
+# Vector with desired order
+
+order.Sites <- c("Salsbury Park", "Triton Cove", "Lilliwaup", "Potlatch", "Twanoh", "Cattle Point", "Lime Kiln", "Friday Harbor")
+
+ASV.nested %>% 
+  select(Miseq_run, Step4.tibble) %>% 
+  unnest(Step4.tibble) %>% 
+  mutate(nReads = as.integer(nReads)) %>% 
+  group_by(sample) %>% 
+  summarise (nReads = sum(nReads)) %>% 
+  separate (sample, into = "event", sep = -3) -> midstep
+
+# As a table
+midstep %>% 
+  group_by(event) %>% 
+  summarise (mean = mean(nReads), max = max(nReads), min = min(nReads), sd = sd(nReads)) %>% 
+transmute (event, data = paste0("(", round(mean,0), " +- ", round(sd,0), ")")) %>% 
+  right_join(Coverage.dataset) %>% 
+  mutate( data = paste0(cases, " ", data)) %>% 
+  arrange(Date) %>% 
+  mutate (Date = paste(lubridate::month(Date, label = T),
+                      lubridate::year(Date),
+                      sep= "'")) %>% 
+  pivot_wider(names_from = Date, values_from = data, id_cols = c(Area, Site),values_fill = list(data = "") ) %>% 
+  slice(match(Site, order.Sites)) %>% 
+  kableExtra::kable() %>% 
+  kableExtra::kable_styling() %>% 
+  kableExtra::collapse_rows(1, valign = "top")
+
+# As a graph
+midstep %>% 
+  right_join(Coverage.dataset) %>% 
+  group_by(event) %>% 
+  mutate(mean = mean(nReads),
+         sum = sum(nReads)) %>% 
+  mutate(Site = factor(Site, levels = order.Sites)) %>% 
+  ggplot(aes(x = Date, 
+             y = nReads)) +
+  geom_boxplot(aes(group=Date),outlier.alpha = 0.5) +
+  geom_point(aes(y = sum,
+                 size = cases,
+                 color = Area)) +
+  scale_y_continuous(labels = scales::comma) +
+  scale_x_date(labels = function(x){paste(lubridate::month(x, label = T),
+                      lubridate::year(x),
+                      sep= "'")}) +
+  labs(y = "#reads", x = "") +
+
+  facet_wrap(~Site , nrow = 5,dir = "v", drop = T) +
+  theme_minimal() +
+  theme(axis.title.y = element_text(angle = 0,vjust = 0.5),
+        legend.position = c(0.6, 0.2))
+
+```

### 2020-08-03 Scripts/Future_climate_conditions.R

@@ -1,0 +1,245 @@
+#library(tidyverse)
+#library(propagate)
+library (fitdistrplus)
+library (tidyverse)
+library (vegan)
+library (tidybayes)
+library (here)
+
+now.env.data <- read_csv(here("Input","env.data.for.log.regression.csv")) %>%
+  filter(pH > 7.4) %>%
+  mutate(
+    Season = case_when(
+      str_detect(sample, "1703|1710|1711|1801|1803") ~ "Winter",
+      TRUE ~ "Summer"
+    ),
+    Area = case_when(
+      str_detect(sample, "CP|LK|FH") ~ "SJI",
+      TRUE ~ "Hood Canal"
+    ))
+
+#calculations based upon Khangaonkar, T., Nugraha, A., Xu, W., & Balaguru, K. (2019). Salish Sea response to global climate change, sea level rise, and future nutrient loads. Journal of Geophysical Research: Oceans, 124. https://doi.org/10.1029/ 2018JC014670
+#and in particular, upon Figure 14 in that paper. 
+
+#1. The paper represents a single model run focusing on the Salish Sea. Hence, there are no available error estimates for changes in mean parameters over time.
+#2. But because the model shows a good deal of spatial variation within the Salish Sea for that single model run, we can do the following:
+#3. Assume spatial variation of means over the year is greater than the likely variation among model runs.
+#4. Use (large) spatial variation as a conservative estimate of within-space variation among model runs. This is the best we can do.
+
+#interpreted from Khangaonkar, Figure 14:
+
+# pH.2000.mean <- 7.81
+# pH.2000.CI25 <- 7.73
+# pH.2000.CI75 <- 7.9
+# pH.2095.mean <- 7.63
+# pH.2095.CI25 <- 7.54
+# pH.2095.CI75 <- 7.78
+# 
+# T.2000.mean <- 10.01
+# T.2000.CI25 <- 9.0
+# T.2000.CI75 <- 11.8
+# T.2095.mean <- 11.52
+# T.2095.CI25 <- 10.6
+# T.2095.CI75 <- 13.5
+# 
+# 
+# #####TO COERCE TO NORMAL DISTRIBUTIONS
+# 
+#     #forcing these into symmetrical distributions by empirically sampling different normal distributions to match
+#     # rnorm(1000, mean = pH.2000.mean, sd = 0.25) %>% 
+#     #   quantile(c(.25, .75))
[…]

### 2020-08-03 Scripts/Models_for_Revisions.R

@@ -1,0 +1,417 @@
+library(tidyverse)
+library(stats)
+library(rstanarm)
+library(loo)
+library(bayesplot)
+library(tidybayes)
+library(modelr)
+library(here)
+library(styler)
+
+#uncomment below, if sourcing at command line (because this apparently doesn't respect here())
+#setwd("/Volumes/GoogleDrive/My Drive/Kelly_Lab/Projects/OA_eDNA/Analysis/Splitting.data.by.lifestyle")
+
+
+####LOAD DATA
+options(mc.cores = parallel::detectCores())
+data.for.stan <- read_csv("Input/data.for.stan.csv")
+now.env.data <- read_csv("Input/env.data.for.log.regression.csv") %>%
+  filter(pH > 7.4) %>%
+  mutate(
+    Season = case_when(
+      str_detect(sample, "1703|1710|1711|1801|1803") ~ "Winter",
+      TRUE ~ "Summer"
+    ),
+    Area = case_when(
+      str_detect(sample, "CP|LK|FH") ~ "SJI",
+      TRUE ~ "Hood Canal"
+    ))
+  
+######FUNCTIONS
+### DEFINE linear models to switch back and forth between standardized and natural-scale data
+value2std <- function(df){lm(std~value, data = df)}
+std2value <- function(df){lm(value~std, data = df)}
+
+ModelPlot <- function(df, variable) {
+  require(tidyverse)
+  temp.plot <-
+    df %>%
+    mutate(Posterior = list(apply(posterior_linpred(df$model[[1]], transform = T), 2, median))) %>%
+    select(-model) %>%
+    unnest() %>%
+    ggplot(aes(x = eval(as.symbol(variable)), y = binary, color = Area)) +
+    geom_point() +
+    geom_smooth(aes(y = Posterior)) +
+    facet_grid(Area ~ .) +
+    ggtitle(label = df$taxa[1]) +
+    xlab(variable) +
+    ylab("Probability of Occurrence")
+  
+  return(temp.plot)
+}
+#
+ProjectionPlot <- function(taxon, modelset = mod.proj.phylum) {
+  temp.plot <- modelset %>%
+    mutate(meanRichness = round(meanRichness, 1)) %>%
+    group_by(phylum, Area) %>%
+    mutate(ScaledMeanRichness = meanRichness / max(meanRichness)) %>%
+    filter(phylum == taxon) %>%
[…]

### 2020-08-03 Scripts/tibble.to.matrix.R

@@ -1,0 +1,167 @@
+# A function that takes a long dataframe and turns it into a number of different objects:
+
+# community table tibble.to.table
+# 
+# distance matrix tibble.to.matrix
+# 
+# Permanova object tibble.to.permanova
+# 
+# Environmental table tibble.to.env
+
+tibble.to.table <- function (long.table,taxon,Abundance, sample.name, transformation, ...) {
+  
+  sample.name = rlang::enquo(sample.name)
+  taxon = rlang::enquo(taxon) # quote taxon so it works with select
+ 
+  Abundance = rlang::enquo(Abundance)
+  
+  
+  # So now the eDNA index works let's see if we can use decostand
+  
+  
+  
+  cols = long.table %>% ungroup %>% dplyr::select(!!taxon) %>% distinct() %>% pull() # capture the unique values of the taxa
+  
+  
+  long.table %>%
+    ungroup %>% 
+    select(!!taxon, !!Abundance, !!sample.name, ...) %>% # select at least three columns: nReads, taxa and the sampleID
+    spread (key = !!taxon, value = !!Abundance, fill = 0) -> matrix_1 # Make a wide table, like vegan likes
+  
+  env <- colnames(matrix_1)[!colnames(matrix_1) %in% cols] # Which column has all the sample info
+  
+  dplyr::select (matrix_1, env) -> env # The values of the sample info in the order they appear in the wide table
+  
+  samples <- pull (matrix_1, !!sample.name)
+  dplyr::select (matrix_1, cols) -> matrix_1 # select only the spp
+  data.matrix(matrix_1) -> matrix_1 # make it a data matrix
+  
+  dimnames(matrix_1)[[1]] <- samples # name it with the sample info
+  return (matrix_1)
+  
+  }
+
+tibble_to_matrix <- function (long.table,taxon,Abundance, sample.name, distance = "bray", transformation, ...) {
+  
+  sample.name = rlang::enquo(sample.name)
+  taxon = rlang::enquo(taxon) # quote taxon so it works with select
+  choice = rlang::enquo(distance)
+  Abundance = rlang::enquo(Abundance)
+  
+  
+  # So now the eDNA index works let's see if we can use decostand
+  
+  
+  
+  cols = long.table %>% ungroup %>% dplyr::select(!!taxon) %>% distinct() %>% pull() # capture the unique values of the taxa
+  
+  
+  long.table %>%
+    ungroup %>% 
[…]

### 2020-08-13 Scripts/All.figures.that.go.in.the.MS.no.seasons.Rmd
All ready to go

@@
 
-This Markdown will make the part of the analysis that goes in the Manuscript:
+This Markdown will make the part of the analysis that goes in the Manuscript: The starting point are the Abundance Table (number of reads per ASV per sample) and the key file (the ASV file has the Hash of the sequences, and the key has the actual DNA sequence).
+If you want to process the raw fastq files to obtain these two files, you have to use the pipeline available in http::/github.com/ramongallego/demultiplexer_for_dada2
 
@@
 library (tidyverse)
-library (cowplot)
 library (concaveman)
@@
 library (ggrepel)
-library (factoextra)
-library (nnet)
+library (factoextra) # Needed for visualizing clusters in Fig2
+library (nnet) # Needed for multinomial regression in Fig 2
 library (here)
 library (conflicted)
+
 conflict_prefer("filter", "dplyr")
@@
 #functions for future Salish Sea
-models <- readRDS(here("Input","full_model_output_rpk_TpH.standardized_nonSeasonal2020-06-24_3262.RDS"))
+models <- readRDS(here("Input","finalModel_20200807.RDS"))
 
@@
 
-stds_to_value <- read_rds(here("Input", "std_to_real.rds"))
-
 ```
@@
 
-```{r }
+```{r loading data}
 data.plankton <- read_csv(here("Input","Combined_Biol_Env_Plankton.csv"))
@@
 
-env.data <- read_csv(here("Input","env.data.for.log.regression.csv"))
-
-#quality-control data
-    env.data %>%
-      filter(pH > 7) %>%
-      mutate(
-        Season = case_when(
-          str_detect(sample, "1703|1710|1711|1801|1803") ~ "Winter",
-          TRUE ~ "Summer"
-        ),
-        Area = case_when(
-          str_detect(sample, "CP|LK|FH") ~ "SJI",
-          TRUE ~ "Hood Canal"
-        )
-      ) -> now.env.data
-
-now.env.data %>%
-  gather(-Season, -sample, -Area,  key = "Predictor", value = "value") %>%
-  group_by(Area, Predictor) %>%   #RPK: this was missing, which messed up the std calculations 
-  mutate(std = vegan::decostand(value, method = "standardize")) -> now.env.data.with.std
-
-#### TODO: remove
-
-# Dataset to go between standarized and real values
-now.env.data.with.std %>%
-  group_by(Area, Predictor) %>%
-  # summarise(mean = mean(value),
-  #           sd = sd(value)) %>%
-  # group_by(Area, Predictor) %>%
-  nest() %>% 
-  summarise(tovalue = map(data, ~ lm(value ~ std, data = .x) ),
-            tostd   = map(data, ~ lm(std ~ value, data = .x)))-> transforming.df
-
-transforming.df[1,4] %>% pull -> test
-broom::tidy(test[[1]])
-extract <- function (lm) {broom::tidy(lm)$estimate}
-
-extract(test[[1]])
-
-transforming.df %>% 
-  mutate(to.values = map(tovalue, extract),
-         to.std= map(tostd, extract)) ->transforming.df
+# How to go from real data from standardized data and back
+value2std <- function(df){lm(std~value, data = df)}
+std2value <- function(df){lm(value~std, data = df)}
+
+stds_to_value <- data.plankton %>% 
+  select(sample = event, Area, mean.TA, mean.DIC = mean.DIC.new.sal,Salinity = Salinity.new ,Temperature, Omega.aragonite, pH = pH_new ) %>% 
+  distinct() %>%  
+  filter(pH > 7.4) %>% 
+  ungroup() %>% 
+  pivot_longer(cols = c(-sample, -Area),
+               names_to = "key",
+               values_to = "value") %>% 
+          group_by(key, Area) %>%   #calculating different scale for each area; standardizing to different means
+          mutate(std = vegan::decostand(value, method = "standardize")) %>%  
+          ungroup() %>% 
+          mutate(Area = as.factor(Area)) %>% 
+          filter(key %in% c("Temperature", "pH"))  %>% 
+          nest(dataset = c(sample, value, std)) %>% 
+          mutate(toStd = map(dataset, value2std),
+                 toValue = map(dataset, std2value))
 
@@
 ```{r Future scenarios}
-Year.area.dataset <- function(Area, Year, n){
-
-    Temp.conv <- stds_to_value[which (stds_to_value$Area == Area & stds_to_value$key == "Temperature"),"toValue"] %>% pull %>% .[[1]]
-    pH.conv   <- stds_to_value[which (stds_to_value$Area == Area & stds_to_value$key == "pH")         ,"toValue"] %>% pull %>% .[[1]]
-
-    one<-  data.frame(
-              Area = Area,
-              Year = Year,
-              Temperature = r_Temperature_year_STD(Year, Area, n, 2017),
-              pH          = r_pH_year_STD(Year, Area, n, 2017)
-              )
-
-    one$TemperatureCelsius = sapply(one$Temperature,function(x){Temp.conv$coefficients[1]+Temp.conv$coefficients[2]*x})
-    one$pH_units            = sapply(one$pH,function(x){pH.conv$coefficients[1]+pH.conv$coefficients[2]*x})
-
-    return(one)
-
-}
-
-j <- Year.area.dataset("Hood Canal", 2095, 5000)
-k <- Year.area.dataset("Hood Canal", 2017, 5000)
-l <- Year.area.dataset("SJI", 2095, 5000)
-m <- Year.area.dataset("SJI", 2017, 5000)
-
-simulated.scenario <- rbind(j,k,l,m)
-
-rm(j,k,l,m)
-
-simulated.scenario %>%
+future.scenario <- tibble(Area = rep(c("Hood Canal", "SJI"), 6),
+                  Year = rep(c(2017, 2091:2095), each = 2))  %>% 
+  mutate(Temperature = map2(Area, Year, ~  r_Temperature_year(area = .x,year =  .y, 1000)),
+         pH = map2(Area, Year, ~  r_pH_year(area = .x,year =  .y, 1000)))
+future.scenario %>% unnest(cols = c(Temperature, pH)) -> future.scenario
+
+# Make 2095 the 95% interval of the 2091-2095 scenarios 
+future.scenario %>% 
+  mutate(Year = case_when(Year == 2017 ~ 2017, 
+                          TRUE        ~ 2095)) %>%
   group_by(Year, Area) %>% 
   nest() %>% 
-  mutate(
-    values = map (data, function (.x) {
-      list(temp = fitdistr(.x$TemperatureCelsius ,
+  mutate(values = map (data, function (.x) {
+                              list(temp = fitdistr(.x$Temperature ,
                                                  densfun = "normal"),
-           ph = fitdistr(.x$pH_units ,
+                                   ph = fitdistr(.x$pH ,
                                                  densfun = "normal")) 
-            }
-          ),
-
+            }),
          data = map2(data,values,  function (.x, .y){
                               .x %>%
-        mutate(pT = pnorm(TemperatureCelsius, mean  = .y[["temp"]]$estimate[1], sd = .y[["temp"]]$estimate[2]),
-               ppH =pnorm(pH_units, mean  = .y[["ph"]]$estimate[1], sd = .y[["ph"]]$estimate[2]) )
-      })
-    ) %>% 
+                                mutate(pT = pnorm(Temperature, 
+                                                  mean  = .y[["temp"]]$estimate[1],
+                                                  sd = .y[["temp"]]$estimate[2]),
+                                       ppH =pnorm(pH,
+                                                  mean  = .y[["ph"]]$estimate[1], 
+                                                  sd = .y[["ph"]]$estimate[2]) )
+           })) %>%
+
   unnest(data) %>%
-  mutate(keep = case_when(pT < 0.05 | ppH < 0.05 | pT > 0.95 | ppH > 0.95 ~ "Discard",
+
+  mutate(keep = case_when(pT < 0.03 | ppH < 0.03 | pT > 0.98 | ppH > 0.98 ~ "Discard",
                           TRUE                   ~ "keep")) %>% 
-  filter (keep == "keep")-> simData.nocrazys
-
-simData.nocrazys %>% 
+  filter (keep == "keep") -> future.scenario
+
+future.scenario %>% 
   select(-values) %>% write_csv(here("Output","simulation.from.std.functions.cleaned.csv"))
@@
 
+We are going to use this future simulation in two ways:
+
+  * As it is, to explore the changes in suitability for each taxa between present and future sceanrios
+  * As a means to establish a continuous grid of Temperature and pH, so we can visualize how broader changes might occur.
+
+```{r create the grid}
+
+future.scenario %>% 
+  group_by(  Area) %>% 
+  # filter (keep == "keep") %>%   
+  summarise_at(c("Temperature", "pH"), list(max = max, min = min)) %>%
+
+  group_by ( Area) %>% 
+  nest() %>% 
+  mutate(data = map (data, ~ expand_grid(Temperature = seq(round(.x$Temperature_min,1), round(.x$Temperature_max, 1),by = 0.1),
+                                         pH = seq(round(.x$pH_min,1), round(.x$pH_max,1), by = 0.1)))) %>% 
+  unnest(data) -> full.grid
+```
+
+# Add the standardized values for each dataset
+
+Because the standarization was done for each region separately, we have to convert them into standard values seapratedly as well.
+
+```{r add standard values to both ds}
+
+convertT <- function( values, area){
+  stds_to_value %>% 
+    filter (Area == area, key == "Temperature") %>% 
+    pull(toStd) -> mod
+    predict(mod[[1]],newdata = data.frame (value = values))
+}
+convertpH <- function( values, area){
+  stds_to_value %>% 
+    filter (Area == area, key == "pH") %>% 
+    pull(toStd) -> mod
+    predict(mod[[1]],newdata = data.frame (value = values))
+}
+
+full.grid %>% 
+  rename(TemperatureCelsius = Temperature,
+         pH_units           = pH) %>% 
+  group_by(Area) %>% 
+  nest() %>% 
+  mutate(data =  map2(Area, data, ~ .y %>% 
+                            mutate(Temp.std =convertT( area = .x, values =.y$TemperatureCelsius),
+                                   pH.std   =convertpH(area = .x, values =.y$pH_units) ))) %>% 
+  unnest(data) -> full.grid
+
+future.scenario %>% 
+  select(Area, Year, TemperatureCelsius = Temperature, pH_units = pH) %>% 
+  group_by(Area) %>% 
+  nest() %>% 
+  mutate(data =  map2(Area, data, ~ .y %>% 
+                            mutate(Temp.std =convertT( area = .x, values =.y$TemperatureCelsius),
+                                   pH.std   =convertpH(area = .x, values =.y$pH_units) ))) %>% 
+  unnest(data) -> future.scenario
+```
+
 ### Save future scenarios
@@
 
-simData.nocrazys %>%
-
-saveRDS( file = here("Output","all.sim.data_climateScenarios_no.outliers.RDS"))
-
-simData.nocrazys %>% 
+full.grid %>%
+
+saveRDS( file = here("Output","Temp.pH.grid.RDS"))
+
+future.scenario %>% 
+  write_rds(file = here("Output", "future.simulation.rds"))
+
+future.scenario %>% 
   group_by( Area, Year) %>% 
@@
 
-We have the datasets and some plots already made -
+We have the datasets and the map already made
 
@@
 ```
-### Panel B Environmental variation
+### Panel B  Richness variation
+
+```{r Species richness}
+
+data.plankton %>% 
+  left_join(phylum.info) %>% 
+  group_by(event, phylum) %>% 
+  summarise (Richness = n_distinct(taxa)) %>% 
+  group_by(event) %>% 
+  mutate(Total.Richness = sum(Richness)) %>% 
+  separate(event, into = c("Site", "Date"), sep = "_") %>% 
+  mutate(Area = case_when(Site %in% c("CP", "LK", "FH") ~ "San Juan Island",
+                          TRUE                          ~ "Hood Canal"),
+         Site = fct_relevel(Site, "CP", "LK", "FH","SA", "TR", "LL","PO", "TW")) -> data.for.richness
+
+data.for.richness %>%   
+ggplot(aes(x = Site, y= Total.Richness )) +
+  geom_boxplot(aes(fill = Area)) +
+  lims (y= c(0,max(data.for.richness$Total.Richness))) +
+  labs (y = "Species richness") -> left.rich.plot
+
+data.for.richness %>% 
+  filter (phylum %in% c("Bacillariophyta", "Dinophyceae")) %>% 
+  ggplot(aes(x = Site, y= Richness )) +
+  geom_boxplot(aes(fill = Area)) +
+  facet_wrap(~phylum, nrow = 2) +
+  labs (y = "") -> right.rich.plot
+
+left.rich.plot + right.rich.plot + plot_layout(tag_level = "new") -> boxplot.richness
+```
+
+### Panel C Environmental variation
 
@@
 
-```{r}
-now.env.data %>% 
-  select(TA = mean.TA, DIC = mean.DIC, everything(), -contains("%")) %>% 
+```{r envRidges}
+data.plankton %>% 
+  select(sample = event, Area, TA = mean.TA, DIC = mean.DIC.new.sal,Salinity = Salinity.new ,Temperature, Omega.aragonite, pH = pH_new ) %>% 
+  distinct() %>%  
+  filter(pH > 7.4) %>% 
+  ungroup() %>% 
   filter(!is.na(TA)) %>% 
-  separate(sample, into= c("Site","Date"), remove = F )-> env.data.no.outliers
-
-env.data.no.outliers %>% 
-  pivot_longer(cols = c(-Area, -Site, -Date, -Season, -sample), names_to =   "Predictor", values_to  = "value") -> env.data.to.keep
-
-env.data.to.keep %>% 
+  separate(sample, into= c("Site","Date"), remove = F, sep = "_" ) %>% 
+  pivot_longer(cols = c(-Area, -Site, -Date,  -sample), names_to =   "Predictor", values_to  = "value")  %>% 
   mutate(Site = fct_relevel(Site, "SA", "TR", "LL","PO", "TW"),
-         Predictor = fct_relevel(Predictor, "DIC", "TA", "pH")) -> env.data.to.keep
-
-env.data.to.keep %>% 
-  mutate(Area = case_when (Area == "SJI" ~ "San Juan Island",
-                           TRUE          ~ Area)) %>% 
-  mutate(Month = month(ymd(paste0(Date, "01")), label = T, abbr = F)) %>% 
+         Predictor = fct_relevel(Predictor, "DIC", "TA", "pH")) %>% 
+  mutate(Month = month(ymd(Date), label = T, abbr = F)) %>% 
   filter (Predictor %in% c("DIC", "TA", "pH", "Temperature", "Salinity")) %>% 
@@
   facet_wrap(~fct_relevel(Predictor, "DIC", "TA", "pH", "Temperature", "Salinity"), scales = "free", nrow = 1) +
-  theme(legend.direction ="horizontal") -> ridges
-
-legend <- get_legend(
-  # create some space to the left of the legend
-  ridges + theme(legend.box.margin = margin(0, 0, 0, 12))
-)
- legend
+  theme(legend.direction ="horizontal", axis.title.x = element_blank()) -> ridges
+
 ridges + guides(fill = "none") -> ridges
@@
 
-### Panel c: Richness variation
-
-```{r}
-boxplot.richness <- readRDS(here("Input","richness.rds"))
-
-boxplot.richness +
-  ylim (0,60) -> boxplot.richness
-```
 ### All together
 
-THe top will be the environmental gradients, the lower panels will have the map and the boxplot
-```{r}
-plot_grid(plotlist = list(map, legend, boxplot.richness),
-                     align = "v",
-                     ncol = 3, 
-          axis = "t",
-          rel_widths = c(1,0.5,0.6))   -> bottom.half.plot1
-
-plot_grid(plotlist = list(ridges, bottom.half.plot1),
-          align = "hv",
-          ncol = 1, 
-          nrow = 2,
-          rel_heights = c(1,0.8),
-          rel_widths = 1)
-
-plot_grid(plotlist = list(legend, boxplot.richness),
-                     align = "hv",
-                     ncol = 1, 
-          axis = "t",
-          rel_heights  = c(.5,1),
-          labels = c(NA,"B" )) -> bottom.right
-
-plot_grid (plotlist = list(map, bottom.right),
-           align = "hv",
-           nrow = 1,
-           rel_heights = 1,
-           rel_widths = c(.6,0.4),
-           labels = c("A", NA)) -> bottom.plot
-
-plot_grid(bottom.plot, ridges,
-          align = "h",
-          nrow = 2,
-          rel_heights = c(0.6, 0.4),
-          labels = c(NA,"C")) +
-  ggsave(filename = here("Figures","Manuscript","Fig1.png"), width = 14, height = 14, dpi = "retina")
+THe top will be the environmental gradients, the lower panels will have the map and the boxplot. Create a custom palette of colors for the two regions
+
+```{r assmbling Fig1}
+Site.palette <- c("#d8b365", "#5ab4ac")
+
+layout <- "
+AAABBB
+AAABBB
+AAACCC
+DDDDDD
+DDDDDD"
+
+wrap_plots(A = map ,
+           B = boxplot.richness,
+           C = guide_area(),
+           D = ridges + labs(x = ""),
+            design = layout,
+            heights = c(10,10,1,10,10),
+           guides = "collect")  +
+  plot_annotation(tag_levels = c("A", "1"), tag_sep = ".")  & 
+  theme_minimal() &
+  theme(legend.direction = "horizontal",
+        strip.background = element_rect(fill = "#2E3440"),
+        strip.text = element_text(face = "bold", color = "#D8DEE9") ) &
+  scale_fill_manual(values = Site.palette) -> Fig1
+
+Fig1
+ggsave(Fig1, filename = here("Figures","Manuscript","Fig1.png"), width = 14, height = 12, dpi = "retina")
 
@@
   separate (event, into = c("Site", "Date"), sep = "_") %>% 
-  mutate(Season = case_when(str_detect(Date,"17-03|17-10|17-11|18-01|18-03") ~ "Winter",
-                             TRUE ~                                         "Summer"),
-         Area = case_when(Site %in% c("CP","LK", "FH") ~ "San Juan Island",
+  mutate(Area = case_when(Site %in% c("CP","LK", "FH") ~ "San Juan Island",
                           TRUE                         ~ "Hood Canal")) %>% 
@@
 
-Step by step
+Step by step. Create a column for each stage of the CAP analysis - and stop to check the signifficance of the CAP
 ```{r CAP1}
@@
 
+Now we want to see if there are clusters associated with this.
+A first step is to choose how many clusters do we need to form.
+There are several methods for evaluating the goodness of fit of the number of clusters.
 So the best numbers of clusters are
@@
 ### Panel A: CAP analysis with environmnental correlation and cluster info
+
 ```{r Hand made CAP}
@@
 
-  all.sim.data %>% 
-    mutate (Area = fct_recode(Area, "San Juan Island" = 
-                                "SJI")) %>% 
+full.grid %>% 
+  mutate (Area = str_replace(Area, replacement = "San Juan Island", pattern =  "SJI")) %>%
+  rename(Temperature = TemperatureCelsius,
+         pH_new = pH_units) %>% 
     group_by(Area) %>% 
     nest() %>% 
-    mutate(pH.range = map(data, ~ seq(round(min(.x$pH_units),1), 
-                                      round(max(.x$pH_units),1),
-                                      length.out = 12)),
-           T.range =  map(data, ~ seq(round(min(.x$TemperatureCelsius),1), 
-                                      round(max(.x$TemperatureCelsius),1),
-                                      length.out = 22)),
-           output.range = map2(pH.range, T.range, ~ expand_grid(pH_new = .x, Temperature = .y))) %>%
-
     left_join(clusters) %>% 
-    mutate(whichComm.res = map2(output.range,multi.mod,~ .x %>% 
+    mutate(whichComm.res = map2(data,multi.mod,~ .x %>% 
                                   mutate(comm = as.numeric (apply(predict(.y, type = "probs", newdata = .x),1, which.max)) )
@@
 
-## Get the data for all seasons and areas
+## Get the data for all  areas
  SIMPER.clusters %>%
@@
     theme_light() + theme(axis.text.y = element_text(face = "bold.italic", colour ="#3B4252", hjust = 0 ),
-                          strip.background = element_rect(fill = "#3B4252"),
-                          strip.text = element_text(face = "bold", color = "#EBCB8B"))} 
+                          strip.background = element_rect(fill = "#2E3440"),
+        strip.text = element_text(face = "bold", color = "#D8DEE9")
+        # strip.background = element_rect(fill = "#3B4252"),
+        # strip.text = element_text(face = "bold", color = "#EBCB8B")
+        )} 
   ))   -> right_most
@@
 
-We use the climate projection to predict changes in suitability
-
-```{r}
-
-all.sims.for.projection <- simData.nocrazys %>% 
-  select(Area, Year, Temperature, TemperatureCelsius, pH, pH_units)
+We use the climate projection to predict changes in suitability. First panel is a visualization of the suitability of planktonic taxa, highlighting those species with a more pronounced change.
+
+```{r Prepare dataset}
+models$data %>% 
+  distinct(Area, taxa, taxonAreaIndex) -> keepers
+
+expanded.future <- future.scenario %>% 
+  select(Area, Year, Temperature = Temp.std, TemperatureCelsius, pH = pH.std, pH_units)
 
 ## Get
-models %>% 
-  transmute(taxa, 
-            post.draws = map2 (model, taxa, ~ posterior_predict(object = .x,
-                                                                newdata = all.sims.for.projection,
-                                                                draws = 100) %>% 
-                                 as_tibble() %>% 
-                                 t(.) %>% 
-                                 as_tibble() %>% 
-                                 rownames_to_column("sim.row") )) -> posterior.draws.all
+
+expand_grid(models$data %>% distinct(taxa), expanded.future) -> expanded.future
+
+left_join(expanded.future, keepers) -> expanded.future
+```
+
+```{r Calculate the 100 draws}
+posterior_predict(object = models,
+                   newdata = expanded.future,
+                   draws = 100) -> posterior.predict.all.taxa
 
@@
 
-posterior.draws.all %>% 
- left_join(phylum.info %>% dplyr::select(phylum, taxa)) %>% 
-  group_by(phylum) %>% # Add phylum information
-  nest() %>% 
-   mutate (data = map (data, ~ .x %>% 
-                        unnest() %>% 
-                        pivot_longer(cols = c(-taxa, -sim.row),
-                                    names_to = "draw.row", values_to = "presence") %>% 
-  group_by(taxa, sim.row) %>% 
-  summarise(prob.of.presence = mean(presence)) %>% 
-  left_join(all.sims.for.projection %>%  
-              rownames_to_column("sim.row"))
-
-  )) -> posterior.probabilities.by.taxa
-```
-Now we have a nested dataframe per each phylum. Not that we need it. But the nested part is a dataframe with the mean result of the 100 draws for each environmental point.
+expanded.future %>% left_join(phylum.info) -> expanded.future
+
+expanded.future$present <- colMeans(posterior.predict.all.taxa)
+
+expanded.future$sd <- apply(posterior.predict.all.taxa,2,sd )
+```
+
+Now choose for each phyla, the taxa with more dramatic change
+
+```{r}
+expanded.future %>% 
+  left_join(phylum.info) %>% 
+  nest(-phylum) %>% 
+  # ungroup() %>% 
+  # filter (phylum %in%c("Dinophyceae", "Bacillariophyta")) %>% 
+  mutate(suma = map (data,~.x %>% group_by(taxa,Area, Year) %>%
+                       summarise(tot = sum(present)) %>%
+                       ungroup() %>% 
+                       pivot_wider(names_from = Year, values_from = tot) %>% 
+                        group_by(taxa) %>% mutate(diff = `2017` - `2095`)
+                     )) %>% 
+  select(phylum, suma) %>% 
+  unnest(suma) %>%
+  group_by(phylum) %>% arrange(desc(abs(diff))) %>% 
+  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
+  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
+                           species == "NA" ~  paste0(genus, " sp."),
+                           TRUE            ~ species)) -> taxa.and.diff
+
+taxa.and.diff %>%
+  slice(1:5) %>%
+  arrange(diff) %>% pull(taxa) -> ahorasi
+
+taxa.and.diff %>% 
+  slice(1:5) %>%
+  arrange(diff) %>% pull(label) -> reorder.labels
+```
+
+```{r Fig3left}
+
+expanded.future %>% 
+  filter(taxa %in% ahorasi) %>%
+  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
+  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
+                           species == "NA" ~  paste0(genus, " sp."),
+                           TRUE            ~ species))    %>%
+  mutate(Area = as.character(Area)) %>%
+  mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
+                          TRUE          ~ Area)) %>%
+  mutate(Year = fct_rev(as.factor(Year) )) %>% 
+  nest(-phylum) -> input.for.left
+
+input.for.left %>% 
+  mutate( plot.by.taxa = map2 (data, phylum, ~ .x %>%
+                                 ggplot(aes(x= present,
+                                            y = fct_relevel(label,reorder.labels))) +
+                                 scale_fill_brewer(name = "Year",  type = "qual", palette = 4, direction = 1,
+                                                   breaks=c("2017","2095")) +
+                                 labs (x = "Probability of presence",
+                                       y = "") +
+                                 facet_wrap(~Area) +
+                                 theme_minimal() +
+                                 theme(strip.text.y = element_text(angle = 0),
+                                       axis.text.y = element_text(face = "italic"),
+                                       legend.direction = "horizontal",
+                                       legend.position  = "bottom",
+                                       plot.title.position = "plot",
+                                       plot.title = element_text(face = "bold",vjust =  -5 ),
+                                       axis.title.x.top  = element_text(face = "bold",vjust = 3) ) +
+                                 ggtitle(.y)  +
+                                 scale_x_continuous(limits = c(0,1),position = "top") )) %>%
+  pull(plot.by.taxa) %>% 
+  set_names(input.for.left$phylum) -> fig3.left
+
+map(fig3.left, function(x) x+ geom_boxplot(aes(fill = as.factor(Year)))) -> fig3.left
+
+fig3.left["Dinophyceae"]
+```
 
 ### Richness by phyla with time
- First create the square matrix of points - the ggplot geometries requires one entry per grid - so we have to comply
-```{r create full grid}
-
-all.sims.for.projection  %>% 
-  group_by(  Area) %>% 
-  # filter (keep == "keep") %>%   
-  summarise_at(c("TemperatureCelsius", "pH_units"), list(max = max, min = min)) %>%
-
-  group_by ( Area) %>% 
-  nest() %>% 
-  mutate(data = map (data, ~ expand_grid(TemperatureCelsius = seq(round(.x$TemperatureCelsius_min,1), round(.x$TemperatureCelsius_max, 1),by = 0.1),
-                                         pH_units = seq(round(.x$pH_units_min,1), round(.x$pH_units_max,1), by = 0.1)))) %>% 
-  unnest(data) -> full.grid
-```
-
-```{r add standarized temperatures to the grid}
-
+ First prepare the grid - expand it to include the taxa
+```{r}
 full.grid %>% 
-  ungroup() %>% 
-  rownames_to_column("Case") %>% 
-  rename(Temperature = TemperatureCelsius,
-         pH = pH_units) %>%
-
-  pivot_longer(cols = c(Temperature, pH),
-               names_to  = "Predictor",
-               values_to = "value") %>% 
-  group_by(Area, Predictor) %>% 
-  nest() %>% 
-  left_join(transforming.df) %>% 
-  mutate(data = map2(data, to.std, ~.x %>% 
-                       mutate(std = .y[1]+.y[2]*value ))) %>% 
-  select(Area, Predictor, data) %>% 
-  unnest(data) %>% 
-  pivot_wider(names_from = Predictor,
-              values_from = c(std,value)) %>% 
-  rename(Temperature = std_Temperature,
-         pH = std_pH,
-         TemperatureCelsius = value_Temperature,
-         pH_units           = value_pH) -> full.grid
-```
-
-`full.grid` has the temperature and pH values we want to project our models to - so now we use it with the `models` output.
-
-```{r calculate the 100 draws from the posterior}
-models %>% 
-  ungroup() %>%  #rpk added
-  transmute(taxa, 
-            post.draws = map2 (model, taxa, ~ posterior_predict(object = .x,
+  select(Area,  Temperature = Temp.std, TemperatureCelsius, pH = pH.std, pH_units) -> full.grid
+
+expand_grid(models$data %>% distinct(taxa), full.grid) -> full.grid
+
+left_join(full.grid, keepers) -> full.grid
+```
+
+Now calculate the posterior from each point in the grid
+```{r}
+posterior_predict(object = models,
                    newdata = full.grid,
-                                                                draws = 100) %>% 
-                                 as_tibble() %>% 
-                                 t(.) %>% 
-                                 as_tibble() %>% 
-                                 rownames_to_column("sim.row") )) -> posterior.draws.raster
-```
-
-```{r calculate richness}
-posterior.draws.raster %>% 
- left_join(phylum.info %>% dplyr::select(phylum, taxa)) %>% 
+                   draws = 100) -> posterior.predict.all.taxa
+
+```
+
+Add the phylum information and calculate the richness at each point
+```{r}
+full.grid %>% left_join(phylum.info) -> full.grid
+
+full.grid$present <- colMeans(posterior.predict.all.taxa)
+
+full.grid$sd <- apply(posterior.predict.all.taxa,2,sd )
+
+full.grid %>% 
+  group_by(Area, Temperature = TemperatureCelsius, pH = pH_units, phylum) %>% 
+  summarise(richness = sum(present),
+            sd       = sum(sd^2)) %>% 
+#  left_join(richness.by.phylum) %>% 
   group_by(phylum) %>% 
-  nest() %>% 
-  mutate(data = map (data, ~ .x %>% unnest(cols= post.draws) %>% 
-                       pivot_longer(cols = c(-taxa, -sim.row),
-                                    names_to = "draw.row", values_to = "presence") %>% 
-                       group_by(draw.row, sim.row) %>% 
-                       summarise(tot = sum(presence)) %>% 
-                       group_by(sim.row) %>% 
-                       summarise_at(.vars = "tot", .funs = list(Richness = mean,sd = sd)))) %>% 
-  unnest(data) -> ready_plot
-
-ready_plot %>% 
-  left_join(full.grid %>%  
-              rownames_to_column("sim.row")) -> ready_plot
-
-# Normailzed the data for each phyla
-
-ready_plot %>% 
-  group_by(phylum) %>% 
-  mutate(st.richness = Richness / max(Richness)) -> ready_plot
-
-```
-
-```{r plot3 function}
-plot3 <- function(tibble,string , response) {
+  mutate(st.richness = richness/max(richness)) -> richness.by.phylum
+```
+
+```{r plot4 function}
+plot4 <- function(tibble, response) {
 
@@
                             TRUE          ~ Area)) %>% 
-    group_by(TemperatureCelsius, pH_units, Area) %>%
+    mutate(Temperature = round(Temperature, 1),
+           pH = round(pH, 1)) %>% 
+    group_by(Temperature, pH, Area) %>%
 
@@
 
-    ggplot(aes(x = round(TemperatureCelsius,10),
-             y = round(pH_units, 10))) +
+    ggplot(aes(x = Temperature,
+             y = pH) ) +
 
@@
   scale_fill_distiller(name = "Standarized Richness", type = "seq", palette = "Spectral",limits=c(0,1) ) +
-
-  # geom_density2d(data = simData %>%
-  #                   filter (Year %in% c(2015, 2095)),
-  #                 aes(x= TemperatureCelsius, y = pH_units, group= Year, color = as.factor(Year)), bins = 5, alpha = 0.5) +
 
@@
     scale_x_continuous(position = "top")
-
 }
@@
 
-```{r}
-ready_plot %>% 
+```{r plot4 in play}
+richness.by.phylum %>% 
   nest(-phylum) %>% 
-  mutate(plot = map2(data, phylum, ~ plot3(.x, .y, response = st.richness)))  -> list.of.richness.plots
-
-```
-
-```{r}
+  mutate(plot = map(data, ~ plot4(.x, response = st.richness)))  -> list.of.richness.plots
 
@@
 
-list.of.richness.plots %>% pull(plot) -> list.of.raster.plots
-
-list.of.raster.plots <- set_names(list.of.raster.plots, nm = list.of.phyla)
+list.of.richness.plots %>% pull(plot) -> fig3.right
+
+fig3.right <- set_names(fig3.right, nm = list.of.phyla)
+
+fig3.right["Dinophyceae"]
+fig3.right["Bacillariophyta"]
 
@@
 ```{r select some }
-posterior.probabilities.by.taxa %>% 
-  ungroup() %>% 
-  filter (phylum %in%c("Dinophyceae", "Bacillariophyta")) %>% 
-  mutate(suma = map (data,~.x %>% group_by(taxa,Area, Year) %>%
-                       summarise(tot = sum(prob.of.presence)) %>%
-                       ungroup() %>% 
-                       pivot_wider(names_from = Year, values_from = tot) %>% 
-                        group_by(taxa) %>% mutate(diff = `2017` - `2095`)
-                     )) %>% 
-  select(phylum, suma) %>% 
-  unnest(suma) %>%
-  group_by(phylum) %>% arrange(desc(abs(diff))) %>% 
-  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
-  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
-                           species == "NA" ~  paste0(genus, " sp."),
-                           TRUE            ~ species)) -> taxa.and.diff
-
-taxa.and.diff %>%
-  slice(1:5) %>%
-  arrange(diff) %>% pull(taxa) -> ahorasi
-
-taxa.and.diff %>% 
-  slice(1:5) %>%
-  arrange(diff) %>% pull(label) -> reorder.labels
-
-# Write to file the posterior probabilietes by taxa so Ryan can play with it
-# posterior.probabilities.by.taxa %>% 
-#   unnest(data) %>% 
-#   write_csv(here("Output", "posterior.probability.by.taxa.csv"))
-```
-
-```{r the final one}
-
-posterior.probabilities.by.taxa %>%
-
-ungroup() %>%
-filter (phylum %in%c("Dinophyceae", "Bacillariophyta")) %>%
-mutate( plot.by.taxa = map2 (data, phylum, ~ .x %>%
-filter(taxa %in% ahorasi) %>% 
-separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
-mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
-species == "NA" ~  paste0(genus, " sp."),
-TRUE            ~ species))    %>%
-mutate(Area = as.character(Area)) %>%
-mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
-TRUE          ~ Area)) %>%
-mutate(Year = fct_rev(as.factor(Year) )) %>%
-ggplot(aes(x= prob.of.presence,
-y = fct_relevel(label,reorder.labels))) +
-#   geom_violin(aes(fill= (Year)), weight = 2, adjust = 1, draw_quantiles =  0.5, scale = "count")+
-scale_fill_brewer(name = "Year",  type = "qual", palette = 4, direction = 1, breaks=c("2017","2095")) +
-labs (x = "Probability of presence",
-y = "") +
-facet_wrap(~Area, ) +
-theme_minimal() +
-theme(strip.text.y = element_text(angle = 0),
-axis.text.y = element_text(face = "italic"),
-legend.direction = "horizontal",
-legend.position  = "bottom",
-plot.title.position = "plot",
-plot.title = element_text(face = "bold",vjust =  -5 ),
-axis.title.x.top  = element_text(face = "bold",vjust = 3) ) +
-ggtitle(.y)  +
-scale_x_continuous(limits = c(0,1),position = "top") )) %>%
-pull(plot.by.taxa) %>% set_names(.$phylum) -> fig3.left
-
-map(fig3.left, function(x) x+ geom_boxplot(aes(fill = as.factor(Year)))) -> fig3.left
-
-posterior.probabilities.by.taxa %>%
-ungroup() %>%
-filter (phylum %in%c("Dinophyceae", "Bacillariophyta")) %>%
-  pull(phylum) %>%
-  set_names(fig3.left, .) -> fig3.left
-
-list.of.raster.plots[c("Dinophyceae", "Bacillariophyta")] -> fig3.right
 
@@
 ```{r}
-posterior.probabilities.by.taxa %>% 
-  ungroup() %>% 
-  # filter (phylum %in%c("Dinophyceae", "Bacillariophyta")) %>% 
-  mutate(suma = map (data,~.x %>% group_by(taxa,Area, Year) %>%
-                       summarise(tot = sum(prob.of.presence)) %>%
-                       ungroup() %>% 
-                       pivot_wider(names_from = Year, values_from = tot) %>% 
-                        group_by(taxa) %>% mutate(diff = `2017` - `2095`)
-                     )) %>% 
-  select(phylum, suma) %>% 
-  unnest(suma) %>%
-  group_by(phylum) %>% arrange(desc(abs(diff))) %>% 
-  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
-  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
-                           species == "NA" ~  paste0(genus, " sp."),
-                           TRUE            ~ species)) -> taxa.and.diff
-
-taxa.and.diff %>%
-  slice(1:5) %>%
-  arrange(diff) %>% pull(taxa) -> ahorasi
-
-taxa.and.diff %>% 
-  slice(1:5) %>%
-  arrange(diff) %>% pull(label) -> reorder.labels
-```
-
-```{r}
-posterior.probabilities.by.taxa %>%
-
-ungroup() %>%
-mutate( plot.by.taxa = map2 (data, phylum, ~ .x %>%
-#filter(taxa %in% ahorasi) %>% 
-separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
-mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
-species == "NA" ~  paste0(genus, " sp."),
-TRUE            ~ species))    %>%
-mutate(Area = as.character(Area)) %>%
-mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
-TRUE          ~ Area)) %>%
-mutate(Year = fct_rev(as.factor(Year) )) %>%
-ggplot(aes(x= prob.of.presence,
-# y = fct_relevel(label,reorder.labels))) +
- y = label)) + 
-#   geom_violin(aes(fill= (Year)), weight = 2, adjust = 1, draw_quantiles =  0.5, scale = "count")+
-scale_fill_brewer(name = "Year",  type = "qual", palette = 4, direction = 1, breaks=c("2017","2095")) +
-labs (x = "Probability of presence",
-y = "") +
-facet_wrap(~Area, ) +
-theme_minimal() +
-theme(strip.text.y = element_text(angle = 0),
-axis.text.y = element_text(face = "italic"),
-legend.direction = "horizontal",
-legend.position  = "bottom",
-plot.title.position = "plot",
-plot.title = element_text(face = "bold",vjust =  -5 ),
-axis.title.x.top  = element_text(face = "bold",vjust = 3) ) +
-ggtitle(.y)  +
-scale_x_continuous(limits = c(0,1),position = "top") )) %>%
-
-pull(plot.by.taxa)   -> Suppl.Material.plots
-
-map(Suppl.Material.plots, function(x) x+ geom_boxplot(aes(fill = as.factor(Year)))) -> Suppl.Material.plots
-
-posterior.probabilities.by.taxa %>%
-ungroup() %>%
-  pull(phylum) %>%
-  set_names(Suppl.Material.plots, .) -> Suppl.Material.plots
-
-list.of.raster.plots[[1]]
-
-names(Suppl.Material.plots) == names(list.of.raster.plots)
-
-map2(Suppl.Material.plots, list.of.raster.plots, 
+
+map2(fig3.left, fig3.right, 
      function(.x,.y){
 
-       .x  +plot_spacer()+ .y + plot_layout(widths = c(4,1,4)) -> plot
-
-        return(plot)
+       .x  + plot_spacer()+ .y + plot_layout(widths = c(4,1,4))
 
@@
 
-names (Suppl.Material.plots.final) <- names(Suppl.Material.plots)
-map2 (Suppl.Material.plots.final,names(Suppl.Material.plots.final),   ~ ggsave(plot = .x, filename = here("Figures","Supplementary",paste0("Supp.fig",.y,".png")),device = "png",
-                units = "in", dpi = "retina", width = 20, height = 12)
-                                                                               )
-```
+paths <- paste0("Supp.fig.",names(fig3.left),".png")
+
+pmap(.l = list(Suppl.Material.plots.final, paths),
+     .f = function(a,b) a +  ggsave(filename = here("Figures","Supplementary", b),
+                                    device = "png",
+                                    units = "in",
+                                    dpi = "retina",
+                                    width = 20,
+                                    height = 12))
+
+```

### 2020-08-13 Scripts/From.asv.to.data.plankton.Rmd
All ready to go

---
title: "From ASV to data.plankton"
output: html_notebook
---

This script will take the ASV table from the pipeline, and combine it with three metadata sources:
  * Taxonomical annotation
  * Biomineralization & lifestyle; these two have already been combined in the Rmarkdown "taxonomy.biom.food.env.Rmd"
  * Environmental data

And will create the final dataset of plankton species, with their Normalized reads and the environmental data associated with them

```{r}
library (tidyverse)
library (here)
source(here("Scripts", "eDNAindex.R"))
ASV.table  <- read_csv(here("Input", "ASV_table_all_together.csv"))
Annotation <- read_csv(here("Input", "hash.annotated.csv"))
env.data   <- read_csv(here("Input", "env.data.updated.csv"))
```

```{r}
ASV.table %>% 
  filter(!Hash %in%  (Annotation %>% 
           filter (str_detect(species, "Homo")) %>% 
           pull(Hash))) -> ASV.table

left_join(ASV.table, Annotation) %>% 
  unite(family, genus, species, col = "taxa", sep = "|") %>% 
  filter(taxa != "NA|NA|NA") %>% 
  select(taxa, sample, nReads, benthos) %>% 
  group_by(taxa, benthos, sample) %>% 
  summarise(nReads = sum(nReads)) -> ASV.by.taxa

ASV.by.taxa %>% 
  separate(sample, into = c("Site", "biol"), sep = "_", remove = F) %>% 
  separate(biol, into = c("biol", "tech.rep"), sep = "\\.") %>% 
  separate(biol, into = c("date", "biol.rep"), sep = -1) %>% 
  unite (Site, date, col = "event") -> ASV.by.taxa

ASV.by.taxa %>%

  filter(!str_detect(event,"201710|201703" ) , benthos != "BEN") %>%

  eDNAindex(Sample_column = event,
            OTU_column = taxa,
            Counts_column = nReads,
            Biological.replicate = biol.rep) -> ASV.plankton.event

ASV.plankton.event %>% 
  separate(event, into = c("Site", "Date")) %>% 
  mutate(Date = ymd(Date, truncated = T)) ->  ASV.plankton.event

env.data %>%  
   select( -contains("DO")) %>% 
   drop_na() %>% 
   right_join (ASV.plankton.event) %>% 
   unite (Site, Date, col = "event", sep = "_") %>% 
  write_csv(here("Input", "Combined_Biol_Env_Plankton.csv"))
```

### 2020-08-13 Scripts/Prepare.data.for.stan.Rmd
All ready to go

---
title: "Prepare data for stan"
output: html_notebook
---

Turn the data into the format needed for stan.

```{r}
library(tidyverse)

plankton <- read_csv(here("Input","Combined_Biol_Env_Plankton.csv"))
phylum.info <- read_csv(here("Input","higher_taxonomy.csv")) %>% 
  unite(family, genus, species, sep = "|", col = "taxa") %>% 
  select(phylum, taxa)

plankton %>% 
  mutate(sample = str_remove_all(event, "-"),
         sample = str_remove(sample, "01$")) %>% 
  select(Area, sample, taxa, DIC = mean.DIC, Salinity = Salinity.new, pH = pH_new, Temperature) %>% 
  mutate(binary = 1) %>% 
  pivot_wider(names_from = taxa, values_from = binary, values_fill = list(binary = 0)) %>% 
  pivot_longer(cols = 7:236, names_to = "taxa", values_to="binary") %>% 
  filter(sample != "LL_201711") %>% 
  filter(!str_detect(taxa, "Gallus")) %>%
  left_join(phylum.info) %>% write_csv(here("Input","data.for.stan.csv"))

```

### 2020-08-13 Scripts/taxonomy.biom.food.env.Rmd
All ready to go

---
title: "Annotating mineralization and environment"
output: html_notebook
---

Loading the all.fams.csv, combine it with the biomineralization data.  And also annotate it with the info of whether they are planktonic or benthic taxa.
The workflow would be to create unique

```{r load packages and database}

library(tidyverse)
library(googlesheets)
library(taxize)
library(rentrez)
library(here)
source(here("Scripts","sequential.join.r"))

# set_entrez_key("")
# Sys.getenv("ENTREZ_KEY")

all.fams <- read_csv(here("Input","all.fams.tax.csv"))

all.hashes.found <- read_csv(here("Input","all.taxonomy.20190130.csv"))

biom <- read_csv(here("Input","Biomineralization.csv"))

```

So the workflow is going to be to try and get the information from each taxonomical rank. So I will create a function that takes 2 datasets, and binds them sequentially. For each join, it will add a column with the information, the taxonomical level at where it got it, and the literature source for that information.

```{r}
biom %>% 
  unite(Genus, Species,col = "species", sep = " ", remove = F) %>% 
  select(-Species) -> biom

biom %>% distinct (AdultMinerology)

#Aragonite -calcite and Calcite - aragonite

```

Let's see how many of the families are represented there

```{r}

summary(all.fams$family %in% biom$Family)

biom %>% 
  filter(Family %in% all.fams$Family) %>% 
  group_by(Family) %>% 
  summarise(n.sp  = n_distinct(Genus),
            n.ELS = n_distinct(EarlyLifeMinerology),
            n.ALS = n_distinct(AdultMinerology)) %>% 
  filter(n.ELS > 1 | n.ALS > 1) %>%  pull(Family) -> fams.with.diff.miner

```

We don't have information from 227 families, And from 15 families we have more than one way of mineralizing either larvae or adult. The approach we have followed is - if there is an annotation for its species, we take that; if there is a genus and if not, family.

```{r}
biom %>% 
  select(family = Family, genus = Genus, species, EarlyLifeMinerology, AdultMinerology, everything()) -> biom

sequential_join(df1 = all.hashes.found,
                 df2 = biom,
                 var.to.join.1 = "species" ,
                 var.to.join.2 = "genus" ,
                 var.to.join.3 = "family" ,
                 confidence = reliability.minerology,
                EarlyLifeMinerology, AdultMinerology) -> database.with.minerology

# For all families, add the order and class so we can get the in-between taxonomy

lineages=list(NA)
for (i in 1:length(unique( all.fams$family))){
        tempx=classification(unique( all.fams$family)[i], db="ncbi")
    lineages[i]=ifelse(isTRUE(tempx), NA, tempx)
}

unique( all.fams$family)[-which(is.na(lineages))] -> fam.to.update

lineages[-which(is.na(lineages))] -> lineages

bind_rows(lineages, .id = "family") %>% 
  group_by(family) %>% 
  nest() %>% 
  mutate(family = fam.to.update) %>% 
  mutate(order = map(data, ~ .x %>%
                        #filter(rank == "order") %>% pull(name)))
                        filter(rank != "no rank") %>%
                        select(-id) %>% 
                        spread(key = rank, value = name))) %>% 
  unnest(order) %>% 
  select(family, order, class) %>% 
  left_join(all.fams, .) -> all.fams

unique( all.fams$family)

# Now join the database with the taxonomy

database.with.minerology %>% 
  left_join(all.fams, by = "family") %>% 
  select(Hash, kingdom, phylum, class, order, family, genus, species, contains("tot."), contains("Minerology")) -> database.with.minerology

```
### Fill a few gaps in the taxonomy

Genus Dictyocha is a Stramenopiles frrom the phyllum Dictyochophyceae

```{r}
database.with.minerology %>% 
  filter(is.na(phylum), !is.na(genus)) %>% distinct(genus) %>% pull -> genus.to.lineage

genus.to.lineage

lineages=list(NA)
for (i in 1:length(genus.to.lineage)){
        tempx=classification(genus.to.lineage[i], db="ncbi")
    lineages[i]=ifelse(isTRUE(tempx), NA, tempx)
}

 bind_rows(lineages, .id = "genus") %>% # NOW CHANGE STRAMENOPILES TO PHYLUM
   mutate (rank = case_when(name == "Stramenopiles" ~ "kingdom",
                            TRUE ~ rank)) %>% 
   group_by(genus) %>%
   nest %>%
   mutate(genus = genus.to.lineage) %>%
   mutate(data2 = map(data, ~.x %>%
                        select(-id) %>%
                        filter(rank != "no rank") %>% 
                        spread(key = rank, value = name) ) ) %>%  
   unnest (data2) %>% 
   mutate(phylum = case_when(is.na(phylum) ~ class,
                             TRUE ~ phylum),
          phylum = case_when(is.na(phylum) ~ order,
                             TRUE ~ phylum)) %>% 
   select(genus, order, class, phylum, kingdom) -> fill.the.gaps
```

Now merge the new info

```{r}

database.with.minerology %>%  select(-kingdom, -phylum, -order, -class) %>% inner_join(fill.the.gaps, by = "genus") -> step1

database.with.minerology %>% anti_join(fill.the.gaps, by = "genus") %>% bind_rows(step1) -> database.with.minerology
```

# Step 2: add the info from trophic level

Trying to remember where did I get this file from. I think it comes from the MBON project. Check with Ryan. It comes from Collin

```{r}

trophic <- read_csv("/Users/ramongallego/GoogleDrive/Kelly_Lab/Projects/OA_eDNA/Data/TrophicData/subnetxtrophic_ordered_20181112_20181130.csv")

```

Check how many families are present in the dataset

```{r}
trophic %>%
  select(family = Rank5,

         order = Rank4,

         class = Rank3,

         phylum = Rank2,

         trophic_level_fv, group_fv) -> trophic

trophic %>% 
  filter(is.na(Rank5)) %>%  distinct(Rank4) %>% pull -> NAs.family

lineages=list(NA)
for (i in 1:length(NAs.family)){
        tempx=classification(NAs.family[i], db="ncbi")
    lineages[i]=ifelse(isTRUE(tempx), NA, tempx)
}

NAs.family[-which(is.na(lineages))] -> NAs.family

lineages[-which(is.na(lineages))] -> lineages

bind_rows(lineages, .id = "Rank4") %>% 
  group_by(Rank4) %>% 
  nest() %>% 
  mutate(Rank4 =  NAs.family) %>% # Make it an order
  mutate(last = map_chr (data, ~.x %>% last() %>% pull(name)))

```

No info for 256 families

```{r}
summary(all.fams$phylum %in% trophic$Rank4) # 4 by class

summary(all.fams$phylum %in% trophic$Rank3) # 51 by class

summary(all.fams$phylum %in% trophic$Rank2) # 299 by phyllum
```

Let's do a weird_join as before
```{r}

sequential_join(database.with.minerology,
                trophic,
                var.to.join.1 = "family",
                var.to.join.2 = "class",
                var.to.join.3 = "phylum",
                confidence = reliability.trophic,
                trophic_level_fv, group_fv) -> database_step.2

database_step.2 %>% 
  filter(reliability.trophic == "class")

```

# Step 3: Benthic or planktonic

OK, now the last step is to look for sources of life style, whether benthic or planktonic.

```{r load dino}

# First, a database from Gomez 2012

dinoflag <- read_csv("dinoflag_annotations.csv")
dinoflag %>% 
  select(family = Family, everything()) -> dinoflag
```

```{r load tara_oceans}
# Second, the dataset from the tara_oceans dataset - if it was on the first couple of meters, can we 
# assume it is a planktonic species?

tara_oceans <- read_delim("Tara_oceans.tsv", 
    "\t", escape_double = FALSE, trim_ws = TRUE)

tara_oceans %>% 
  select(lineage, taxogroup, chloroplast, trophism) %>% 
  distinct() -> tara_oceans
tara_oceans %>% 
  filter(str_detect(lineage, pattern ="Amoebozoa" ))

 database.step.2 %>% distinct(kingdom) %>% arrange(kingdom) %>% filter(kingdom != "Metazoa") %>% pull() -> kingdoms.to.find.out

# Check how many species from our dataset are present in the Tara Oceans

tara_oceans %>% 
  filter(str_detect(lineage, pattern = paste(unique(database.step.2$species), collapse = "|"))) # Five spp

tara_oceans %>% 
  filter(str_detect(lineage, pattern = paste(unique(database.step.2$genus), collapse = "|"))) %>% 
  filter(str_detect(lineage, pattern = "Annelida"))# 451 genus

tara_oceans %>% 
  filter(str_detect(lineage, pattern = paste(unique(database.step.2$family), collapse = "|"))) # 289 families
```

```{r}
database_step.2 %>% 
  select(-n) -> database_step.2
sequential_join(database_step.2,
                dinoflag,
                var.to.join.1 = "species",
                var.to.join.2 = "genus",
                var.to.join.3 = "family",
                confidence = reliability.benthic,
                benthic, oceanic) -> database_step.3

```

How many did get assigned?

```{r}
database_step.3 %>% 
  group_by(is.na(benthic)) %>% 
  summarise(nhash = n(),
            tot = sum(tot.reads))
```

58 Hashes, 362 thousand reads

```{r load massana2015}
Massana2015 <- read_csv("Massana2015.csv")

Massana2015 %>% 
  select(species = `Closest match to described species`, everything()) %>% 
  separate(species, into = "genus", sep = " ", remove = F) -> Massana2015

summary(med.step$genus %in% Massana2015$genus)

Massana2015 %>%
  group_by(genus) %>% 
  nest(species,Source) %>% 
  mutate(Ben = map_dbl (data, ~ sum(str_count(.x$Source, "Benthos"))),
         Plk = map_dbl (data, ~ sum(str_count(.x$Source, "plankton"))),
         Source = case_when(Ben == 0  ~ "PLK",
                            Plk == 0  ~ "BEN",
                            TRUE      ~ "Both")) %>% 
  select(genus, Source) -> Massana2015
```

```{r join Massana}
sequential_join(df1 = database_step.3,df2 = Massana2015, var.to.join.1 = "genus", var.to.join.2 = "",var.to.join.3 = "",confidence = confidence.source, Source) -> database_step.4

```

Let's see what is left to assign

```{r}
database_step.3 %>% 
  filter(is.na(benthic)) %>% 
  select(benthic,kingdom,phylum, class, order, family, genus, species ) %>% 
  distinct() %>% 
  arrange(kingdom)

```

Look for protist info from Forster2016. From there, turns out there is a database for planktonic protist - so let's look at it
```{r PFR2 }
PFR2_Metadata <- read_delim("PFR2_Metadata.txt", "\t", escape_double = FALSE, trim_ws = TRUE)

PFR2_Metadata %>% 
  select(Genus, Species) %>% 
  unite(Genus, Species, col = "Species", sep = " ", remove = F) %>% 
  distinct(Species, .keep_all = T) -> PFR2_Metadata

lineages=list(NA)
for (i in 1:length(PFR2_Metadata$Genus)){
        tempx=classification(PFR2_Metadata$Genus[i], db="ncbi")
    lineages[i]=ifelse(isTRUE(tempx), NA, tempx)
}

PFR2_Metadata$Genus[-which(is.na(lineages))] -> good.names

lineages[-which(is.na(lineages))] -> lineages

bind_rows(lineages, .id = "genus") %>% 
  group_by(genus) %>% 
  nest() %>% 
  mutate(genus = good.names) %>% 
  mutate(data2 = map (data, ~ .x %>% 
                        filter(rank %in% c("family", "order", "phylum")) %>% 
                        select(-id) %>% 
                        spread(key = rank, value = name))) %>%

  unnest(data2) -> test

# do they share any taxa?

summary(med.step$species %in% PFR2_Metadata$Species) # No species shared

unique(med.step$genus)[(unique(med.step$genus) %in% PFR2_Metadata$Genus)] # no genus shared -

unique (test$family) %in% unique(database_step.4$family)

# No Shared taxa here

```

None of them matched our samples
So, this seems like really complicated. Let's change the course of action. Print out the unique taxa/metadata as a csv file and fill the data with references.

```{r}
med.step %>%
  select(-Hash, -starts_with("tot.")) %>%
  distinct %>%
  select(kingdom, phylum, family, genus, species, benthic, benthic.reliability, everything()) %>%
  arrange(desc(kingdom )) %>% filter(is.na(family))
#   write_csv("taxa.for.life.style.csv")
```

Done some editing - see if there is a big chunk I can do from the Massana work

```{r}
med.step.summ <- read_csv("taxa.for.life.style.csv")

med.step.summ %>% 
  filter (str_detect(phylum, "Florideoph"))

med.step.summ %>% 
  filter(is.na(benthic)) %>% 
  left_join(Massana2015, by = "genus") %>% 
  filter(!is.na(Source)) %>% 
  mutate(benthic = Source,
         benthic.reliability = "genus",
         source.benthic = "Massana2015") %>% 
  select(  - Source ) -> to.bind

med.step.summ %>% #filter(is.na(family))
  filter(!is.na(benthic)) %>% 
  bind_rows(to.bind) %>% 
  bind_rows(med.step.summ %>% 
              filter(is.na(benthic)) %>%
              anti_join(Massana2015, by = "genus")) %>%
  write_csv("taxa.for.life.style2.csv")

```
Emily Jacobs-Palmer found another source of benthic / pelagic diatoms

```{r loading Kasim2006}

Kasim2006 <- read_csv("Kasim_Mukai2006")

# Separate No. and Name of species
Kasim2006 %>% 
  separate(`No. Name of species`, into = c("No", "Species"), sep = " ", extra = "merge") %>% 
  separate(Species, into = c("genus","species"), sep = " ", remove = F) %>%
  unite(genus, species, col = "species", sep = " ", remove = F) %>% 
  mutate(Source = case_when(Group == "B" ~ "BEN",
                            Group == "P" ~ "PLK")) -> Kasim2006

#Do species from the same genus share lifestyle?

Kasim2006 %>% 
  group_by(genus) %>% 
  summarise(n = n_distinct(Source)) %>% 
  filter (n>1) # No,so I can simplify the table

Kasim2006.gen <- Kasim2006 %>% 
  group_by(genus) %>% 
  summarise(Source = first(Source))

# Done - now get the lineages of those genus

lineages=list(NA)

for (i in 1:length(Kasim2006.gen$genus)){
        tempx=classification(Kasim2006.gen$genus[i], db="ncbi")
    lineages[i]=ifelse(isTRUE(tempx), NA, tempx)
}

Kasim2006.gen$genus[-which(is.na(lineages))] -> genus.to.update

lineages[-which(is.na(lineages))] -> lineages

bind_rows(lineages, .id = "genus") %>% 
  group_by(genus) %>% 
  nest() %>% 
  mutate(genus = genus.to.update) %>% 
  mutate(data2 = map (data, ~ .x %>% 
                        filter(rank %in% c("family", "order", "phylum")) %>% 
                        select(-id) %>% 
                        spread(key = rank, value = name))) %>%

  unnest(data2) -> test

Kasim2006.gen %>% 
  inner_join(test %>% select(-data), by = "genus") -> Kasim2006.gen

Kasim2006.gen %>% 
  select(Source.Kasim = Source, everything()) -> Kasim2006.gen

```

Now do the sequential join with Kasim

```{r join Kasim}
database_step.4 %>% 
  select(-n) -> database_step.4

sequential_join(df1 = database_step.4,
                df2 = Kasim2006.gen,
                var.to.join.1 = "genus",
                var.to.join.2 = "family",
                var.to.join.3 = "order",
                confidence = "source.Kasim",
                Source.Kasim) -> database_step.5

# Let's load our personal database - keeping only the benthic info

med.step.2 <- read_csv("taxa.for.life.style2.csv")

summary(med.step.2$family %in% database_step.5$family)

summary(is.na(med.step.2$family))

med.step.2 %>% 
  filter(is.na(family)) %>% 
  distinct(genus) %>% 
  semi_join(database_step.5, by = "genus")

# So all the families are there, and for those that are not, we have the genus - we can join it with database5
```

```{r join csv with data5}

# 1 remove n

database_step.5 %>% 
  select(-n) -> database_step.5

med.step.2 %>% 
  select(Source.csv = benthic, everything()) %>% 
  sequential_join(df1 = database_step.5, 
                  df2 = ., 
                  var.to.join.1 = "species", 
                  var.to.join.2 = "genus", 
                  var.to.join.3 = "family", 
                  confidence = "Summary.benthic",
                  Source.csv,source.benthic ) %>% distinct() -> database_step.5

```
Now that I have all the benthic information in a table, let's deal withthe disagreemnts

```{r differences in benthic assignment}

# Step1
# for which phylum do we have an answer

database_step.5 %>%
  filter(!is.na(benthic) | !is.na(Source) | !is.na(Source.Kasim) | !is.na(Source.csv)) %>% 
  distinct(phylum) # $43 phyla

# Step2
# disagreements

database_step.5 %>%
  #filter(!is.na(benthic) | !is.na(Source) | !is.na(Source.Kasim) | !is.na(Source.csv)) %>% 
  unite(benthic, Source, Source.Kasim, Source.csv, sep = "|", col = all.benthos) %>% 
  mutate(agreed.benthos = case_when(str_detect(all.benthos, "PLK") ~ "PLK",
                                    str_detect(all.benthos, "BEN") ~ "BEN",
                                    str_detect(all.benthos, "Both")~ "Both" ,
                                    TRUE                           ~ "None")) %>% 
  select(Hash, kingdom, phylum, class, order, family, genus, species, contains("minerology"), contains("_fv"), benthos = agreed.benthos
         ) -> to.export

```

Collapse some values of the group_fv variable

```{r fixing group_fv}
write_csv(to.export, path = "hash.annotated.csv")

read_csv("hash.annotated.csv") %>%

  unite (family, genus, species, col ="taxa", sep = "|", remove = F) %>%

  mutate(group_fv = case_when(group_fv == "daitom" ~ "diatom",
                              group_fv == "copepods" ~ "copepod",
                              class == "Clitellata"  ~ "earthworm",
                              phylum == "Nematoda"   ~ "worm",
                              str_detect(taxa, "Harpacticoida") ~"copepod",
                              str_detect(taxa, "Polychaeta") ~"polychaete",
                              group_fv == "amoeboid" ~ "amoebozoa",
                              str_detect(taxa, "Parvamoeba") ~ "amoebozoa",
                              str_detect (group_fv, "echino") ~ "echinoderm",
                              str_detect (group_fv, "flagellate") ~ "flagellate",
                              str_detect(taxa, "Dictyocha|Pseudochatt") ~ "flagellate",
                              str_detect (group_fv, "hydrozoa") ~ "hydrozoan",
                              str_detect (taxa, "Campanulariidae") ~ "hydrozoan",
                              !is.na(group_fv) ~ group_fv,
                              str_detect(taxa, "Dinophysiaceae|Prorocentra|Gonyaulacaceae|Gymnodiniaceae|Kareniaceae|Thoracosphaeraceae") ~ "dinoflagellates",
                              str_detect(taxa, "Skeletonemataceae|Bacillariaceae|Stephanodiscus|Naviculace|Chaetoceros") ~ "diatom",
                              str_detect(taxa, "Cymbomonas|Haramonas|Trebouxia|Pycnoc") ~ "algae, green",
                              str_detect(taxa, "Rhodomelace") ~ "algae, red",
                              kingdom == "Rhodophyta" ~ "algae, red",
                              str_detect(taxa, "Aspergillaceae|Phytop|Haptoglo|Pythia|Pyramimonas|Saproleg|Apodachlya") ~ "fungi",
                              kingdom == "Stramenopiles" ~ "fungi" )) %>% 
  select(-taxa) -> to.export

  write_csv(to.export, path = "hash.annotated.csv")

```
```{r aragonite calcite}
 to_export <- read_csv("hash.annotated.csv")

to_export %>% 
 # slice(1:10) %>% 
  mutate_at(vars(contains("Minerology")), function (x) case_when(x == "Calcite, Aragonite" ~ "Aragonite, Calcite", 
                                                                 x == "Caclite, Aragonite" ~ "Aragonite, Calcite",
                                                                 TRUE                      ~ x)) %>% 
 write_csv("hash.annotated.csv")
```
Fixing some mistakes

```{r}
to_export <- read_csv("hash.annotated.csv")

# Fix the polychaete Sccocirrus
to_export [which (to_export$species == "Saccocirrus sp. CMC01"), c(9:11, 13)] <- list("Non-CaCO3 mineral", "Non-CaCO3 mineral", "family", "polychaete")
# fix some undtermined
to_export[which (to_export$Hash == "fc9fed56b0203902b5eff4eca28b358d4419813a"),12:14] <- list("fungi", "fungi", "None"  )

to_export[which (to_export$Hash == "0124f9c1e949536ce587cd038df6e6f635a8a7d8"),12:14] <- list("3", "copepod", "PLK"  )    
to_export[which (to_export$Hash %in%c("459872f12aa1cf78e8eafa97db881ba6ef6af564",
                                      "b074016d069e564387f67ad18a6e5bc181726411",
                                      "60e1a14acfe5ff45c2b2dbf5b29bb067632a3bf3",
                                      "d6a7fde26da9282448c961966b6d50ab554e76a5",
                                      "2f8238e8b72201b38768d46b32b1266ddc892f0c") ),12:14] <- list("3, worm", "polychaete", "BEN"  )

to_export[which (to_export$Hash %in%c(  "3b86a8af70ee3ae734e65aed41f36d6169e3f60a",
                                        "a14d493abcb3b525eb449d241bd2564ca50eba6f",
                                        "448756e5f9def73770c548b6b772d9d38d4f5bc2") ),12:14] <- list("3, insect", "undetermined", "None"  )                         
# Fill in some taxonomy

  to_export[which(to_export$species == "Harpacticoida sp. DNZ085"), 2:5]  <- list("Metazoa", "Arthropoda", "Hexanauplia", "Harpacticoida")

  to_export[which(to_export$species == "Polychaeta sp. BOLD:ACQ0474"), 2:4] <- list("Metazoa","Annelida","Polychaeta")
  to_export[which(to_export$species == "Polychaeta sp. NB-Po356"), 2:4] <- list("Metazoa","Annelida","Polychaeta")
   to_export[which(to_export$species == "Psocoptera sp. BOLD:ACB0984"), 2:4] <- list("Metazoa","Arthropoda","Insecta")

# Some copepods too   
   to_export[which(to_export$species %in%c("Centropages abdominalis","Nitokra spinipes",
                                        "Stenhelia pubescens") ), 13] <- "copepod"

 to_export[which(to_export$class == "Polychaeta"), "group_fv"] <- "polychaete"

 to_export %>% 
    write_csv(here("Input", "hash.annotated.csv"))

```

```{r write.out}

to.export %>% 
  filter (is.na(group_fv)) %>% 
  select (-Hash) %>% 
  distinct() %>% 
  arrange(kingdom, phylum)

```

```{r summary stats}

database_step.5 %>%
  #filter(!is.na(benthic) | !is.na(Source) | !is.na(Source.Kasim) | !is.na(Source.csv)) %>% 
  unite(benthic, Source, Source.Kasim, Source.csv, sep = "|", col = all.benthos) %>% 
  mutate(agreed.benthos = case_when(str_detect(all.benthos, "PLK") ~ "PLK",
                                    str_detect(all.benthos, "BEN") ~ "BEN",
                                    str_detect(all.benthos, "Both")~ "Both" ,
                                    TRUE                           ~ "None")) %>% 
  group_by(agreed.benthos) %>% 
  summarise(nReads = sum(tot.reads),
            nHash = n()            )

```

### 2020-08-13 Scripts/Future_climate_conditions.R

@@ -6,17 +6,11 @@ library (vegan)
 library (tidybayes)
 library (here)
 
-now.env.data <- read_csv(here("Input","env.data.for.log.regression.csv")) %>%
-  filter(pH > 7.4) %>%
-  mutate(
-    Season = case_when(
-      str_detect(sample, "1703|1710|1711|1801|1803") ~ "Winter",
-      TRUE ~ "Summer"
-    ),
-    Area = case_when(
-      str_detect(sample, "CP|LK|FH") ~ "SJI",
-      TRUE ~ "Hood Canal"
-    ))
+now.env.data <- read_csv(here("Input","Combined_Biol_Env_Plankton.csv")) %>% 
+  select(event, Area, mean.TA, mean.DIC = mean.DIC.new.sal,Salinity = Salinity.new ,Temperature, Omega.aragonite, pH = pH_new ) %>% 
+  distinct() %>%  
+  filter(pH > 7.4) 
+
 
 #calculations based upon Khangaonkar, T., Nugraha, A., Xu, W., & Balaguru, K. (2019). Salish Sea response to global climate change, sea level rise, and future nutrient loads. Journal of Geophysical Research: Oceans, 124. https://doi.org/10.1029/ 2018JC014670
 #and in particular, upon Figure 14 in that paper. 
@@ -102,18 +96,18 @@ now.env.data <- read_csv(here("Input","env.data.for.log.regression.csv")) %>%
         pH.delta.mean.model <- lm(deltapH ~ Year, data = data.frame(
           deltapH = c(0, -0.18),
           Year = c(2000, 2095)
-        ))
+        )) 
         
-        
+        predict(pH.delta.mean.model, newdata = data.frame(Year = 2020:2080)) %>% fitdist(distr = "norm") -> increase.model.params
         return(data.frame(
           mean =  ph.normal.params$estimate[1] +  #observed
             predict(pH.delta.mean.model, newdata = data.frame(Year = year)), #predicted change
-          sd = ph.normal.params$estimate[2]   #observed
+          sd = ph.normal.params$estimate[2] + increase.model.params$estimate[2]  #observed
         ))
       }
     }
-    #get_pH_distribution_simple(2089, "SJI")  #example
-    
+    #get_pH_distribution_simple(2017, "SJI")  #example
+   
     
     
     
@@ -130,7 +124,7 @@ now.env.data <- read_csv(here("Input","env.data.for.log.regression.csv")) %>%
[…]

### 2020-08-13 Scripts/Models_for_Revisions.R

@@ -1,417 +1,0 @@
-library(tidyverse)
-library(stats)
-library(rstanarm)
-library(loo)
-library(bayesplot)
-library(tidybayes)
-library(modelr)
-library(here)
-library(styler)
-
-#uncomment below, if sourcing at command line (because this apparently doesn't respect here())
-#setwd("/Volumes/GoogleDrive/My Drive/Kelly_Lab/Projects/OA_eDNA/Analysis/Splitting.data.by.lifestyle")
-
-
-####LOAD DATA
-options(mc.cores = parallel::detectCores())
-data.for.stan <- read_csv("Input/data.for.stan.csv")
-now.env.data <- read_csv("Input/env.data.for.log.regression.csv") %>%
-  filter(pH > 7.4) %>%
-  mutate(
-    Season = case_when(
-      str_detect(sample, "1703|1710|1711|1801|1803") ~ "Winter",
-      TRUE ~ "Summer"
-    ),
-    Area = case_when(
-      str_detect(sample, "CP|LK|FH") ~ "SJI",
-      TRUE ~ "Hood Canal"
-    ))
-  
-######FUNCTIONS
-### DEFINE linear models to switch back and forth between standardized and natural-scale data
-value2std <- function(df){lm(std~value, data = df)}
-std2value <- function(df){lm(value~std, data = df)}
-
-ModelPlot <- function(df, variable) {
-  require(tidyverse)
-  temp.plot <-
-    df %>%
-    mutate(Posterior = list(apply(posterior_linpred(df$model[[1]], transform = T), 2, median))) %>%
-    select(-model) %>%
-    unnest() %>%
-    ggplot(aes(x = eval(as.symbol(variable)), y = binary, color = Area)) +
-    geom_point() +
-    geom_smooth(aes(y = Posterior)) +
-    facet_grid(Area ~ .) +
-    ggtitle(label = df$taxa[1]) +
-    xlab(variable) +
-    ylab("Probability of Occurrence")
-  
-  return(temp.plot)
-}
-#
-ProjectionPlot <- function(taxon, modelset = mod.proj.phylum) {
-  temp.plot <- modelset %>%
-    mutate(meanRichness = round(meanRichness, 1)) %>%
-    group_by(phylum, Area) %>%
-    mutate(ScaledMeanRichness = meanRichness / max(meanRichness)) %>%
-    filter(phylum == taxon) %>%
[…]

### 2020-08-13 Scripts/eDNAindex.R

@@ -1,0 +1,74 @@
+# Usage eDNAindex(x, sample, taxa, nReads) all elements without quotes
+
+
+eDNAindex<- function(df, Sample_column, OTU_column, Counts_column, Biological.replicate ){ # 
+  
+  require(dplyr)
+  require(rlang)
+  
+  
+  Sample_column <- rlang::enquo(Sample_column)
+  OTU_column    <- rlang::enquo(OTU_column)
+  Counts_column <- rlang::enquo(Counts_column)
+  Biological.replicate <- rlang::enquo(Biological.replicate)
+
+  
+  if (quo_name(Biological.replicate) %in% colnames(df)){
+    
+    print ("Averaging ratios between Biological replicates")
+    
+    df %>% 
+      
+      group_by(!!Sample_column, !!OTU_column, !!Biological.replicate) %>%
+      
+      summarise ( sumreads = sum(!!Counts_column))  %>%  # This sums technical replicates
+
+      group_by(!!Sample_column,!!Biological.replicate) %>%
+
+      mutate (Tot = sum(sumreads),
+              Row.prop = sumreads / Tot)  %>%                      # This creates the proporttion on each biological replicate
+
+      group_by(!!Sample_column) %>%
+
+      mutate (nreps = length(unique(!!Biological.replicate))) %>%
+
+      group_by(!!Sample_column, !!OTU_column) %>%
+
+      summarise (mean.prop = sum (Row.prop) / max(nreps))   %>%
+
+      group_by (!!OTU_column) %>%
+
+      mutate (Colmax = max (mean.prop),
+              Normalized.reads = mean.prop / Colmax) %>%
+
+      dplyr::select( -Colmax, -mean.prop) -> output
+    return(output)
+  }
+
+
+  print("Calculating eDNAindex directly")
+
+  # IF THERE ARE TECHNICAL REPLICATES, WE NEED TO SUM THOSE VALUES FIRST
+  #return(!!Technical_replicate)
+
+
+  df %>%
+
+    group_by(!!Sample_column, !!OTU_column) %>%
+
+    summarise (sumreads = sum(!!Counts_column)) %>% # In case the sample column is of a higher group than the nrows
+
+    group_by(!!Sample_column) %>%
+
+    mutate (Tot = sum(sumreads),
+            Row.prop = sumreads / Tot) %>%
+
+    group_by (!!OTU_column) %>%
+
+    mutate (Colmax = max (Row.prop),
[…]

### 2020-08-13 Scripts/furtherModeling.R

@@ -1,0 +1,222 @@
+library (lme4)
+library (here)
+library (tidyverse)
+library (tidybayes)
+library (rstanarm)
+library (rstan)
+
+data.for.stan <- read_csv(here("Input", "data.for.stan.csv"))
+rstan_options(auto_write = TRUE)
+options(mc.cores = parallel::detectCores())
+
+set.seed(108)
+taxVec <- unique(data.for.stan$taxa)
+subsetTax <- sample(taxVec, 20)
+
+stan.input <- data.for.stan %>% 
+  filter(taxa %in% subsetTax) %>% 
+  group_by(Area) %>% 
+  mutate(Temperature = vegan::decostand(Temperature, method = "standardize"),
+         pH = vegan::decostand(pH, method = "standardize"),
+         Salinity = vegan::decostand(Salinity, method = "standardize")) %>%
+  group_by(taxa) %>%
+  mutate(Noccur = sum(binary)) %>%
+  filter(Noccur > 7 & Noccur < 69) %>%  #must be present or absent in at least 10% of 76 samples in dataset
+  dplyr::select(-c(sample, phylum, DIC, Noccur)) %>% 
+  ungroup() %>% 
+  unite(Area, taxa, col = "AreaTaxa", remove = F) %>% 
+  mutate(sampleIndex = 1:nrow(.),
+         taxonIndex = match(taxa, unique(taxa)),
+         areaIndex = match(Area, unique(Area)),
+         seasonIndex = match(Season, unique(Season)),
+         taxonAreaIndex = match(AreaTaxa, unique(AreaTaxa))) %>%   #unique combinations of taxon and area
+  dplyr::select(-AreaTaxa) 
+
+
+##rstanarm
+      t_prior <- student_t(df = 7, location = 0, scale = 2.5)
+      
+      #model fitting one intercept per unique area/taxon combination, plus
+      #a different slope for the effect of Temperature and of pH on each taxon.
+      mod1.arm <- stan_glmer(binary ~ 0 + (0 + Temperature + pH | taxa) + (1 | taxonAreaIndex), 
+                             data = stan.input, 
+                             family = "binomial",
+                             prior = t_prior,
+                             prior_intercept = t_prior, 
+                             chains = 2,
+                             adapt_delta = 0.99)
+      
+      #model fitting one intercept per taxon, plus
+      #a different slope for the effect of Temperature and of pH and of area on each taxon.
[…]

### 2020-08-13 Scripts/sequential.join.r

@@ -1,0 +1,72 @@
+library(tidyverse)
+pkgconfig::set_config("dplyr::na_matches" = "never")
+sequential_join <- function(df1, df2,
+                            var.to.join.1,
+                            var.to.join.2="",
+                            var.to.join.3="",
+                            confidence,
+                            ... ){
+  interesting.vars <- rlang::enquos(...)
+  join.col <- rlang::ensym(var.to.join.1) # To use it with select
+  join.quo <- rlang::enquo(var.to.join.1) # To use it as right hand side of mutate
+  varname <- rlang::ensym(confidence)     # To use it as left hand side of mutate
+  
+  anti_join(df1, df2, by = var.to.join.1) -> left_over # what does not have a match
+  
+  
+  df2 %>% 
+    select(..., !!join.col) %>% # Keep only: key variable, interesting variables
+    inner_join(df1,., by = var.to.join.1) %>% 
+    mutate(!!varname := (!!join.quo)) -> step.1 # The confidence level of the join 
+  
+  ### Now we do the second step join.   
+  
+  if(var.to.join.2 !=""){                      # So it only works if there are more than 1 key variables
+    join.col <- rlang::ensym(var.to.join.2)
+    
+    join.quo <- rlang::enquo(var.to.join.2)
+    
+    anti_join(left_over, df2, by = var.to.join.2) -> left_over2 # Whatever doesn't have a match on either 1 or 2 
+    
+    
+    df2 %>% 
+      select(..., !!join.col) %>% 
+      
+      distinct()  %>% # To stop duplication of column information - maybe change it later on to a group_by - mutate - case_when - both
+      add_count(!!join.col) %>% 
+      mutate_at(.vars = vars(...), .funs = function(x) ifelse (.$n == 1, x, "undetermined")) %>% 
+      distinct() %>% 
+      inner_join(left_over, .) %>%
+      mutate(!!varname := (!!join.quo)) -> step.2
+    
+    
+    left_over <- left_over2
+    step.1 <- bind_rows(step.1, step.2)
+    
+    
+  } 
+  #}
+  if(var.to.join.3!=""){  
+    
+    join.col <- rlang::ensym(var.to.join.3)
+    
+    join.quo <- rlang::enquo(var.to.join.3)
+    
+    anti_join(left_over, df2, by = var.to.join.3) -> left_over2
+    
+    df2 %>% 
+      select(..., !!join.col) %>% 
+      
[…]

### 2020-08-13 Manuscript/Response.to.reviewers.tex
Added the manuscript folder

\documentclass[11pt]{article}

\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{gensymb}
\usepackage[square,numbers]{natbib}
\usepackage[legalpaper, margin=1in]{geometry}
\usepackage{lineno}
\usepackage{xcolor}

\begin{document}
\begin{linenumbers}

\textit{\tiny	This manuscript was compiled on \today }	
\section*{}

\textbf{ 
 Environmental DNA Metabarcoding Reveals Winners and Losers of Global Change in Coastal Waters}

\textbf {Response to reviewers}

\vspace*{1em}

We are thankful for the feedback, comments and suggestions Our manuscript RSPB-2019-2901  "Environmental DNA Metabarcoding Reveals Winners and Losers of Global Change in Coastal Waters" has received. Find below our response to their suggestions, and how we implement them in our manuscript.

\vspace*{1em}
\textbf{\large{ Reviewer 1}}:

\begin{itemize}

\item{1.      The manuscript needs to be restructured so that it follows a more conventional Intro, Methods, Results, Discussion format. In its current format, the results are difficult to follow as some of the methods that are crucial in interpreting the data are not detailed until the end of the MS. Please revise so that Methods come before Results. }

\textbf{Response:}

We have now reorganized the manuscript sections following a standard structure. As a result, a few sentences in the results section that introduced the procedures follow to model and project the likelihood of taxon presence have now been merged  into the Methods section:

line 107: Within a geographic area and season, temperature and pH were most strongly associated with community-level changes. We therefore developed a hierarchical logistical regression model relating the presence of each taxon to temperature and pH (as fixed effects) {\color{red} -salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}.}

\item{2.      This is certainly not the first eDNA metabarcoding paper to track shifts in community composition across environmental gradients, yet there are almost no references to other studies that have used this approach, nor how eDNA metabarcoding is being applied more broadly as a monitoring tool. Pleas incorporate this literature into the Introduction. }

\textbf{Response:}

We now include in the Introduction references to previous efforts to use eDNA metabarcoding for community composition changes assessment linked to environmental gradients, and biomonitoring of marine fauna in general

Line 78: {\color{red}A growing body of evidence backs up the efficacy of eDNA metabarcoding for monitoring biodiversity (see a review in \cite{ruppert2019past}), and this approach has been successfully used to detect community composition variation across environmental changes in aquatic  \cite{emilson2017dna}, estuarine \cite{chariton2015metabarcoding, lallias2015environmental}, and marine ecosystems \cite{berry2019marine}}

\item{3.      The manuscript suffers from a lack of transparent results pertaining to the sequencing and bioinformatic methods. How many samples failed to pass QC? How many reads per sample were the analyses based on? How many samples did you collect from each site? I think this manuscript will benefit from including a table in the main text that summarizes some of these data.}

\textbf{Response:}

It was the intent of the authors to have raw data and scripts available for the reviewers and the readership when published. We have now created an open github repository with all the code and the raw fastq files are in a google drive link.

\item{4.      The title needs to be more specific as this study only focusses on plankton communities. Please consider including Plankton in the title to give the readers a better sense of what the paper is actually about.}

\textbf{Response:}

The reviewer is right in pointing out that the dataset used here is limited to  planktonic taxa. We have reworded our title, which now reads {\color{red} Environmental DNA Metabarcoding Reveals Winners and Losers of Global Change in Coastal Plankton}

\item{5.      The figures do a good job at displaying the study site and data, however they are very busy and figure legends/axes are very difficult to read in print. Please consider revising figures to make them more legible and digestible. For example, perhaps Figure 3 only needs to include taxa that show significant shifts? }

\textbf{Response:}

We agree with the reviewer and we have rethought and redesigned figure 3 to highlight the taxa that shows meaningful shifts - focusing on the most varying diatoms and dinoflagellates we reveal the magnitude and repercussion of the changes in suitability. Panel C has been moved to the supplementary material

\end{itemize}

\vspace*{1em}
\textbf{\large{ Reviewer 2}}:

\begin{itemize}
\item{Figures: the colors red and blue are used both to distinguish time periods (2019/2095) and locality (San Juan/Hood Canal) in different figures. Figure 2 legend doesn?t say which color is which (presumably they are the same as in Figure 1). Use consistent colors for categories across figures and use different colors for time period and locality, if possible.}

\textbf{Response:}
    We realize that the shared color scales can lead to confusion. This has now been fixed, and we employ now a different color scales for temporal and spatial variables.

\item{Figure 2 is presented as for Hood Canal in summer but differs from the image in the supplemental information for that same place and time period. The points and axes in the CAP plot are the same, but the plots in sections B and C differ. What?s the difference between Fig 2 and the (ostensibly same) figure in the supplement? If these differ, how do we interpret the others in the supplement? The legend doesn?t explain the colors in Fig 2C.}

\textbf{Response:}

Figure 2 should be exactly the same as the first part of the Supplemental figure - A thorough review of the code that generates the supplemental figures revealed a small difference in the input dataset. We thank the reviewer for noticing this, and now the code that generates each of the figures is consistent.

\item{As Figure 3 is divided into sections A, B, and C, these ought to be referred to in the text (see below). The text is very small and hard to read in the review version I have.}

\textbf{Response:}

In an effort to declutter Figure 3, and following suggestions from reviewer 1, we have focused the figure in those species showing a more dramatic change in suitability. The new version of the figure has two panels (A and B) and we refer to them specifically in the text

\item{Lines 74-76 ?with different points along the environmental gradient simultaneously showing differences equivalent to those predicted between present-day and future oceans?: This seems like an important point that perhaps should be included in the Introduction.}

\textbf{Response:}

Agreed. The advantage of the Hood Canal in this regard is paramount, and we now have included this concept in the introduction.  Line 52 onwards: {\color{red} [in close geographic proximity]. This gradients are pronounced and show values of pH and temperature expected worldwide by the end of the century}.

\item{Line 96: Specify Figure 2A}

\textbf{Response:}

TODO

\item{Line 99: The value of predicting which cluster (?community?) a given set of environmental parameters will fall within is not entirely clear to me. Is there a consistent set of taxa that make of up these clusters across time and space? Can specific groups of taxa be deduced from the ?community? number? Do the environmental variables accurately predict which set of taxa will be present? This section needs a bit more clarity.}

\textbf{Response:}

We think the relationship between clusters and the environmental variables are important in a few aspects. First, as a means of ground truthing our findings, it has great value as the clusters were detected blind to the environmental data, and the CAP analysis return the distinctiveness of these clusters in that space - Figure 2B highlights the translation of those modified axis of the CAP plot into the real values of temperature and pH. Second, these clusters (maybe communities is a more committed term that requires a longer dataset in which to confirm co-ocurrence of members) are a combination of species that are unique to them plus changes in relative presence - not a clear cut between total abence of a species but definitely a gradient in the relative importance of the species present.
We have reworded the section

TODO

\item{Line 102-104: According to my interpretation of Fig. 2C, it seems like the colder water communities actually had *lower* values of the given taxa, as the differences are all negative relative to other communities. It may help to specify in the text which characteristics are specific to which communities (communities 2 and 3 both span the lower temperature area of Fig. 2B).}

\item{Line 112-117: The authors base much of their calculation on a set of simulated temperature and pH values for different regions in the Salish Sea in 2095. The underlying data was not available so it?s not clear to me exactly how they subdivide the regional variation in the model predictions from Khangaonkar et al. (2019) (see comments on the Methods section as well).}

\item{Line 118-120: Please use figure subsections (A-C) when referring to Figure 3. Line 188 should reference Fig. 3A, line 119 should refernce Fig. 3C, and line 120 should reference Fig. 3B (the text erroneously refers to middle and bottom panels, when it ought to be bottom and middle panels, respectively.}

\item{Line 121: Refer to the relevant section of Figure 3 (A, B, or C)}

\item{Line 121-133: When referring to specific taxa as winners or losers, I was expecting to see them in Figure 3B, but not all of them were there. Or if it did appear, it was only at one site, but it was mentioned as being an overall winner or loser.}

\item{Line 122-123: According to Figure 3B, Thalassionema, Navicula, and Nitzschia appear to refer specifically to San Juan. If so, please state this.}

\item{Line 123: Coscinodiscus and Ditylum seem to refer to Hood Canal. This ought to specified as before.}

\item{Line 129: Write out Emiliania. The ?sp.? after Alexandrium should not be italicized. This sentence might read better if written: ?The coccolithophore Emilinania huxleyi and the dinoflagellate Alexandrium sp. both find more suitable habitat? (rather than ?each finds?)}

\item{Line 136-137: This is minor, but this section (or methods) might benefit from a brief mention of why ANOVA was used in one case and Wilcoxon in another.}

\item{Line 138-141: It seems like this paragraph could be moved to the discussion section.}

\item{Line 147-148: Need a citation for Coscinodiscus as key fish food.}

\item{Line 178: Spell out harmful algal bloom the first time it?s used.}

\item{Line 192: Might be helpful to specifically mention carbon sequestration as an important aspect of sinking rates.}

\item{Line 204-205: Citation needed for fish mortality. Eutrophication seems like a cause of blooms, not a result of them. What are the potential cascade effects on benthic-pelagic coupling? The term here on its own isn?t really saying anything.}

\item{Line 208-214: This is a big aspect of why this study is valuable! I think it?s worth discussing this in the introduction.}

\item{Line 237: Should be Scripps Institution of Oceanography}

\item{Line 241: Should be Figure 1C}

\item{Line 267-268: Which index of abundance from Kelly et al. (2019) was used? That paper presents several and though it declares one to be more useful, I think it would be helpful in this manuscript to state which one was used.}

\item{Line 273-282: It?s hard to discern based on the limited code and data provided exactly how the authors subsampled the different subregions of Puget Sound, since the paper they base their data on has overall values for the whole area. I think this section could use a bit of clarification, and all the code and raw data should be given (the provided github link is broken).}

\end{itemize}
 \end{linenumbers}
\end{document}

### 2020-08-13 Manuscript/What.lies.ahead.procb_20191205.tex
Added the manuscript folder

\documentclass[11pt]{article}

\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{gensymb}
\usepackage[square,numbers]{natbib}
\usepackage[legalpaper, margin=1in]{geometry}
\usepackage{lineno}

\begin{document}
\begin{linenumbers}

\textit{\tiny	This manuscript was compiled on \today }	
\section{Title}

\textbf{ 
 Environmental DNA Metabarcoding Reveals Winners and Losers of Global Change in Coastal Waters }
\vspace*{1em}

\textbf{Ram\'{o}n Gallego \textsuperscript{$a,b,1$}, Emily Jacobs-Palmer \textsuperscript{$b$}, Kelly Cribari \textsuperscript{$b$}, Ryan P. Kelly}\textsuperscript{$b$}
\vspace{1em}

\textbf{$^a$} {\small NRC Research Associateship Program,
Northwest Fisheries Science Center,
National Marine Fisheries Service,
National Oceanic and Atmospheric Administration. 2725 Montlake Boulevard East, Seattle, WA 98112}

\textbf{$^b$} {\small School of Marine and Environmental Affairs. University of Washington. 3707 Brooklyn Ave NE Seattle, WA 98105}

{\textsuperscript{1}To whom correspondence should be addressed. E-mail: rgallego@uw.edu}

\vspace{1em}

\textbf{Author contributions}: RG led sampling, laboratory work, data analysis, and writing. RPK conceived of the project and assisted in sampling, analysis, and writing, and received funding for the work. EJP and KC assisted in sampling, laboratory work, and analysis.

\vspace{1em}

{\small The authors declare no conflict of interest.}

%%%%%%%
%%%%%%%

\section{Abstract}

Studies of the ecological effects of global change often focus on one or few species at a time. Consequently, we know relatively little about the changes underway at real-world scales of biological communities, which typically have hundreds or thousands of interacting species. Here, we use monthly samples of environmental DNA to survey over 222 planktonic taxa along a gradient of temperature, salinity, dissolved oxygen, and carbonate chemistry in nearshore marine habitat. The result is a high-resolution picture of changes in ecological communities using a technique replicable across a wide variety of ecosystems. We estimate community-level differences associated with time, space and environmental variables, and use these results to forecast near-term community changes due to warming and ocean acidification. We find distinct communities in warmer and more acidified conditions in each of the areas and seasons sampled, and a strong reduction in richness of current diatom assemblages. Taxa finding greater suitable habitat in near-future waters are more taxonomically varied and include the ubiquitous coccolithophore \textit{Emiliana huxleyi} and the harmful \textit{Alexandrium sp.} These results suggest foundational changes for nearshore food webs under near-future conditions.

%%%%%%%
%%%%%%%

\section{Keywords}
Ocean Acidification $|$ environmental DNA $|$ metabarcoding $|$ ecosystem response $|$ climate change
%%%%%%%
%%%%%%%

\section{Main Text}

%%%%%%%%% Background %%%%%%%%%%%%%%%%%

\subsection{Background}

As ocean acidification and warming continue apace, changes in the marine environment are having an effect on many species' metabolism, development, growth and reproduction success \cite{kroeker2010meta, espinel2018effects, bednarsek2019systematic, clements2018eating}, very likely altering food webs \cite{Riebesell:2018aa, Busch2013Potential, heiden2019impact} and species' interactions in ways that are poorly understood. Laboratory or mesocosm-based manipulation experiments have documented a wide variety of biological responses under projected climate scenarios of \textit{p}CO$_2$, pH, solar radiation, salinity and temperature \cite{Gao2012,delorme2014temperature, kurihara2008effects}, showing an array of species-specific responses among particular taxa of interest. However, information regarding multi-species or community-wide responses to these stressors is far more limited \cite{kroeker2011divergent, jokiel2008ocean}. The scarcity of such data is likely attributable to the difficulty of simultaneously tracking the responses of many species in the field, and to the difficulty of identifying natural systems that adequately reflect the environmental gradients under study.

Two natural CO$_2$ seeps in nearshore marine habitats -- one in Italy and one in Papua New Guinea -- have demonstrated shifts in benthic communities associated with especially acute acidification in the present day, previewing those we might expect at a more global scale under future conditions \cite{kroeker2011divergent, fabricius2015situ}. But beyond these exceptional sites, it is difficult to measure changes in ecological communities associated with the relatively subtle shifts in nearshore ocean chemistry observed to date, particularly in light of naturally large spatial and temporal variation in these communities. The Puget Sound in Washington, USA, offers a gradient of carbonate chemistry parameters and other environmental conditions in close geographic proximity. Complex bathymetry, water circulation patterns, and nearshore landforms create intertidal sites exposed to large variations in temperature, \textit{p}CO$_2$, pH, and related parameters \cite{khangaonkar2019salish}, creating an opportunity to test the effect of these measures on marine communities under conditions expected worldwide in the near future \cite{murray2015inland}, and time-series sampling across the spatial gradient lets us control for site- and season-specific effects. This study system therefore provides a powerful means of modeling community-level responses to changing environmental conditions.

Even given the appropriate environmental gradients, tracking the biological responses of many taxa simultaneously remains challenging. Environmental DNA (eDNA) metabarcoding \cite{holman2019detection, kelly2016genetic} addresses this problem by amplifying a common gene region out of DNA present in a water sample; the technique can detect hundreds to thousands of taxa per sample, potentially with species-level identification.

% Paragraph 4: Summary of intentions

Here we use a temporal and spatial series of metabarcoding samples to track changes in nearshore ecological communities associated with differences in pH, water temperature, and other environmental variables. We use broad-spectrum PCR primers \cite{leray_new_2013} to target eukaryotes specifically, identifying the likely effects of future climate scenarios on suites of planktonic taxa.

\section{Methods}

\subsection*{Sampling}
We collected water samples to assess eDNA communities in two regions of the Salish Sea (Washington, USA): San Juan Island and the Hood Canal. These sites experience substantial variation in water chemistry and other environmental conditions despite geographic proximity (ca. 300km; Fig. 1). We sampled eight sites monthly for approximately 1.5 years (March 2017 to August 2018), taking three 1L samples (biological replicates; ca. 10m apart) each month at each site (261 bottle samples total). Each sample was filtered through a 0.45 $\mu$m  cellulose filter, and the filter preserved in Longmire buffer until DNA extraction \cite{renshaw2015room}. Concurrently, we collected one 120 ml water sample from each site and poisoned it with 0.1 ml of saturated HgCl$_2$ for carbonate chemistry analysis, following \cite{riebesell2011guide}. We also collected \textit{in situ} measurements of temperature, salinity and dissolved oxygen using a handheld multiprobe (Hanna Instruments, USA) and a portable refractometer. We note that many unmeasured variables influence planktonic communities (e.g., nutrients, sunlight, wave energy), but that our set of measured parameters clearly distinguished communities and was adequate for our purposes.

We characterized sample carbonate chemistry by measuring Total Alkalinity (TA; open-cell automated titration based on a 876 Dosimat plus (Metrohm AG) as part of a custom system assembled by Andrew Dickson (UCSD) and used in the laboratory of Alex Gagnon at UW) and Dissolved Inorganic Carbon (DIC; Apollo Instruments, USA; CO$_2$ extraction system with 10\% (v/v) phosphoric acid). Both measurements were calibrated and validated with certified reference material from the Scripps Oceanographic Institute. Using DIC and TA, we calculated pH and the remaining carbonate system parameters using the R package `seacarb` \cite{seacarb}.

Our sampled areas differed in the environmental variables driving changes in carbonate chemistry. San Juan Island was less seasonally variable than the Hood Canal in every measured parameter (Figure 1B); the island is more directly affected by summer coastal upwelling as a function of bathymetry and circulation patterns \cite{murray2015inland}, and this appears to be the dominant influence on carbonate chemistry there. By contrast, photosynthesis and respiration likely drive much of the carbonate chemistry variation in the Hood Canal (See Supporting Information).

\subsection*{eDNA sequencing and bioinformatic processes}
We purified DNA from each filtered sample using a Phenol Chloroform Isoamyl Alcohol protocol, following \cite{renshaw2015room}. A 1/10 to 1/100 dilution of the extract was used as template for a PCR reaction targeting a 313bp fragment of cytochrome oxidase I \cite{leray_new_2013}. PCR reactions were performed in triplicate and sequenced individually to quantify the stochasticity of PCR reactions on a mixed template sample, and we attached secondary indexing tags using a two-step PCR process \cite{odonnellPrimers}. PCR conditions and protocols for sample identification followed \cite{kelly2018tides}, and batches of 49 to 178 multiplexed samples were sequenced using MiSeq v2-500 or v3-600 sequencing kits using manufacturer protocols. On each sequencing run, we added triplicate samples consisting on DNA obtained from species not present in the marine environment under study (Red Kangaroo (\textit{Macropus rufus}) and Ostrich (\textit{Struthio camelus})) to establish quality controls of sample assignment and quantify levels of 'tag-jumping' or sample-cross-talk \cite{schnell2015tag}.

Code for all quality-screening and bioinformatics is available in the Supporting Information. Briefly, we used a Unix script that calls secondary programs for primer-trimming and preliminary quality-control \cite{martin2011cutadapt, Callahan:2016aa} we estimated the likely composition of each sample using DADA2. This approach avoids clustering, such that we retained all of the amplicon sequence variants (ASVs, \textit{i.e.}, unique sequences); we subsequently carried out secondary quality-control and decontamination following \cite{kelly2018tides}. We then assigned sequences to known taxa using phylogenetic tree placement with insect v1.1 \cite{insect}. We conservatively kept only taxa annotated at the level of taxonomic family, genus, or species, so we could reliably infer taxon natural history. Using published literature and online databases, we placed every recovered taxon into a benthic/planktonic category and focused our analysis on the planktonic community (see Supporting Information).

By treating amplification efficiency as consistent within a given taxon, we created an index of abundance for each taxon across space and time \cite{kelly2019understanding}, using pooled data from technical replicates and mean proportions across biological replicates. The multivariate tests (PERMANOVA, CAP, k-means clustering) used these index values as the species abundances and Bray-Curtis as the dissimilarity measurement between samples.

\subsection*{Year 2095 Environmental Scenario and Biological Responses}

We estimated the distribution of environmental parameters for the overall Salish Sea in 2095 from the results of \cite{khangaonkar2019salish}, which estimated an annual mean increase in temperature of 1.51 \textdegree C and mean pH decrease of 0.18 for the Salish Sea as a whole. We then subsampled this distribution to reflect the characteristic environments of different areas (San Juan Island and Hood Canal) and seasons (Summer and Winter, with April - September coded as ``Summer''). To do this, we fit beta parameters to describe the fraction of the overall observed data (2017 - 2018) reflected in each subset; we then subsampled the 2095 distribution accordingly, under the assumption that the Hood Canal in Summer (for example) would occupy the same fraction of the overall annual Salish Sea temperature and pH distributions in 2095 as it did in 2017. See Supporting Information for details of sampling parameters.

For community-level projections, we coded community-cluster identity (Figure 2) as an unordered response variable in a multinomial logistic model, with temperature, pH, area, and season as predictor variables. We calculated the probability of each community, given these predictors, using the package `nnet` \cite{nnet2002}. Most-common combinations of T and pH were estimated from 2-dimensional density plots for each year.

Within a geographic area and season, temperature and pH were most strongly associated with community-level changes. We therefore developed a hierarchical logistical regression model relating the presence of each taxon to temperature and pH (as fixed effects), allowing slope to vary by geographic area and intercept to vary by season, to capture large-scale seasonal differences associated with sunlight, photosynthesis, and related factors. For each taxon, we fit these models using the Bayesian generalized linear mixed effects functions in R package rstanarm \cite{rstanarm} for R. See Supporting Information for analytical code.

Given the estimated sea-surface temperatures and pH values for 2095 and observed data from 2017, we then used taxon-specific logistic regression models to estimate the suitability of habitat for each taxon in the future scenario. For each point in the pH - temperature grid, we calculated species richness as the mean result of 100 independent draws of the posterior distributions of all modeled taxa, grouping the individual taxa models by phylum.

We can only model responses of taxa present in our data set. That is, we may predict that the number of (for example) diatom species present will decline relative to those present today, but our data do not allow us to predict whether new species will immigrate from elsewhere or how species might evolutionarily adapt to future conditions. It is beyond the scope of our work to account for the latter, and furthermore, because of the extreme uncertainty of evolutionary responses, the predictions of species distribution models are often interpreted without considering adaptation or phenotypic plasticity \cite{moore2015present}.

%%%%% RESULTS %%%%%

\subsection*{Results}
\underline{Variation in Carbonate Chemistry and in Ecological Communities}

%%%%% Figure 1: map and environmental conditions
\begin{figure}%[tbhp]
\centering
\includegraphics[width=1 \linewidth]{Figures_for_MS/Fig1.png}
\caption{\footnotesize {\textbf{A} Sampling locations in intertidal areas of the Hood Canal (Red) and San Juan Island (Blue). \textbf{B} Planktonic richness (unique taxa) per sampling locality across seasons, as reflected by the eDNA COI assay. \textbf{C} The observed environmental profiles of these two regions reflect a broad range of environmental conditions, with the Hood Canal resembling future conditions in temperate areas worldwide. Shown: Dissolved Inorganic Carbon (DIC, $\mu$moles/kg); Total Alkalinity (TA); Temperature ($\degree$C) and Salinity (PSU). Individual sampling sites shown as facets.}}
\label{fig:Fig1}
\end{figure}
%%%%%

We collected intertidal environmental and genetic data from bottle samples each month between March 2017 and August 2018 (Figure 1A). Despite geographic proximity and similar overall species composition (127 of the 222 planktonic taxa were found in both regions and accounted for 98\% of the sequences), the areas under study -- San Juan Island and the Hood Canal -- varied widely in pH, temperature, and other environmental parameters (Figure 1C), with different points along the environmental gradient simultaneously showing differences equivalent to those predicted between present-day and future oceans \cite{bopp2013multiple}.

Metabarcoding analysis of eDNA samples generated more than 52.5M sequences from 745 samples. These samples represented biological and technical replicates from 86 unique sampling events. After bioinformatics quality-control (See Methods) the dataset included $\sim${}41M sequences, from 4782 unique amplicon sequence variants (ASVs). Of these, 1374 ASVs (22.8M reads) could be annotated to a taxonomic level of Family or lower. These $\sim$ 500 taxa from 43 phyla were split according to their natural history and habitat (benthic vs. planktonic; see Supporting Information). Because we expect planktonic taxa to vary with water mass \cite{kelly2018tides} and therefore with bottle-sampled carbonate chemistry, here we focus on only the planktonic taxa (N = 222). These taxa showed a seasonal richness gradient between study areas, consistent with documented biodiversity clines in the area (Fig. 1B) \cite{dethier_decadal_2008}.

Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to season (Summer vs. Winter; PERMANOVA F-model = 14.4; p $<$ 0.001) and geographic Area (Hood Canal vs. San Juan Island; F-model = 15.4; p $<$ 0.001). We therefore performed a constrained analysis of principal components (CAP) for samples within each combination of Season and Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig. 2 shows results for Hood Canal in summer as an example; full analysis in the Supporting Information). Independent of environmental parameters, we separately clustered samples by pairwise Bray-Curtis dissimilarities among metabarcoding communities (k-means; N = 3 or 4) to identify groups of samples that were similar to one another with respect to biological community. Each biological cluster (colored hulls, Fig. 2) occupied a unique area of environmental parameter space. Planktonic communities therefore varied predictably with water temperature, salinity, and pH, across a range of those parameters likely to be encountered in many near-term future-ocean scenarios \cite{murray2015inland}. 
 Multinomial logistic regression yielded predictions of the most-likely community for any combination of environmental parameters (Fig. 2B).

These communities were distinguished by changes in the relative abundances of many taxa: in the Hood Canal in summer for example (Fig. 2C), the colder water communities showed higher values of \textit{Centropages} sp., \textit{Nitzschia longissima}, \textit{Prorocentrum} sp., \textit{Chrysochromulina} sp., \textit{Navicula} sp. and \textit{Chaetoceros} sp. Warmer waters were dominated by \textit{Bathycoccus prasinos} and \textit{Cymbomonas} sp., while at lower pH values, marine communities showed \textit{Litchtheimia ramosa}, \textit{Micromonas pusilla} at higher relative abundances.

\underline{Climate envelopes and future distributions}

The low-pH planktonic communities (cluster 3, Fig. 2) occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls. For example, by 2095 we expect summer in the Hood Canal to be T $>$ 12 and pH $<$ 8.2 -- where cluster 3 is the most likely community -- to occur more than two-thirds of the time (68\%).

%Under a common environmental condition, for example, in Hood Canal in summer, we expect such a planktonic community to go from occurring less than 1\% of the time in 2017 (0.001\% at T $=$ 15$\degree$C and pH $=$ 8.4; conditions more extreme than this occur ca. 50\% of the time) to occurring 37.5\% of the time under an equivalently-common condition in 2095 (84\% at T $=$ 22$\degree$C and pH $=$ 8.05).

%%%%%Figure 2:  CAP and community composition

%%%%%

\begin{figure}%[tbhp]
\centering
\includegraphics[width=1\linewidth]{Figures_for_MS/Figure2.2019-11-06.png}
\caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal. \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows). \textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model. \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis). See Supporting Information for full analysis.}}
\label{fig:Fig2}
\end{figure}

 To explore the suitability of different environmental conditions for each taxon, we modeled the likelihood of taxon presence as a function of area and season (random effects) and temperature and pH (fixed effects).  Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}. Using predicted conditions for Puget Sound for the year 2095 \cite{khangaonkar2019salish}, we then projected the probability of presence of each taxon across these conditions over 100 draws of the posterior probabilities of all taxa. We used these draws to 1) show the change in the probability of presence of each taxon for each scenario (i.e. 2017 vs 2095, Fig.3 top panel); 2) calculate the  taxon richness within taxonomic groups across the pH-temperature continuum (Fig. 3 center panel),% 2) show the consequent variation in richness for higher taxa across the pH - temperature continuum (Figure 3B),
  and 3) visualized the change in suitability for each taxon in our dataset individually (selected taxa in Fig. 3 bottom panel).

Diatoms show the steepest richness decline under future conditions (Figure 3); the probability of occurrence decreases markedly for diverse diatom taxa including both pennate (e.g., \textit{Thalassionema}, \textit{Navicula}, \textit{Nitzschia}) and centric (e.g., \textit{Coscinodiscus}, \textit{Ditylum}) body forms. These declines in diatom richness were more accentuated at lower pH values, both in the warm and colder seasons. Other taxa likely to find less-suitable habitat in the future include \textit{Karlodinium} (dinoflagellate; San Juan Island), and \textit{Phaeocystis} and red-tide-associated \textit{Chattonella} (haptophyte and raphidophyte, respectively; Hood Canal).

Likely winners under future conditions are more widely scattered among higher taxonomic groups. The coccolithophore \textit{E. huxleyi} and the dinoflagellate \textit{Alexandrium sp.} each finds more suitable habitat in both of our study areas. Among others, \textit{Chaetoceros} (diatom) and two small hydrozoans (Cnidaria) likely increase in San Juan Island, and the potentially fish-killing heterokont flagellate \textit{Pseudochattonella} increases in the warmer Hood Canal. See Supporting Information for a complete list of taxon-specific projections.

Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at richness variation within phyla; we find no change in median richness for the year 2095 relative to the present (ANOVA; taxon richness within phyla by year, p $=$ 0.3). Diatoms, however, show consistent and significant declines in richness (Wilcoxon p $< 10^{-15}$).

Taxa surveyed are a function of our metabarcoding PCR primers \cite{leray_new_2013} and reflect the current status of genetic databases, rather than a complete sampling of the planktonic community; we view these results as a cross-section of common taxa useful for understanding the biological effects of ocean conditions.

%%%%%
 \begin{figure*}
 \centering
 \includegraphics[width=1\linewidth]{Figures_for_MS/Figure3.20191105.png}
\caption{\footnotesize{Forecasted changes in plankton in the Salish Sea. \textbf{A} Probability densities for the occurrence of individual taxa (species- and genus-level) for 2017 (blue) and 2095 (red); data are mean probabilities over 100 model draws, and variance in probability is due to differences in underlying environmental conditions. Four prominent higher-taxa are shown: Arthropods, Dinoflagellates (Dinophyceae), Diatoms (Bacillariophyta), and Haptophytes.  \textbf{B} Individual taxa with the largest changes in probabilities of occurrence between 2017 and 2095, color-coded by higher taxonomic group, for Hood Canal (left) and San Juan Island (right). Median probabilities plotted. \textbf{C} Relative taxon richness (raster color, warmer colors are more taxon-rich) for each of these same higher-taxa, for plausible ranges of pH and Temperature. Envelopes of observed (2017, blue) and modeled (2095, red) annual conditions in the Salish Sea shown for reference. Hood Canal and San Juan Island plotted separately to illustrate environmental differences between them.}}

\label{fig:Fig3}
 \end{figure*}
%%%%%

The bulk of our projected community changes result from now-rare conditions occurring more frequently in the future. For example, in summer in the Hood Canal at present, we expect surface waters to have pH $<$ 7.9 and T $>$ 19$\degree$C only 1\% of the time. In 2095, we expect these conditions 18 times more frequently (i.e., 18\% of the time). At these values of T and pH, our model predicts the harmful \textit{Alexandrium} sp. to occur more often than not (mean frequency of occurrence $=$ 0.7). By contrast, the large centric diatom \textit{Coscinodiscus} -- a key source of carbon for zooplankton and small fishes -- occurs only one-third of the time under these same conditions (mean frequency $=$ 0.34).

\subsection*{Discussion}

Temperate surface oceans worldwide average approximately 14$\degree$C and pH of 8.1 \cite{bopp2013multiple}, and will change substantially in this century \citep[mean $\Delta$ T 2.5$\degree$C, $\Delta$ pH -0.35 globally; RCP 8.5;][]{gattuso2015contrasting}.  Here we document communities exposed to this same range of projected conditions in the present day, along an environmental gradient only ca. 200km wide, allowing us to project future ocean communities from a robust set of underlying observations. Our results reflect patterns in a diverse selection of species from nearshore marine communities in the Salish Sea, consisting of 222 planktonic taxa obtained from the metabarcoding analysis of 227 discrete samples across 77 space-time points (eight sites, 1.5 years). We find that changes in the composition of biological communities closely mirrored the variation in pH and temperature, with clear winners (e.g., \textit{Emiliana huxleyi}, \textit{Alexandrium}, and others) and losers (many, but not all, diatoms) likely to shift the structure and function of future marine communities.

A vast amount of evidence suggests climate-associated effects on marine species, and broad patterns of sensitivity are discernible within major taxonomic groups  \citep[][among many others]{gattuso2015contrasting, stock2017reconciling}. However, because the strength and direction of these effects are variable and species-specific \cite{kroeker2013impacts}, very little is known about community-level impacts. Our work illustrates the nearshore planktonic communities that can thrive in low pH - high temperature conditions; such communities are therefore likely to become more prevalent under future conditions.

The large number of species and broad set of environmental conditions we sampled yield substantial inferential power despite lacking the the degree of experimental control present in a laboratory or mesocosm.

Among the taxa surveyed, diatoms are of particular interest for their ubiquity in the world's oceans and their important roles in marine food webs \cite{armbrust2009life, wasmund2017diatom}, as well as in ecological and evolutionary theory \cite{margalef11978}. Our model suggests that diatoms will decrease in richness between the present and 2095, particularly in the Hood Canal summer months, when extreme temperatures are more likely. Although the most prevalent response among diatoms is a decrease in suitability, some substantial variability in responses exists within the group. For example, the centric diatom \textit{Coscinodiscus} spp., which is a food source for \textit{Acartia} spp. copepods \cite{jansen2008copepods} and many other animal species, will see future suitable habitat only in colder waters such as those in San Juan Island, while \textit{Skeletonema} spp. and the HAB-forming species \textit{Pseudo-nitzschia} spp. will see their habitat suitability remain constant or slightly increased, especially at low pH levels.

More strikingly, we see a dramatic increase in suitable environment for the HAB-forming dinoflagellate \textit{Alexandrium} sp., which can substantially harm local ecosystems \cite{colin2002latitudinal} and economies \cite{anderson2000estimated}. This increase is particularly high in the summer months of the Hood Canal, when pH is low and temperatures are are high. Both archaeological and experimental evidence suggest \textit{Alexandrium} sp. blooms with warmer temperatures \cite{mudie2002palynological}, and models \cite{moore2015present} also predict an increase in bloom-favorable conditions for \textit{Alexandrium} sp. in future oceans.

 Our results therefore suggest a possible change in relative dominance between diatoms and other phytoplankton species such as dinoflagellates, consistent with those seen at ecological regime shifts found elsewhere \cite{wasmund2017diatom, hinder2012changes}. Such a shift could affect ecosystems in many ways; even under the assumption that the surviving taxa would maintain the primary production levels, for example, the smaller cell-size of dinoflagellates and the differential sinking rates of the two groups would likely alter regional patterns of nutrient cycling\cite{xiao2018warming, anderson2002harmful, bienfang1984sinking}. Although the north Atlantic has shown an increase in diatom abundance\cite{hinder2012changes}, the increase in wind stress and associated mixing in the water column in the open ocean is unlikely to occur in the Hood Canal, where stratification is the strongest in the Salish Sea \cite{moore2008descriptive}. Furthermore, locally focused models support an increase in dinoflagellate dominance with climate change, particularly during summer months \cite{khangaonkar2019salish}.

 Our model also suggests increased environmental suitability for the coccolithophore \textit{Emiliania huxleyi}. There is evidence supporting increased calcification and respiration rates with higher pCO$_2$ levels \cite{iglesias2008phytoplankton} for this ubiquitous species, although the many strains of this species and its adaptive capacity make it difficult to predict longer-term effects with confidence \cite{bolton2016decrease}.

 Changes in environmental conditions and associated shifts in planktonic communities will likely reshape ecosystems and food webs, although some environmental processes may be conserved even as the particular taxa change. A switch from a diatom-dominated ecosystem to one in which dinoflagellate blooms extend in space and time could provoke cascade effects \cite{xiao2018warming} including fish mortality, eutrophication \cite{anderson2002harmful}, and bentho-pelagic coupling \citep[][]{bienfang1984sinking}. Beyond the phylum-specific patterns, the increase in suitable habitat for HAB-forming species will alone be an engine for ecosystem change \cite{trainer2019pelagic, wells2018harmful}.

One general challenge for model-based work is a tendency to extrapolate from observed conditions in ways that are often untestable -- by necessity, projections frequently operate outside the range of parameters on which the model was trained \cite{fitzpatrick2009projection}. Our study system lets us avoid this pitfall, in that our observed conditions encompass much of the environmental range predicted for future temperate oceans. That is, the changes we predict for the year 2095 do not primarily come from extreme values of pH and temperature, but rather from presently-rare conditions becoming more common.

Our observations are strong evidence of the kinds of changes likely in future marine communities, and they offer testable predictions about the magnitude and direction of effects on focal species.

%%% Material y metodos

\section{Acknowledgments}

 We are grateful to the UW Center for Environmental Genomics and Linda Park's lab at NOAA Fisheries for access to lab equipment and associated expertise and support. Thanks to Micah Horwith and the Washington State Department of Natural Resources for in-field support, to Terrie Klinger for feedback on an earlier draft, to Stephanie Moore for advice on phytoplankton generally, to Ana Ram\'{o}n-Laca for lab help, to Dan Drinan for field help, and to Alex Gagnon for access to carbonate-chemistry expertise and equipment.
 % Display the acknowledgments section

% Bibliography
\bibliography{refs_for_ms}
\bibliographystyle{abbrvnat}
 \end{linenumbers}
\end{document}

### 2020-08-13 Manuscript/What.lies.ahead.reviewed.tex
Added the manuscript folder

--- 2020-08-13 Manuscript/What.lies.ahead.procb_20191205.tex
@@
 \usepackage{lineno}
+\usepackage{xcolor}
 
@@
 We estimate community-level differences associated with time, space and environmental variables, and use these results to forecast near-term community changes due to warming and ocean acidification.
-We find distinct communities in warmer and more acidified conditions in each of the areas and seasons sampled, and a strong reduction in richness of current diatom assemblages.
-Taxa finding greater suitable habitat in near-future waters are more taxonomically varied and include the ubiquitous coccolithophore \textit{Emiliana huxleyi} and the harmful \textit{Alexandrium sp.} These results suggest foundational changes for nearshore food webs under near-future conditions.
+We find distinct communities in warmer and more acidified conditions in both areas sampled, and a strong reduction in richness of current diatom assemblages.
+Taxa finding greater suitable habitat in near-future waters are more taxonomically varied and include the ubiquitous coccolithophore \textit{Emiliania huxleyi} and the harmful \textit{Alexandrium sp.} These results suggest foundational changes for nearshore food webs under near-future conditions.
 
+%% Emiliania huxleyi 
 %%%%%%%
@@
 Environmental DNA (eDNA) metabarcoding \cite{holman2019detection, kelly2016genetic} addresses this problem by amplifying a common gene region out of DNA present in a water sample; the technique can detect hundreds to thousands of taxa per sample, potentially with species-level identification.
+% Adding stuff on eDNA metabarcoding for community change.
+
+{\color{red}A growing body of evidence backs up the efficacy of eDNA metabarcoding for monitoring biodiversity (see a review in \cite{ruppert2019past}), and this approach has been successfully used to detect community composition variation across environmental changes in aquatic  \cite{emilson2017dna}, estuarine \cite{chariton2015metabarcoding, lallias2015environmental}, and marine ecosystems \cite{berry2019marine}}
 
@@
 
+%
+
 \section{Methods}
+\subsection*{Sampling}
 
-\subsection*{Sampling}
 We collected water samples to assess eDNA communities in two regions of the Salish Sea (Washington, USA): San Juan Island and the Hood Canal.
@@
 Within a geographic area and season, temperature and pH were most strongly associated with community-level changes.
-We therefore developed a hierarchical logistical regression model relating the presence of each taxon to temperature and pH (as fixed effects), allowing slope to vary by geographic area and intercept to vary by season, to capture large-scale seasonal differences associated with sunlight, photosynthesis, and related factors.
+We therefore developed a hierarchical logistical regression model relating the presence of each taxon to temperature and pH (as fixed effects) {\color{red} -salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}.
+We allowed} slope to vary by geographic area and intercept to vary by season, to capture large-scale seasonal differences associated with sunlight, photosynthesis, and related factors.
 For each taxon, we fit these models using the Bayesian generalized linear mixed effects functions in R package rstanarm \cite{rstanarm} for R.
@@
 \end{figure}
-
+%Move this to the methods%
  To explore the suitability of different environmental conditions for each taxon, we modeled the likelihood of taxon presence as a function of area and season (random effects) and temperature and pH (fixed effects).

### 2020-08-27 Scripts/Manuscript.Main.Analysis.Rmd
I think this is done

--- 2020-08-13 Scripts/All.figures.that.go.in.the.MS.no.seasons.Rmd
@@
 
-saveRDS( file = here("Output","Temp.pH.grid.RDS"))
+write_rds( path  = here("Output","Temp.pH.grid.RDS"))
 
 future.scenario %>% 
-  write_rds(file = here("Output", "future.simulation.rds"))
+  write_rds(path = here("Output", "future.simulation.rds"))
 
@@
 ```{r}
-all.sim.data <- simData.nocrazys  #to see simulation data
 
@@
 
-SIMPER.clusters %>% 
-  mutate(summary.simper = map(summary.simper,function(.x) bind_rows(.x[1:length(.x)], .id = "comparison") %>% rownames_to_column("taxa"))) %>% pull(summary.simper) %>% set_names(SIMPER.clusters$Area) %>% bind_rows(.id = "Area") %>% separate(taxa, into = c("taxa",NA), sep = "\\...") %>%
-  filter (p < 0.05) %>% 
-  write_csv(here("Output","simper.all.csv"))
-
 ## Get the data for all  areas
@@
    pull(summary.simper)-> list.of.simpers
-names(list.of.simpers) <- paste(SIMPER.clusters$Area, SIMPER.clusters$Season, sep = "_")
+names(list.of.simpers) <- SIMPER.clusters$Area
 
@@
 
-```{r}
+```{r choose taxa with more change}
 expanded.future %>% 
@@
   nest(-phylum) %>% 
-  # ungroup() %>% 
-  # filter (phylum %in%c("Dinophyceae", "Bacillariophyta")) %>% 
   mutate(suma = map (data,~.x %>% group_by(taxa,Area, Year) %>%
-                       summarise(tot = sum(present)) %>%
+                       summarise(prop = mean(present)) %>%
                        ungroup() %>% 
-                       pivot_wider(names_from = Year, values_from = tot) %>% 
-                        group_by(taxa) %>% mutate(diff = `2017` - `2095`)
+                       pivot_wider(names_from = Year, values_from = prop) %>% 
+                        group_by(taxa) %>% mutate(diff = `2017` - `2095`) %>% 
+                       pivot_wider(names_from = Area, values_from = c(`2017`, `2095`, diff)) %>% 
+                       mutate(sum = sum(abs(`diff_Hood Canal`), abs(diff_SJI)))
                      )) %>% 
@@
   unnest(suma) %>% 
-  group_by(phylum) %>% arrange(desc(abs(diff))) %>% 
+  group_by(phylum) %>% arrange(desc(sum)) %>% 
   separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
@@
 taxa.and.diff %>% 
-  slice(1:5) %>%
-  arrange(diff) %>% pull(taxa) -> ahorasi
+  filter(str_detect(label, paste(c("Centropages", "Nitzschia lo", "Nitzschia pro","Prorocentrum", "Emili", "Alexandrium", "Chaetoceros soc","Stephanodiscus","Coscinodiscus", "Karlodi",
+                                   #"Navicula cry","Navicula gla",
+                                   "Chryso", "Bathy", "Gym", "Li"), collapse = "|"))) %>% select(taxa, label)-> second.sel
 
-taxa.and.diff %>% 
-  slice(1:5) %>%
-  arrange(diff) %>% pull(label) -> reorder.labels
 ```
@@
 expanded.future %>% 
-  filter(taxa %in% ahorasi) %>%
   separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
@@
                           TRUE          ~ Area)) %>%
-  mutate(Year = fct_rev(as.factor(Year) )) %>% 
-  nest(-phylum) -> input.for.left
+  mutate(Year = fct_rev(as.factor(Year) )) -> input.for.left
 
 input.for.left %>%
+   group_by(phylum, Temperature, pH, Year, Area ) %>%
+  summarise(present = sum(present)) %>%
+  group_by(phylum,Area) %>% 
+  mutate(present = present/max(present)) %>% 
+  mutate (label = phylum) -> phylum.stats
+
+phylum.stats %>% 
+  nest(-phylum) %>% 
+  mutate( plot.by.phyl = map2 (data, phylum, ~ .x %>%
+  ggplot(aes()) +
+  scale_fill_brewer(name = "Year",  type = "qual", palette = 4, direction = 1,
+                    breaks=c("2017","2095")) +
+  labs (x = "Standardized richness",
+        y = "") +
+  facet_wrap(~Area) +
+  theme_minimal() +
+  theme (strip.text.y = element_text(angle = 0),
+        axis.text.y = element_text(face = "italic"), 
+        #plot.background   = element_rect(fill = "#E5E9F0"),
+        legend.direction = "horizontal",
+        legend.position  = "bottom",
+        plot.title.position = "plot",
+        plot.title = element_text(face = "bold",vjust =  -5 ),
+        axis.title.x.top  = element_text(face = "bold",vjust = 3) ) +
+        geom_histogram(aes( x= present,after_stat(ncount),
+                            fill = as.factor(Year),
+                            group= as.factor(Year)),
+                       binwidth = 0.025,
+                       colour = "black",
+                       position = "identity", 
+                       alpha = 0.75) +
+        scale_x_continuous(position = "top", limits = c(0,1)))) %>% 
+  select(-data)-> summary.plot
+
+input.for.left %>% 
+  semi_join(second.sel, by = "taxa") %>%
+  nest(-phylum) %>% 
   mutate( plot.by.taxa = map2 (data, phylum, ~ .x %>%
                                  ggplot(aes(x= present,
-                                            y = fct_relevel(label,reorder.labels))) +
+                                           y = label)) +
                                  scale_fill_brewer(name = "Year",  type = "qual", palette = 4, direction = 1,
@@
                                  ggtitle(.y)  +
-                                 scale_x_continuous(limits = c(0,1),position = "top") )) %>%
-  pull(plot.by.taxa) %>% 
-  set_names(input.for.left$phylum) -> fig3.left
+                                 geom_boxplot(aes(fill = as.factor(Year)), outlier.alpha = 0.5) +
+                                 scale_x_continuous(position = "top") )) -> fig3.left
 
-map(fig3.left, function(x) x+ geom_boxplot(aes(fill = as.factor(Year)))) -> fig3.left
+fig3.left %>%  right_join(summary.plot) -> left.plot.all
+left.plot.all %>% pull(plot.by.phyl) %>% set_names(nm = left.plot.all$phylum) -> left.plot.top
+left.plot.all %>% pull(plot.by.taxa) %>% set_names(nm = left.plot.all$phylum) -> left.plot
 
-fig3.left["Dinophyceae"]
 ```
@@
   scale_color_brewer(name = "Year",  type = "qual", palette = 4, direction = -1, breaks=c("2017","2095")) +
-  ggforce::geom_mark_hull(data = all.sims.for.projection %>%
+  ggforce::geom_mark_hull(data = future.scenario %>%
                             ungroup() %>%
@@
                  alpha = 0.5) +
-     facet_grid(.~Area) +
+     facet_grid(.~Area, scales = "free_x",space = "free_x") +
   guides(color = "none") +
@@
 
-fig3.right["Dinophyceae"]
-fig3.right["Bacillariophyta"]
-
 ```
@@
 Following reviewer's suggestions, we are limiting the information displayed in the main Figures to two of most important phyla for coastal ecosystems dinamics: Diatoms and Dinoflagellates.
-
-```{r select some }
-
+```{r}
 layout <- "
 AA
+BB
+##
+CC
+DD
+##
+EE
+GG
+HH
+"
+wrap_plots(A = left.plot[["Dinophyceae"]] + labs(title = "A.1", tag = "Dinophyceae") ,
+           B = left.plot.top[["Dinophyceae"]] + guides(fill = "none")+ labs(title  ="A.2"),
+          # F = plot_spacer(),
+           C = left.plot[["Bacillariophyta"]] + labs(title ="B.1", tag = "Bacillariophyta"),
+           D = left.plot.top[["Bacillariophyta"]] + guides(fill = "none")+ labs(title ="B.2"),
+           E = left.plot[["Haptophyceae"]] + labs(title ="C.1", tag = "Haptophyceae"),
+           G = left.plot.top[["Haptophyceae"]] + guides(fill = "none")+ labs(title  ="C.2"),
+           H = guide_area(), design = layout, heights = c(3,2,1,3,2,1,3,2,1), guides = "collect") &
+  theme(plot.tag.position = c(0.1,1),
+        plot.tag = element_text(face = "bold")) -> left
+
+layout2 <- "
 AA
-FF
+AA
+##
+CC
+CC
+##
 EE
 EE
-GG
+HH
 "
+wrap_plots(A = fig3.right[["Dinophyceae"]] + labs(title  ="A.3"),
+          # F = plot_spacer(),
+           C = fig3.right[["Bacillariophyta"]] + labs(title ="B.3"),
+           E = fig3.right[["Haptophyceae"]]+ labs(title ="C.3"),
+           H = guide_area(), design = layout2, heights = c(3,2,1,3,2,1,3,2,1), guides = "collect")&
+  theme(plot.tag.position = c(0.1,1),
+        plot.title  = element_text(face = "bold")) -> right
 
-wrap_plots(A = fig3.left[["Dinophyceae"]] +   labs(tag = "A"),
-           F = plot_spacer(),
-           E = fig3.left[["Bacillariophyta"]] ,
-           G = guide_area(),
-            design = layout, heights = c(5,5,1,5,5,5), guides = "collect") -> left
-
-layout2 <- "
-#BB
-#BB
-JJJ
-#HH
-#HH
-III
-"
-wrap_plots(B = fig3.right[["Dinophyceae"]] +   labs(tag = "B"),
-           J = plot_spacer(),
-           H = fig3.right[["Bacillariophyta"]] ,
-           I = guide_area(),
-            design = layout2,heights = c(5,5,1,5,5,5),
-           widths = c(1,2,2), guides = "collect") -> right
-
-(left|right)+plot_layout(widths = c(2,3))  -> plot 
-   ggsave(plot , filename = here("Figures","Manuscript", paste0("Top_fig3_",Sys.Date(),".Other.sel.png")), units = "in", dpi = "retina", width = 20, height = 12)
+(left|right)+plot_layout(widths = c(2,2))  -> plot 
+   ggsave(plot , filename = here("Figures","Manuscript", paste0("Top_fig3_",Sys.Date(),".Other.sel.png")), units = "in", dpi = "retina", width = 20, height = 14)
 
@@
 ```{r}
+input.for.left %>% 
+  nest(-phylum) %>% 
+  mutate( plot.by.taxa = map2 (data, phylum, ~ .x %>%
+                                 ggplot(aes(x= present,
+                                           y = label)) +
+                                 scale_fill_brewer(name = "Year",  type = "qual", palette = 4, direction = 1,
+                                                   breaks=c("2017","2095")) +
+                                 labs (x = "Probability of presence",
+                                       y = "") +
+                                 facet_wrap(~Area) +
+                                 theme_minimal() +
+                                 theme(strip.text.y = element_text(angle = 0),
+                                       axis.text.y = element_text(face = "italic"),
+                                       legend.direction = "horizontal",
+                                       legend.position  = "bottom",
+                                       plot.title.position = "plot",
+                                       plot.title = element_text(face = "bold",vjust =  -5 ),
+                                       axis.title.x.top  = element_text(face = "bold",vjust = 3) ) +
+                                 ggtitle(.y)  +
+                                 geom_boxplot(aes(fill = as.factor(Year)), outlier.alpha = 0.5) +
+                                 scale_x_continuous(position = "top") )) -> fig3.left
 
-map2(fig3.left, fig3.right, 
+fig3.left %>% 
+  left_join(list.of.richness.plots, by = "phylum") -> fig3.left
+
+fig3.left %>% pull(plot) %>% set_names(fig3.left$phylum) -> rights
+fig3.left %>% pull(plot.by.taxa) %>% set_names(fig3.left$phylum) -> lefts
+
+map2(lefts, rights, 
      function(.x,.y){
@@
 
-paths <- paste0("Supp.fig.",names(fig3.left),".png")
+paths <- paste0("Supp.fig.",names(lefts),".png")
 

### 2020-08-27 Manuscript/What.lies.ahead.reviewed.tex

@@
 We estimate community-level differences associated with time, space and environmental variables, and use these results to forecast near-term community changes due to warming and ocean acidification.
-We find distinct communities in warmer and more acidified conditions in both areas sampled, and a strong reduction in richness of current diatom assemblages.
+We find distinct communities in warmer and more acidified conditions, and a strong reduction in richness of current diatom assemblages.
 Taxa finding greater suitable habitat in near-future waters are more taxonomically varied and include the ubiquitous coccolithophore \textit{Emiliania huxleyi} and the harmful \textit{Alexandrium sp.} These results suggest foundational changes for nearshore food webs under near-future conditions.
@@
 
-{\color{red}A growing body of evidence backs up the efficacy of eDNA metabarcoding for monitoring biodiversity (see a review in \cite{ruppert2019past}), and this approach has been successfully used to detect community composition variation across environmental changes in aquatic  \cite{emilson2017dna}, estuarine \cite{chariton2015metabarcoding, lallias2015environmental}, and marine ecosystems \cite{berry2019marine}}
+{\color{red}A growing body of evidence supports the efficacy of eDNA metabarcoding for monitoring biodiversity (see a review in \cite{ruppert2019past}), and this approach has been successfully used to detect community composition variation across environmental changes in aquatic  \cite{emilson2017dna}, estuarine \cite{chariton2015metabarcoding, lallias2015environmental}, and marine ecosystems \cite{berry2019marine, djurhuus2019eDNA}}
 
@@
 Both measurements were calibrated and validated with certified reference material from the Scripps Oceanographic Institute.
-Using DIC and TA, we calculated pH and the remaining carbonate system parameters using the R package `seacarb` \cite{seacarb}.
+Using DIC and TA, we calculated pH and the remaining carbonate system parameters using the R package `seacarb' \cite{seacarb}.
 
@@
 
-Code for all quality-screening and bioinformatics is available in the Supporting Information.
+Code for all quality-screening and bioinformatics is available in the Supporting Information, implemented in Unix and R \cite{r_core_development_team_r:_2013}.
 Briefly, we used a Unix script that calls secondary programs for primer-trimming and preliminary quality-control \cite{martin2011cutadapt, Callahan:2016aa} we estimated the likely composition of each sample using DADA2.
 This approach avoids clustering, such that we retained all of the amplicon sequence variants (ASVs, \textit{i.e.}, unique sequences); we subsequently carried out secondary quality-control and decontamination following \cite{kelly2018tides}.
-We then assigned sequences to known taxa using phylogenetic tree placement with insect v1.1 \cite{insect}.
-We conservatively kept only taxa annotated at the level of taxonomic family, genus, or species, so we could reliably infer taxon natural history.
+We then assigned sequences to known taxa using phylogenetic tree placement with insect v1.1 \cite{insect}; where insect could not place individual taxa, we supplemented assignment with a COI database (CITE CRUX).
+We conservatively kept only taxa annotated at the level of taxonomic family, genus, or species, so we could reliably infer taxon natural history under the assuming that taxa within the same family shared broad natural-history characteristics.
 Using published literature and online databases, we placed every recovered taxon into a benthic/planktonic category and focused our analysis on the planktonic community (see Supporting Information).
@@
 
-\subsection*{Year 2095 Environmental Scenario and Biological Responses}
+\subsection*{Year-2095 Environmental Scenario and Biological Responses}
 
 We estimated the distribution of environmental parameters for the overall Salish Sea in 2095 from the results of \cite{khangaonkar2019salish}, which estimated an annual mean increase in temperature of 1.51 \textdegree C and mean pH decrease of 0.18 for the Salish Sea as a whole.
-We then subsampled this distribution to reflect the characteristic environments of different areas (San Juan Island and Hood Canal) and seasons (Summer and Winter, with April - September coded as ``Summer'').
+{\color{red}[ADD DETAIL HERE]}
+
+%We then subsampled this distribution to reflect the characteristic environments of different areas (San Juan Island and Hood Canal) and seasons (Summer and Winter, with April - September coded as ``Summer'').
 To do this, we fit beta parameters to describe the fraction of the overall observed data (2017 - 2018) reflected in each subset; we then subsampled the 2095 distribution accordingly, under the assumption that the Hood Canal in Summer (for example) would occupy the same fraction of the overall annual Salish Sea temperature and pH distributions in 2095 as it did in 2017.
@@
 
-For community-level projections, we coded community-cluster identity (Figure 2) as an unordered response variable in a multinomial logistic model, with temperature, pH, area, and season as predictor variables.
-We calculated the probability of each community, given these predictors, using the package `nnet` \cite{nnet2002}.
+For community-level projections, we coded community-cluster identity (Figure 2) as an unordered response variable in a multinomial logistic model, with temperature, pH, and area (Hood Canal vs.
+San Juan Island), as predictor variables.
+{\color{red} Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}, and because salinity was correlated with temperature in our dataset, it was not an important predictor variable and we subsequently dropped it from our models.} We calculated the probability of each community, given these predictors, using the package `nnet` \cite{nnet2002}.
 Most-common combinations of T and pH were estimated from 2-dimensional density plots for each year.
 
-Within a geographic area and season, temperature and pH were most strongly associated with community-level changes.
-We therefore developed a hierarchical logistical regression model relating the presence of each taxon to temperature and pH (as fixed effects) {\color{red} -salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}.
-We allowed} slope to vary by geographic area and intercept to vary by season, to capture large-scale seasonal differences associated with sunlight, photosynthesis, and related factors.
+Within a geographic area, temperature and pH were most strongly associated with community-level changes, although these conditions varied substantially between the Hood Canal and San Juan Island sites.
+We therefore developed a hierarchical logistical regression model relating the presence of each taxon to temperature and pH, in which the slopes of temperature and pH effects varied by taxon, and each taxon had a unique intercept that was allowed to vary by geographic area.
 For each taxon, we fit these models using the Bayesian generalized linear mixed effects functions in R package rstanarm \cite{rstanarm} for R.
-See Supporting Information for analytical code.
+Model selection using WAIC supported this as the preferred model over several similar ones (see Supporting Information).
 
 Given the estimated sea-surface temperatures and pH values for 2095 and observed data from 2017, we then used taxon-specific logistic regression models to estimate the suitability of habitat for each taxon in the future scenario.
-For each point in the pH - temperature grid, we calculated species richness as the mean result of 100 independent draws of the posterior distributions of all modeled taxa, grouping the individual taxa models by phylum.
+For each point in the pH - temperature grid, we calculated species richness as the mean result of 100 independent draws of the posterior distributions of all modeled taxa.
 
@@
 \caption{\footnotesize {\textbf{A} Sampling locations in intertidal areas of the Hood Canal (Red) and San Juan Island (Blue).
-\textbf{B} Planktonic richness (unique taxa) per sampling locality across seasons, as reflected by the eDNA COI assay.
+\textbf{B} Planktonic richness (unique taxa) per sampling locality, as reflected by the eDNA COI assay.
 \textbf{C} The observed environmental profiles of these two regions reflect a broad range of environmental conditions, with the Hood Canal resembling future conditions in temperate areas worldwide.
 Shown: Dissolved Inorganic Carbon (DIC, $\mu$moles/kg); Total Alkalinity (TA); Temperature ($\degree$C) and Salinity (PSU).
-Individual sampling sites shown as facets.}}
+Individual sampling sites shown as facets.
+{\color{red} CHANGE y-scale for Dinophyceae}}}
 \label{fig:Fig1}
@@
 We collected intertidal environmental and genetic data from bottle samples each month between March 2017 and August 2018 (Figure 1A).
-Despite geographic proximity and similar overall species composition (127 of the 222 planktonic taxa were found in both regions and accounted for 98\% of the sequences), the areas under study -- San Juan Island and the Hood Canal -- varied widely in pH, temperature, and other environmental parameters (Figure 1C), with different points along the environmental gradient simultaneously showing differences equivalent to those predicted between present-day and future oceans \cite{bopp2013multiple}.
+Despite geographic proximity and similar overall species composition (127 of the 222 planktonic taxa were found in both regions and accounted for 98\% of the sequences), the areas under study -- San Juan Island and the Hood Canal -- varied widely in pH, temperature, and other environmental parameters (Figure 1C), with a smooth gradient in conditions along the Hood Canal, and San Juan Island more closely resembling full marine conditions.
+Different points along the environmental continuum simultaneously showed differences equivalent to those predicted between present-day and future oceans \cite{bopp2013multiple}.
 
@@
 These taxa showed a seasonal richness gradient between study areas, consistent with documented biodiversity clines in the area (Fig.
-1B) \cite{dethier_decadal_2008}.
+1B.1) \cite{dethier_decadal_2008}.
 
-Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to season (Summer vs.
-Winter; PERMANOVA F-model = 14.4; p $<$ 0.001) and geographic Area (Hood Canal vs.
+Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to geographic Area (Hood Canal vs.
 San Juan Island; F-model = 15.4; p $<$ 0.001).
-We therefore performed a constrained analysis of principal components (CAP) for samples within each combination of Season and Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig.
+{\color{red} RPK: still up-to-date?
+no adonis code in the Rmd file that I saw} .
+We therefore performed a constrained analysis of principal components (CAP) for samples within each Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig.
 2 shows results for Hood Canal in summer as an example; full analysis in the Supporting Information).
@@
  Multinomial logistic regression yielded predictions of the most-likely community for any combination of environmental parameters (Fig. 2B).
+
+{\color{red} RPK: I think everything below needs updating, according to some final analysis}
 
@@
 \centering
-\includegraphics[width=1\linewidth]{Figures_for_MS/Figure2.2019-11-06.png}
+\includegraphics[width=1\linewidth]{Figures_for_MS/Figure2.2020-08-12.png}
 \caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal.
@@
  \centering
- \includegraphics[width=1\linewidth]{Figures_for_MS/Figure3.20191105.png}
+ \includegraphics[width=1\linewidth]{Figures_for_MS/Top_fig3_2020-08-12.Other.sel.png}
 \caption{\footnotesize{Forecasted changes in plankton in the Salish Sea.

### 2020-08-28 Manuscript/What.lies.ahead.reviewed.tex
Started updating the ms

@@
 \usepackage{xcolor}
+\usepackage{soul}
 
@@
 This approach avoids clustering, such that we retained all of the amplicon sequence variants (ASVs, \textit{i.e.}, unique sequences); we subsequently carried out secondary quality-control and decontamination following \cite{kelly2018tides}.
-We then assigned sequences to known taxa using phylogenetic tree placement with insect v1.1 \cite{insect}; where insect could not place individual taxa, we supplemented assignment with a COI database (CITE CRUX).
+We then assigned sequences to known taxa using phylogenetic tree placement with insect v1.1 \cite{insect}; where insect could not place individual taxa, we supplemented assignment with a COI database \cite{curd2019anacapa}.
 We conservatively kept only taxa annotated at the level of taxonomic family, genus, or species, so we could reliably infer taxon natural history under the assuming that taxa within the same family shared broad natural-history characteristics.
@@
 By treating amplification efficiency as consistent within a given taxon, we created an index of abundance for each taxon across space and time \cite{kelly2019understanding}, using pooled data from technical replicates and mean proportions across biological replicates.
-The multivariate tests (PERMANOVA, CAP, k-means clustering) used these index values as the species abundances and Bray-Curtis as the dissimilarity measurement between samples.
+
+\subsection*{Present scenario community analyses}
+
+{\color{red}
+We performed a series of multivariate tests to investigate the community changes across the environmental space we observed.
+We used the index of abundance as the species abundances and Bray-Curtis as the dissimilarity measurement between samples.
+A global PERMANOVA test (Bray-Curtis ~ Area) further justified the subsequent separated  analysis of the Hood Canal and San Juan Island samples; a Constrained Analysis of Principal Components (CAP, \cite{Anderson}) investigated the distinctiveness of plankton communities across the gradient of temperature, pH and salinity.
+
+Independent of environmental parameters, we separately clustered samples by pairwise Bray-Curtis dissimilarities among metabarcoding communities (k-means; N = 3) to identify groups of samples that were similar to one another with respect to biological community.
+The SIMPER procedure from the r package \textit{vegan} revealed the taxa most contributing to between-cluster differences.
+
+For community-level projections, we coded community-cluster identity (Figure 2) as an unordered response variable in a multinomial logistic model, with temperature, pH, and area (Hood Canal vs.
+San Juan Island), as predictor variables.
+Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}, and because salinity was correlated with temperature in our dataset, it was not an important predictor variable and we subsequently dropped it from our models.
+We calculated the probability of each community, given these predictors, using the r package \textit{nnet} \cite{nnet2002}.
+}
 
@@
 We estimated the distribution of environmental parameters for the overall Salish Sea in 2095 from the results of \cite{khangaonkar2019salish}, which estimated an annual mean increase in temperature of 1.51 \textdegree C and mean pH decrease of 0.18 for the Salish Sea as a whole.
-{\color{red}[ADD DETAIL HERE]}
-
+{\color{red}We fit our current environmental observations to a normal distribution, and used the parameters of that distribution alongside a linear variation of temperature and pH between the present and the 2095 scenario to generate 1000 simulations of each year scenario.
+The scenario labelled as 2095 is the result of 5 simulations for the years 2091-2095.
+We removed outliers from these simulations by keeping the 95\% percentile of observations.
+See the R code in the supplementary information to follow the process of generating these datasets (lines 98-135).
 %We then subsampled this distribution to reflect the characteristic environments of different areas (San Juan Island and Hood Canal) and seasons (Summer and Winter, with April - September coded as ``Summer'').
@@
 
-For community-level projections, we coded community-cluster identity (Figure 2) as an unordered response variable in a multinomial logistic model, with temperature, pH, and area (Hood Canal vs.
+\st{For community-level projections, we coded community-cluster identity (Figure 2) as an unordered response variable in a multinomial logistic model, with temperature, pH, and area (Hood Canal vs.
 San Juan Island), as predictor variables.
-{\color{red} Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}, and because salinity was correlated with temperature in our dataset, it was not an important predictor variable and we subsequently dropped it from our models.} We calculated the probability of each community, given these predictors, using the package `nnet` \cite{nnet2002}.
+We calculated the probability of each community, given these predictors, using the package `nnet` .
 Most-common combinations of T and pH were estimated from 2-dimensional density plots for each year.
+}}
 
@@
 \includegraphics[width=1 \linewidth]{Figures_for_MS/Fig1.png}
-\caption{\footnotesize {\textbf{A} Sampling locations in intertidal areas of the Hood Canal (Red) and San Juan Island (Blue).
-\textbf{B} Planktonic richness (unique taxa) per sampling locality, as reflected by the eDNA COI assay.
-\textbf{C} The observed environmental profiles of these two regions reflect a broad range of environmental conditions, with the Hood Canal resembling future conditions in temperate areas worldwide.
-Shown: Dissolved Inorganic Carbon (DIC, $\mu$moles/kg); Total Alkalinity (TA); Temperature ($\degree$C) and Salinity (PSU).
+\caption{\footnotesize {\textbf{A} Sampling locations in intertidal areas of the Hood Canal ({\color{red}dark gold}) and San Juan Island ({\color{red}turquoise}).
+\textbf{B} Planktonic richness (unique taxa) per sampling locality, as reflected by the eDNA COI assay.{\color{red}\textbf{B.1} Taxa from all phyla; \textbf{B.2} Diatoms(above) and dinoflagellates (below) shown separately} \textbf{C} The observed environmental profiles of these two regions reflect a broad range of environmental conditions, with the Hood Canal resembling future conditions in temperate areas worldwide.
+Shown: Dissolved Inorganic Carbon (DIC, $\mu$moles/kg); Total Alkalinity (TA, $\mu$moles/kg); Temperature ($\degree$C) and Salinity (PSU).
 Individual sampling sites shown as facets.
-{\color{red} CHANGE y-scale for Dinophyceae}}}
+{\color{red} CHANGE y-scale for Dinophyceae DONE}}}
 \label{fig:Fig1}
@@
 Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to geographic Area (Hood Canal vs.
-San Juan Island; F-model = 15.4; p $<$ 0.001).
+San Juan Island; F-model = 5.4; p $<$ 0.0001).
 {\color{red} RPK: still up-to-date?
-no adonis code in the Rmd file that I saw} .
+no adonis code in the Rmd file that I saw.
+RGS: that's true - I added new chunk of code line 382 \& updated results}.
 We therefore performed a constrained analysis of principal components (CAP) for samples within each Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig.
-2 shows results for Hood Canal in summer as an example; full analysis in the Supporting Information).
-Independent of environmental parameters, we separately clustered samples by pairwise Bray-Curtis dissimilarities among metabarcoding communities (k-means; N = 3 or 4) to identify groups of samples that were similar to one another with respect to biological community.
+2 shows results for Hood Canal; full analysis in the Supporting Information).
+{\color{red}\st{Independent of environmental parameters, we separately clustered samples by pairwise Bray-Curtis dissimilarities among metabarcoding communities (k-means; N = 3) to identify groups of samples that were similar to one another with respect to biological community.}}
  Each biological cluster (colored hulls, Fig.
-2) occupied a unique area of environmental parameter space.
+2.A) occupied a unique area of environmental parameter space.
 Planktonic communities therefore varied predictably with water temperature, salinity, and pH, across a range of those parameters likely to be encountered in many near-term future-ocean scenarios \cite{murray2015inland}.
@@
 
-{\color{red} RPK: I think everything below needs updating, according to some final analysis}
+{\color{red} RPK: I think everything below needs updating, according to some final analysis. TODO: RGS}
 
-These communities were distinguished by changes in the relative abundances of many taxa: in the Hood Canal in summer for example (Fig.
-2C), the colder water communities showed higher values of \textit{Centropages} sp., \textit{Nitzschia longissima}, \textit{Prorocentrum} sp., \textit{Chrysochromulina} sp., \textit{Navicula} sp.
+These communities were distinguished by changes in the relative abundances of many taxa: in the Hood Canal for example (Fig.
+2C), the colder water and higher pH communities showed higher values of \textit{Centropages} sp., \textit{Nitzschia longissima}, \textit{Prorocentrum} sp., \textit{Chrysochromulina} sp., \textit{Navicula} sp.
 and \textit{Chaetoceros} sp.
@@
 
-The low-pH planktonic communities (cluster 3, Fig.
-2) occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls.
+The low-pH and {\color{red}high-temperature} planktonic communities (cluster 3, Fig.
+2{\color{red}A and B}) occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls.
 For example, by 2095 we expect summer in the Hood Canal to be T $>$ 12 and pH $<$ 8.2 -- where cluster 3 is the most likely community -- to occur more than two-thirds of the time (68\%).
@@
 \centering
-\includegraphics[width=1\linewidth]{Figures_for_MS/Figure2.2020-08-12.png}
+\includegraphics[width=1\linewidth]{	Figures_for_MS/Figure2.2020-08-27.png}
 \caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal.
@@
  \centering
- \includegraphics[width=1\linewidth]{Figures_for_MS/Top_fig3_2020-08-12.Other.sel.png}
+ \includegraphics[width=1\linewidth]{Figures_for_MS/Top_fig3_2020-08-27.Other.sel.png}
 \caption{\footnotesize{Forecasted changes in plankton in the Salish Sea.

### 2020-09-02 Manuscript/What.lies.ahead.reviewed.tex
edits to manuscript text; have not yet gotten to Results

@@
 We estimate community-level differences associated with time, space and environmental variables, and use these results to forecast near-term community changes due to warming and ocean acidification.
-We find distinct communities in warmer and more acidified conditions, and a strong reduction in richness of current diatom assemblages.
-Taxa finding greater suitable habitat in near-future waters are more taxonomically varied and include the ubiquitous coccolithophore \textit{Emiliania huxleyi} and the harmful \textit{Alexandrium sp.} These results suggest foundational changes for nearshore food webs under near-future conditions.
+We find distinct communities in warmer and more acidified conditions, with overall reduced richness in diatom assemblages and increased richness in dinoflagellates.
+Individual taxa finding greater suitable habitat in near-future waters are more taxonomically varied and include the ubiquitous coccolithophore \textit{Emiliania huxleyi} and the harmful dinoflagellate \textit{Alexandrium sp.} These results suggest foundational changes for nearshore food webs under near-future conditions.
 
@@
 Environmental DNA (eDNA) metabarcoding \cite{holman2019detection, kelly2016genetic} addresses this problem by amplifying a common gene region out of DNA present in a water sample; the technique can detect hundreds to thousands of taxa per sample, potentially with species-level identification.
-% Adding stuff on eDNA metabarcoding for community change.
-
-{\color{red}A growing body of evidence supports the efficacy of eDNA metabarcoding for monitoring biodiversity (see a review in \cite{ruppert2019past}), and this approach has been successfully used to detect community composition variation across environmental changes in aquatic  \cite{emilson2017dna}, estuarine \cite{chariton2015metabarcoding, lallias2015environmental}, and marine ecosystems \cite{berry2019marine, djurhuus2019eDNA}}
+A growing body of evidence supports the efficacy of eDNA metabarcoding for monitoring biodiversity (see a review in \cite{ruppert2019past}), and this approach has been successfully used to detect community composition variation across environmental changes in aquatic  \cite{emilson2017dna}, estuarine \cite{chariton2015metabarcoding, lallias2015environmental}, and marine ecosystems \cite{berry2019marine, djurhuus2019eDNA}.
 
@@
 
-Here we use a temporal and spatial series of metabarcoding samples to track changes in nearshore ecological communities associated with differences in pH, water temperature, and other environmental variables.
+Here we use series of metabarcoding samples taken across space and time to track changes in nearshore ecological communities associated with differences in pH, water temperature, and other environmental variables.
 We use broad-spectrum PCR primers \cite{leray_new_2013} to target eukaryotes specifically, identifying the likely effects of future climate scenarios on suites of planktonic taxa.
@@
 \subsection*{eDNA sequencing and bioinformatic processes}
-We purified DNA from each filtered sample using a Phenol Chloroform Isoamyl Alcohol protocol, following \cite{renshaw2015room}.
-A 1/10 to 1/100 dilution of the extract was used as template for a PCR reaction targeting a 313bp fragment of cytochrome oxidase I \cite{leray_new_2013}.
+We purified DNA from each filtered sample using a Phenol- Chloroform-Isoamyl Alcohol protocol, following \cite{renshaw2015room}.
+After reducing inhibition via a 1/10 to 1/100 dilution, the extract was used as template for a PCR reaction targeting a 313bp fragment of cytochrome oxidase I \cite{leray_new_2013}.
 PCR reactions were performed in triplicate and sequenced individually to quantify the stochasticity of PCR reactions on a mixed template sample, and we attached secondary indexing tags using a two-step PCR process \cite{odonnellPrimers}.
 PCR conditions and protocols for sample identification followed \cite{kelly2018tides}, and batches of 49 to 178 multiplexed samples were sequenced using MiSeq v2-500 or v3-600 sequencing kits using manufacturer protocols.
-On each sequencing run, we added triplicate samples consisting on DNA obtained from species not present in the marine environment under study (Red Kangaroo (\textit{Macropus rufus}) and Ostrich (\textit{Struthio camelus})) to establish quality controls of sample assignment and quantify levels of 'tag-jumping' or sample-cross-talk \cite{schnell2015tag}.
+On each sequencing run, we added triplicate samples consisting on DNA obtained from species not present in the marine environment under study (Red Kangaroo (\textit{Macropus rufus}) and Ostrich (\textit{Struthio camelus})) to establish quality controls of sample assignment and to quantify levels of 'tag-jumping' or sample-cross-talk \cite{schnell2015tag}.
 
@@
 This approach avoids clustering, such that we retained all of the amplicon sequence variants (ASVs, \textit{i.e.}, unique sequences); we subsequently carried out secondary quality-control and decontamination following \cite{kelly2018tides}.
-We then assigned sequences to known taxa using phylogenetic tree placement with insect v1.1 \cite{insect}; where insect could not place individual taxa, we supplemented assignment with a COI database \cite{curd2019anacapa}.
+We then assigned sequences to known taxa using phylogenetic tree placement with \textit{insect} v1.1 \cite{insect}; where \textit{insect} could not place individual taxa, we supplemented assignment by classification against a custom COI database using \textit{anacapa} \cite{curd2019anacapa} and \textit{bowtie2} \cite{langmead2012fast}.
 We conservatively kept only taxa annotated at the level of taxonomic family, genus, or species, so we could reliably infer taxon natural history under the assuming that taxa within the same family shared broad natural-history characteristics.
@@
 By treating amplification efficiency as consistent within a given taxon, we created an index of abundance for each taxon across space and time \cite{kelly2019understanding}, using pooled data from technical replicates and mean proportions across biological replicates.
+{\color{red} RPK: which analyses use the index, rather than presence-absence?
+make that clear here.}
 
@@
 
-{\color{red}
-We performed a series of multivariate tests to investigate the community changes across the environmental space we observed.
-We used the index of abundance as the species abundances and Bray-Curtis as the dissimilarity measurement between samples.
-A global PERMANOVA test (Bray-Curtis ~ Area) further justified the subsequent separated  analysis of the Hood Canal and San Juan Island samples; a Constrained Analysis of Principal Components (CAP, \cite{Anderson}) investigated the distinctiveness of plankton communities across the gradient of temperature, pH and salinity.
+We measured community changes across environmental space using multivariate analyses.
+We used the index of eDNA abundance to calculate Bray-Curtis dissimilarities  between samples, and estimated the effects of temperature, pH, and salinity on community composition using Constrained Analysis of Principal Components (CAP, \cite{anderson2003canonical}; `capscale' function in the R package \textit{vegan} \cite{vegan}).
 
-Independent of environmental parameters, we separately clustered samples by pairwise Bray-Curtis dissimilarities among metabarcoding communities (k-means; N = 3) to identify groups of samples that were similar to one another with respect to biological community.
-The SIMPER procedure from the r package \textit{vegan} revealed the taxa most contributing to between-cluster differences.
+Independent of environmental parameters, we separately clustered samples by pairwise Bray-Curtis dissimilarities (k-means; N = 3) to identify groups of samples that were similar to one another with respect to biological community.
+The SIMPER procedure in \textit{vegan} revealed the taxa most contributing to between-cluster differences.
 
@@
 We calculated the probability of each community, given these predictors, using the r package \textit{nnet} \cite{nnet2002}.
-}
 
@@
 We estimated the distribution of environmental parameters for the overall Salish Sea in 2095 from the results of \cite{khangaonkar2019salish}, which estimated an annual mean increase in temperature of 1.51 \textdegree C and mean pH decrease of 0.18 for the Salish Sea as a whole.
-{\color{red}We fit our current environmental observations to a normal distribution, and used the parameters of that distribution alongside a linear variation of temperature and pH between the present and the 2095 scenario to generate 1000 simulations of each year scenario.
-The scenario labelled as 2095 is the result of 5 simulations for the years 2091-2095.
-We removed outliers from these simulations by keeping the 95\% percentile of observations.
-See the R code in the supplementary information to follow the process of generating these datasets (lines 98-135).
-%We then subsampled this distribution to reflect the characteristic environments of different areas (San Juan Island and Hood Canal) and seasons (Summer and Winter, with April - September coded as ``Summer'').
-To do this, we fit beta parameters to describe the fraction of the overall observed data (2017 - 2018) reflected in each subset; we then subsampled the 2095 distribution accordingly, under the assumption that the Hood Canal in Summer (for example) would occupy the same fraction of the overall annual Salish Sea temperature and pH distributions in 2095 as it did in 2017.
-See Supporting Information for details of sampling parameters.
+We fit a normal distribution to our 2017 environmental observations to create baseline conditions, and modeled the change in mean parameter values between 2017 and 2095 as a linear function of time.
+We then used the modeled distribution of environmental parameters to generate 1000 simulations of each year scenario.
+The scenario labeled as 2095 is the set of parameters falling within the 95\% percentile in simulations for the years 2091-2095.
+See the R code in the supplementary information (lines 98-135).
 
-\st{For community-level projections, we coded community-cluster identity (Figure 2) as an unordered response variable in a multinomial logistic model, with temperature, pH, and area (Hood Canal vs.
-San Juan Island), as predictor variables.
-We calculated the probability of each community, given these predictors, using the package `nnet` .
-Most-common combinations of T and pH were estimated from 2-dimensional density plots for each year.
-}}
-
-Within a geographic area, temperature and pH were most strongly associated with community-level changes, although these conditions varied substantially between the Hood Canal and San Juan Island sites.
-We therefore developed a hierarchical logistical regression model relating the presence of each taxon to temperature and pH, in which the slopes of temperature and pH effects varied by taxon, and each taxon had a unique intercept that was allowed to vary by geographic area.
+To model biological responses to present and future scenarios, we used a hierarchical logistical regression model relating the presence of each taxon to temperature and pH, in which the slopes of temperature and pH effects varied by taxon, and each taxon had a unique intercept that was allowed to vary by geographic area.
 For each taxon, we fit these models using the Bayesian generalized linear mixed effects functions in R package rstanarm \cite{rstanarm} for R.
@@
 
-We collected intertidal environmental and genetic data from bottle samples each month between March 2017 and August 2018 (Figure 1A).
 Despite geographic proximity and similar overall species composition (127 of the 222 planktonic taxa were found in both regions and accounted for 98\% of the sequences), the areas under study -- San Juan Island and the Hood Canal -- varied widely in pH, temperature, and other environmental parameters (Figure 1C), with a smooth gradient in conditions along the Hood Canal, and San Juan Island more closely resembling full marine conditions.

### 2020-09-03 Manuscript/What.lies.ahead.reviewed.tex
Change in caption Fig2. Removed sd from calculations. Change Fig 3 axis labels. Added Rmd for model test

@@
 \subsection*{eDNA sequencing and bioinformatic processes}
-We purified DNA from each filtered sample using a Phenol- Chloroform-Isoamyl Alcohol protocol, following \cite{renshaw2015room}.
+We purified DNA from each filtered sample using a Phenol-Chloroform-Isoamyl Alcohol protocol, following \cite{renshaw2015room}.
 After reducing inhibition via a 1/10 to 1/100 dilution, the extract was used as template for a PCR reaction targeting a 313bp fragment of cytochrome oxidase I \cite{leray_new_2013}.
@@
 By treating amplification efficiency as consistent within a given taxon, we created an index of abundance for each taxon across space and time \cite{kelly2019understanding}, using pooled data from technical replicates and mean proportions across biological replicates.
-{\color{red} RPK: which analyses use the index, rather than presence-absence?
+{\color{red} RGS: We used this index of abundance to calculate the multivariate community analysis, while species responses used the presence/absence dataset.
+RPK: which analyses use the index, rather than presence-absence?
 make that clear here.}
@@
 \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows).
-\textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model.
+\textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model {\color{red}(salinity was not represented as it was tightly correlated with temperature)}.
 \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis).

### 2020-09-03 Scripts/Manuscript.Main.Analysis.Rmd
Change in caption Fig2. Removed sd from calculations. Change Fig 3 axis labels. Added Rmd for model test

@@
   geom_boxplot(aes(fill = Area)) +
-  facet_wrap(~phylum, nrow = 2) +
+  facet_wrap(~phylum, nrow = 2, scales = "free_y") +
   labs (y = "") -> right.rich.plot
@@
 To reuse for future analysis, create a function that does the fist step on a multivariate analysis: Create a horizontal matrix, a compatible matrix with the environmental data, and a dissimilarity object.
+
+We start with a global dataset PERMANOVA, to show that the communities from each region are different to each other and require independent analyisis
+
+```{r Global PERMANOVA}
+data.plankton %>% 
+  separate (event, into = c("Site", "Date"), sep = "_") %>% 
+  mutate(Area = case_when(Site %in% c("CP","LK", "FH") ~ "San Juan Island",
+                          TRUE                         ~ "Hood Canal")) %>% 
+   unite (Site, Date, col = "event", sep = "_") %>% 
+  mutate(GLOBAL = "Global") %>% 
+  nest(-GLOBAL) %>% 
+  mutate(comm.matrix = map (data, ~ tibble.to.table(.x, taxon = taxa, Abundance = Normalized.reads, sample.name = event)),
+         bc = map(data, ~tibble_to_matrix(.x, taxon = taxa,Abundance = Normalized.reads, sample.name = event, distance = "bray",transformation = "" )),
+         env = map (data, ~tibble_to_env(.x, taxon = taxa, Abundance = Normalized.reads, sample.name = event, everything()) %>% 
+                       select(-contains("%")) %>% 
+                       separate(event, into = c("Site", "Date"), sep = "_")),
+         PERMANOVA = map2(bc, env, function(.x, .y){
+           adonis(formula = .x ~ Area,data = .y, permutations = 9999
+                    )
+
+         })) %>% pull(PERMANOVA)
+```
 
@@
 
-expanded.future$sd <- apply(posterior.predict.all.taxa,2,sd )
 ```
@@
                     breaks=c("2017","2095")) +
-  labs (x = "Standardized richness",
+  labs (x = "Phylum-wide standardized richness",
         y = "") +
@@
   group_by(Area, Temperature = TemperatureCelsius, pH = pH_units, phylum) %>% 
-  summarise(richness = sum(present),
-            sd       = sum(sd^2)) %>% 
+  summarise(richness = sum(present)) %>% 
 #  left_join(richness.by.phylum) %>% 
@@
 
-# Assembling Figure 3
+### Assembling Figure 3
 

### 2020-09-03 Scripts/Supplement_ModelTesting.Rmd
Change in caption Fig2. Removed sd from calculations. Change Fig 3 axis labels. Added Rmd for model test

---
title: "Supplement: Testing Logistic Regression Models"
header-includes:
   - \usepackage{gensymb}    
   - \usepackage{fontspec}
   - \setmainfont{Open Sans}
output:
  pdf_document:
    fig_caption: yes
    keep_tex: true
    number_sections: no
    latex_engine: xelatex
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
library(rstanarm)
  options(mc.cores = parallel::detectCores())
library(loo)
library (here)
###LOAD DATA
      data.for.stan <- read_csv(here("Input","data.for.stan.csv"))

#create random subset of taxa on which to do preliminary model testing
      allTax <- unique(data.for.stan$taxa)
      set.seed(108)
      subsetTax <- allTax[sample(1:length(allTax), 50)]

#prep stan input
stan.input <- data.for.stan %>% 
      filter(taxa %in% subsetTax) %>% 
      group_by(Area) %>% 
      mutate(Temperature = vegan::decostand(Temperature, method = "standardize"),
             pH = vegan::decostand(pH, method = "standardize"),
             Salinity = vegan::decostand(Salinity, method = "standardize")) %>%
      group_by(taxa) %>%
      mutate(Noccur = sum(binary)) %>%
      filter(Noccur > 7 & Noccur < 69) %>%  #must be present or absent in at least 10% of 76 samples in dataset
      dplyr::select(-c(sample, phylum, DIC, Noccur)) %>% 
      ungroup() %>% 
      unite(Area, taxa, col = "AreaTaxa", remove = F) %>% 
      mutate(sampleIndex = 1:nrow(.),
             taxonIndex = match(taxa, unique(taxa)),
             areaIndex = match(Area, unique(Area)),
             seasonIndex = match(Season, unique(Season)),
             taxonAreaIndex = match(AreaTaxa, unique(AreaTaxa))) %>%   #unique combinations of taxon and area
      dplyr::select(-AreaTaxa)

```

We aimed to develop a quantitative model relating the probability of a taxon's presence at a site as a function of the environmental conditions at that site. Creating and tuning separate models for each of more than 200 taxa would have been analytically and practially unwieldy, and moreover would have missed the benefits of partially pooling information across taxa. Consequently, we developed a hierarchial logistic regression model in which the effects (slopes) of environmental parameters were allowed to vary by taxon.

Given the available continuous explanatory variables (Temperature, pH, and Salinity) and the discrete location data (geographic Area: Hood Canal vs. San Juan Island), a threshold decision was whether Area should feature in the model.  PERMANOVA analysis (see main text) indicated strongly different biological communities across the two Areas, and as a result, plausible models could include an intercept term that varied by Area and taxon. Conversely, it was implausible that the relationship between individual taxon presence and a given environmental variable (say, Temperature) would vary between Areas.

For all model fitting, we used \textit{rstanarm}, precompiled functions from the Stan language for Bayesian modeling, which implement Hamiltonian Monte Carlo parameter searches.

# Preliminary Model Testing and Model Selection

We conducted a preliminary model test on a suite of models meeting the above criteria using a random subset of taxa drawn from the overall dataset, further restricting this subset to those taxa occurring in $>10 \%$ and $< 90 \%$ of samples, ensuring a degree of variability in taxon presence on which to train the model. Furthermore, we standardized environmental variables to have mean = 0 and sd = 1 in order to provide sensible priors and to improve computational efficiency.

```{r, echo = F}
      t_prior <- student_t(df = 7, location = 0, scale = 2.5)

      #models fitting one intercept per unique area/taxon combination, plus
      #a different slope for the effect of environmental variables on each taxon.
      mod1.3vars <- stan_glmer(binary ~ 0 + (0 + Temperature + pH + Salinity | taxa) + (1 | taxonAreaIndex), 
                             data = stan.input, 
                             family = "binomial",
                             prior = t_prior,
                             prior_intercept = t_prior, 
                             chains = 2,
                             adapt_delta = 0.99,
                             refresh = 0)

      mod1.2vars <- stan_glmer(binary ~ 0 + (0 + Temperature + pH | taxa) + (1 | taxonAreaIndex), 
                             data = stan.input, 
                             family = "binomial",
                             prior = t_prior,
                             prior_intercept = t_prior, 
                             chains = 2,
                             adapt_delta = 0.99,
                             refresh = 0)

      mod1.1vars.pH <- stan_glmer(binary ~ 0 + (0 + pH | taxa) + (1 | taxonAreaIndex), 
                             data = stan.input, 
                             family = "binomial",
                             prior = t_prior,
                             prior_intercept = t_prior, 
                             chains = 2,
                             adapt_delta = 0.99,
                             refresh = 0)

      mod1.1vars.Temperature <- stan_glmer(binary ~ 0 + (0 + Temperature | taxa) + (1 | taxonAreaIndex), 
                             data = stan.input, 
                             family = "binomial",
                             prior = t_prior,
                             prior_intercept = t_prior, 
                             chains = 2,
                             adapt_delta = 0.99,
                             refresh = 0)

      #models fitting one intercept per taxon, plus
      #a different slope for the effect of environmental variables on each taxon. No Area term.
      mod2.3vars <- stan_glmer(binary ~ 0 + (0 + Temperature + pH + Salinity | taxa) + (1 | taxa), 
                             data = stan.input, 
                             family = "binomial",
                             prior = t_prior,
                             prior_intercept = t_prior, 
                             chains = 2,
                             adapt_delta = 0.99,
                             refresh = 0)

      mod2.2vars <- stan_glmer(binary ~ 0 + (0 + Temperature + pH | taxa) + (1 | taxa), 
                             data = stan.input, 
                             family = "binomial",
                             prior = t_prior,
                             prior_intercept = t_prior, 
                             chains = 2,
                             adapt_delta = 0.99,
                             refresh = 0)

      mod2.1vars.pH <- stan_glmer(binary ~ 0 + (0 + pH | taxa) + (1 | taxa), 
                              data = stan.input, 
                             family = "binomial",
                             prior = t_prior,
                             prior_intercept = t_prior, 
                             chains = 2,
                             adapt_delta = 0.99,
                             refresh = 0)

      mod2.1vars.Temperature <- stan_glmer(binary ~ 0 + (0 + Temperature | taxa) + (1 | taxa), 
                             data = stan.input, 
                             family = "binomial",
                             prior = t_prior,
                             prior_intercept = t_prior, 
                             chains = 2,
                             adapt_delta = 0.99,
                             refresh = 0)

      #models fitting a different slope for the effect of environmental variables on each taxon, with no intercept term.
      mod3.3vars <- stan_glmer(binary ~ 0 + (0 + Temperature + pH + Salinity | taxa), 
                             data = stan.input, 
                             family = "binomial",
                             prior = t_prior,
                             prior_intercept = t_prior, 
                             chains = 2,
                             adapt_delta = 0.99,
                             refresh = 0)

      mod3.2vars <- stan_glmer(binary ~ 0 + (0 + Temperature + pH | taxa), 
                             data = stan.input, 
                             family = "binomial",
                             prior = t_prior,
                             prior_intercept = t_prior, 
                             chains = 2,
                             adapt_delta = 0.99,
                             refresh = 0)

      mod3.1vars.pH <- stan_glmer(binary ~ 0 + (0 + pH | taxa), 
                             data = stan.input, 
                             family = "binomial",
                             prior = t_prior,
                             prior_intercept = t_prior, 
                             chains = 2,
                             adapt_delta = 0.99,
                             refresh = 0)

      mod3.1vars.Temperature <- stan_glmer(binary ~ 0 + (0 + Temperature | taxa), 
                             data = stan.input, 
                             family = "binomial",
                             prior = t_prior,
                             prior_intercept = t_prior, 
                             chains = 2,
                             adapt_delta = 0.99,
                             refresh = 0)

      #compare model results
      loo_compare(waic(mod1.3vars), 
                  waic(mod1.2vars), 
                  waic(mod1.1vars.pH), 
                  waic(mod1.1vars.Temperature), 
                  waic(mod2.3vars), 
                  waic(mod2.2vars), 
                  waic(mod2.1vars.pH), 
                  waic(mod2.1vars.Temperature), 
                  waic(mod3.3vars), 
                  waic(mod3.2vars), 
                  waic(mod3.1vars.pH), 
                  waic(mod3.1vars.Temperature)
                  )

```

The best-fit model included effects of Temperature and pH that varied by taxon, and an intercept term that varied for each unique taxon-Area combination. Formally, each observation is modeled as a single (Bernoulli) draw from a binomial distribution, with probability $p$. This probability $p$ depends, in turn, on a linear combination of the explanatory variables, using a logit link.

$$Presence \sim Bernoulli(p_i)$$

$$logit(p_i) = \alpha_{j} + Temperature * \beta_{1_k} + pH * \beta_{2_k}$$

Where $i$ indexes the individual observations in the dataset, $j$ indexes unique combinations of taxon and geographic Area, and $k$ indexes taxon identity. The parameters $\alpha$, $\beta_1$, and $\beta_2$ represent the intercept, effect of Temperature, and effect of pH, respectively. Priors and hyperpriors were drawn from a regularizing Student's t distribution (df = 7, location = 0, scale = 2.5).

# Final Model Fit and Illustration

Finally, we fit this model to the overall dataset -- again, keeping only taxa with at least 10\% variability in presence across samples.

```{r}
stan.input <- data.for.stan %>% 
      group_by(Area) %>% 
      mutate(Temperature = vegan::decostand(Temperature, method = "standardize"),
             pH = vegan::decostand(pH, method = "standardize"),
             Salinity = vegan::decostand(Salinity, method = "standardize")) %>%
      group_by(taxa) %>%
      mutate(Noccur = sum(binary)) %>%
      filter(Noccur > 7 & Noccur < 69) %>%  #must be present or absent in at least 10% of 76 samples in dataset
      dplyr::select(-c(sample, phylum, DIC, Noccur)) %>% 
      ungroup() %>% 
      unite(Area, taxa, col = "AreaTaxa", remove = F) %>% 
      mutate(sampleIndex = 1:nrow(.),
             taxonIndex = match(taxa, unique(taxa)),
             areaIndex = match(Area, unique(Area)),
             seasonIndex = match(Season, unique(Season)),
             taxonAreaIndex = match(AreaTaxa, unique(AreaTaxa))) %>%   #unique combinations of taxon and area
      dplyr::select(-AreaTaxa)

finalModel <- stan_glmer(binary ~ 0 + (0 + Temperature + pH | taxa) + (1 | taxonAreaIndex), 
                             data = stan.input, 
                             family = "binomial",
                             prior = t_prior,
                             prior_intercept = t_prior, 
                             chains = 4,
                             adapt_delta = 0.99,
                             refresh = 0)
```

The plot below illustrates the model fit, with observed data in black and modeled probabilities in blue.

```{r, echo = F}
stan.input$postpredict <- posterior_predict(finalModel) %>% colMeans()

stan.input %>% 
  left_join(data.for.stan[,c("taxa", "phylum")]) %>% 
        filter(phylum == "Bacillariophyta") %>% 
        filter(taxa %in% unique(taxa)[1:5]) %>% 
        ggplot(aes(x = pH, y = binary)) +
          geom_point() +
          geom_point(aes(x = pH, y = postpredict), color = "blue") +
          facet_grid(taxa ~ Area) +
          ylab("Prob of Presence") +
          xlab("pH (Standardized)") +
          ggtitle("Selected Diatom Taxa")

      stan.input %>% 
        filter(taxa == "Gonyaulacaceae|Alexandrium|NA") %>% 
        #filter(taxa %in% unique(taxa)[1:5]) %>% 
        dplyr::select(pH, Temperature, taxa, Area, binary, postpredict) %>% 
        pivot_longer(c(pH, Temperature)) %>% 
        ggplot(aes(x = value, y = binary)) +
        geom_point() +
        geom_point(aes(x = value, y = postpredict), color = "blue") +
        facet_grid(name ~ Area) +
        ylab("Prob of Presence") +
        xlab("pH (Standardized)") +
        ggtitle("Alexandrium")
```

### 2020-09-03 Manuscript/What.lies.ahead.reviewed.tex
edited and updated up to line number 170 in the tex file, which is line 226 in the PDF

@@
 By treating amplification efficiency as consistent within a given taxon, we created an index of abundance for each taxon across space and time \cite{kelly2019understanding}, using pooled data from technical replicates and mean proportions across biological replicates.
-{\color{red} RGS: We used this index of abundance to calculate the multivariate community analysis, while species responses used the presence/absence dataset.
-RPK: which analyses use the index, rather than presence-absence?
-make that clear here.}
+We used this index of abundance in the multivariate community analysis, and used binary (presence/absence) data to capture species-level responses to environmental conditions.
 
@@
 For each taxon, we fit these models using the Bayesian generalized linear mixed effects functions in R package rstanarm \cite{rstanarm} for R.
-Model selection using WAIC supported this as the preferred model over several similar ones (see Supporting Information).
+Model selection using WAIC \cite{watanabe2010asymptotic} supported this as the preferred model over several similar ones (see Supporting Information for model comparison information and code) and helped to avoid model overfitting and maintain out-of-sample predictive power.
 
-Given the estimated sea-surface temperatures and pH values for 2095 and observed data from 2017, we then used taxon-specific logistic regression models to estimate the suitability of habitat for each taxon in the future scenario.
-For each point in the pH - temperature grid, we calculated species richness as the mean result of 100 independent draws of the posterior distributions of all modeled taxa.
+Given the sea-surface temperatures and pH values for 2017 (observed) and 2095 (estimated) and taxon-specific logistic regression models, we then evaluated the suitability of habitat for each taxon in the future scenario.
+For each point in the pH - temperature grid, we calculated a species' probability of presence as the mean of 100 independent draws of the posterior model response.
+We estimated species richness as the sum of taxa with $>$ 50\% probability of occurrence at a given set of environmental parameters.
 
@@
 \includegraphics[width=1 \linewidth]{Figures_for_MS/Fig1.png}
-\caption{\footnotesize {\textbf{A} Sampling locations in intertidal areas of the Hood Canal ({\color{red}dark gold}) and San Juan Island ({\color{red}turquoise}).
-\textbf{B} Planktonic richness (unique taxa) per sampling locality, as reflected by the eDNA COI assay.{\color{red}\textbf{B.1} Taxa from all phyla; \textbf{B.2} Diatoms(above) and dinoflagellates (below) shown separately} \textbf{C} The observed environmental profiles of these two regions reflect a broad range of environmental conditions, with the Hood Canal resembling future conditions in temperate areas worldwide.
+\caption{\footnotesize {\textbf{A} Sampling locations in intertidal areas of the Hood Canal (dark gold) and San Juan Island (turquoise).
+\textbf{B} Planktonic richness (unique taxa) per sampling locality, as reflected by the eDNA COI assay.
+Boxplots represent the variability in richness across all time points for samples taken at the indicated site.
+\textbf{B.1} Taxa from all phyla; \textbf{B.2} Diatoms(above) and dinoflagellates (below) shown separately; note change in the scale of the y-axis.
+\textbf{C} The observed environmental profiles of these two regions reflect a broad range of environmental conditions, with the Hood Canal resembling future conditions in temperate areas worldwide.
 Shown: Dissolved Inorganic Carbon (DIC, $\mu$moles/kg); Total Alkalinity (TA, $\mu$moles/kg); Temperature ($\degree$C) and Salinity (PSU).
-Individual sampling sites shown as facets.
-{\color{red} CHANGE y-scale for Dinophyceae DONE}}}
+Individual sampling sites shown as vertical facets.}}
 \label{fig:Fig1}
@@
 San Juan Island; F-model = 5.4; p $<$ 0.0001).
-{\color{red} RPK: still up-to-date?
-no adonis code in the Rmd file that I saw.
-RGS: that's true - I added new chunk of code line 382 \& updated results}.
 We therefore performed a constrained analysis of principal components (CAP) for samples within each Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig.
-2 shows results for Hood Canal; full analysis in the Supporting Information).
-{\color{red}\st{Independent of environmental parameters, we separately clustered samples by pairwise Bray-Curtis dissimilarities among metabarcoding communities (k-means; N = 3) to identify groups of samples that were similar to one another with respect to biological community.}}
+2; for clarity, results for Hood Canal shown; full analysis in the Supporting Information).
+
  Each biological cluster (colored hulls, Fig.
@@
 
-{\color{red} RPK: I think everything below needs updating, according to some final analysis. TODO: RGS}
-
-These communities were distinguished by changes in the relative abundances of many taxa: in the Hood Canal for example (Fig.
-2C), the colder water and higher pH communities showed higher values of \textit{Centropages} sp., \textit{Nitzschia longissima}, \textit{Prorocentrum} sp., \textit{Chrysochromulina} sp., \textit{Navicula} sp.
-and \textit{Chaetoceros} sp.
-Warmer waters were dominated by \textit{Bathycoccus prasinos} and \textit{Cymbomonas} sp., while at lower pH values, marine communities showed \textit{Litchtheimia ramosa}, \textit{Micromonas pusilla} at higher relative abundances.
-
-\underline{Climate envelopes and future distributions}
-
-The low-pH and {\color{red}high-temperature} planktonic communities (cluster 3, Fig.
-2{\color{red}A and B}) occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls.
-For example, by 2095 we expect summer in the Hood Canal to be T $>$ 12 and pH $<$ 8.2 -- where cluster 3 is the most likely community -- to occur more than two-thirds of the time (68\%).
+These communities were distinguished by changes in the relative abundances of a wide variety of taxa.
+In the Hood Canal for example, the colder water and higher-pH community (community 1 in Fig.
+2C) showed higher eDNA indexes of diatoms \textit{Thalassiosira nordenskioeldii}, \textit{Nitzschia cf.
+promare}, and \textit{Ditylum brightwellii} -- among other species -- relative to a higher-temperature community (community 3 in Fig.
+2C).
+Communities 2 and 3 had relatively higher values for taxa from a mix of groups including \textit{Emiliana huxleyi}, \textit{Minutocellus polymorphus}, \textit{Micromonas pusilla}, and \textit{Bathycoccus prasinos}.
+Community 3 (Fig.
+2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls.
+For example, by 2095 we expect summer in the Hood Canal to be T $>$ 12 and pH $<$ 8.2 -- where community 3 is the most likely community -- to occur more than two-thirds of the time (68\%).
+{\color{red} RAMON -- can you see if this 68\% stat is still true, and also if you can incorporate the similar expectations that are commented-out directly below, updating their \% if necessary?}  \\
 
@@
 \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows).
-\textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model {\color{red}(salinity was not represented as it was tightly correlated with temperature)}.
+\textbf{B} Most-likely cluster as a function of temperature and pH given a multinomial logistic model (salinity was uninformative, being tightly correlated with temperature).
 \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis).
@@
 \end{figure}
-%Move this to the methods%
- To explore the suitability of different environmental conditions for each taxon, we modeled the likelihood of taxon presence as a function of area and season (random effects) and temperature and pH (fixed effects).
-Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}.
-Using predicted conditions for Puget Sound for the year 2095 \cite{khangaonkar2019salish}, we then projected the probability of presence of each taxon across these conditions over 100 draws of the posterior probabilities of all taxa.
-We used these draws to 1) show the change in the probability of presence of each taxon for each scenario (i.e.
-2017 vs 2095, Fig.3 top panel); 2) calculate the  taxon richness within taxonomic groups across the pH-temperature continuum (Fig.
-3 center panel),% 2) show the consequent variation in richness for higher taxa across the pH - temperature continuum (Figure 3B),
-  and 3) visualized the change in suitability for each taxon in our dataset individually (selected taxa in Fig. 3 bottom panel).
+
+\noindent\underline{Climate envelopes and future distributions}
+
+To explore the suitability of different environmental conditions for each taxon, we modeled the likelihood of taxon presence as a function of temperature, pH, and geographic Area as described in the Methods.
+Salinity was not informative in our models, as it is highly correlated with temperature in our dataset and is moreover predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}.
+Model projections let us show the change in the probability of presence of each individual taxon for 2095 vs 2017 (Fig.
+3 A.1, B.1, C.1), estimate richness for larger taxonomic grounds as a whole for these two climates (Fig.
+3 A.2, B.2, C.2), and estimate richness within taxonomic groups across the pH-temperature continuum (Fig.
+3 A.3, B.3, C.3).
+
+{\color{red} RAMON - I edited everything above this point. From here down, the text all needs updating to reflect the results in the current figures. }
 

### 2020-09-04 Manuscript/What.lies.ahead.reviewed.tex
Waiting for Wilcoxon

@@
  Each biological cluster (colored hulls, Fig.
-2.A) occupied a unique area of environmental parameter space.
+2A) occupied a unique area of environmental parameter space.
 Planktonic communities therefore varied predictably with water temperature, salinity, and pH, across a range of those parameters likely to be encountered in many near-term future-ocean scenarios \cite{murray2015inland}.
@@
 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls.
-For example, by 2095 we expect summer in the Hood Canal to be T $>$ 12 and pH $<$ 8.2 -- where community 3 is the most likely community -- to occur more than two-thirds of the time (68\%).
-{\color{red} RAMON -- can you see if this 68\% stat is still true, and also if you can incorporate the similar expectations that are commented-out directly below, updating their \% if necessary?}  \\
+For example, by 2095 we expect summer in the Hood Canal to be T $>$ 12 and pH $<$ {\color{red} 7.9  \st{8.2}} -- where community 3 is the most likely community -- to occur more than {\color{red} four-fifths \st{two-thirds}} of the time ({\color{red} 83\%  \st{68\%}}).
+{\color{red} RAMON -- can you see if this 68\% stat is still true, and also if you can incorporate the similar expectations that are commented-out directly below, updating their \% if necessary?} RYAN: It is a tricky sentence as we no longer estimate seasons independently.
+Cluster 3 is now mode prevalent on pH < 7.8 and Temperature > 12.5; and that condition appears now 83 \% of the time in our simulation.
+\\
 
@@
 \centering
-\includegraphics[width=1\linewidth]{	Figures_for_MS/Figure2.2020-08-27.png}
+\includegraphics[width=1\linewidth]{	Figures_for_MS/Figure2.2020-09-03.png}
 \caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal.
@@
 
-Diatoms show the steepest richness decline under future conditions (Figure 3); the probability of occurrence decreases markedly for diverse diatom taxa including both pennate (e.g., \textit{Thalassionema}, \textit{Navicula}, \textit{Nitzschia}) and centric (e.g., \textit{Coscinodiscus}, \textit{Ditylum}) body forms.
-These declines in diatom richness were more accentuated at lower pH values, both in the warm and colder seasons.
-Other taxa likely to find less-suitable habitat in the future include \textit{Karlodinium} (dinoflagellate; San Juan Island), and \textit{Phaeocystis} and red-tide-associated \textit{Chattonella} (haptophyte and raphidophyte, respectively; Hood Canal).
+Diatoms show the steepest richness decline under future conditions (Figure 3 B.2); the probability of occurrence decreases markedly for diverse diatom taxa including both pennate (e.g., {\color{red} \st{\textit{Thalassionema}, \textit{Navicula},}} \textit{Nitzschia} ) and centric (e.g., \textit{Coscinodiscus},  {\color{red} \textit{Stephanodiscus}}) body forms.
+These declines in diatom richness were more accentuated at lower pH values{\color{red} and higher temperatures}.
+Other taxa likely to find less-suitable habitat in the future include the dinoflagellates {\color{red}\textit{Karlodinium} and \textit{Gymnodiniaceae} sp.\st{ and red-tide-associated \textit{Chattonella} (haptophyte and raphidophyte, respectively; Hood Canal).} }
 
 Likely winners under future conditions are more widely scattered among higher taxonomic groups.
-The coccolithophore \textit{E.
-huxleyi} and the dinoflagellate \textit{Alexandrium sp.} each finds more suitable habitat in both of our study areas.
-Among others, \textit{Chaetoceros} (diatom) and two small hydrozoans (Cnidaria) likely increase in San Juan Island, and the potentially fish-killing heterokont flagellate \textit{Pseudochattonella} increases in the warmer Hood Canal.
+The haptophytes \textit{E.
+huxleyi} and \textit{Chrysochromulina} sp.
+the dinoflagellate \textit{Alexandrium sp.} each finds more suitable habitat in both of our study areas.
+Among others, \textit{Chaetoceros} (diatom) and {\color{red} many} hydrozoans (Cnidaria) likely increase in San Juan Island, and the potentially fish-killing heterokont flagellate \textit{Pseudochattonella} increases in {\color{red} both study areas}.
 See Supporting Information for a complete list of taxon-specific projections.
-
+{\color{red}RYAN - I finished here today, will talk to you ab Wilcoxon} 
 Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at richness variation within phyla; we find no change in median richness for the year 2095 relative to the present (ANOVA; taxon richness within phyla by year, p $=$ 0.3).
@@
  \centering
- \includegraphics[width=1\linewidth]{Figures_for_MS/Top_fig3_2020-08-27.Other.sel.png}
+ \includegraphics[width=1\linewidth]{Figures_for_MS/Top_fig3_2020-09-03.Other.sel.png}
 \caption{\footnotesize{Forecasted changes in plankton in the Salish Sea.

### 2020-09-04 Scripts/Manuscript.Main.Analysis.Rmd
Waiting for Wilcoxon

@@
 
+And keep some that are more abundant on each cluster too
+
 ```{r}
 taxa.to.follow %>% 
-  filter (p<0.01, average > 0.01) %>%  
+  filter (average > 0.01) %>%  
   arrange(desc(average)) %>% 
-  group_by(Area, Comparison) %>% 
+  group_by(Area, Comparison, ava > avb) %>% 
     slice(1:5) %>%
@@
     ggplot(.x, aes(y = average,
-             x = fct_reorder(label2, average))) +
+             x = fct_reorder(label2, phylum))) +
   geom_col(aes(fill = diff>0), size = 0.5, color = "black") +
@@
 pmap(.l = list(Suppl.Material.plots.final, paths),
-     .f = function(a,b) a +  ggsave(filename = here("Figures","Supplementary", b),
+     .f = function(a,b) a + ggsave(plot = a,
+                               filename = here("Figures","Supplementary", b),
                                     device = "png",
@@
 ```
+
+#### Bonus track 2: changes in richness, tested
+
+```{r}
+input.for.left %>% 
+  nest(-phylum) %>% 
+  #  slice(1) %>%
+    mutate(data = map (data, ~ .x %>% 
+                         )) %>% unnest(data)
+                         pivot_longer(cols = c(-taxa, -sim.row),
+                                      names_to = "draw.row",
+                                      values_to = "presence")  %>% 
+                         group_by(draw.row, sim.row) %>% 
+                         summarise(tot = sum(presence)) %>% 
+                         group_by(sim.row) %>% 
+                         summarise_at(.vars = "tot", .funs = list(Richness = mean,sd = sd)) %>% 
+  left_join(all.sims.for.projection %>%  
+              rownames_to_column("sim.row")) ))%>%
+
+   mutate(lm = map(data, ~ lm(Richness ~ Year, data = .x)),
+         lm.results = map (lm, broom::tidy))  -> results.lm
+
+### Using a wilcox test to see if there is a difference between Years
+
+results.lm %>% 
+  mutate(wilcoxon = map (data, ~ wilcox.test(Richness ~ Year, data = .x) ),
+         tidy.w   = map (wilcoxon, broom::tidy)) -> results.lm

### 2020-09-04 Manuscript/Response.to.reviewers.tex

@@
 
-We have now reorganized the manuscript sections following a standard structure.
-As a result, a few sentences in the results section that introduced the procedures follow to model and project the likelihood of taxon presence have now been merged  into the Methods section:
-
-line 107: Within a geographic area and season, temperature and pH were most strongly associated with community-level changes.
-We therefore developed a hierarchical logistical regression model relating the presence of each taxon to temperature and pH (as fixed effects) {\color{red} -salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}.}
+We have now reorganized the manuscript sections following a more standard structure.
+As a result, a few sentences in the results section that introduced the procedures follow to model and project the likelihood of taxon presence have now been merged into the Methods section, and the whole of the document should now reflect a better distinction between Methods and Results.
 
@@
 
-We now include in the Introduction references to previous efforts to use eDNA metabarcoding for community composition changes assessment linked to environmental gradients, and biomonitoring of marine fauna in general
-
-Line 78: {\color{red}A growing body of evidence backs up the efficacy of eDNA metabarcoding for monitoring biodiversity (see a review in \cite{ruppert2019past}), and this approach has been successfully used to detect community composition variation across environmental changes in aquatic  \cite{emilson2017dna}, estuarine \cite{chariton2015metabarcoding, lallias2015environmental}, and marine ecosystems \cite{berry2019marine}}
+We now include in the Introduction references to previous efforts to use eDNA metabarcoding for community composition changes assessment linked to environmental gradients, and biomonitoring of marine fauna in general.
 
@@
 
-It was the intent of the authors to have raw data and scripts available for the reviewers and the readership when published.
-We have now created an open github repository with all the code and the raw fastq files are in a google drive link.
+We had intended to strike a balance between readability (on one hand) and details that are important to specialists (on the other).
+Plainly, the first version of the manuscript erred too far on the side of brevity, and we have now included many such details in the main text.
+
+Moreover, due to a technical error, the raw data and scripts were not available to reviewers in the first submission -- our sincere apologies.
+All data and scripts are now available in an open github repository (raw fastq files are in a google drive link, because of their size).
+We have strived for transparency and repeatability throughout.
 
@@
 
-The reviewer is right in pointing out that the dataset used here is limited to  planktonic taxa.
-We have reworded our title, which now reads {\color{red} Environmental DNA Metabarcoding Reveals Winners and Losers of Global Change in Coastal Plankton}
+The reviewer is correct in pointing out that the dataset used here is limited to  planktonic taxa, but we worry that including `plankton' in the title would unduly limit the appeal of a paper that is aimed at a more general audience.
+We have tried to be clear in the main text which taxa are, and are not, under study here.
 
@@
 
-We agree with the reviewer and we have rethought and redesigned figure 3 to highlight the taxa that shows meaningful shifts - focusing on the most varying diatoms and dinoflagellates we reveal the magnitude and repercussion of the changes in suitability.
-Panel C has been moved to the supplementary material
+We agree with the reviewer and we have rethought and redesigned figures 2 and 3 for simplicity and impact.
+
+In fact, this reviewer's concern -- which another reviewer shared -- led us to simplify our analysis in general, dropping a seasonality term from our model.
+This helped clarify the core findings, and we are grateful for the feedback.
 
@@
 \textbf{Response:}
-    We realize that the shared color scales can lead to confusion. This has now been fixed, and we employ now a different color scales for temporal and spatial variables.
+
+    Among many other revisions to our figures, this has now been fixed, and we employ now a different color scales for temporal and spatial variables.
 
@@
 
-\item{Lines 74-76 ?with different points along the environmental gradient simultaneously showing differences equivalent to those predicted between present-day and future oceans?: This seems like an important point that perhaps should be included in the Introduction.}
+\item{Lines 74-76: with different points along the environmental gradient simultaneously showing differences equivalent to those predicted between present-day and future oceans?: This seems like an important point that perhaps should be included in the Introduction.}
 
@@
 
-Agreed.
-The advantage of the Hood Canal in this regard is paramount, and we now have included this concept in the introduction.
-Line 52 onwards: {\color{red} [in close geographic proximity].
-This gradients are pronounced and show values of pH and temperature expected worldwide by the end of the century}.
+Agreed. The advantage of the Hood Canal in this regard is paramount, and we now have included this concept in the introduction and several times in the main text.
 
@@
 
-TODO
+Done.
 
@@
 
-We think the relationship between clusters and the environmental variables are important in a few aspects.
+In sum, yes: these clusters are coherent groups of species and predictably occur in different environmental conditions. We have revised the main text to be more clear in this regard.
+
+We think the relationship between clusters and the environmental variables is important in a few aspects.
 First, as a means of ground truthing our findings, it has great value as the clusters were detected blind to the environmental data, and the CAP analysis return the distinctiveness of these clusters in that space - Figure 2B highlights the translation of those modified axis of the CAP plot into the real values of temperature and pH.
-Second, these clusters (maybe communities is a more committed term that requires a longer dataset in which to confirm co-ocurrence of members) are a combination of species that are unique to them plus changes in relative presence - not a clear cut between total abence of a species but definitely a gradient in the relative importance of the species present.
-We have reworded the section
+Second, these clusters (maybe communities is a more committed term that requires a longer dataset in which to confirm co-ocurrence of members) do indeed reflect changes in the relative importance of the species present.
 
-TODO
+{\color{red} RAMON -- if we should use `cluster' instead of `community' we will need to change that in the main text. }
 
@@
 
+{\color{red} TODO - check }
+
 \item{Line 112-117: The authors base much of their calculation on a set of simulated temperature and pH values for different regions in the Salish Sea in 2095.
@@
 (2019) (see comments on the Methods section as well).}
+
+We have now simplified our approach to these data in an effort to create a more straightforward and transparent analysis, by merely linking the observed (2017) and projected (2095) means in a linear model, assuming a constant variance (observed) over time.
+The underlying model data for Khangaonkar et al.
+(2019) are not available -- we based our analysis only on their published paper.
+
+Clearly, climate science and oceanography are massively complex undertakings with many and important subtleties, some of which we have glossed over here.
+Our purpose was not to generate a definitive climate projection, but rather to use a readily available estimate of the magnitude and directions of expected change to give a sense of how communities are likely to change.
 
@@
 
+Done.
+
 \item{Line 121: Refer to the relevant section of Figure 3 (A, B, or C)}
+
+Done.
 
@@
 
+{\color{red} TODO - check }
+
 \item{Line 122-123: According to Figure 3B, Thalassionema, Navicula, and Nitzschia appear to refer specifically to San Juan. If so, please state this.}
 
+{\color{red} TODO - check }
+
 \item{Line 123: Coscinodiscus and Ditylum seem to refer to Hood Canal. This ought to specified as before.}
+
+{\color{red} TODO - check }
 
@@
 
+{\color{red} TODO - check }
+
 \item{Line 136-137: This is minor, but this section (or methods) might benefit from a brief mention of why ANOVA was used in one case and Wilcoxon in another.}
+
+{\color{red} TODO - check }
 
@@
 
+{\color{red} TODO - check }
+
 \item{Line 147-148: Need a citation for Coscinodiscus as key fish food.}
+
+{\color{red} TODO - check }
 
@@
 
+{\color{red} TODO - check }
+
 \item{Line 192: Might be helpful to specifically mention carbon sequestration as an important aspect of sinking rates.}
+
+{\color{red} TODO - check }
 
@@
 
+{\color{red} TODO - check }
+
 \item{Line 208-214: This is a big aspect of why this study is valuable! I think it?s worth discussing this in the introduction.}
+
+{\color{red} TODO - check }
 
@@
 
+{\color{red} TODO - check }
+
 \item{Line 241: Should be Figure 1C}
+
+{\color{red} TODO - check }
 
@@
 
+{\color{red} TODO - check }
+
 \item{Line 273-282: It?s hard to discern based on the limited code and data provided exactly how the authors subsampled the different subregions of Puget Sound, since the paper they base their data on has overall values for the whole area.
 I think this section could use a bit of clarification, and all the code and raw data should be given (the provided github link is broken).}
+
+{\color{red} TODO - check }
 

### 2020-09-04 Manuscript/What.lies.ahead.reviewed.tex
CHanges in richness with time

@@
 {\color{red}RYAN - I finished here today, will talk to you ab Wilcoxon} 
-Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at richness variation within phyla; we find no change in median richness for the year 2095 relative to the present (ANOVA; taxon richness within phyla by year, p $=$ 0.3).
-Diatoms, however, show consistent and significant declines in richness (Wilcoxon p $< 10^{-15}$).
+Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at richness variation within phyla; we find no change in median richness for the year 2095 relative to the present in the Hood Canal(ANOVA; taxon richness  by year, p $=$ 0.7).
+Diatoms, however, show consistent and significant declines in richness in both Areas (Wilcoxon p $< 10^{-16}$).
 

### 2020-09-04 Scripts/Manuscript.Main.Analysis.Rmd
CHanges in richness with time

@@
 
+Using a global ANOVA to see if there are differences between Years (2095-2017) in both Areas. Then a  wilcoxon test to see if there is a difference  on each phylum
 ```{r}
+
+## GLOBAL TEST
+
 input.for.left %>% 
-  nest(-phylum) %>% 
+  mutate(sim.row = rep(1:10873, n_distinct(taxa))) -> input.richness
+
+input.richness %>% 
+  nest(-Area) %>% 
+  mutate(data = map (data, ~ .x %>% 
+                         group_by(Year,  sim.row) %>% 
+                         summarise( Richness = sum(present)) ),
+         lm = map(data, ~ lm(Richness ~ Year, data = .x)),
+         tidy_lm = map(lm, broom::tidy)) %>%
+  unnest(tidy_lm)
+
+input.richness %>% 
+  nest(-phylum,) %>% 
   #  slice(1) %>%
     mutate(data = map (data, ~ .x %>% 
-                         )) %>% unnest(data)
-                         pivot_longer(cols = c(-taxa, -sim.row),
-                                      names_to = "draw.row",
-                                      values_to = "presence")  %>% 
-                         group_by(draw.row, sim.row) %>% 
-                         summarise(tot = sum(presence)) %>% 
-                         group_by(sim.row) %>% 
-                         summarise_at(.vars = "tot", .funs = list(Richness = mean,sd = sd)) %>% 
-  left_join(all.sims.for.projection %>%  
-              rownames_to_column("sim.row")) ))%>%
-
-   mutate(lm = map(data, ~ lm(Richness ~ Year, data = .x)),
-         lm.results = map (lm, broom::tidy))  -> results.lm
-
-### Using a wilcox test to see if there is a difference between Years
-
-results.lm %>% 
-  mutate(wilcoxon = map (data, ~ wilcox.test(Richness ~ Year, data = .x) ),
-         tidy.w   = map (wilcoxon, broom::tidy)) -> results.lm
+                         group_by(Year,  sim.row) %>% 
+                         summarise( Richness = sum(present)) )) %>% 
+   mutate(wilcoxon = map (data, ~ wilcox.test(Richness ~ Year, data = .x) ))  -> results.lm
+
+results.lm %>%  filter (phylum == "Bacillariophyta") %>% pull(wilcoxon) 
+```

### 2020-09-04 Scripts/Manuscript.Main.Analysis.Rmd
Wilcox

@@
   mutate(data = map (data, ~ .x %>% 
+                       mutate(present = case_when(present < 0.5 ~ 0,
+                                                  TRUE          ~ 1)) %>% 
                          group_by(Year,  sim.row) %>% 
                          summarise( Richness = sum(present)) ),
-         lm = map(data, ~ lm(Richness ~ Year, data = .x)),
-         tidy_lm = map(lm, broom::tidy)) %>%
-  unnest(tidy_lm)
+         wilcoxon = map(data, ~ wilcox.test(Richness ~ Year, data = .x, alternative = "greater")),
+         tidy_wilcox = map(wilcoxon, broom::tidy)) %>%
+  pull(wilcoxon)
+
+## We cannot
 
 input.richness %>% 
-  nest(-phylum,) %>% 
+  nest(-phylum,-Area) %>% 
   #  slice(1) %>%
     mutate(data = map (data, ~ .x %>% 
+                         mutate(present = case_when(present < 0.5 ~ 0,
+                                                  TRUE          ~ 1)) %>% 
                          group_by(Year,  sim.row) %>% 
                          summarise( Richness = sum(present)) )) %>% 
-   mutate(wilcoxon = map (data, ~ wilcox.test(Richness ~ Year, data = .x) ))  -> results.lm
+   mutate(wilcoxon = map (data, ~ wilcox.test(Richness ~ Year, data = .x, alternative = "greater") ))  -> results.lm
 

### 2020-09-04 Manuscript/What.lies.ahead.reviewed.tex
Wilcoxon with proper richness

@@
 For each point in the pH - temperature grid, we calculated a species' probability of presence as the mean of 100 independent draws of the posterior model response.
-We estimated species richness as the sum of taxa with $>$ 50\% probability of occurrence at a given set of environmental parameters.
+{\color{red} For richness estimations in future scenarios, we used each of the 100 draws of the posteriors in each environmental point as a sample and tallied the number of species detected on each draw.
+The mean value across the 100 draws was the input for a Wilcoxon test of differences in species richness between the 2017 and 2095 scenarios in each region.
+We performed the Wilcoxon test globally (total species richness) and on each phylum independently }
 

### 2020-09-04 Scripts/Manuscript.Main.Analysis.Rmd
Wilcoxon with proper richness

--- 2020-09-04 Scripts/Manuscript.Main.Analysis.Rmd
@@
 
-Using a global ANOVA to see if there are differences between Years (2095-2017) in both Areas. Then a  wilcoxon test to see if there is a difference  on each phylum
+First calculate the richness in each point as following.
+In each environmental point, the 100 draws represent if a taxa was found or not.
+We sum the draws of all taxa so we have 100 estimation of richness for each point - we then take the mean of those 100 draws and we have a estimation of richness at that point.
+
+```{r get the richness per phylum}
+
+split.matrices.by.taxa <- list()
+
+for (i in 1:n_distinct(expanded.future$taxa)){
+  start.col = 1 + (i-1)*10873
+  fin.col = start.col + 10872
+
+  split.matrices.by.taxa[[i]] <- posterior.predict.all.taxa[,start.col:fin.col]
+  }
+
+richness.fun <- function(x) {
+    Reduce('+', x) 
+    }
+
+distinct(expanded.future,taxa) %>% 
+  mutate(draws = split.matrices.by.taxa) %>% 
+  left_join(phylum.info %>% select(taxa, phylum)) %>% 
+  group_by(phylum) %>% 
+  summarise(draws =  Reduce('+',draws)) -> richness.by.draw.by.phylum.by.point
+
+ richness.by.draw.by.phylum.by.point %>% 
+   nest() %>% 
+   mutate(data = map (data, function(.x) colMeans(as.matrix(.x)) %>% 
+                        as_tibble %>% 
+                        rownames_to_column("sim.row"))) %>% 
+   unnest()-> richness.by.phylum.by.point
+
+ expanded.future %>%
+   ungroup() %>% 
+   slice (1:10873) %>% 
+   rownames_to_column("sim.row") %>% 
+   select(Area, Year, sim.row) %>% 
+   right_join(richness.by.phylum.by.point) -> richness.by.phylum.by.point
+
+```
+
+Now the Global - richness by point
+
 ```{r}
+distinct(expanded.future,taxa) %>% 
+  mutate(draws = split.matrices.by.taxa) %>% 
+  summarise(draws =  Reduce('+',draws)) -> richness.by.draw.by.point
+
+richness.by.draw.by.point %>% 
+   nest() %>% 
+   mutate(data = map (data, function(.x) colMeans(as.matrix(.x)) %>% 
+                        as_tibble %>% 
+                        rownames_to_column("sim.row"))) %>% 
+   unnest()-> richness.by.point
+
+expanded.future %>%
+   ungroup() %>% 
+   slice (1:10873) %>% 
+   rownames_to_column("sim.row") %>% 
+   select(Area, Year, sim.row) %>% 
+   right_join(richness.by.point) -> richness.by.point
+```
+
+Because we are comparing two 
+```{r}
 
@@
 
-input.for.left %>% 
-  mutate(sim.row = rep(1:10873, n_distinct(taxa))) -> input.richness
-
-input.richness %>% 
+richness.by.point %>% 
   nest(-Area) %>% 
-  mutate(data = map (data, ~ .x %>% 
-                         group_by(Year,  sim.row) %>% 
-                         summarise( Richness = sum(present)) ),
-         lm = map(data, ~ lm(Richness ~ Year, data = .x)),
-         tidy_lm = map(lm, broom::tidy)) %>%
-  unnest(tidy_lm)
-
-input.richness %>% 
-  nest(-phylum,) %>% 
-  #  slice(1) %>%
-    mutate(data = map (data, ~ .x %>% 
-                         group_by(Year,  sim.row) %>% 
-                         summarise( Richness = sum(present)) )) %>% 
-   mutate(wilcoxon = map (data, ~ wilcox.test(Richness ~ Year, data = .x) ))  -> results.lm
+  mutate(wilcoxon = map(data, ~ wilcox.test(value ~ Year, data = .x)),
+         tidy_wilcox = map(wilcoxon, broom::tidy)) %>%
+  pull(wilcoxon)
+
+## We cannot
+
+richness.by.phylum.by.point %>% 
+   nest(-phylum,-Area) %>% 
+   mutate(wilcoxon = map (data, ~ wilcox.test(value ~ Year, data = .x) ))  -> results.lm
 

### 2020-09-04 Manuscript/What.lies.ahead.reviewed.tex
Added Fig3 captio and results of Wilcox

@@
 See Supporting Information for a complete list of taxon-specific projections.
-{\color{red}RYAN - I finished here today, will talk to you ab Wilcoxon} 
-Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at richness variation within phyla; we find no change in median richness for the year 2095 relative to the present in the Hood Canal(ANOVA; taxon richness  by year, p $=$ 0.7).
-Diatoms, however, show consistent and significant declines in richness in both Areas (Wilcoxon p $< 10^{-16}$).
+
+Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at richness variation within phyla; {\color{red} we find no change in median richness for the year 2095 relative to the present in the Hood Canal(taxon richness  by year, Wilcoxon p $=$ 0.7).
+Diatoms, however, show consistent and significant declines in richness in that region (Wilcoxon p $< 10^{-16}$).
+}
 
@@
  \includegraphics[width=1\linewidth]{Figures_for_MS/Top_fig3_2020-09-03.Other.sel.png}
-\caption{\footnotesize{Forecasted changes in plankton in the Salish Sea.
-\textbf{A} Probability densities for the occurrence of individual taxa (species- and genus-level) for 2017 (blue) and 2095 (red); data are mean probabilities over 100 model draws, and variance in probability is due to differences in underlying environmental conditions.
-Four prominent higher-taxa are shown: Arthropods, Dinoflagellates (Dinophyceae), Diatoms (Bacillariophyta), and Haptophytes.
-\textbf{B} Individual taxa with the largest changes in probabilities of occurrence between 2017 and 2095, color-coded by higher taxonomic group, for Hood Canal (left) and San Juan Island (right).
-Median probabilities plotted.
-\textbf{C} Relative taxon richness (raster color, warmer colors are more taxon-rich) for each of these same higher-taxa, for plausible ranges of pH and Temperature.
+\caption{\footnotesize{Forecasted changes in plankton in the Salish Sea {\color{red} for Dinoflagellates (panel \textbf{A}), Diatoms (\textbf{B}) and Haptophytes (\textbf{C}).
+Each panels shows:  (\textbf{.1}) probability densities for the occurrence of selected taxa (species- and genus-level) for 2017 (blue) and 2095 (red); data are mean probabilities over 100 model draws, and variance in probability is due to differences in underlying environmental conditions.
+(\textbf{.2}) Changes in relative species richness within each phylum across the simulated scenario.
+(\textbf{.3}) Relative taxon richness (raster color, warmer colors are more taxon-rich) for each of these same higher-taxa, for plausible ranges of pH and Temperature.
 Envelopes of observed (2017, blue) and modeled (2095, red) annual conditions in the Salish Sea shown for reference.
-Hood Canal and San Juan Island plotted separately to illustrate environmental differences between them.}}
+Hood Canal and San Juan Island plotted separately to illustrate environmental differences between them.}}}
 

### 2020-09-04 Manuscript/What.lies.ahead.reviewed.tex
replied to comments

@@
 For each point in the pH - temperature grid, we calculated a species' probability of presence as the mean of 100 independent draws of the posterior model response.
-{\color{red} For richness estimations in future scenarios, we used each of the 100 draws of the posteriors in each environmental point as a sample and tallied the number of species detected on each draw.
+For richness estimations in future scenarios, we used each of the 100 draws of the posteriors in each environmental point as a sample and tallied the number of species detected on each draw.
 The mean value across the 100 draws was the input for a Wilcoxon test of differences in species richness between the 2017 and 2095 scenarios in each region.
-We performed the Wilcoxon test globally (total species richness) and on each phylum independently }
+We performed the Wilcoxon test globally (total species richness) and on each phylum independently.
 
@@
 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls.
-For example, by 2095 we expect summer in the Hood Canal to be T $>$ 12 and pH $<$ {\color{red} 7.9  \st{8.2}} -- where community 3 is the most likely community -- to occur more than {\color{red} four-fifths \st{two-thirds}} of the time ({\color{red} 83\%  \st{68\%}}).
+For example, by 2095 we expect summer in the Hood Canal to be T $>$ 12\degree C and pH $<$ {\color{red} 7.9  \st{8.2}} -- where community 3 is the most likely community -- to occur more than {\color{red} four-fifths \st{two-thirds}} of the time ({\color{red} 83\%  \st{68\%}}).
 {\color{red} RAMON -- can you see if this 68\% stat is still true, and also if you can incorporate the similar expectations that are commented-out directly below, updating their \% if necessary?} RYAN: It is a tricky sentence as we no longer estimate seasons independently.
 Cluster 3 is now mode prevalent on pH < 7.8 and Temperature > 12.5; and that condition appears now 83 \% of the time in our simulation.
-\\
+{\color{blue} RAMON - ah, yes -- forgot about seasons.
+Could we just have some analogous thing where we say what fraction of the time we expect each community, at each area, in 2017 vs 2095?.} \\
 
@@
 
-{\color{red} RAMON - I edited everything above this point. From here down, the text all needs updating to reflect the results in the current figures. }
-
-Diatoms show the steepest richness decline under future conditions (Figure 3 B.2); the probability of occurrence decreases markedly for diverse diatom taxa including both pennate (e.g., {\color{red} \st{\textit{Thalassionema}, \textit{Navicula},}} \textit{Nitzschia} ) and centric (e.g., \textit{Coscinodiscus},  {\color{red} \textit{Stephanodiscus}}) body forms.
-These declines in diatom richness were more accentuated at lower pH values{\color{red} and higher temperatures}.
-Other taxa likely to find less-suitable habitat in the future include the dinoflagellates {\color{red}\textit{Karlodinium} and \textit{Gymnodiniaceae} sp.\st{ and red-tide-associated \textit{Chattonella} (haptophyte and raphidophyte, respectively; Hood Canal).} }
+Diatoms (Bacillariophyta) show the steepest richness decline under future conditions (Figure 3 B.2); the probability of occurrence decreases markedly for diverse diatom taxa including both pennate (e.g., \textit{Nitzschia}) and centric (e.g., \textit{Coscinodiscus}, \textit{Stephanodiscus}) body forms.
+These declines in diatom richness were more accentuated at lower pH values and higher temperatures.
+Other taxa likely to find less-suitable habitat in the future include dinoflagellates \textit{Karlodinium} and Gymnodiniaceae sp.
 
@@
 the dinoflagellate \textit{Alexandrium sp.} each finds more suitable habitat in both of our study areas.
-Among others, \textit{Chaetoceros} (diatom) and {\color{red} many} hydrozoans (Cnidaria) likely increase in San Juan Island, and the potentially fish-killing heterokont flagellate \textit{Pseudochattonella} increases in {\color{red} both study areas}.
+Among others not shown in Fig 3, \textit{Chaetoceros} (diatom) and many hydrozoans (Cnidaria) likely increase in San Juan Island, and the potentially fish-killing heterokont flagellate \textit{Pseudochattonella} increases in both study areas.
 See Supporting Information for a complete list of taxon-specific projections.
 
-Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at richness variation within phyla; {\color{red} we find no change in median richness for the year 2095 relative to the present in the Hood Canal(taxon richness  by year, Wilcoxon p $=$ 0.7).
+Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at richness variation within phyla; {\color{red} we find no change in median richness for the year 2095 relative to the present in the Hood Canal (taxon richness by year, Wilcoxon p $=$ 0.7).
 Diatoms, however, show consistent and significant declines in richness in that region (Wilcoxon p $< 10^{-16}$).
-}
+} {\color{blue} RAMON - can you clarify here?
+there should be comparisons for A.
+2017 vs 2095 in overall richness, for both HC and SJI.
+Then B.
+same for Diatom richness, dinoflagellate richness, or anything else you find interesting.}
 

### 2020-09-04 Scripts/Manuscript.Main.Analysis.Rmd
replied to comments

@@
 ```{r}
-distinct(expanded.future,taxa) %>% 
-  mutate(draws = split.matrices.by.taxa) %>% 
-  summarise(draws =  Reduce('+',draws)) -> richness.by.draw.by.point
-
-richness.by.draw.by.point %>% 
-   nest() %>% 
-   mutate(data = map (data, function(.x) colMeans(as.matrix(.x)) %>% 
-                        as_tibble %>% 
-                        rownames_to_column("sim.row"))) %>% 
-   unnest()-> richness.by.point
+# distinct(expanded.future,taxa) %>% 
+#   mutate(draws = split.matrices.by.taxa) %>% 
+#   summarise(draws =  Reduce('+',draws)) -> richness.by.draw.by.point
+# 
+# richness.by.draw.by.point %>% 
+#    nest() %>% 
+#    mutate(data = map (data, function(.x) colMeans(as.matrix(.x)) %>% 
+#                         as_tibble %>% 
+#                         rownames_to_column("sim.row"))) %>% 
+#    unnest()-> richness.by.point
+
+#overall change in richness
+expanded.future %>% 
+  group_by(Year, Temperature, pH, Area) %>% 
+  summarise(richness = sum(present)) %>% 
+  group_by(Year, Area) %>% 
+  summarise(median(richness))
+#diatom change in richness
+expanded.future %>% 
+  filter(phylum == "Bacillariophyta") %>% 
+  group_by(Year, Temperature, pH, Area) %>% 
+  summarise(richness = sum(present)) %>% 
+  group_by(Year, Area) %>% 
+  summarise(median(richness))
 

### 2020-09-05 Scripts/Manuscript.Main.Analysis.Rmd
Wilcoxon made easy

--- 2020-09-04 Scripts/Manuscript.Main.Analysis.Rmd
@@
   select(Area, Year, Temperature = Temp.std, TemperatureCelsius, pH = pH.std, pH_units) 
-
+nsims <- nrow(future.scenario)
 ## Get
@@
 ```{r Fig3left}
-
 expanded.future %>% 
@@
 input.for.left %>% 
-   group_by(phylum, Temperature, pH, Year, Area ) %>%
+  group_by(taxa) %>% 
+  nest %>% 
+  mutate(data = map (data, ~.x %>% 
+                       rownames_to_column("sim.row"))) %>% 
+  unnest -> input.for.left
+
+```
+
+```{r Fig3left botto}
+input.for.left %>%
+   group_by(phylum, sim.row,Temperature, pH, Year, Area ) %>%
   summarise(present = sum(present)) %>%
@@
   select(-data)-> summary.plot
-
+```
+
+```{r Fig3left top}
 input.for.left %>% 
@@
 full.grid$present <- colMeans(posterior.predict.all.taxa)
-
-full.grid$sd <- apply(posterior.predict.all.taxa,2,sd )
 
@@
 
-Using a global ANOVA to see if there are differences between Years (2095-2017) in both Areas. Then a  wilcoxon test to see if there is a difference  on each phylum
+First calculate the richness in each point as following.
+In each environmental point, the 100 draws represent if a taxa was found or not.
+We sum the draws of all taxa so we have 100 estimation of richness for each point - we then take the mean of those 100 draws and we have a estimation of richness at that point.
+
+```{r get the richness per phylum}
+input.for.left %>%
+   group_by(phylum, sim.row,Temperature, pH, Year, Area ) %>%
+  summarise(richness = sum(present)) -> richness.by.phylum.by.point
+
+# And from here - make the global
+
+richness.by.phylum.by.point %>% 
+  group_by (Area,Year, sim.row) %>% 
+  summarise(richness = sum(richness)) -> richness.by.point
+
+```
+
+Now the Global - richness by point
+
+```{r}
+
+#overall change in richness
+richness.by.point %>%
+  group_by(Year, Area) %>%
+  summarise(median(richness))
+#diatom change in richness
+richness.by.phylum.by.point %>%
+  filter(phylum == "Bacillariophyta") %>%
+  group_by(Year, Area) %>%
+  summarise(median(richness))
+```
+
+Because we are comparing two 
 ```{r}
@@
 
-input.for.left %>% 
-  mutate(sim.row = rep(1:10873, n_distinct(taxa))) -> input.richness
-
-input.richness %>% 
+richness.by.point %>% 
   nest(-Area) %>% 
-  mutate(data = map (data, ~ .x %>% 
-                       mutate(present = case_when(present < 0.5 ~ 0,
-                                                  TRUE          ~ 1)) %>% 
-                         group_by(Year,  sim.row) %>% 
-                         summarise( Richness = sum(present)) ),
-         wilcoxon = map(data, ~ wilcox.test(Richness ~ Year, data = .x, alternative = "greater")),
-         tidy_wilcox = map(wilcoxon, broom::tidy)) %>%
-  pull(wilcoxon)
+  mutate(wilcoxon = map(data, ~ wilcox.test(richness ~ Year, data = .x)),
+         tidy_wilcox = map_df(wilcoxon, broom::tidy))
 
@@
 
-input.richness %>% 
+richness.by.phylum.by.point %>% 
    nest(-phylum,-Area) %>% 
-  #  slice(1) %>%
-    mutate(data = map (data, ~ .x %>% 
-                         mutate(present = case_when(present < 0.5 ~ 0,
-                                                  TRUE          ~ 1)) %>% 
-                         group_by(Year,  sim.row) %>% 
-                         summarise( Richness = sum(present)) )) %>% 
-   mutate(wilcoxon = map (data, ~ wilcox.test(Richness ~ Year, data = .x, alternative = "greater") ))  -> results.lm
-
-results.lm %>%  filter (phylum == "Bacillariophyta") %>% pull(wilcoxon) 
-```
+   mutate(wilcoxon = map (data, ~ wilcox.test(richness ~ Year, data = .x) ),
+          wiloc.tidy= map_df(wilcoxon, broom::tidy))  -> results.lm
+
+results.lm %>%  filter (phylum == "Bacillariophyta")
+```

### 2020-09-05 Manuscript/What.lies.ahead.reviewed.tex
Wilcox results done

@@
 
-Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at richness variation within phyla; {\color{red} we find no change in median richness for the year 2095 relative to the present in the Hood Canal (taxon richness by year, Wilcoxon p $=$ 0.7).
-Diatoms, however, show consistent and significant declines in richness in that region (Wilcoxon p $< 10^{-16}$).
-} {\color{blue} RAMON - can you clarify here?
+Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at overall richness variation; {\color{red} we find no change in median richness for the year 2095 relative to the present in the Hood Canal (taxon richness by year, Wilcoxon p $=$ 0.7); while higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 2.1, Wilcoxon p $=$ 0).
+Diatoms, however, show consistent and significant declines in richness in the Hood Canal (Wilcoxon p $< 10^{-16}$), while the changes on the San Juan Island are undetectable (p $=$ 0.8).
+Dinoflagellates see their richness increase in both regions with the future scenario(Wilcoxon p $< 10^{-50}$) }.{\color{blue} RAMON - can you clarify here?
 there should be comparisons for A.
@@
 Then B.
-same for Diatom richness, dinoflagellate richness, or anything else you find interesting.}
+same for Diatom richness, dinoflagellate richness, or anything else you find interesting.RYAN - See what you think}
 

### 2020-09-05 Scripts/Manuscript.Main.Analysis.Rmd
Wilcox results done

@@
   select(Area, Year, Temperature = Temp.std, TemperatureCelsius, pH = pH.std, pH_units)
-nsims <- nrow(future.scenario)
+
 ## Get
@@
   summarise(median(richness))
+richness.by.phylum.by.point %>%
+  filter(phylum == "Dinophyceae") %>%
+  group_by(Year, Area) %>%
+  summarise(median(richness))
 ```
@@
    mutate(wilcoxon = map (data, ~ wilcox.test(richness ~ Year, data = .x) ),
-          wiloc.tidy= map_df(wilcoxon, broom::tidy))  -> results.lm
-
-results.lm %>%  filter (phylum == "Bacillariophyta")
-```
+          wiloc.tidy= map(wilcoxon, broom::tidy))  -> results.lm
+
+results.lm %>%  filter (phylum == "Bacillariophyta") %>% unnest(wiloc.tidy)
+
+results.lm %>%  filter (phylum == "Dinophyceae") %>% unnest(wiloc.tidy)
+
+results.lm %>% 
+  unnest(wiloc.tidy) %>%
+
+  ggplot(aes(x = phylum , y = `p.value`)) +
+  geom_point(aes(color = Area))
+```

### 2020-09-06 Manuscript/What.lies.ahead.reviewed.tex
minor changes, still trying to figure out wilcox weirdness and to figure out how to easily calculate fraction of time in the future different enviro conditions occur

--- 2020-09-04 Manuscript/What.lies.ahead.reviewed.tex
@@
 
+Taxa surveyed are a function of our metabarcoding PCR primers \cite{leray_new_2013} and reflect the current status of genetic databases, rather than a complete sampling of the planktonic community; we view these results as a cross-section of common taxa useful for understanding the biological effects of ocean conditions.
+
 We can only model responses of taxa present in our data set.
@@
 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls.
-For example, by 2095 we expect summer in the Hood Canal to be T $>$ 12\degree C and pH $<$ {\color{red} 7.9  \st{8.2}} -- where community 3 is the most likely community -- to occur more than {\color{red} four-fifths \st{two-thirds}} of the time ({\color{red} 83\%  \st{68\%}}).
-{\color{red} RAMON -- can you see if this 68\% stat is still true, and also if you can incorporate the similar expectations that are commented-out directly below, updating their \% if necessary?} RYAN: It is a tricky sentence as we no longer estimate seasons independently.
-Cluster 3 is now mode prevalent on pH < 7.8 and Temperature > 12.5; and that condition appears now 83 \% of the time in our simulation.
-{\color{blue} RAMON - ah, yes -- forgot about seasons.
-Could we just have some analogous thing where we say what fraction of the time we expect each community, at each area, in 2017 vs 2095?.} \\
+For example, by 2095 we expect summer in the Hood Canal to be T $>$ 12\degree C and pH $<$ {7.9} -- where community 3 is the most likely community -- to occur more than {four-fifths} of the time (83\%}).
+{\color{blue} I couldn't find where you were doing this calculation, so I couldn't do the fractions of time we expect the other conditions.
+It would be good to say that we go from condition 1 being XX\% of the time in 2017 to YY\% of the time in 2095.}
+
+\\
 
@@
 
-Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at richness variation within phyla; {\color{red} we find no change in median richness for the year 2095 relative to the present in the Hood Canal (taxon richness by year, Wilcoxon p $=$ 0.7).
-Diatoms, however, show consistent and significant declines in richness in that region (Wilcoxon p $< 10^{-16}$).
-} {\color{blue} RAMON - can you clarify here?
-there should be comparisons for A.
-2017 vs 2095 in overall richness, for both HC and SJI.
-Then B.
-same for Diatom richness, dinoflagellate richness, or anything else you find interesting.}
-
-Taxa surveyed are a function of our metabarcoding PCR primers \cite{leray_new_2013} and reflect the current status of genetic databases, rather than a complete sampling of the planktonic community; we view these results as a cross-section of common taxa useful for understanding the biological effects of ocean conditions.
+Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at overall richness variation; {\color{red} we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, averaged over all environmental conditions, Wilcoxon p $=$ 0.5); while higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 2.1, Wilcoxon p $=$ 0).
+Diatoms, however, show small but significant declines in richness in the Hood Canal (0.5 species, Wilcoxon p $< 10^{-16}$), while the changes on the San Juan Island are undetectable (p $=$ 0.8).
+Dinoflagellates see their richness increase in both regions with the future scenario(Wilcoxon p $< 10^{-50}$) }.
+{\color{blue} I re-did some of these in the Rmd file (starting at line 1129, doing it line-by-line to convince myself what was going on.
+See line 1157 in particular: for diatoms, there is no change between years, but yet the wilcox p value is very low.
+I can't figure out what's happening there.}
 
@@
 %%%%%
+
+{\color{blue} The paragraph below is important, and links to the question I had above: where are we doing this calculation in the code?
+and is the below paragraph still true?
+Plus, a reviewer wanted a citation of the Coscinodiscus statement.}
 

### 2020-09-06 Scripts/Manuscript.Main.Analysis.Rmd
minor changes, still trying to figure out wilcox weirdness and to figure out how to easily calculate fraction of time in the future different enviro conditions occur

@@
   summarise(median(richness))
+
+#we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, averaged over all environmental conditions, Wilcoxon p $=$ 0.5);
+
+wilcox.test(x = richness.by.point %>%
+                  filter(Year == 2017, Area == "Hood Canal") %>% 
+                    pull(richness),
+            y = richness.by.point %>%
+                  filter(Year == 2095, Area == "Hood Canal") %>% 
+                    pull(richness))
+
+#while higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 2.1, Wilcoxon p $=$ 0).
+
+wilcox.test(x = richness.by.point %>%
+                  filter(Year == 2017, Area == "San Juan Island") %>% 
+                    pull(richness),
+            y = richness.by.point %>%
+                  filter(Year == 2095, Area == "San Juan Island") %>% 
+                    pull(richness))
+
+#Diatoms, however, show small but significant declines in richness in the Hood Canal (0.5 species, Wilcoxon p $< 10^{-16}$),
+
+wilcox.test(x = richness.by.phylum.by.point %>%
+                  filter(Year == 2017, Area == "Hood Canal", phylum == "Bacillariophyta") %>% 
+                    pull(richness),
+            y = richness.by.phylum.by.point %>%
+                  filter(Year == 2095, Area == "Hood Canal", phylum == "Bacillariophyta") %>% 
+                    pull(richness))
+
+#while the changes on the San Juan Island are undetectable (p $=$ 0.8).
+
+wilcox.test(x = richness.by.phylum.by.point %>%
+                  filter(Year == 2017, Area == "San Juan Island", phylum == "Bacillariophyta") %>% 
+                    pull(richness),
+            y = richness.by.phylum.by.point %>%
+                  filter(Year == 2095, Area == "San Juan Island", phylum == "Bacillariophyta") %>% 
+                    pull(richness))
+
+#Dinoflagellates see their richness increase in both regions with the future scenario(Wilcoxon p $< 10^{-50}$)
+
 ```

### 2020-09-06 Manuscript/What.lies.ahead.reviewed.tex
intermediate changes for Moncho's review. still working on Results.

@@
 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls.
-For example, by 2095 we expect summer in the Hood Canal to be T $>$ 12\degree C and pH $<$ {7.9} -- where community 3 is the most likely community -- to occur more than {four-fifths} of the time (83\%}).
-{\color{blue} I couldn't find where you were doing this calculation, so I couldn't do the fractions of time we expect the other conditions.
-It would be good to say that we go from condition 1 being XX\% of the time in 2017 to YY\% of the time in 2095.}
+For example, we expect the Hood Canal to be T $>$ 12\degree C and pH $<$ {7.9} -- where community 3 is the most likely community -- more than four times as often in 2095 than in 2017 (4.9 \% vs 20.7\%).
+{\color{blue} I couldn't find where you were doing this calculation, so I made one up -- see lines 1134 and following in the Rmd.
+BUT ALSO : community 3 isn't sensitive to pH, so what are we even saying here?}
 
@@
 
-Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at overall richness variation; {\color{red} we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, averaged over all environmental conditions, Wilcoxon p $=$ 0.5); while higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 2.1, Wilcoxon p $=$ 0).
+Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at overall richness variation; {\color{red} we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, averaged over all environmental conditions, Wilcoxon p $=$ 0.5); while higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 2.1, Wilcoxon p $< 10^{-16}$).
 Diatoms, however, show small but significant declines in richness in the Hood Canal (0.5 species, Wilcoxon p $< 10^{-16}$), while the changes on the San Juan Island are undetectable (p $=$ 0.8).
-Dinoflagellates see their richness increase in both regions with the future scenario(Wilcoxon p $< 10^{-50}$) }.
+Dinoflagellates see their richness increase in both regions with the future scenario(Wilcoxon p $< 10^{-16}$) }.
 {\color{blue} I re-did some of these in the Rmd file (starting at line 1129, doing it line-by-line to convince myself what was going on.
-See line 1157 in particular: for diatoms, there is no change between years, but yet the wilcox p value is very low.
+See line 1179 in particular: for diatoms, there is no change between years, but yet the wilcox p value is very low.
 I can't figure out what's happening there.}
@@
 {\color{blue} The paragraph below is important, and links to the question I had above: where are we doing this calculation in the code?
-and is the below paragraph still true?
-Plus, a reviewer wanted a citation of the Coscinodiscus statement.}
+I went through line by line; double check calculations -- see lines 1207 and following in the Rmd.
+Plus, a reviewer wanted a citation of the Coscinodiscus statement, so I added some.}
 
-The bulk of our projected community changes result from now-rare conditions occurring more frequently in the future.
-For example, in summer in the Hood Canal at present, we expect surface waters to have pH $<$ 7.9 and T $>$ 19$\degree$C only 1\% of the time.
-In 2095, we expect these conditions 18 times more frequently (i.e., 18\% of the time).
+{\color{blue}The bulk of our projected community changes result from now-rare conditions occurring more frequently in the future.
+For example, in in the Hood Canal at present, we expect surface waters to have pH $<$ 7.9 and T $>$ 19$\degree$C only 1\% of the time.
+In 2095, we expect these conditions 6 times more frequently (i.e., 6\% of the time).
 At these values of T and pH, our model predicts the harmful \textit{Alexandrium} sp.
-to occur more often than not (mean frequency of occurrence $=$ 0.7).
-By contrast, the large centric diatom \textit{Coscinodiscus} -- a key source of carbon for zooplankton and small fishes -- occurs only one-third of the time under these same conditions (mean frequency $=$ 0.34).
+to occur more often than not (mean frequency of occurrence $=$ 0.83).
+By contrast, the large centric diatom \textit{Coscinodiscus} -- a potentially key source of carbon for zooplankton and small fishes \cite{pasquaud2010determination, zamon2002tidal} with effects on dissolved oxygen and other water-column characteristics \cite{manabe1991bloom} -- occurs only one-third of the time under these same conditions (mean frequency $=$ 0.35).}
 

### 2020-09-06 Scripts/Manuscript.Main.Analysis.Rmd
intermediate changes for Moncho's review. still working on Results.

@@
 
+#For example, we expect the Hood Canal to be T $>$ 12\degree C and pH $<$ {7.9} -- where community 3 is the most likely community -- more than four times as often in 2095 than in 2017 (4.9 \% vs 20.7\%) )
+
+stds_to_value$toValue[[3]]
+stds_to_value$toValue[[4]]
+
+richness.by.phylum.by.point %>% 
+  ungroup() %>% 
+  dplyr::select(Area, Temperature, pH, Year) %>% 
+  distinct() %>% 
+  filter(Area == "Hood Canal") %>% 
+  mutate(Temperature = 14.995 + 4.299*Temperature,
+         pH = 8.2157 + 0.1813*pH) %>% 
+  group_by(Year) %>% 
+  add_count() %>% 
+  summarise(condition = sum(Temperature > 12 & pH < 7.9),
+            totalcount = n()) %>% 
+  mutate(percent_time = condition / totalcount)
+
 #we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, averaged over all environmental conditions, Wilcoxon p $=$ 0.5);
@@
 
-#while higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 2.1, Wilcoxon p $=$ 0).
+#while higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 2.1, Wilcoxon p $<$ 10 ^{-16}).
 
@@
 
-#Dinoflagellates see their richness increase in both regions with the future scenario(Wilcoxon p $< 10^{-50}$)
+#Dinoflagellates see their richness increase in both regions with the future scenario(Wilcoxon p $< 10^{-16}$)
+
+wilcox.test(x = richness.by.phylum.by.point %>%
+                  filter(Year == 2017, Area == "Hood Canal", phylum == "Dinophyceae") %>% 
+                    pull(richness),
+            y = richness.by.phylum.by.point %>%
+                  filter(Year == 2095, Area == "Hood Canal", phylum == "Dinophyceae") %>% 
+                    pull(richness))
+
+wilcox.test(x = richness.by.phylum.by.point %>%
+                  filter(Year == 2017, Area == "San Juan Island", phylum == "Dinophyceae") %>% 
+                    pull(richness),
+            y = richness.by.phylum.by.point %>%
+                  filter(Year == 2095, Area == "San Juan Island", phylum == "Dinophyceae") %>% 
+                    pull(richness))
+
+#For example, in in the Hood Canal at present, we expect surface waters to have pH $<$ 7.9 and T $>$ 19$\degree$C only 1\% of the time.
+In 2095, we expect these conditions 6 times more frequently (i.e., 6\% of the time).
+
+stds_to_value$toValue[[3]]
+stds_to_value$toValue[[4]]
+
+richness.by.phylum.by.point %>% 
+  ungroup() %>% 
+  dplyr::select(Area, Temperature, pH, Year) %>% 
+  distinct() %>% 
+  filter(Area == "Hood Canal") %>% 
+  mutate(Temperature = 14.995 + 4.299*Temperature,
+         pH = 8.2157 + 0.1813*pH) %>% 
+  group_by(Year) %>% 
+  add_count() %>% 
+  summarise(condition = sum(Temperature > 19 & pH < 7.9),
+            totalcount = n()) %>% 
+  mutate(percent_time = condition / totalcount)
+
+#At these values of T and pH, our model predicts the harmful \textit{Alexandrium} sp. to occur more often than not (mean frequency of occurrence $=$ 0.833)
+
+expanded.future %>% 
+  filter(Area == "Hood Canal", Year == 2095, TemperatureCelsius > 19, pH_units < 7.90, taxa == "Gonyaulacaceae|Alexandrium|NA") %>% 
+  dplyr::select(present) %>% 
+  summarise(mean(present))
+
+#By contrast, the large centric diatom \textit{Coscinodiscus} -- a potentially key source of carbon for zooplankton and small fishes \cite{pasquaud2010determination, zamon2002tidal} with effects on dissolved oxygen and other water-column characteristics \cite{manabe1991bloom} -- occurs only one-third of the time under these same conditions (mean frequency $=$ 0.35)
+expanded.future %>% 
+  filter(Area == "Hood Canal", Year == 2095, TemperatureCelsius > 19, pH_units < 7.90, taxa == "Coscinodiscaceae|Coscinodiscus|Coscinodiscus wailesii") %>% 
+  dplyr::select(present) %>% 
+  summarise(mean(present))
 

### 2020-09-06 Manuscript/What.lies.ahead.reviewed.tex
minor changes to TEX file

@@
 
+{\color{blue} needs RAMON review}
 Among the taxa surveyed, diatoms are of particular interest for their ubiquity in the world's oceans and their important roles in marine food webs \cite{armbrust2009life, wasmund2017diatom}, as well as in ecological and evolutionary theory \cite{margalef11978}.
-Our model suggests that diatoms will decrease in richness between the present and 2095, particularly in the Hood Canal summer months, when extreme temperatures are more likely.
+Our model suggests that diatoms will decrease in richness between the present and 2095, particularly in the Hood Canal, where extreme temperatures are more common.
 Although the most prevalent response among diatoms is a decrease in suitability, some substantial variability in responses exists within the group.
@@
 copepods \cite{jansen2008copepods} and many other animal species, will see future suitable habitat only in colder waters such as those in San Juan Island, while \textit{Skeletonema} spp.
-and the HAB-forming species \textit{Pseudo-nitzschia} spp.
+and the harmful algal bloom (HAB)-forming species \textit{Pseudo-nitzschia} spp.
 will see their habitat suitability remain constant or slightly increased, especially at low pH levels.
+{\color{blue} is this in supplement?}
 
@@
 A switch from a diatom-dominated ecosystem to one in which dinoflagellate blooms extend in space and time could provoke cascade effects \cite{xiao2018warming} including fish mortality, eutrophication \cite{anderson2002harmful}, and bentho-pelagic coupling \citep[][]{bienfang1984sinking}.
-Beyond the phylum-specific patterns, the increase in suitable habitat for HAB-forming species will alone be an engine for ecosystem change \cite{trainer2019pelagic, wells2018harmful}.
+Beyond the phylum-specific patterns, the increase in suitable habitat for harmful algae species will alone be an engine for ecosystem change \cite{trainer2019pelagic, wells2018harmful}.
 

### 2020-09-09 Scripts/Manuscript.Main.Analysis.Rmd
Updated at Wed

@@
 
-plot(layout5)
 p1 + 
@@
 
+```{r show change, echo = F}
+explore.phylum <- function(phyl){
+
+richness.by.phylum.by.point %>%
+  filter(phylum == phyl) %>% 
+  group_by(Area, Year) %>% 
+  summarise (grp.median = median (richness)) -> medians
+
+richness.by.phylum.by.point %>%
+  filter(phylum == phyl) %>% 
+  ggplot(aes())+
+  geom_histogram(bins = 100, aes(x =  richness, group = Year, color = Year, after_stat(ndensity)),fill = "white", position = "identity") +
+  geom_vline(data = medians, aes(xintercept = grp.median, color = Year), linetype = "dashed") +
+  facet_wrap (~Area)
+
+}
+```
+
 Now the Global - richness by point
@@
 ```{r}
-
 #overall change in richness
@@
   summarise(median(richness))
+explore.phylum("Bacillariophyta")
+
 richness.by.phylum.by.point %>%
@@
   summarise(median(richness))
+explore.phylum("Dinophyceae")
 
@@
             totalcount = n()) %>% 
-  mutate(percent_time = condition / totalcount)
-
+  mutate(percent_time = 100 *condition / totalcount)
+```
+
+In general, the prevalence of cluster 3 in the Hood Canal will increase from 54.2% of the time to 75% of the time,
+```{r echo = F}
+# Add the cluster information to the projected scenario
+
+expanded.future %>%
+
+  mutate (Area = str_replace(Area, replacement = "San Juan Island", pattern =  "SJI")) %>%
+  select(Area, Year, Temperature = TemperatureCelsius,
+         pH_new = pH_units) %>% 
+  distinct() %>% 
+    group_by(Area) %>% 
+    nest() %>% 
+    left_join(clusters) -> expand.with.clusters
+
+# Project which one on each environmental simulation
+expand.with.clusters %>% 
+    mutate(whichComm.res = map2(data,multi.mod,~ .x %>% 
+                                  mutate(comm = as.numeric (apply(predict(.y, type = "probs", newdata = .x),1, which.max)) )
+              ) ) %>% unnest(whichComm.res) -> expand.with.clusters
+# Calculate the prevalence of each cluster
+expand.with.clusters %>%
+  group_by(Area, Year, comm) %>% 
+  tally() %>%  
+  group_by(Area, Year) %>% 
+  add_count() %>%
+  mutate(prev = n/nn) %>%
+  select(Area, Year, comm, prev) %>% pivot_wider(names_from = comm, values_from = prev)
+#Note that clusters are different on each Area - cluster #3 from Hood Canal is different from cluster #3 from the San Juans
+```
+
+```{r echo = F}
 #we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, averaged over all environmental conditions, Wilcoxon p $=$ 0.5);
@@
                   filter(Year == 2095, Area == "Hood Canal") %>% 
-                    pull(richness))
+                    pull(richness), conf.int = T)
 
@@
                   filter(Year == 2095, Area == "San Juan Island") %>% 
-                    pull(richness))
+                    pull(richness), conf.int = T)
 
@@
                   filter(Year == 2095, Area == "Hood Canal", phylum == "Bacillariophyta") %>% 
-                    pull(richness))
-
-#while the changes on the San Juan Island are undetectable (p $=$ 0.8).
+                    pull(richness), conf.int = T)
+
+#while the changes on the San Juan Island are undetectable (difference in median switch of 0.05, with  p $=$ 0.01).
 
@@
                   filter(Year == 2095, Area == "San Juan Island", phylum == "Bacillariophyta") %>% 
-                    pull(richness))
+                    pull(richness), conf.int = T)
 
@@
                   filter(Year == 2095, Area == "Hood Canal", phylum == "Dinophyceae") %>% 
-                    pull(richness))
+                    pull(richness),
+            conf.int= TRUE)
 
@@
                   filter(Year == 2095, Area == "San Juan Island", phylum == "Dinophyceae") %>% 
-                    pull(richness))
+                    pull(richness),
+            conf.int= TRUE)
 
@@
 
-```
-
-Because we are comparing two 
-```{r}
-
-## GLOBAL TEST
-
-richness.by.point %>% 
-  nest(-Area) %>% 
-  mutate(wilcoxon = map(data, ~ wilcox.test(richness ~ Year, data = .x)),
-         tidy_wilcox = map_df(wilcoxon, broom::tidy))
-
-## We cannot
-
-richness.by.phylum.by.point %>% 
-   nest(-phylum,-Area) %>% 
-   mutate(wilcoxon = map (data, ~ wilcox.test(richness ~ Year, data = .x) ),
-          wiloc.tidy= map(wilcoxon, broom::tidy))  -> results.lm
-
-results.lm %>%  filter (phylum == "Bacillariophyta") %>% unnest(wiloc.tidy)
-
-results.lm %>%  filter (phylum == "Dinophyceae") %>% unnest(wiloc.tidy)
-
-results.lm %>% 
-  unnest(wiloc.tidy) %>%
-
-  ggplot(aes(x = phylum , y = `p.value`)) +
-  geom_point(aes(color = Area))
-```
+# Responses in diatoms are not all in the same direction - Pseudonitzschia spp shows higher probability of presence at low pH values, and low pH values are more common in the 2095 scenario
+
+expanded.future %>% 
+  group_by(Area) %>% 
+  filter(  str_detect(taxa, "Pseudo-nitzschia") ) %>% 
+  ggplot(aes(x = pH_units, y = present)) +
+  geom_smooth(aes(color = as.factor(Year), group = Year)) +
+  facet_grid(~Area) + 
+  ggtitle("Pseudo-nitzschia spp. response to pH")
+
+expanded.future %>%
+
+  mutate (Area = str_replace(Area, replacement = "San Juan Island", pattern =  "SJI")) %>%
+  select(Area, Year, Temperature = TemperatureCelsius,
+         pH_new = pH_units) %>% 
+  distinct() %>% 
+    group_by(Area) %>% 
+    nest() %>% 
+    left_join(clusters) %>% 
+    mutate(whichComm.res = map2(data,multi.mod,~ .x %>% 
+                                  mutate(comm = as.numeric (apply(predict(.y, type = "probs", newdata = .x),1, which.max)) )
+              ) ) %>% unnest(whichComm.res)
+```

### 2020-09-09 Scripts/Supplement_ModelTesting.Rmd
adding modeltesting fit data so user can load it directly rather than rerunning; adding accompanying knitted PDF for supplement.

@@
 ---
-title: "Supplement: Testing Logistic Regression Models"
+title: "Testing Logistic Regression Models (Supplement to: Environmental DNA Metabarcoding Reveals Winners and Losers of Global Change in Coastal Waters)"
+author: "Gallego et al."
 header-includes:
@@
    - \usepackage{fontspec}
-   - \setmainfont{Open Sans}
+   - \setmainfont{Times New Roman}
 output:
@@
     fig_caption: yes
-    keep_tex: true
     number_sections: no
@@
 
-```{r}
+```{r, echo = F, include = F}
 library(tidyverse)
@@
 
+#if loading data from previous model runs
+load(file = "ModelTest_fits.Rdata")
+
 ```
@@
 
-```{r, echo = F}
-      t_prior <- student_t(df = 7, location = 0, scale = 2.5)
-
-      #models fitting one intercept per unique area/taxon combination, plus
-      #a different slope for the effect of environmental variables on each taxon.
-      mod1.3vars <- stan_glmer(binary ~ 0 + (0 + Temperature + pH + Salinity | taxa) + (1 | taxonAreaIndex), 
-                             data = stan.input, 
-                             family = "binomial",
-                             prior = t_prior,
-                             prior_intercept = t_prior, 
-                             chains = 2,
-                             adapt_delta = 0.99,
-                             refresh = 0)
-
-      mod1.2vars <- stan_glmer(binary ~ 0 + (0 + Temperature + pH | taxa) + (1 | taxonAreaIndex), 
-                             data = stan.input, 
-                             family = "binomial",
-                             prior = t_prior,
-                             prior_intercept = t_prior, 
-                             chains = 2,
-                             adapt_delta = 0.99,
-                             refresh = 0)
-
-      mod1.1vars.pH <- stan_glmer(binary ~ 0 + (0 + pH | taxa) + (1 | taxonAreaIndex), 
-                             data = stan.input, 
-                             family = "binomial",
-                             prior = t_prior,
-                             prior_intercept = t_prior, 
-                             chains = 2,
-                             adapt_delta = 0.99,
-                             refresh = 0)
-
-      mod1.1vars.Temperature <- stan_glmer(binary ~ 0 + (0 + Temperature | taxa) + (1 | taxonAreaIndex), 
-                             data = stan.input, 
-                             family = "binomial",
-                             prior = t_prior,
-                             prior_intercept = t_prior, 
-                             chains = 2,
-                             adapt_delta = 0.99,
-                             refresh = 0)
-
-      #models fitting one intercept per taxon, plus
-      #a different slope for the effect of environmental variables on each taxon. No Area term.
-      mod2.3vars <- stan_glmer(binary ~ 0 + (0 + Temperature + pH + Salinity | taxa) + (1 | taxa), 
-                             data = stan.input, 
-                             family = "binomial",
-                             prior = t_prior,
-                             prior_intercept = t_prior, 
-                             chains = 2,
-                             adapt_delta = 0.99,
-                             refresh = 0)
-
-      mod2.2vars <- stan_glmer(binary ~ 0 + (0 + Temperature + pH | taxa) + (1 | taxa), 
-                             data = stan.input, 
-                             family = "binomial",
-                             prior = t_prior,
-                             prior_intercept = t_prior, 
-                             chains = 2,
-                             adapt_delta = 0.99,
-                             refresh = 0)
-
-      mod2.1vars.pH <- stan_glmer(binary ~ 0 + (0 + pH | taxa) + (1 | taxa), 
-                              data = stan.input, 
-                             family = "binomial",
-                             prior = t_prior,
-                             prior_intercept = t_prior, 
-                             chains = 2,
-                             adapt_delta = 0.99,
-                             refresh = 0)
-
-      mod2.1vars.Temperature <- stan_glmer(binary ~ 0 + (0 + Temperature | taxa) + (1 | taxa), 
-                             data = stan.input, 
-                             family = "binomial",
-                             prior = t_prior,
-                             prior_intercept = t_prior, 
-                             chains = 2,
-                             adapt_delta = 0.99,
-                             refresh = 0)
-
-      #models fitting a different slope for the effect of environmental variables on each taxon, with no intercept term.
-      mod3.3vars <- stan_glmer(binary ~ 0 + (0 + Temperature + pH + Salinity | taxa), 
-                             data = stan.input, 
-                             family = "binomial",
-                             prior = t_prior,
-                             prior_intercept = t_prior, 
-                             chains = 2,
-                             adapt_delta = 0.99,
-                             refresh = 0)
-
-      mod3.2vars <- stan_glmer(binary ~ 0 + (0 + Temperature + pH | taxa), 
-                             data = stan.input, 
-                             family = "binomial",
-                             prior = t_prior,
-                             prior_intercept = t_prior, 
-                             chains = 2,
-                             adapt_delta = 0.99,
-                             refresh = 0)
-
-      mod3.1vars.pH <- stan_glmer(binary ~ 0 + (0 + pH | taxa), 
-                             data = stan.input, 
-                             family = "binomial",
-                             prior = t_prior,
-                             prior_intercept = t_prior, 
-                             chains = 2,
-                             adapt_delta = 0.99,
-                             refresh = 0)
-
-      mod3.1vars.Temperature <- stan_glmer(binary ~ 0 + (0 + Temperature | taxa), 
-                             data = stan.input, 
-                             family = "binomial",
-                             prior = t_prior,
-                             prior_intercept = t_prior, 
-                             chains = 2,
-                             adapt_delta = 0.99,
-                             refresh = 0)
+```{r, echo = F, message = F, warning = F}
+#uncomment if running de novo; otherwise, load data from file, above
+      # t_prior <- student_t(df = 7, location = 0, scale = 2.5)
+      # 
+      # #models fitting one intercept per unique area/taxon combination, plus
+      # #a different slope for the effect of environmental variables on each taxon.
+      # mod1.3vars <- stan_glmer(binary ~ 0 + (0 + Temperature + pH + Salinity | taxa) + (1 | taxonAreaIndex), 
+      #                        data = stan.input, 
+      #                        family = "binomial",
+      #                        prior = t_prior,
+      #                        prior_intercept = t_prior, 
+      #                        chains = 2,
+      #                        adapt_delta = 0.99,
+      #                        refresh = 0)
+      # 
+      # mod1.2vars <- stan_glmer(binary ~ 0 + (0 + Temperature + pH | taxa) + (1 | taxonAreaIndex), 
+      #                        data = stan.input, 
+      #                        family = "binomial",
+      #                        prior = t_prior,
+      #                        prior_intercept = t_prior, 
+      #                        chains = 2,
+      #                        adapt_delta = 0.99,
+      #                        refresh = 0)
+      # 
+      # mod1.1vars.pH <- stan_glmer(binary ~ 0 + (0 + pH | taxa) + (1 | taxonAreaIndex), 
+      #                        data = stan.input, 
+      #                        family = "binomial",
+      #                        prior = t_prior,
+      #                        prior_intercept = t_prior, 
+      #                        chains = 2,
+      #                        adapt_delta = 0.99,
+      #                        refresh = 0)      
+      # 
+      # mod1.1vars.Temperature <- stan_glmer(binary ~ 0 + (0 + Temperature | taxa) + (1 | taxonAreaIndex), 
+      #                        data = stan.input, 
+      #                        family = "binomial",
+      #                        prior = t_prior,
+      #                        prior_intercept = t_prior, 
+      #                        chains = 2,
+      #                        adapt_delta = 0.99,
+      #                        refresh = 0)      
+      # 
+      # #models fitting one intercept per taxon, plus
+      # #a different slope for the effect of environmental variables on each taxon. No Area term.
+      # mod2.3vars <- stan_glmer(binary ~ 0 + (0 + Temperature + pH + Salinity | taxa) + (1 | taxa), 
+      #                        data = stan.input, 
+      #                        family = "binomial",
+      #                        prior = t_prior,
+      #                        prior_intercept = t_prior, 
+      #                        chains = 2,
+      #                        adapt_delta = 0.99,
+      #                        refresh = 0)
+      # 
+      # mod2.2vars <- stan_glmer(binary ~ 0 + (0 + Temperature + pH | taxa) + (1 | taxa), 
+      #                        data = stan.input, 
+      #                        family = "binomial",
+      #                        prior = t_prior,
+      #                        prior_intercept = t_prior, 
+      #                        chains = 2,
+      #                        adapt_delta = 0.99,
+      #                        refresh = 0)
+      # 
+      # mod2.1vars.pH <- stan_glmer(binary ~ 0 + (0 + pH | taxa) + (1 | taxa), 
+      #                         data = stan.input, 
+      #                        family = "binomial",
+      #                        prior = t_prior,
+      #                        prior_intercept = t_prior, 
+      #                        chains = 2,
+      #                        adapt_delta = 0.99,
+      #                        refresh = 0)
+      # 
+      # mod2.1vars.Temperature <- stan_glmer(binary ~ 0 + (0 + Temperature | taxa) + (1 | taxa), 
+      #                        data = stan.input, 
+      #                        family = "binomial",
+      #                        prior = t_prior,
+      #                        prior_intercept = t_prior, 
+      #                        chains = 2,
+      #                        adapt_delta = 0.99,
+      #                        refresh = 0)
+      # 
+      # #models fitting a different slope for the effect of environmental variables on each taxon, with no intercept term.
+      # mod3.3vars <- stan_glmer(binary ~ 0 + (0 + Temperature + pH + Salinity | taxa), 
+      #                        data = stan.input, 
+      #                        family = "binomial",
+      #                        prior = t_prior,
+      #                        prior_intercept = t_prior, 
+      #                        chains = 2,
+      #                        adapt_delta = 0.99,
+      #                        refresh = 0)
+      # 
+      # mod3.2vars <- stan_glmer(binary ~ 0 + (0 + Temperature + pH | taxa), 
+      #                        data = stan.input, 
+      #                        family = "binomial",
+      #                        prior = t_prior,
+      #                        prior_intercept = t_prior, 
+      #                        chains = 2,
+      #                        adapt_delta = 0.99,
+      #                        refresh = 0)
+      # 
+      # mod3.1vars.pH <- stan_glmer(binary ~ 0 + (0 + pH | taxa), 
+      #                        data = stan.input, 
+      #                        family = "binomial",
+      #                        prior = t_prior,
+      #                        prior_intercept = t_prior, 
+      #                        chains = 2,
+      #                        adapt_delta = 0.99,
+      #                        refresh = 0)
+      # 
+      # mod3.1vars.Temperature <- stan_glmer(binary ~ 0 + (0 + Temperature | taxa), 
+      #                        data = stan.input, 
+      #                        family = "binomial",
+      #                        prior = t_prior,
+      #                        prior_intercept = t_prior, 
+      #                        chains = 2,
+      #                        adapt_delta = 0.99,
+      #                        refresh = 0)      
+      #
 
@@
 
-$$logit(p_i) = \alpha_{j} + Temperature * \beta_{1_k} + pH * \beta_{2_k}$$
+$$logit(p_i) = \alpha_{j} + Temperature_i * \beta_{1_k} + pH_i * \beta_{2_k}$$
 
@@
 
-```{r}
-stan.input <- data.for.stan %>% 
-      group_by(Area) %>% 
-      mutate(Temperature = vegan::decostand(Temperature, method = "standardize"),
-             pH = vegan::decostand(pH, method = "standardize"),
-             Salinity = vegan::decostand(Salinity, method = "standardize")) %>%
-      group_by(taxa) %>%
-      mutate(Noccur = sum(binary)) %>%
-      filter(Noccur > 7 & Noccur < 69) %>%  #must be present or absent in at least 10% of 76 samples in dataset
-      dplyr::select(-c(sample, phylum, DIC, Noccur)) %>% 
-      ungroup() %>% 
-      unite(Area, taxa, col = "AreaTaxa", remove = F) %>% 
-      mutate(sampleIndex = 1:nrow(.),
-             taxonIndex = match(taxa, unique(taxa)),
-             areaIndex = match(Area, unique(Area)),
-             seasonIndex = match(Season, unique(Season)),
-             taxonAreaIndex = match(AreaTaxa, unique(AreaTaxa))) %>%   #unique combinations of taxon and area
-      dplyr::select(-AreaTaxa)
-
-finalModel <- stan_glmer(binary ~ 0 + (0 + Temperature + pH | taxa) + (1 | taxonAreaIndex), 
-                             data = stan.input, 
-                             family = "binomial",
-                             prior = t_prior,
-                             prior_intercept = t_prior, 
-                             chains = 4,
-                             adapt_delta = 0.99,
-                             refresh = 0)
+```{r, echo = F, include = F}
+#uncomment if running de novo; otherwise, load data from file, above
+# stan.input <- data.for.stan %>% 
+#       group_by(Area) %>% 
+#       mutate(Temperature = vegan::decostand(Temperature, method = "standardize"),
+#              pH = vegan::decostand(pH, method = "standardize"),
+#              Salinity = vegan::decostand(Salinity, method = "standardize")) %>%
+#       group_by(taxa) %>%
+#       mutate(Noccur = sum(binary)) %>%
+#       filter(Noccur > 7 & Noccur < 69) %>%  #must be present or absent in at least 10% of 76 samples in dataset
+#       dplyr::select(-c(sample, phylum, DIC, Noccur)) %>% 
+#       ungroup() %>% 
+#       unite(Area, taxa, col = "AreaTaxa", remove = F) %>% 
+#       mutate(sampleIndex = 1:nrow(.),
+#              taxonIndex = match(taxa, unique(taxa)),
+#              areaIndex = match(Area, unique(Area)),
+#              seasonIndex = match(Season, unique(Season)),
+#              taxonAreaIndex = match(AreaTaxa, unique(AreaTaxa))) %>%   #unique combinations of taxon and area
+#       dplyr::select(-AreaTaxa) 
+# 
+# finalModel <- stan_glmer(binary ~ 0 + (0 + Temperature + pH | taxa) + (1 | taxonAreaIndex), 
+#                              data = stan.input, 
+#                              family = "binomial",
+#                              prior = t_prior,
+#                              prior_intercept = t_prior, 
+#                              chains = 4,
+#                              adapt_delta = 0.99,
+#                              refresh = 0)
 ```
@@
 
-```{r, echo = F}
+```{r, echo = F, message = F, warning = F}
 stan.input$postpredict <- posterior_predict(finalModel) %>% colMeans()
 
-stan.input %>% 
+tmp <- stan.input %>% 
   left_join(data.for.stan[,c("taxa", "phylum")]) %>% 
         filter(phylum == "Bacillariophyta") %>% 
-        filter(taxa %in% unique(taxa)[1:5]) %>% 
+        filter(taxa %in% unique(taxa)[1:5]) 
+#tmp$taxa <- gsub(".+\\|", "", tmp$taxa) #for ease of reading labels, remove family name
+tmp$taxa <- gsub("\\|", "\n", tmp$taxa) #for ease of reading labels, insert newlines
+tmp$taxa <- gsub(" ", "\n", tmp$taxa) #for ease of reading labels, insert newlines
+        tmp %>% 
         ggplot(aes(x = pH, y = binary)) +
@@
           geom_point(aes(x = pH, y = postpredict), color = "blue") +
+          geom_smooth(aes(x = pH, y = postpredict), color = "blue", method = "lm") +
           facet_grid(taxa ~ Area) +
+          theme(aspect.ratio=1/3) +
           ylab("Prob of Presence") +
@@
 
+```
+
+```{r, echo = F}
       stan.input %>% 
@@
         geom_point(aes(x = value, y = postpredict), color = "blue") +
+        geom_smooth(aes(x = value, y = postpredict), color = "blue", method = "lm") +
         facet_grid(name ~ Area) +
@@
 ```
+
+```{r Store Data if Desired, echo  = F, include = F}
+#save.image(file = "ModelTest_fits.Rdata")
+```

### 2020-09-09 Manuscript/Internal_Review.tex

--- 2020-09-06 Manuscript/What.lies.ahead.reviewed.tex
@@
 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls.
-For example, we expect the Hood Canal to be T $>$ 12\degree C and pH $<$ {7.9} -- where community 3 is the most likely community -- more than four times as often in 2095 than in 2017 (4.9 \% vs 20.7\%).
-{\color{blue} I couldn't find where you were doing this calculation, so I made one up -- see lines 1134 and following in the Rmd.
-BUT ALSO : community 3 isn't sensitive to pH, so what are we even saying here?}
+For example, we expect the Hood Canal in 2095 to present the conditions in which community 3 is the most likely community 75\% of the time, an increase of 22\% compared to 2017(see Supplemental Material).
 
-\\
-
-%Under a common environmental condition, for example, in Hood Canal in summer, we expect such a planktonic community to go from occurring less than 1\% of the time in 2017 (0.001\% at T $=$ 15$\degree$C and pH $=$ 8.4; conditions more extreme than this occur ca.
+% Under a common environmental condition, for example, in Hood Canal in summer, we expect such a planktonic community to go from occurring less than 1\% of the time in 2017 (0.001\% at T $=$ 15$\degree$C and pH $=$ 8.4; conditions more extreme than this occur ca.
 50\% of the time) to occurring 37.5\% of the time under an equivalently-common condition in 2095 (84\% at T $=$ 22$\degree$C and pH $=$ 8.05).
@@
 
-Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at overall richness variation; {\color{red} we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, averaged over all environmental conditions, Wilcoxon p $=$ 0.5); while higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 2.1, Wilcoxon p $< 10^{-16}$).
-Diatoms, however, show small but significant declines in richness in the Hood Canal (0.5 species, Wilcoxon p $< 10^{-16}$), while the changes on the San Juan Island are undetectable (p $=$ 0.8).
-Dinoflagellates see their richness increase in both regions with the future scenario(Wilcoxon p $< 10^{-16}$) }.
-{\color{blue} I re-did some of these in the Rmd file (starting at line 1129, doing it line-by-line to convince myself what was going on.
-See line 1179 in particular: for diatoms, there is no change between years, but yet the wilcox p value is very low.
-I can't figure out what's happening there.}
+Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at overall richness variation; we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, 95\% confidence interval in median species richness  -0.08, +0.04; Wilcoxon p $=$ 0.5); while higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 1.84-2.2, Wilcoxon p $< 10^{-16}$).
+Diatoms, however, show small but significant declines in richness in the Hood Canal (0.47-0.57 species, Wilcoxon p $< 10^{-16}$), while the changes on the San Juan Island are undetectable (median change 0-0.08 p $=$ 0.04).
+Dinoflagellates see their richness increase in both regions with the future scenario(Wilcoxon p $< 10^{-16}$).
 
@@
  \includegraphics[width=1\linewidth]{Figures_for_MS/Top_fig3_2020-09-03.Other.sel.png}
-\caption{\footnotesize{Forecasted changes in plankton in the Salish Sea {\color{red} for Dinoflagellates (panel \textbf{A}), Diatoms (\textbf{B}) and Haptophytes (\textbf{C}).
+\caption{\footnotesize{Forecasted changes in plankton in the Salish Sea for Dinoflagellates (panel \textbf{A}), Diatoms (\textbf{B}) and Haptophytes (\textbf{C}).
 Each panels shows:  (\textbf{.1}) probability densities for the occurrence of selected taxa (species- and genus-level) for 2017 (blue) and 2095 (red); data are mean probabilities over 100 model draws, and variance in probability is due to differences in underlying environmental conditions.
@@
 Envelopes of observed (2017, blue) and modeled (2095, red) annual conditions in the Salish Sea shown for reference.
-Hood Canal and San Juan Island plotted separately to illustrate environmental differences between them.}}}
+Hood Canal and San Juan Island plotted separately to illustrate environmental differences between them.}}
 
@@
 
-{\color{blue} The paragraph below is important, and links to the question I had above: where are we doing this calculation in the code?
-I went through line by line; double check calculations -- see lines 1207 and following in the Rmd.
-Plus, a reviewer wanted a citation of the Coscinodiscus statement, so I added some.}
-
-{\color{blue}The bulk of our projected community changes result from now-rare conditions occurring more frequently in the future.
+The bulk of our projected community changes result from now-rare conditions occurring more frequently in the future.
 For example, in in the Hood Canal at present, we expect surface waters to have pH $<$ 7.9 and T $>$ 19$\degree$C only 1\% of the time.
@@
 to occur more often than not (mean frequency of occurrence $=$ 0.83).
-By contrast, the large centric diatom \textit{Coscinodiscus} -- a potentially key source of carbon for zooplankton and small fishes \cite{pasquaud2010determination, zamon2002tidal} with effects on dissolved oxygen and other water-column characteristics \cite{manabe1991bloom} -- occurs only one-third of the time under these same conditions (mean frequency $=$ 0.35).}
+By contrast, the large centric diatom \textit{Coscinodiscus} -- a potentially key source of carbon for zooplankton and small fishes \cite{pasquaud2010determination, zamon2002tidal} with effects on dissolved oxygen and other water-column characteristics \cite{manabe1991bloom} -- occurs only one-third of the time under these same conditions (mean frequency $=$ 0.35).
 
@@
 
-{\color{blue} needs RAMON review}
 Among the taxa surveyed, diatoms are of particular interest for their ubiquity in the world's oceans and their important roles in marine food webs \cite{armbrust2009life, wasmund2017diatom}, as well as in ecological and evolutionary theory \cite{margalef11978}.
@@
 and the harmful algal bloom (HAB)-forming species \textit{Pseudo-nitzschia} spp.
-will see their habitat suitability remain constant or slightly increased, especially at low pH levels.
-{\color{blue} is this in supplement?}
+will see their habitat suitability remain constant or slightly increased, especially at low pH levels (see Supplemental Material).
 

### 2020-09-09 Manuscript/What.lies.ahead.reviewed.tex

@@
 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls.
-For example, we expect the Hood Canal to be T $>$ 12\degree C and pH $<$ {7.9} -- where community 3 is the most likely community -- more than four times as often in 2095 than in 2017 (4.9 \% vs 20.7\%).
-{\color{blue} I couldn't find where you were doing this calculation, so I made one up -- see lines 1134 and following in the Rmd.
+For example, we expect the Hood Canal {\color{red} in 2095 to present the conditions in which community 3 is the most likely community 75\% of the time, an increase of 22\% compared to 2017(see Supplemental Material).} \st{T $>$ 12\degree C and pH $<$ {7.9} -- where community 3 is the most likely community -- more than four times as often in 2095 than in 2017 (4.9 \% vs 20.7\%).} {\color{blue} I couldn't find where you were doing this calculation, so I made one up -- see lines 1134 and following in the Rmd.
 BUT ALSO : community 3 isn't sensitive to pH, so what are we even saying here?}
 
-\\
-
-%Under a common environmental condition, for example, in Hood Canal in summer, we expect such a planktonic community to go from occurring less than 1\% of the time in 2017 (0.001\% at T $=$ 15$\degree$C and pH $=$ 8.4; conditions more extreme than this occur ca.
+% Under a common environmental condition, for example, in Hood Canal in summer, we expect such a planktonic community to go from occurring less than 1\% of the time in 2017 (0.001\% at T $=$ 15$\degree$C and pH $=$ 8.4; conditions more extreme than this occur ca.
 50\% of the time) to occurring 37.5\% of the time under an equivalently-common condition in 2095 (84\% at T $=$ 22$\degree$C and pH $=$ 8.05).
@@
 
-Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at overall richness variation; {\color{red} we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, averaged over all environmental conditions, Wilcoxon p $=$ 0.5); while higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 2.1, Wilcoxon p $< 10^{-16}$).
-Diatoms, however, show small but significant declines in richness in the Hood Canal (0.5 species, Wilcoxon p $< 10^{-16}$), while the changes on the San Juan Island are undetectable (p $=$ 0.8).
+Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at overall richness variation; {\color{red} we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, 95\% confidence interval in median species richness  -0.08, +0.04; Wilcoxon p $=$ 0.5); while higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 1.84-2.2, Wilcoxon p $< 10^{-16}$).
+Diatoms, however, show small but significant declines in richness in the Hood Canal (0.47-0.57 species, Wilcoxon p $< 10^{-16}$), while the changes on the San Juan Island are undetectable (median change 0-0.08 p $=$ 0.04).
 Dinoflagellates see their richness increase in both regions with the future scenario(Wilcoxon p $< 10^{-16}$) }.
@@
 I went through line by line; double check calculations -- see lines 1207 and following in the Rmd.
-Plus, a reviewer wanted a citation of the Coscinodiscus statement, so I added some.}
+Plus, a reviewer wanted a citation of the Coscinodiscus statement, so I added some.
+RYAN- I agree with the statement.
+Now we do these calculations in the script.
+I will add that the last bits of the prevalence of Alexandrium and Coscinoduscus are impactful and supercool}
 
-{\color{blue}The bulk of our projected community changes result from now-rare conditions occurring more frequently in the future.
+{\color{red}The bulk of our projected community changes result from now-rare conditions occurring more frequently in the future.
 For example, in in the Hood Canal at present, we expect surface waters to have pH $<$ 7.9 and T $>$ 19$\degree$C only 1\% of the time.

### 2020-09-09 Scripts/Supplement_ModelTesting.Rmd
adding model testing data and changes to make Rmd more easily accessible without rerunning models denovo

@@
 
-#if loading data from previous model runs
-load(file = "ModelTest_fits.Rdata")
-
 ```
@@
 ```{r, echo = F, message = F, warning = F}
+load("preliminaryModels.Rdata")
+
+      t_prior <- student_t(df = 7, location = 0, scale = 2.5)
+      #models fitting one intercept per unique area/taxon combination, plus
+      #a different slope for the effect of environmental variables on each taxon.
+
 #uncomment if running de novo; otherwise, load data from file, above
-      # t_prior <- student_t(df = 7, location = 0, scale = 2.5)
-      # 
-      # #models fitting one intercept per unique area/taxon combination, plus
-      # #a different slope for the effect of environmental variables on each taxon.
+
        # mod1.3vars <- stan_glmer(binary ~ 0 + (0 + Temperature + pH + Salinity | taxa) + (1 | taxonAreaIndex),
@@
       # 
-
+      # 
       #compare model results
@@
                   )
-
+#optionally save model fits      
+#save(list = ls(pattern = "mod"), file = "preliminaryModels.Rdata")      
 ```
@@
 ```{r, echo = F, include = F}
+load("bestModelFit.Rdata") #if loading from existing data
+stan.input <- data.for.stan %>%
+      group_by(Area) %>%
+      mutate(Temperature = vegan::decostand(Temperature, method = "standardize"),
+             pH = vegan::decostand(pH, method = "standardize"),
+             Salinity = vegan::decostand(Salinity, method = "standardize")) %>%
+      group_by(taxa) %>%
+      mutate(Noccur = sum(binary)) %>%
+      filter(Noccur > 7 & Noccur < 69) %>%  #must be present or absent in at least 10% of 76 samples in dataset
+      dplyr::select(-c(sample, phylum, DIC, Noccur)) %>%
+      ungroup() %>%
+      unite(Area, taxa, col = "AreaTaxa", remove = F) %>%
+      mutate(sampleIndex = 1:nrow(.),
+             taxonIndex = match(taxa, unique(taxa)),
+             areaIndex = match(Area, unique(Area)),
+             seasonIndex = match(Season, unique(Season)),
+             taxonAreaIndex = match(AreaTaxa, unique(AreaTaxa))) %>%   #unique combinations of taxon and area
+      dplyr::select(-AreaTaxa)
+
 #uncomment if running de novo; otherwise, load data from file, above
-# stan.input <- data.for.stan %>% 
-#       group_by(Area) %>% 
-#       mutate(Temperature = vegan::decostand(Temperature, method = "standardize"),
-#              pH = vegan::decostand(pH, method = "standardize"),
-#              Salinity = vegan::decostand(Salinity, method = "standardize")) %>%
-#       group_by(taxa) %>%
-#       mutate(Noccur = sum(binary)) %>%
-#       filter(Noccur > 7 & Noccur < 69) %>%  #must be present or absent in at least 10% of 76 samples in dataset
-#       dplyr::select(-c(sample, phylum, DIC, Noccur)) %>% 
-#       ungroup() %>% 
-#       unite(Area, taxa, col = "AreaTaxa", remove = F) %>% 
-#       mutate(sampleIndex = 1:nrow(.),
-#              taxonIndex = match(taxa, unique(taxa)),
-#              areaIndex = match(Area, unique(Area)),
-#              seasonIndex = match(Season, unique(Season)),
-#              taxonAreaIndex = match(AreaTaxa, unique(AreaTaxa))) %>%   #unique combinations of taxon and area
-#       dplyr::select(-AreaTaxa) 
-# 
 # finalModel <- stan_glmer(binary ~ 0 + (0 + Temperature + pH | taxa) + (1 | taxonAreaIndex),
@@
 #                              refresh = 0)
+#optionally save
+#save(finalModel, file = "bestModelFit.Rdata")
 ```
@@
 
-```{r, echo = F}
+```{r, echo = F, message = F, warning = F}
       stan.input %>% 

### 2020-09-09 Scripts/Manuscript.Main.Analysis.Rmd
adding modeltesting fit data so user can load it directly rather than rerunning; adding accompanying knitted PDF for supplement.
minor text changes and removing color comments throughout.

@@
 
-#while the changes on the San Juan Island are undetectable (difference in median switch of 0.05, with  p $=$ 0.01).
+#while the changes on the San Juan Island are negligible (difference in median switch of 0.05, with  p $=$ 0.01).
 
@@
             conf.int= TRUE)
+
+wilcox.test(x = richness.by.phylum.by.point %>%
+                  filter(Year == 2017, Area == "Hood Canal", phylum == "Dinophyceae") %>% 
+                    pull(richness),
+            y = richness.by.phylum.by.point %>%
+                  filter(Year == 2095, Area == "Hood Canal", phylum == "Dinophyceae") %>% 
+                    pull(richness))
 

### 2020-09-09 Manuscript/What.lies.ahead.reviewed.tex

@@
 For each point in the pH - temperature grid, we calculated a species' probability of presence as the mean of 100 independent draws of the posterior model response.
-For richness estimations in future scenarios, we used each of the 100 draws of the posteriors in each environmental point as a sample and tallied the number of species detected on each draw.
+For each point, the sum of mean probabilities across species provided richness estimations.
 The mean value across the 100 draws was the input for a Wilcoxon test of differences in species richness between the 2017 and 2095 scenarios in each region.
@@
 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls.
-For example, we expect the Hood Canal {\color{red} in 2095 to present the conditions in which community 3 is the most likely community 75\% of the time, an increase of 22\% compared to 2017(see Supplemental Material).} \st{T $>$ 12\degree C and pH $<$ {7.9} -- where community 3 is the most likely community -- more than four times as often in 2095 than in 2017 (4.9 \% vs 20.7\%).} {\color{blue} I couldn't find where you were doing this calculation, so I made one up -- see lines 1134 and following in the Rmd.
-BUT ALSO : community 3 isn't sensitive to pH, so what are we even saying here?}
+For example, we expect the Hood Canal in 2095 to have the conditions in which community 3 is the most likely community 75\% of the time, an increase of 22\% compared to 2017(see Supplemental Material).
 
@@
 
-Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at overall richness variation; {\color{red} we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, 95\% confidence interval in median species richness  -0.08, +0.04; Wilcoxon p $=$ 0.5); while higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 1.84-2.2, Wilcoxon p $< 10^{-16}$).
-Diatoms, however, show small but significant declines in richness in the Hood Canal (0.47-0.57 species, Wilcoxon p $< 10^{-16}$), while the changes on the San Juan Island are undetectable (median change 0-0.08 p $=$ 0.04).
-Dinoflagellates see their richness increase in both regions with the future scenario(Wilcoxon p $< 10^{-16}$) }.
-{\color{blue} I re-did some of these in the Rmd file (starting at line 1129, doing it line-by-line to convince myself what was going on.
-See line 1179 in particular: for diatoms, there is no change between years, but yet the wilcox p value is very low.
-I can't figure out what's happening there.}
+Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at overall richness variation; we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, 95\% confidence interval in median species richness  -0.08, +0.04; Wilcoxon p $=$ 0.5); while higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 1.84-2.2, p $< 10^{-16}$).
+Diatoms, in particular, show small but significant declines in richness in the Hood Canal (0.47-0.57 species, p $< 10^{-16}$), while the changes on the San Juan Island are negligible (median change 0-0.08 p $=$ 0.04).
+Dinoflagellates see their richness increase in both regions with the future scenario (median change 0.14-0.18 Hood Canal; 0.21-0.25 San Juan Island; p $< 10^{-16}$ for each).
 
@@
  \includegraphics[width=1\linewidth]{Figures_for_MS/Top_fig3_2020-09-03.Other.sel.png}
-\caption{\footnotesize{Forecasted changes in plankton in the Salish Sea {\color{red} for Dinoflagellates (panel \textbf{A}), Diatoms (\textbf{B}) and Haptophytes (\textbf{C}).
-Each panels shows:  (\textbf{.1}) probability densities for the occurrence of selected taxa (species- and genus-level) for 2017 (blue) and 2095 (red); data are mean probabilities over 100 model draws, and variance in probability is due to differences in underlying environmental conditions.
-(\textbf{.2}) Changes in relative species richness within each phylum across the simulated scenario.
-(\textbf{.3}) Relative taxon richness (raster color, warmer colors are more taxon-rich) for each of these same higher-taxa, for plausible ranges of pH and Temperature.
+\caption{\footnotesize{Forecasted changes in plankton in the Salish Sea for Dinoflagellates (panel \textbf{A}), Diatoms (\textbf{B}) and Haptophytes (\textbf{C}).
+Each panel shows:  (\textbf{1}) probability densities for the occurrence of selected taxa (species- and genus-level) for 2017 (blue) and 2095 (red); data are mean probabilities over 100 model draws, and variance in probability is due to differences in underlying environmental conditions.
+(\textbf{2}) Changes in relative species richness within each phylum across the simulated scenario.
+(\textbf{3}) Relative taxon richness (raster color, warmer colors are more taxon-rich) for each of these same higher-taxa, for plausible ranges of pH and Temperature.
 Envelopes of observed (2017, blue) and modeled (2095, red) annual conditions in the Salish Sea shown for reference.
-Hood Canal and San Juan Island plotted separately to illustrate environmental differences between them.}}}
+Hood Canal and San Juan Island plotted separately to illustrate environmental differences between them.}}
 
@@
 
-{\color{blue} The paragraph below is important, and links to the question I had above: where are we doing this calculation in the code?
-I went through line by line; double check calculations -- see lines 1207 and following in the Rmd.
-Plus, a reviewer wanted a citation of the Coscinodiscus statement, so I added some.
-RYAN- I agree with the statement.
-Now we do these calculations in the script.
-I will add that the last bits of the prevalence of Alexandrium and Coscinoduscus are impactful and supercool}
-
-{\color{red}The bulk of our projected community changes result from now-rare conditions occurring more frequently in the future.
+The bulk of our projected community changes result from now-rare conditions occurring more frequently in the future.
 For example, in in the Hood Canal at present, we expect surface waters to have pH $<$ 7.9 and T $>$ 19$\degree$C only 1\% of the time.
@@
  Our results therefore suggest a possible change in relative dominance between diatoms and other phytoplankton species such as dinoflagellates, consistent with those seen at ecological regime shifts found elsewhere \cite{wasmund2017diatom, hinder2012changes}.
-Such a shift could affect ecosystems in many ways; even under the assumption that the surviving taxa would maintain the primary production levels, for example, the smaller cell-size of dinoflagellates and the differential sinking rates of the two groups would likely alter regional patterns of nutrient cycling\cite{xiao2018warming, anderson2002harmful, bienfang1984sinking}.
-Although the north Atlantic has shown an increase in diatom abundance\cite{hinder2012changes}, the increase in wind stress and associated mixing in the water column in the open ocean is unlikely to occur in the Hood Canal, where stratification is the strongest in the Salish Sea \cite{moore2008descriptive}.
+Such a shift could affect ecosystems in many ways; even under the assumption that the surviving taxa would maintain the primary production levels, for example, the smaller cell-size of dinoflagellates and the differential sinking rates of the two groups would likely alter regional patterns of nutrient cycling \cite{xiao2018warming, anderson2002harmful, bienfang1984sinking}.
+Although the north Atlantic has shown an increase in diatom abundance \cite{hinder2012changes}, the increase in wind stress and associated mixing in the water column in the open ocean is unlikely to occur in the Hood Canal, where stratification is the strongest in the Salish Sea \cite{moore2008descriptive}.
 Furthermore, locally focused models support an increase in dinoflagellate dominance with climate change, particularly during summer months \cite{khangaonkar2019salish}.

### 2020-09-09 Manuscript/Response.to.reviewers.tex
Updated some text figures and response to reviewee

@@
 
-In sum, yes: these clusters are coherent groups of species and predictably occur in different environmental conditions. We have revised the main text to be more clear in this regard.
+In sum, yes: these clusters are coherent groups of species and predictably occur in different environmental conditions.
+We have revised the main text to be more clear in this regard.
+{\color{red} RYAN Maybe add that some of the taxa are present in many clusters, and their differ in their relative proportin }
 
@@
 
-{\color{red} RAMON -- if we should use `cluster' instead of `community' we will need to change that in the main text. }
+{\color{red} RAMON -- if we should use `cluster' instead of `community' we will need to change that in the main text.
+RYAN: Now that we think more carefully about it, it should be clusters of samples - so you use it when you talk about samples - and community of organism: which I agree should only be applied when you have  a deep understanding of the species composition.}
 
@@
 
-{\color{red} TODO - check }
+{\color{red} RGS: Figure 2C includes now taxa with the higher contribution to the dissimilarity between clusters, highlighting both positive and negative contributions.  }
 
@@
 
-{\color{red} TODO - check }
+{\color{red} RGS: We thank the reviewer for clarifying this point.
+We have now decluttered Figure 3; and highlight taxa that are now mentioned in the text: the Dinoflagellates Karlodinium, Gymnodinaceae and Alexandrium; the Diatoms Stephanodiscus, Nitzschia, Coscinodiscus and Chaetoceros; and the Haptophytes Emiliaina and Chrysochromulina spp.}
 
@@
 
-{\color{red} TODO - check }
+{\color{red} RGS: Our effort has been on visualizing species with more pronounced change, not to highlight species that respond in contrasting ways in the San Juan Islands and the Hood Canal.
+Most taxa react in the same direction in both regions; with a few exceptions in which species are almost absent on one study area.
+}
 
@@
 
-{\color{red} TODO - check }
+{\color{red} RGS: Coscinodiscus sees its probability of presence reduced on both Areas.
+The same happens for D.
+brightwelli, although our approach for a more readable figure has seen this taxa drop into the Supplementary figures}
 
@@
 
-{\color{red} TODO - check }
+{\color{red} RGS: We have updated that sentence, which now reads: "The haptophytes \textit{Emiliania huxleyi} and \textit{Chrysochromulina} sp.
+the dinoflagellate \textit{Alexandrium} sp.
+all find more suitable habitat in both of our study areas." Thanks for the suggestion.
+}
 
@@
 
-{\color{red} TODO - check }
+{\color{red} RGS: We thank the reviewer for pointing this out: We have rethought this and applied Wilcoxon tests for both global and phylum-specific changes in diversity.
+As we are comparing pairs of data we prefer not to assume normal distributions of the datasets or the differences between them, Wilcoxon test was a better choice.
+}
 
@@
 
-{\color{red} TODO - check }
+{\color{red} RYAN: It is now in the methods, but it has a Discussion feel to it }
 
@@
 
-{\color{red} TODO - check }
+{\color{red} RGS: Agreed: We now cite the work of J. Zamon (2002) and S. Pasquaud and colleagues (2010) to support the role of Coscinodiscus in the food chain }
 
@@
 
-{\color{red} TODO - check }
+{\color{red} RGS: Done }
 

### 2020-09-09 Manuscript/What.lies.ahead.reviewed.tex
Updated some text figures and response to reviewee

@@
 We performed the Wilcoxon test globally (total species richness) and on each phylum independently.
-
+% Reviewer suggests that the paragraoh below belongs to the discussion
+{\color{red} RYAN: Move paragraph below to discussion?}
 Taxa surveyed are a function of our metabarcoding PCR primers \cite{leray_new_2013} and reflect the current status of genetic databases, rather than a complete sampling of the planktonic community; we view these results as a cross-section of common taxa useful for understanding the biological effects of ocean conditions.
@@
 2C).
-Communities 2 and 3 had relatively higher values for taxa from a mix of groups including \textit{Emiliana huxleyi}, \textit{Minutocellus polymorphus}, \textit{Micromonas pusilla}, and \textit{Bathycoccus prasinos}.
+Communities 2 and 3 had relatively higher values for taxa from a mix of groups including \textit{Emiliania huxleyi}, \textit{Minutocellus polymorphus}, \textit{Micromonas pusilla}, and \textit{Bathycoccus prasinos}.
 Community 3 (Fig.
@@
 \centering
-\includegraphics[width=1\linewidth]{	Figures_for_MS/Figure2.2020-09-03.png}
+\includegraphics[width=1\linewidth]{	Figures_for_MS/Fig2.png}
 \caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal.
@@
 Likely winners under future conditions are more widely scattered among higher taxonomic groups.
-The haptophytes \textit{E.
-huxleyi} and \textit{Chrysochromulina} sp.
-the dinoflagellate \textit{Alexandrium sp.} each finds more suitable habitat in both of our study areas.
+The haptophytes \textit{Emiliania huxleyi} and \textit{Chrysochromulina} sp.
+the dinoflagellate \textit{Alexandrium} sp.
+all find more suitable habitat in both of our study areas.
 Among others not shown in Fig 3, \textit{Chaetoceros} (diatom) and many hydrozoans (Cnidaria) likely increase in San Juan Island, and the potentially fish-killing heterokont flagellate \textit{Pseudochattonella} increases in both study areas.
@@
  \centering
- \includegraphics[width=1\linewidth]{Figures_for_MS/Top_fig3_2020-09-03.Other.sel.png}
+ \includegraphics[width=1\linewidth]{Figures_for_MS/Fig3.png}
 \caption{\footnotesize{Forecasted changes in plankton in the Salish Sea for Dinoflagellates (panel \textbf{A}), Diatoms (\textbf{B}) and Haptophytes (\textbf{C}).
@@
 Our results reflect patterns in a diverse selection of species from nearshore marine communities in the Salish Sea, consisting of 222 planktonic taxa obtained from the metabarcoding analysis of 227 discrete samples across 77 space-time points (eight sites, 1.5 years).
-We find that changes in the composition of biological communities closely mirrored the variation in pH and temperature, with clear winners (e.g., \textit{Emiliana huxleyi}, \textit{Alexandrium}, and others) and losers (many, but not all, diatoms) likely to shift the structure and function of future marine communities.
+We find that changes in the composition of biological communities closely mirrored the variation in pH and temperature, with clear winners (e.g., \textit{Emiliania huxleyi}, \textit{Alexandrium}, and others) and losers (many, but not all, diatoms) likely to shift the structure and function of future marine communities.
 

### 2020-09-10 Manuscript/Response.to.reviewers.tex
response

@@
 If these differ, how do we interpret the others in the supplement?
-The legend doesn?t explain the colors in Fig 2C.}
+The legend doesn't explain the colors in Fig 2C.}
 
@@
 
-{\color{red} TODO - check }
+{\color{red}RGS: Agreed - it will definitely have an effect on the `fate` of the photosynthesized carbon- a quickly sunk carbon will join marine sediments instead of being recycled in teh upper layers of the water masses.
+We now cite the work of Polimene and colleagues on the role of phytoplankton and carbon sequestration  }
 
@@
 
-{\color{red} TODO - check }
+{\color{red}RGS: In their work, D.
+Anderson and colleagues cite a range of consequences of increased HAB blooms, including fish mortality and anoxia - and not eutrophization as we wrongly pointed out.
+Thanks for noticing.
+On the second argument - we saw potential consequences in two ways: by affecting recruiting of benthic species with pelagic larvae with reduced diatom diversity/biomass; and by changing the nature of sinking matter for filter feeders.
+We realize that bentho pelagic coupling is not the only source of carbon for benthic communities in the area of study so (RYAN, check) we have decided to remove this assertion from the discussion }
 
@@
 
-{\color{red} TODO - check }
+{\color{red} We thank the reviewer for this suggestion }
 
@@
 
-{\color{red} TODO - check }
+{\color{red} Done }
 
@@
 
-{\color{red} TODO - check }
+{\color{red} RGS: We used the eDNAindex (the proportion of reads assigned to each taxa per sample, scaled to the maximum proportion that taxa achieved).
+We provide the R implementation of this index in the Supplementary material, and have noted so in the text (line 128) }
 
@@
 
-{\color{red} TODO - check }
+{\color{red} RGS: We apologize for the broken github code - We have clarified how we simulated the future projection in the Supplementary Information - Also see response to reviewer1 on line 122 - 137 of this letter  }
 

### 2020-09-10 Manuscript/What.lies.ahead.reviewed.tex
response

@@
 We characterized sample carbonate chemistry by measuring Total Alkalinity (TA; open-cell automated titration based on a 876 Dosimat plus (Metrohm AG) as part of a custom system assembled by Andrew Dickson (UCSD) and used in the laboratory of Alex Gagnon at UW) and Dissolved Inorganic Carbon (DIC; Apollo Instruments, USA; CO$_2$ extraction system with 10\% (v/v) phosphoric acid).
-Both measurements were calibrated and validated with certified reference material from the Scripps Oceanographic Institute.
+Both measurements were calibrated and validated with certified reference material from the Scripps Institution of Oceanography.
 Using DIC and TA, we calculated pH and the remaining carbonate system parameters using the R package `seacarb' \cite{seacarb}.
@@
 Our sampled areas differed in the environmental variables driving changes in carbonate chemistry.
-San Juan Island was less seasonally variable than the Hood Canal in every measured parameter (Figure 1B); the island is more directly affected by summer coastal upwelling as a function of bathymetry and circulation patterns \cite{murray2015inland}, and this appears to be the dominant influence on carbonate chemistry there.
+San Juan Island was less seasonally variable than the Hood Canal in every measured parameter (Figure 1C); the island is more directly affected by summer coastal upwelling as a function of bathymetry and circulation patterns \cite{murray2015inland}, and this appears to be the dominant influence on carbonate chemistry there.
 By contrast, photosynthesis and respiration likely drive much of the carbonate chemistry variation in the Hood Canal (See Supporting Information).
@@
 We used this index of abundance in the multivariate community analysis, and used binary (presence/absence) data to capture species-level responses to environmental conditions.
+{\color{red}An implementation in R of this index is provided in the Supplementary information.}
 
@@
  Our results therefore suggest a possible change in relative dominance between diatoms and other phytoplankton species such as dinoflagellates, consistent with those seen at ecological regime shifts found elsewhere \cite{wasmund2017diatom, hinder2012changes}.
-Such a shift could affect ecosystems in many ways; even under the assumption that the surviving taxa would maintain the primary production levels, for example, the smaller cell-size of dinoflagellates and the differential sinking rates of the two groups would likely alter regional patterns of nutrient cycling \cite{xiao2018warming, anderson2002harmful, bienfang1984sinking}.
+Such a shift could affect ecosystems in many ways; even under the assumption that the surviving taxa would maintain the primary production levels, for example, the smaller cell-size of dinoflagellates and the differential sinking rates of the two groups would likely alter regional patterns of nutrient cycling and carbon sequestration \cite{xiao2018warming, anderson2002harmful, bienfang1984sinking, polimene2017biological}.
 Although the north Atlantic has shown an increase in diatom abundance \cite{hinder2012changes}, the increase in wind stress and associated mixing in the water column in the open ocean is unlikely to occur in the Hood Canal, where stratification is the strongest in the Salish Sea \cite{moore2008descriptive}.
@@
  Changes in environmental conditions and associated shifts in planktonic communities will likely reshape ecosystems and food webs, although some environmental processes may be conserved even as the particular taxa change.
-A switch from a diatom-dominated ecosystem to one in which dinoflagellate blooms extend in space and time could provoke cascade effects \cite{xiao2018warming} including fish mortality, eutrophication \cite{anderson2002harmful}, and bentho-pelagic coupling \citep[][]{bienfang1984sinking}.
+A switch from a diatom-dominated ecosystem to one in which dinoflagellate blooms extend in space and time could provoke cascade effects \cite{xiao2018warming} including fish mortality, anoxia \cite{anderson2002harmful}, and {\color{red} carbon sinking dynamics \st{alter current bentho-pelagic coupling processes} \cite{bienfang1984sinking}}.
 Beyond the phylum-specific patterns, the increase in suitable habitat for harmful algae species will alone be an engine for ecosystem change \cite{trainer2019pelagic, wells2018harmful}.

### 2020-09-12 Scripts/Denoising.all.runs.Rmd
jags ready

@@
 ```{r load libraries, include=FALSE}
- knitr::opts_chunk$set(warning = FALSE)
+ knitr::opts_chunk$set(warning = FALSE, message = F)
 
@@
   all.reps %>% 
-  filter(prob < 0.1 & tot < normparams.reads[1])
+  filter(prob < 0.075 & tot < normparams.reads[1])
 
@@
   unnest()
+
 ```
@@
 cleaned.tibble %>%
+  ungroup %>% 
   summarise(n_distinct(sample), # 770
@@
 
-to_write_discarded <- as.tibble(all_distances[outliers]) %>% rownames_to_column("sample") %>% dplyr::select(sample, 
-                                                                                                     distance_to_centroid = value)
+to_write_discarded <- as.data.frame(all_distances[outliers]) %>% rownames_to_column("sample") %>% dplyr::select(sample, 
+                                                                                                     distance_to_centroid = 2)
 to_write_discarded <- to_write_discarded %>% bind_rows(tibble(sample = discard.1,

### 2020-09-12 Scripts/From.asv.to.data.plankton.Rmd
jags ready

@@
 title: "From ASV to data.plankton"
-output: html_notebook
+output:
+  html_document:
+    df_print: paged
 ---
 
-This script will take the ASV table from the pipeline, and combine it with three metadata sources:
+This script will take the ASV table from the script Denoising.all.runs.Rmd, and combine it with three metadata sources:
   * Taxonomical annotation
@@
 library (here)
+library (lubridate)
 source(here("Scripts", "eDNAindex.R"))

### 2020-09-12 Scripts/Manuscript.Main.Analysis.Rmd
jags ready

@@
 This Markdown will make the part of the analysis that goes in the Manuscript: The starting point are the Abundance Table (number of reads per ASV per sample) and the key file (the ASV file has the Hash of the sequences, and the key has the actual DNA sequence).
-If you want to process the raw fastq files to obtain these two files, you have to use the pipeline available in http::/github.com/ramongallego/demultiplexer_for_dada2
+If you want to process the raw fastq files to obtain these two files, you have to use the pipeline available in http::/github.com/ramongallego/demultiplexer_for_dada2.
+
+It also generates some Supplementary Figures and data: Community composition analysis for the San Juan Islands (only Hood Canal in Figure 2 in the Manuscript) and phylum-specific plots of diversity projections and species-specific responses.
 

### 2020-09-12 Scripts/Rjags.tunning.Rmd
jags ready

---
title: "Rjags parameters"
output: html_notebook
---
Adjusting rjags parameters to see how they affect the prob of presence of a particular ASV in the dataset

The jags model feeds from two main sources:

  * The presence / absence matrix

  * The initial parameters for the true positive rate and the false positive rate

We can modify the matrix by subsetting the data and doing the detection of an ASV in a month or Site - and later keep all hashes that were true (prob of Occ > 0.8) for a Site or Month.

We can modify the initial parameters by hard-coding a maximum value for the false positive rate; or by picking it up from a left-skewed beta distribution
```{r}
library(tidyverse)
library(vegan)
library(rjags)
library(proxy)
library(here)

```

## Custom functions
```{r}
model.round.nested <- function(list.hashes,n){
  require(dplyr)
  tibble.out <- paste("model", n , sep=".")
  tibble.in <- "data"
  tibble.in <- rlang::ensym(tibble.in)
  list.hashes <- mutate (list.hashes, !!tibble.out := map (!!tibble.in,  ~jags_for_presence(.x)))
}
sink("RoyleLink_prior.txt")
										cat("model {
										    # Priors
										     psi ~ dunif(0,1)
										    p11 ~ dunif(0.01,1)
										    p10 ~ dbeta(1,50)

										    # Likelihood 
										    for (i in 1:S){
										    z[i] ~ dbern(psi)
										    p[i] <- z[i]*p11 + (1-z[i])*p10
										    for (j in 1:K){
										    Y[i,j] ~ dbern(p[i])
										    }
										    }
										    } ",fill=TRUE)
										sink()

jags_for_presence <- function(.x){
  .x %>% 
    transmute( model =  map_dbl (data,
                             function(.y, doprint=FALSE, ni=3000,nt=2,nc=1,nb=1000,myparallel=TRUE){

                               .y %>% ungroup %>% dplyr::select(-biol) -> .y # Reduce the tibble to just the presence/abs matrix

                               jags.inits <- function()(list(psi=runif(1,0.05,0.95),p11=runif(1, 0.01,1),p10=rbeta(1,1,50))) # generates three random starting estimates of psi, p11 and p10
                               jags.data <- list (Y= .y,
                                                  S = nrow(.y),
                                                  K = ncol(.y)) 
                               jags.params <- c("psi","p11","p10")
                               model<-jags(data = jags.data, inits = jags.inits, parameters.to.save= jags.params, 
                                           model.file= "RoyleLink_prior.txt", n.thin= nt, n.chains= nc, 
                                           n.iter= ni, n.burnin = nb, parallel=myparallel)

                               psihat <- model$summary["psi","50%"]
                               p11hat <- model$summary["p11","50%"]
                               p10hat <- model$summary["p10","50%"]    
                               modelSummaries <- model$summary

                               nObs   <- max(rowSums(.y))
                               K <- ncol(.y)

                               model.output <- ProbOcc(nObs, psihat, p11hat, p10hat, K)

										           return(model.output)

                             }))

}
```

### Change false positive rate

The Time it takes to calculate the fate for each Hash is enormous. I had to run it several times until I realized that running an Occupancy model for each Hash in a matrix of 261 samples x 3 PCR replicates will take ~ 8 days. To get to that conclusion I run several tests which I have stored in a different Rmarkdown ("test.Occupancy.Rmd").

My conclusions from that exercise are:

  * It is better to run all Miseq runs at once

  * What matters is the pattern of replication, ie in how many biological replicates are there 3 presences in 3 technical replicates, how many 2 presences and how many only once. So we can summarise the pattern of presence of each Hash with a A.B.C pattern, which is how many times apears once, twice and three times. And we can run the model only once per replication pattern

  * When there are few presences in the dataset, the output of the model is extremely variable

So we will start by creating two objects, one is the nested dataframe that has the 261*3 matrices, and the other one is the reduced information with the pattern of replication
```{r}
# Load the dataset
ASV.nested <- read_rds(here("Input","Cleaning.before.Occ.model"))
# Split by site
ASV.nested %>% 
  ungroup() %>% 
  select(Step2.tibble) %>% 
  unnest(Step2.tibble) %>%
  ungroup() %>% 
  separate(sample, into = c("biol","rep"), sep = "\\.") %>%
  separate(biol, into = c("Site", "biol")) %>% 
  group_by(Site) %>% 
  nest() -> nested.by.site

# For each Hash - calculate the pattern of replication

nested.by.site %>% 
  mutate(data = map(data,  function(.y){
        .y %>% 
          mutate (nReads = 1) %>% 
  spread(key = "rep", value = "nReads", fill= 0,drop = F) %>% 
      mutate (ndetections = `1`+`2`+`3`) %>% 
      group_by(ndetections, Hash) %>% 
      summarise(tot = sum(!is.na(ndetections)))

      })) -> nested.by.site
# And finally calculate the pattern of presence
nested.by.site %>% 
  unnest() %>% 
  spread(ndetections, tot,fill = 0) %>% 
  unite(repetition.level,`0`,`1`,`2`,`3`, sep = ".") -> Pattern.of.presence

# Choose one representative
Pattern.of.presence %>% 
  group_by(repetition.level) %>%
  summarise(first = head(Hash,1),
            Site =  head(Site,1)) %>%
  unite(Site,first,col = "key", sep = ".", remove = F)-> unique.by.site

# Subset 
 ASV.nested %>% 
   ungroup() %>% 
   select(Step2.tibble) %>% 
  unnest(Step2.tibble) %>% 
  ungroup() %>% 
  separate(sample, into = c("biol","rep"), sep = "\\.") %>%
  separate(biol, into = c("Site", "biol")) %>% 
  group_by(Site) %>% 
  nest() %>% 
  mutate(data = map(data,
                    function(.y){
                      .y %>% 
                        mutate (nReads = 1) %>% 
                        spread(key = "rep", value = "nReads", fill= 0,drop = F)
  })) %>% 
  unnest() %>% 
  unite(Site, Hash, col = "key", sep = ".", remove = F) %>% 
# Select one  
  filter(key %in% unique.by.site$key) %>% 
  select(-key) %>% group_by(Site, Hash) %>% 
  nest %>% 
  group_by(Site) %>% 
  nest -> like.this 
```
Finally - run it ten times
```{r}
like.this %>% 
  mutate(Hashes = map (data , ~select(.x, Hash))) -> like.this
# Run 10 times
for (i in 1:10){
  like.this <- model.round.nested(like.this, n = i)
}
like.this %>% mutate(model.10chain = map(data, jags_for_presence)) -> all.combos.10chain
```

Now the first step is to compare these values with the other two treatments in which we pooled all sites together

```{r}

all.combos.10chain %>% 
  unnest (Hashes, model.10chain ) %>% 
 # select(Site, Hash, contains("chain")) %>% 
 # filter(Hash %in% test.for.occ.modeling.p10.beta.50$Hash) %>% 
#  gather(contains("model"), key = "Attempt", value = "Output") %>% 
  ggplot(aes(x = Hash, y = model, color = Site))+
  geom_boxplot(position = "dodge",
               outlier.color = "red") + 
  theme(axis.text.x = element_blank(), 
        axis.ticks.x = element_blank()) +
  geom_hline(yintercept = 0.8, color = "red") +
  facet_wrap(~Site, ncol = 1)

```

That looks nice - so this is the process we will follow: Analyze the Prob of Ocurrence for each Hash and Site - and if a Hash has a value in a Site greater than 0.8, then we will keep it.
```{r}
like.this %>% 
  unnest %>% 
  select(Site, Hash, contains("model")) %>% 
  write_csv("Occ.all.hashes.by.site.csv")

all.combos.10chain %>% 
  unnest (Hashes, model.10chain ) %>% 
  write_csv("Occ.all.hashes.10chain.csv")

```

Now let's rejoin the Hash.site combo with the original dataset, and filter out those hashes that has a 25% quantile lower 0.8

```{r}

#Hash.Site.combos <- read_csv("Occ.all.hashes.by.site.csv")
Hash.Site.combos <- read_csv("Occ.all.hashes.10chain.csv")

Hash.Site.combos %>%
  unite(Site,Hash, sep = ".", col = "key") %>% 
#  gather(contains("model"), key = "Attempt", value = "Output") %>% 
#  group_by(key) %>%
 # nest %>%
#  mutate(Q25 = map_dbl(data, 
 #                                   function(.x){
#                                      quantile(.x$Output, probs = 0.25)
 #                                   })) %>%
 # unnest(Q25) %>% #leftjoin this with the pattern of repetition
  left_join(Pattern.of.presence %>%
              unite(Site, Hash, sep = ".", col = "key") %>%
              select(key, repetition.level)) %>% 
  select(repetition.level, model) %>% # and now right join this with all Hashes
  right_join(Pattern.of.presence, by = "repetition.level") %>% 
  group_by (Hash) %>% 
  summarise (real = max (model, na.rm = T)) -> Occ.model.fate

Occ.model.fate %>% 
  arrange((real))

Occ.model.fate %>% write_csv("Occ.fate.csv")

```

### 2020-09-12 Scripts/binomial_richness_demo.R

@@ -1,0 +1,31 @@
+#binomial richness demo
+
+#suppose you have 3 species: A, B, and C
+#each has a different probability of presence
+
+probA <- 0.2
+probB <- 0.8
+probC <- 0.5
+
+
+#if we treat Richness as the number of species with prob > 0.5, the answer is always the same. R = 1. 
+
+#if we treat Richness as probabilistic, we get a distribution of Richnesses that are possible. If we do 100 draws, we get:
+
+Richness <- NA
+for (i in 1:100) {
+  Richness[i] <-
+    sum (
+    rbinom(1,1,probA),
+    rbinom(1,1,probB),
+    rbinom(1,1,probC)
+  )
+}
+
+hist(Richness)
+mean(Richness)
+median(Richness)
+
+
+#so, instead of doing the total number of species at each parameter set w prob > 50%, we should be using all of those 100 draws at 
+#each set of environmental parameters to calculate 100 different richnesses, and taking a mean.

### 2020-09-12 Scripts/Rjags.tunning.Rmd
models by site

@@
 library(here)
+library(unmarked)
+library(jagsUI)
 
@@
   list.hashes <- mutate (list.hashes, !!tibble.out := map (!!tibble.in,  ~jags_for_presence(.x)))
+}
+
+ProbOcc <- function(x, psi, p11, p10, K){
+ (psi*(p11^x)*(1-p11)^(K-x)) / ((psi*(p11^x)*(1-p11)^(K-x))+(((1-psi)*(p10^x))*((1-p10)^(K-x))))
 }
@@
 like.this %>% mutate(model.10chain = map(data, jags_for_presence)) -> all.combos.10chain
+
+all.combos.10chain %>% write_rds(here("Input", "10chain.rds"))
+
+all.combos.10chain %>% unnest(cols = c(Hashes, starts_with("model")), names_repair = "unique") -> row.df
+
+row.df %>% ungroup() %>% rowwise(Site, Hash) %>% summarise(mean = mean(c_across(where(is.numeric))))
 ```

### 2020-09-12 Scripts/Rjags.tunning.Rmd
without site

@@
 output: html_notebook
+editor_options: 
+  chunk_output_type: console
 ---
@@
 ```
+
+## Without Site into the mix
+```{r}
+ASV.nested %>% 
+    ungroup() %>% 
+  select(Step2.tibble) %>% 
+  unnest(Step2.tibble) %>%
+
+  separate(sample, into = c("biol","rep"), sep = "\\.") %>%
+  mutate(nReads = 1) %>% 
+  spread(key = "rep", value = "nReads", fill= 0,drop = F) %>% 
+  group_by(Hash) %>%
+  mutate(ndetections = `1`+`2`+`3`) %>% 
+   group_by(ndetections, Hash) %>%
+  summarise(tot = sum(!is.na(ndetections))) %>% 
+  spread(ndetections, tot,fill = 0) %>% 
+  unite(repetition.level,`1`,`2`,`3`, sep = ".") -> Pattern.of.presence2
+```
+
+```{r}
+Pattern.of.presence2 %>% 
+  group_by(repetition.level) %>% 
+  select(-`0`) %>% 
+  summarise(first = head(Hash, 1)) -> key.fromrep.to.hash
+key.fromrep.to.hash %>% 
+  pull(first) -> Hashes.to.run
+
+```
+
+```{r}
+ASV.nested %>% 
+    ungroup() %>% 
+  select(Step2.tibble) %>% 
+  unnest(Step2.tibble) %>%
+
+  separate(sample, into = c("biol","rep"), sep = "\\.") %>%
+  mutate(nReads = 1) %>% 
+  spread(key = "rep", value = "nReads", fill= 0,drop = F) %>% 
+  filter (Hash %in% Hashes.to.run) %>% 
+  nest(-Hash) -> Nested.Hashes
+
+for (i in 1:10){
+  all.hashes <- model.round.nested(Nested.Hashes, n = i)
+}
+```

### 2020-09-14 Scripts/Rjags.tunning.Rmd
no site
Aaargg

@@
 
+Nested.Hashes %>% 
+  mutate (susp = "Yes") %>% 
+  nest(-susp) -> Nested.Hashes
+
 for (i in 1:10){
-  all.hashes <- model.round.nested(Nested.Hashes, n = i)
+  Nested.Hashes <- model.round.nested(Nested.Hashes, n = i)
 }
+
+write_rds(Nested.Hashes, here("Input", "Occupancy.nosite.rds"))
+
+all.hashes <- read_rds(here("Input", "Occupancy.nosite.rds"))
+
+all.hashes ->test
+
 ```

### 2020-09-22 Scripts/Denoising.all.runs.Rmd
I think we good

@@
  library (seqinr)
- library(patchwork)
+ library (patchwork)
 
@@
   distinct(Hash, .keep_all = T) -> Hash.key
+
+# Write the Hash key from start
+
+Hash.key %>% 
+  write_csv(here("Output", "starting.hash.key.csv"))
+
 ```
@@
 
-occ.results <- read_csv("../Input/Occ.fate.csv")
+occ.results <- read_csv("../Input/Occ.fate.2020.csv")
 
 occ.results %>% 
-  ggplot(aes(x = real)) +
+  ggplot(aes(x = max)) +
   geom_histogram(fill = "blue") +
@@
               summarise (tot = sum(nReads))) %>% 
-  ggplot(aes(x = cut_interval(real, n = 20))) +
-  geom_col(aes(y = tot), fill = "red") +
+  ggplot(aes(x = cut_interval(max, n = 20))) +
+  geom_col(aes(y = tot), fill = "red", position = "stack") +
 
@@
  occ.results %>% 
-  filter(real > 0.8) %>% 
+  filter(max > 0.8) %>% 
   pull (Hash) -> to.keep
@@
 ```
-Ok, so there are 13 samples for which we only have 2 PCR replicates1.   We will get rid of those with only 1, as we can't estimate the PCR bias there.
+Ok, so there are 13 samples for which we only have 2 PCR replicates1.   We will get rid of those with only 1, as we can't estimate the PCR bias there. THis is
 
@@
 
-#names(all.distances.full)
-
-summary(is.na(names(all.distances.full)))
+# Do all samples have a name?
+summary(is.na(names(all.distances.full))) # Yes they do
 ```
@@
 
-as.tibble(subset(melt(as.matrix(all.distances.full)))) -> all.distances.melted
+as_tibble(subset(melt(as.matrix(all.distances.full)))) -> all.distances.melted
 
+# Any mjor screw ups
 summary(is.na(all.distances.melted$value))
@@
 
-sapply(all.distances.to.plot, function(x) summary(is.na(x)))
+sapply(all.distances.to.plot, function(x) summary(is.na(x))) # good boi
 
@@
 
-  ggplot (all.distances.to.plot , aes (fill = Distance.type, x = value)) +
-  geom_histogram (position = "dodge", stat = 'density', alpha = 0.9) +
+  ggplot (all.distances.to.plot ) +
+  geom_histogram (aes (fill = Distance.type, x = value, after_stat(ndensity)), position = "dodge",  alpha = 0.9, bins = 50) +
   facet_wrap( ~ Distance.type) +
   labs (x = "Pairwise dissimilarity", y = "density" ,
-        Distance.type = "Distance")
+        Distance.type = "Distance") +
+    guides (fill = "none")
 ggsave("visual.anova.png", dpi = "retina")
 ```
+
+So our the distribution of dissimilarities is as we expected : lowest in technical replicates, then biological replicates and higher across our study system.
+Now let's see if there are any technical replicates that should be discarded due to their higher dissimilarity.
+We will calculate the distance from each PCR replicate to their group centroid, fit those distances to a normal distribution and discard values that are too high
 
@@
 
+hist(all_distances)
 ```
@@
   geom_raster(aes(x= Date, y = Site, fill = cases))+
-  geom_text(aes(x= Date, y = Site, label = cases))
+  geom_text(aes(x= Date, y = Site, label = cases)) +
+  facet_wrap(~Area, ncol = 1, scales = "free_y")
 

### 2020-09-22 Scripts/Manuscript.Main.Analysis.Rmd
I think we good

@@
 #functions for future Salish Sea
-models <- readRDS(here("Input","finalModel_20200807.RDS"))
+models <- readRDS(here("Input","bestModelFit.rds"))
 
@@
 taxa.to.follow %>% 
+  mutate(Comparison = case_when(Comparison %in% c("1_2", "2_1") ~ "1_2",
+                                Comparison %in% c("1_3", "3_1") ~ "1_3",
+                                Comparison %in% c("3_2", "2_3") ~ "2_3")) %>% 
   ggplot(aes(x = average)) + 

### 2020-09-22 Scripts/Prepare.data.for.stan.Rmd
I think we good

@@
 library(tidyverse)
-
+old<- read_csv(here("Input","data.for.stan.csv"))
 plankton <- read_csv(here("Input","Combined_Biol_Env_Plankton.csv"))

### 2020-09-22 Scripts/Rjags.tunning.Rmd
I think we good

@@
 
-We can modify the matrix by subsetting the data and doing the detection of an ASV in a month or Site - and later keep all hashes that were true (prob of Occ > 0.8) for a Site or Month.
+We can modify the matrix by subsetting the data and doing the detection of an ASV in a month or Site - and later keep all hashes that were true (prob of Occ > 0.8).
 
-We can modify the initial parameters by hard-coding a maximum value for the false positive rate; or by picking it up from a left-skewed beta distribution
 ```{r}
@@
 
+all.combos.10chain <- read_rds(here("Input", "10chain.rds"))
+
 all.combos.10chain %>% unnest(cols = c(Hashes, starts_with("model")), names_repair = "unique") -> row.df
+
+row.df %>% 
+  left_join(unique.by.site, by = c("Hash" = "first", "Site")) %>% 
+  ungroup() %>% 
+  select(-Hash, -Site) %>% 
+  right_join(Pattern.of.presence) %>% 
+  group_by(Hash) %>% tally(name = "max", sort = T)
 
@@
 all.hashes <- read_rds(here("Input", "Occupancy.nosite.rds"))
+```
 
-all.hashes ->test
+```{r}
+all.hashes %>% unnest(cols = c(data, starts_with("model")), names_repair = "unique") -> row1.df
 
+row1.df %>% ungroup() %>% rowwise( Hash) %>% summarise(mean = mean(c_across(where(is.numeric))), 
+                                                       median =median(c_across(where(is.numeric))),
+                                                       max  = max(c_across(where(is.numeric)))) %>%
+  left_join(Pattern.of.presence2, by = "Hash") %>% # this adds the output to the key
+  ungroup %>% select(-Hash) %>% 
+  left_join(Pattern.of.presence2) %>% write_csv(here("Input", "Occ.fate.2020.csv")) # this adds the output to the Hashes
 ```

### 2020-09-22 Scripts/Supplement_ModelTesting.Rmd
I think we good

@@
              areaIndex = match(Area, unique(Area)),
-             seasonIndex = match(Season, unique(Season)),
+             #seasonIndex = match(Season, unique(Season)),
              taxonAreaIndex = match(AreaTaxa, unique(AreaTaxa))) %>%   #unique combinations of taxon and area
@@
 #uncomment if running de novo; otherwise, load data from file, above
-# finalModel <- stan_glmer(binary ~ 0 + (0 + Temperature + pH | taxa) + (1 | taxonAreaIndex),
-#                              data = stan.input,
-#                              family = "binomial",
-#                              prior = t_prior,
-#                              prior_intercept = t_prior,
-#                              chains = 4,
-#                              adapt_delta = 0.99,
-#                              refresh = 0)
-#optionally save
-#save(finalModel, file = "bestModelFit.Rdata")
+finalModel <- stan_glmer(binary ~ 0 + (0 + Temperature + pH | taxa) + (1 | taxonAreaIndex),
+                             data = stan.input,
+                             family = "binomial",
+                             prior = t_prior,
+                             prior_intercept = t_prior,
+                             chains = 4,
+                             adapt_delta = 0.99,
+                             refresh = 0)
+# optionally save
+write_rds(finalModel,path = here("Input","bestModelFit.rds"))
 ```

### 2020-09-24 Manuscript/What.lies.ahead.reviewed.tex
Rmarkdown shows consistency and figures are good to go. Latex document updated

@@
 
+%% All changes due to new run of data are shown with *** before them for ease of search
 Despite geographic proximity and similar overall species composition (127 of the 222 planktonic taxa were found in both regions and accounted for 98\% of the sequences), the areas under study -- San Juan Island and the Hood Canal -- varied widely in pH, temperature, and other environmental parameters (Figure 1C), with a smooth gradient in conditions along the Hood Canal, and San Juan Island more closely resembling full marine conditions.
@@
 
-Metabarcoding analysis of eDNA samples generated more than 52.5M sequences from 745 samples.
+Metabarcoding analysis of eDNA samples generated more than ***50.8M sequences from 778 samples.
 These samples represented biological and technical replicates from 86 unique sampling events.
-After bioinformatics quality-control (See Methods) the dataset included $\sim${}41M sequences, from 4782 unique amplicon sequence variants (ASVs).
-Of these, 1374 ASVs (22.8M reads) could be annotated to a taxonomic level of Family or lower.
+After bioinformatics quality-control \st{(See Methods)} the dataset included $\sim${}***45M sequences, from ***4849 unique amplicon sequence variants (ASVs).
+Of these, ***1364 ASVs (22.6M reads) could be annotated to a taxonomic level of Family or lower.
 These $\sim$ 500 taxa from 43 phyla were split according to their natural history and habitat (benthic vs.
 planktonic; see Supporting Information).
-Because we expect planktonic taxa to vary with water mass \cite{kelly2018tides} and therefore with bottle-sampled carbonate chemistry, here we focus on only the planktonic taxa (N = 222).
+Because we expect planktonic taxa to vary with water mass \cite{kelly2018tides} and therefore with bottle-sampled carbonate chemistry, here we focus on only the planktonic taxa (N = 221).
 These taxa showed a seasonal richness gradient between study areas, consistent with documented biodiversity clines in the area (Fig.
@@
 
-Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to geographic Area (Hood Canal vs.
-San Juan Island; F-model = 5.4; p $<$ 0.0001).
+Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to geographic Area ***(Hood Canal vs.
+San Juan Island; F-model = 1.6184; p $<$ 0.01).
 We therefore performed a constrained analysis of principal components (CAP) for samples within each Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig.
@@
 These communities were distinguished by changes in the relative abundances of a wide variety of taxa.
-In the Hood Canal for example, the colder water and higher-pH community (community 1 in Fig.
-2C) showed higher eDNA indexes of diatoms \textit{Thalassiosira nordenskioeldii}, \textit{Nitzschia cf.
-promare}, and \textit{Ditylum brightwellii} -- among other species -- relative to a higher-temperature community (community 3 in Fig.
-2C).
-Communities 2 and 3 had relatively higher values for taxa from a mix of groups including \textit{Emiliania huxleyi}, \textit{Minutocellus polymorphus}, \textit{Micromonas pusilla}, and \textit{Bathycoccus prasinos}.
+In the Hood Canal for example, the community linked with colder water and higher-pH (community **2 in Fig.
+2C) showed higher eDNA indexes of diatoms ***(\textit{Minutocellus polymorphus}), green algae  (\textit{Bathycoccus prasinos} and \textit{Micromonas pusilla}), and dinoflagellates like \textit{Karlodinium} sp than that on the low pH spectrum.
+In that community (cluster 1) \textit{Emiliania huxleyi}, \textit{Ditylum brightwellii} and the copepod \textit{Paracalanus} sp.
+C are more prevalent.
+The community thriving in the high temperature range (community 3 in Fig.
+2C) shows high values of \textit{Phaeocystis globosa},\textit{Ditylum brightwellii} among other species.
 Community 3 (Fig.
 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls.
-For example, we expect the Hood Canal in 2095 to have the conditions in which community 3 is the most likely community 75\% of the time, an increase of 22\% compared to 2017(see Supplemental Material).
+For example, we expect the Hood Canal in 2095 to have the conditions in which community 3 is the most likely community 66\% of the time, an increase of 11\% compared to 2017(see Supplemental Material).
+On the other hand, conditions that make cluster 2 the most likely community drop from 9\% to 1\% of the time between 2017 and 2095.
 
@@
 \centering
-\includegraphics[width=1\linewidth]{	Figures_for_MS/Fig2.png}
+\includegraphics[width=1\linewidth]{	Figures_for_MS/Fig2}
 \caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal.
@@
 These declines in diatom richness were more accentuated at lower pH values and higher temperatures.
-Other taxa likely to find less-suitable habitat in the future include dinoflagellates \textit{Karlodinium} and Gymnodiniaceae sp.
+Other taxa likely to find less-suitable habitat in the future include dinoflagellates \textit{Karlodinium} sp.
+and Gymnodiniaceae sp.
 
@@
 Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at overall richness variation; we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, 95\% confidence interval in median species richness  -0.08, +0.04; Wilcoxon p $=$ 0.5); while higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 1.84-2.2, p $< 10^{-16}$).
-Diatoms, in particular, show small but significant declines in richness in the Hood Canal (0.47-0.57 species, p $< 10^{-16}$), while the changes on the San Juan Island are negligible (median change 0-0.08 p $=$ 0.04).
+Diatoms, in particular, show small but significant declines in richness in the Hood Canal (0.46-0.55 species, p $< 10^{-16}$), while the changes on the San Juan Island are negligible (median change 0-0.08 p $=$ 0.04).
 Dinoflagellates see their richness increase in both regions with the future scenario (median change 0.14-0.18 Hood Canal; 0.21-0.25 San Juan Island; p $< 10^{-16}$ for each).
@@
  \centering
- \includegraphics[width=1\linewidth]{Figures_for_MS/Fig3.png}
+ \includegraphics[width=1\linewidth]{Figures_for_MS/Fig3}
 \caption{\footnotesize{Forecasted changes in plankton in the Salish Sea for Dinoflagellates (panel \textbf{A}), Diatoms (\textbf{B}) and Haptophytes (\textbf{C}).

### 2020-09-24 Scripts/From.asv.to.data.plankton.Rmd
Rmarkdown shows consistency and figures are good to go. Latex document updated

@@
 source(here("Scripts", "eDNAindex.R"))
-ASV.table  <- read_csv(here("Input", "ASV_table_all_together.csv"))
+ASV.table  <- read_csv(here("Output", "ASV_table_all_together.csv"))
 Annotation <- read_csv(here("Input", "hash.annotated.csv"))
@@
            pull(Hash))) -> ASV.table
+
+# How many Hashes have family or more, how many reads
+left_join(ASV.table, Annotation) %>% 
+  unite(family, genus, species, col = "taxa", sep = "|") %>% 
+  filter(taxa != "NA|NA|NA") %>% 
+  summarise(n_distinct(Hash), sum(nReads), n_distinct(taxa))
 
@@
 
+ASV.plankton.event %>%
+  ungroup() %>% 
+  summarise(n_distinct(taxa))
+
 env.data %>%  
@@
   write_csv(here("Input", "Combined_Biol_Env_Plankton.csv"))
+
 ```

### 2020-09-24 Scripts/Manuscript.Main.Analysis.Rmd
Rmarkdown shows consistency and figures are good to go. Latex document updated

@@
                                                    decostand(method = "standardize", na.action = na.omit),
-                                                 distance = "jaccard", 
-                                                binary = T)
+                                                 distance = "bray", 
+                                                binary = F)
 
@@
          ) -> Clusters.analysis.and.cap
-
+write_rds(Clusters.analysis.and.cap, here("Output", "CAP.halfway.rds"))
 # Write out clusers and CAP points, and raw env data
@@
 taxa.to.follow %>% 
-  filter (average > 0.01) %>%  
+ # filter (average > 0.01) %>%  
   arrange(desc(average)) %>% 
@@
     ggplot(.x, aes(y = average,
-             x = fct_reorder(label2, phylum))) +
+             x = fct_reorder(label2, label2))) +
   geom_col(aes(fill = diff>0), size = 0.5, color = "black") +
@@
 explore.phylum("Dinophyceae")
-
-#For example, we expect the Hood Canal to be T $>$ 12\degree C and pH $<$ {7.9} -- where community 3 is the most likely community -- more than four times as often in 2095 than in 2017 (4.9 \% vs 20.7\%) )
-
-stds_to_value$toValue[[3]]
-stds_to_value$toValue[[4]]
-
-richness.by.phylum.by.point %>% 
+```
+
+### Bonus track 3: Prevalence of clusters in future scenarios
+
+For example, we expect the Hood Canal to be T $>$ 12\degree C and pH $<$ {7.9} -- where community 3 is the most likely community -- more than four times as often in 2095 than in 2017 (4.9 \% vs 20.7\%) )s
+```{r}
+input.for.left %>% 
   ungroup() %>% 
-  dplyr::select(Area, Temperature, pH, Year) %>% 
+  select(Temperature = TemperatureCelsius,
+         pH_new = pH_units, 
+         Area,
+         Year, sim.row) %>%
   distinct() %>% 
-  filter(Area == "Hood Canal") %>% 
-  mutate(Temperature = 14.995 + 4.299*Temperature,
-         pH = 8.2157 + 0.1813*pH) %>% 
-  group_by(Year) %>% 
-  add_count() %>% 
-  summarise(condition = sum(Temperature > 12 & pH < 7.9),
-            totalcount = n()) %>% 
-  mutate(percent_time = 100 *condition / totalcount)
-```
-
-In general, the prevalence of cluster 3 in the Hood Canal will increase from 54.2% of the time to 75% of the time,
+    group_by(Area) %>% 
+    nest() %>% 
+    left_join(clusters) %>% 
+    mutate(whichComm.res = map2(data,multi.mod,~ .x %>% 
+                                  mutate(comm = as.numeric (apply(predict(.y, type = "probs", newdata = .x),1, which.max)) ))) %>% 
+  select(Area, whichComm.res) %>% 
+  unnest(whichComm.res) %>% 
+  add_count(Area, Year, name = "Tot") %>% 
+  add_count(Area, Year, comm, name = "case.counts") %>% 
+  group_by (Area, Year, comm) %>%
+
+  summarise(percent_time = 100 * (case.counts/ Tot)) %>% distinct()
+```
+
+<!-- ```{r} -->
+<!-- stds_to_value$toValue[[3]] -->
+<!-- stds_to_value$toValue[[4]] -->
+
+<!-- richness.by.phylum.by.point %>%  -->
+<!--   ungroup() %>%  -->
+<!--   dplyr::select(Area, Temperature, pH, Year) %>%  -->
+<!--   distinct() %>%  -->
+<!--   filter(Area == "Hood Canal") %>%  -->
+<!--   mutate(Temperature = 14.995 + 4.299*Temperature, -->
+<!--          pH = 8.2157 + 0.1813*pH) %>%  -->
+<!--   group_by(Year) %>%  -->
+<!--   add_count() %>%  -->
+<!--   summarise(condition = sum(Temperature > 12 & pH < 7.9), -->
+<!--             totalcount = n()) %>%  -->
+<!--   mutate(percent_time = 100 *condition / totalcount) -->
+<!-- ``` -->
+
+In general, the prevalence of cluster 3 in the Hood Canal will increase from 55.6% of the time to 66.4% of the time,
 ```{r echo = F}
@@
 
+**Overall richness, Hood Canal**
+We find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, averaged over all environmental conditions, Wilcoxon p $=$ 0.5)
 ```{r echo = F}
-#we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, averaged over all environmental conditions, Wilcoxon p $=$ 0.5);
+#;
 
@@
                     pull(richness), conf.int = T)
-
-#while higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 2.1, Wilcoxon p $<$ 10 ^{-16}).
+```
+
+**Overall richness, San Juan Islands**
+While higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 2.1, Wilcoxon p $<$ 10 ^{-16}).
+```{r echo = F}
+#
 
@@
                     pull(richness), conf.int = T)
-
-#Diatoms, however, show small but significant declines in richness in the Hood Canal (0.5 species, Wilcoxon p $< 10^{-16}$),
-
+```
+
+**Diatom richness, Hood Canal**
+Diatoms, however, show small but significant declines in richness in the Hood Canal (0.5 species, Wilcoxon p $< 10^{-16}$)
+```{r echo = F}
 wilcox.test(x = richness.by.phylum.by.point %>%
@@
                     pull(richness), conf.int = T)
-
-#while the changes on the San Juan Island are negligible (difference in median switch of 0.05, with  p $=$ 0.01).
+```
+
+**Diatom richness, San Juan Islands**
+While the changes on the San Juan Island are negligible (difference in median switch of 0.05, with  p $=$ 0.01).
+```{r echo = F}
+#
 
@@
                     pull(richness), conf.int = T)
-
-#Dinoflagellates see their richness increase in both regions with the future scenario(Wilcoxon p $< 10^{-16}$)
+```
+
+**Dinoflagellates richness, Hood Canal**
+Dinoflagellates see their richness increase in both regions with the future scenario(Wilcoxon p $< 10^{-16}$)
+```{r echo = F}
+#
 
@@
             conf.int= TRUE)
-
-#For example, in in the Hood Canal at present, we expect surface waters to have pH $<$ 7.9 and T $>$ 19$\degree$C only 1\% of the time.
+```
+
+For example, in in the Hood Canal at present, we expect surface waters to have pH $<$ 7.9 and T $>$ 19$\degree$C less than 1\% of the time.
 In 2095, we expect these conditions 6 times more frequently (i.e., 6\% of the time).
-
-stds_to_value$toValue[[3]]
-stds_to_value$toValue[[4]]
+```{r echo = F}
 
@@
             totalcount = n()) %>% 
-  mutate(percent_time = condition / totalcount)
-
-#At these values of T and pH, our model predicts the harmful \textit{Alexandrium} sp. to occur more often than not (mean frequency of occurrence $=$ 0.833)
+  mutate(percent_time = 100*(condition / totalcount))
+```
+
+#### Species prevalence in future climate conditions
+At these values of T and pH, our model predicts the harmful \textit{Alexandrium} sp. to occur more often than not (mean frequency of occurrence $=$ 0.833)
+```{r echo = F}
+#
 
@@
   summarise(mean(present))
-
-#By contrast, the large centric diatom \textit{Coscinodiscus} -- a potentially key source of carbon for zooplankton and small fishes \cite{pasquaud2010determination, zamon2002tidal} with effects on dissolved oxygen and other water-column characteristics \cite{manabe1991bloom} -- occurs only one-third of the time under these same conditions (mean frequency $=$ 0.35)
+```
+
+By contrast, the large centric diatom \textit{Coscinodiscus} -- a potentially key source of carbon for zooplankton and small fishes \cite{pasquaud2010determination, zamon2002tidal} with effects on dissolved oxygen and other water-column characteristics \cite{manabe1991bloom} -- occurs only one-third of the time under these same conditions (mean frequency $=$ 0.35)
+```{r echo = F}
+#
 expanded.future %>% 
@@
 
-# Responses in diatoms are not all in the same direction - Pseudonitzschia spp shows higher probability of presence at low pH values, and low pH values are more common in the 2095 scenario
-
+# 
+```
+
+Responses in diatoms are not all in the same direction -  \textit{Pseudonitzschia} spp shows higher probability of presence at low pH values, and low pH values are more common in the 2095 scenario.
+
+```{r echo = F}
 expanded.future %>% 
@@
 
-expanded.future %>%
-
-  mutate (Area = str_replace(Area, replacement = "San Juan Island", pattern =  "SJI")) %>%
-  select(Area, Year, Temperature = TemperatureCelsius,
-         pH_new = pH_units) %>% 
-  distinct() %>% 
-    group_by(Area) %>% 
-    nest() %>% 
-    left_join(clusters) %>% 
-    mutate(whichComm.res = map2(data,multi.mod,~ .x %>% 
-                                  mutate(comm = as.numeric (apply(predict(.y, type = "probs", newdata = .x),1, which.max)) )
-              ) ) %>% unnest(whichComm.res)
-```
+```

### 2020-09-24 Scripts/insect.all.Rmd
Rmarkdown shows consistency and figures are good to go. Latex document updated

---
title: "Growing the database and annotate sequences"
output: html_notebook
---

This notebook contains all the steps taken to taxonomically annotate the sequences found in the Hood Canal. We will start by loading the hash keys from all sequencing runs and create a fasta file with all unique sequences

```{r load libraries, echo = FALSE}

library (tidyverse)
library (insect)
library (seqinr)
library (googlesheets)
library (here)
```

```{r load objects by the end of the cleaning }

Hash     <- read_csv (here("Output", "Hash_Key_all_together.csv")) %>% 
  select(Hash, Sequence) %>% distinct()
ALL.ASVs <- read_csv (here("Output","ASV_table_all_together.csv"))
tree.2   <- read_rds (here("Input", "classifier.rds"))

# we will also load the previous classification so we don't do the same thing all over again

previous.effort <- read_csv(here("Input", "insect.2019.csv"))
```

So these sequences are in the same direction, and they have some shared and some new sequences.

Let's keep only the unique sequences

That's 4,849 unique sequences. THis is going to take a while

Make them into a DNAbin object for insect

```{r make it into a DNA object}

new.set <- anti_join(Hash, previous.effort, by = c("Hash" = "representative"))

all.hashes.insect <- char2dna(new.set$Sequence)

names (all.hashes.insect) <- new.set$Hash

all.hashes.insect

```

Now, we will load v3 of the metazoan classification tree from insect v1.1.1

```{r taxonomy import}

#tree <- readRDS ("/Users/ramongallego/GoogleDrive/Collaborations/insect/metazoan_COI_marine_v3/classification_tree.rds")

```

And finally, classify all sequences in our bin file

```{r classify}

clasif.hashes <- classify (x = all.hashes.insect, tree = tree, cores = 4)
clasif.hases.new.tree <- classify(x = all.hashes.insect, 
                                  tree = tree.2,
                                  cores = "autodetect")

clasif.hases.new.tree %>% 
  bind_rows(previous.effort) %>% 
  write_csv(here("Input", "last.insect.csv"))

clasif.hases.new.tree %>% 
  unite (family, genus, species, sep = "|", col = "taxa")

clasif.hases.new.tree %>% dplyr::count (rank) %>% arrange(desc(n))
```
Run the translation classification

```{r classiffy translations}
INV <- getGeneticCode(id_or_name2 = "5")

Hash$AA <- as.character(suppressWarnings(translate(subseq(DNAStringSet(Hash$Sequence),start = 2),
                                                        genetic.code = INV,
                                                        no.init.codon = T)))

Hash %>% 
  mutate (stop.codons = str_count(string = AA, pattern = "\\*")) %>% 
  filter(stop.codons == 0) -> to.insect

all.hashes.insect.AA <- char2aa(to.insect$AA)

names (all.hashes.insect.AA) <- to.insect$Hash

all.hashes.insect.AA

```
And now, classify them
```{r classiffy translations2}

AAS.flatten <- lapply(to.insect$AA, char2aa)

names(AAS.flatten) <- to.insect$Hash

possibly_clasify <- possibly(~classify(x = .,
                                      tree = tree.2,
                                      cores = "autodetect",
                                      ping = 1),
                             otherwise = NA_real_)

AA.s.test <- AAS.flatten[51:60]

possibly.AAs <- map(AA.s.test, possibly_clasify)

clasif.hashes.AA <- possibly(classify(x = all.hashes.insect.AA, 
                                  tree = tree.2,
                                  cores = "autodetect", 
                             ping = 1,
                             species = "none"))

test <- all.hashes.insect.AA[19]

test.insect.AA <- classify(x= all.hashes.insect.AA, 
                           tree = tree.2,
                           ping = 1,
                           species = "none")

Hash$AA[19]
# So it breaks with stop codons
```

OK. So now let's save the classification object as an RDS

```{r save it for now}
clasif.hases.new.tree %>% 
  filter(family!= "" & phylum == "") %>% 
  distinct(class) # How many have a valid family but no phylum info

# Add new phylum info

clasif.hases.new.tree %>% 
  mutate(phylum = case_when(phylum != "" ~ phylum,
                            TRUE   ~ class))

saveRDS(clasif.hashes, paste0("hashes.annotated.",Sys.Date(),".rds"))

clasif.hashes <- readRDS("hashes.annotated.20180830.rds")

```

```{r}

thresholds <- list(0.8, 0.85, 0.95)

thresholds.classif <- map(thresholds, ~ classify(x= all.hashes.insect,
                                              tree = tree,
                                              cores = 8,
                                              threshold = .))

names(thresholds.classif) <- thresholds

#ADD THE DEFAULT

thresholds.classif[[4]] <- clasif.hashes

# Name it

names(thresholds.classif)[4] <- "0.9"

saveRDS(thresholds.classif, file =paste0("hashes.annotated.",Sys.Date(),".rds"))

list.of.thres <- readRDS(file ="hashes.annotated.2019-01-17.rds")

l2 <- lapply (list.of.thres, function (x) as.tibble(x))

```
UPDATE on NOV 6
We have added a new run of sequencing so I need to run insect on them too.

## Check the classification

And to see what is the actual resolution of our dataset, let's see how many reads can be classified to species, genus, family and order levels

```{r checking the clasiffication}

clasif.hashes %>% dplyr::count (rank) %>% arrange(desc(n))# an overview of the taxonomic resolution

All.Hashes.Classified[[1]] %>% dplyr::count (rank) %>% arrange(desc(n))

map_dfc (All.Hashes.Classified , ~ pull(., rank)) %>% mutate (representative = All.Hashes.Classified[[1]]$representative)

# a first check: How many sequences did not have a score: Either bc they exactly match a sequence in the tree, or they do not have a translation

clasif.hashes %>% 
  filter (is.na(score)) %>% # 176 Hashes
  left_join(ALL.ASVs, by = c("representative" = "Hash")) %>% 
 # group_by(representative) %>% 
  summarise(tot = sum(nReads)) # 2.55M reads - not only chimeras...

clasif.hashes %>% 
  filter(rank == "" & !is.na(score)) %>% # 2.8k  Hashes
  left_join(ALL.ASVs, by = c("representative" = "Hash")) %>% 
 # group_by(representative) %>% 
  summarise(tot = sum(nReads)) # 37M reads -  holy molly

 clasif.hashes %>% 
   filter(rank !="") %>% # 1.5 k Hashes
   left_join(ALL.ASVs, by = c("representative" = "Hash")) %>% 
 # group_by(representative) %>%
   group_by(sample) %>% 
  summarise(tot = sum(nReads)) %>% 
   ggplot(aes(x=fct_reorder(sample, tot), y = tot))+
   geom_col()# 6M reads

clasif.hashes %>% 
  filter(rank == "")

# Create the species- level dataset

map_dfc (l2, ~(dplyr::count(.,rank)))

map_dfr(l2, ~(dplyr::count(.,rank)), .id = "Threshold" ) -> Summary.thresholds.df# This puts the results in a df

Summary.thresholds.df$rank <- fct_relevel(Summary.thresholds.df$rank, "kingdom", "phylum", "subphylum", "subdivision", "superclass", "class",  "subclass", "subterclass", "infraclass", "superorder" , "order" , "suborder", "infraorder", "superfamily", "family", "subfamily" , "tribe",  "genus", "subgenus", "subsection",  "species", "subspecies", "variety")

Summary.thresholds.df$rank <- fct_recode(Summary.thresholds.df$rank, "Not Assigned" = "")

levels(Summary.thresholds.df$rank)

Summary.thresholds.df %>%
  ggplot (aes (x= rank, y = n, color = Threshold)) +
  geom_line(aes(group = Threshold))

Summary.thresholds.df %>%

  ggplot (aes (x= rank, y = n, fill = Threshold)) +
  geom_col(position = "dodge", color = "black") +
  theme(axis.text.x = element_text(angle = 45,vjust = 0.75))
  ggsave(filename = "Different.thresholds.png", width = 14)

clas.hash.tibble %>% dplyr::count (rank) %>%
  filter (str_detect(rank, ("species|genus|family|order")) ) %>%
  summarise (sum(n)) # 1331 Hashes identified to a level of order or lower

clas.hash.tibble %>% dplyr::count (rank) %>%
  filter (!str_detect(rank, ("species|genus|family|order")) ) %>%
  summarise (sum(n)) # 17k have not being IDed to that level

clas.hash.tibble %>% dplyr::count (rank) %>%
  filter ( rank == "")  %>%
  summarise (sum(n)) # 13k have a big fat 0 as a result

```
# Add the BLAST classiffication

We also classify the reads using a BLAST search, and retrieving the ???

```{r loading BLAST annotations}

all.hashes.blast     <- read_csv ("../../Analysis/Annotation/BLASTed_COI_Gallego_20190117-family.txt", col_names = c("Hash", "Family"))
all.hashes.blast.spp <- read_csv ("../../Analysis/Annotation/BLASTed_COI_Gallego_20190117-species.txt", col_names = c("Hash", "Species"))

all.hashes.blast %>% 
  left_join(all.hashes.blast.spp) -> all.hashes.blast

all.hashes.blast %>% 
  distinct(Species) # 440 unique spp

all.hashes.blast %>% 
  filter (!str_detect(Family, "idae")) %>% 
  select(-Hash) %>% 
  distinct(Family, .keep_all = T)

all.hashes.blast %>% 
  distinct(Family) # 279 unique "family"
 # get the lineages

Taxonomy.spp <- read_csv("/Users/ramongallego/GoogleDrive/Kelly_Lab/Projects/OA_eDNA/Analysis/Analysis.for.WSC/LineageLookupTable.csv") # LIneages for all spp
Taxonomy.fam <-  read_csv("/Users/ramongallego/GoogleDrive/Kelly_Lab/Projects/OA_eDNA/Analysis/Analysis.for.WSC/LineageLookupTable.families.csv")

# do we have them all?

summary (all.hashes.blast$Family %in% Taxonomy.spp$Family) # 3.3k families don't show in the lineages

all.hashes.blast %>% 
  filter (! Family %in% Taxonomy.fam$Family) %>% 
  distinct(Family) %>% 
 # mutate(match = str_which(string = Taxonomy.fam, pattern = .$Family)[1])

  mutate(Present = str_which(string = Taxonomy.fam, pattern = .$Family))#Taxonomy.fam$Phylum[str_which(string = Taxonomy.fam, pattern = .$Family)[1]],

all.hashes.blast %>% 
  left_join(clasif.hashes %>% select(Hash = representative, rank, taxon, score)) %>% 
  left_join(ALL.ASVs) %>% 
  filter (is.na(score)) %>% 
 # group_by(plant = str_detect(Family, "ceae")) %>% 
 # summarise(n.hashes = n_distinct(Hash)) # 51 are plants
  filter(!str_detect(Family, "ceae")) %>% 
  group_by(Family) %>% 
  summarise(tot = sum (nReads))

```

```{r}

all.hashes.blast %>% 
  bind_rows(new.hashes.blast) %>% 
  left_join(all.hashes.blast.spp %>%
              bind_rows(new.hashes.blast.spp), by = "Hash") %>% 
  dplyr::select (Hash,
          Blast.family = "Family",
          Blast.Species = "Species") -> BLAST.to.join

# Now extract the same info from the l2

map_dfc (All.Hashes.Classified , ~ pull(., species)) %>%
  mutate (Hash = All.Hashes.Classified[[1]]$representative) %>% 
  mutate (Insect.Spp = case_when(`0.95` != ""        ~ `0.95`,
                           `0.9`  != ""        ~ `0.9`,
                           `0.85` != ""        ~ `0.85`,
                           TRUE                ~ `0.8`)) %>%
  dplyr::select(Hash,
                Insect.Spp) -> all.hashes.insect.Spp

map_dfc (All.Hashes.Classified , ~ pull(., family)) %>%
  mutate (Hash = All.Hashes.Classified[[1]]$representative) %>% 
  mutate (Insect.Fam = case_when(`0.95` != ""        ~ `0.95`,
                           `0.9`  != ""        ~ `0.9`,
                           `0.85` != ""        ~ `0.85`,
                           TRUE                ~ `0.8`)) %>%
  dplyr::select(Hash,
                Insect.Fam) -> all.hashes.insect.Family

# Bind the two insect objects
all.hashes.insect.Spp %>% 
  left_join (all.hashes.insect.Family) %>% 
  left_join(BLAST.to.join) %>% # Now get a final classification 
  mutate(Species = case_when(Insect.Spp != "" ~ Insect.Spp,
                             TRUE ~ Blast.Species),
         Family = case_when(Insect.Fam != "" ~ Insect.Fam,
                             TRUE ~ Blast.family)) %>%
  dplyr::select(Hash, Family, Species) %>% 
  write_csv(path = paste0("Final.Classification.Hashes",Sys.Date(),".csv"))

all.hashes.insect.Family <- map_df(l2, ~ pull(.,family)) %>%
  rename_all(funs(paste0("Insect_Family_",.)))

all.hashes.insect.Species <- map_df(l2, ~pull(.,species)) %>%
  rename_all(funs(paste0("Insect_Species_",.)))

all.hashes <- bind_cols(all.hashes.insect.Family, all.hashes.insect.Species) %>%
  mutate (Hash = l2[[1]]$representative) %>%
  left_join(all.hashes.blast, .)

all.hashes.blast %>% dplyr::count(Family) # It has family and higher taxonomical info

all.hashes.blast %>% dplyr::count (Species)

```

### Now let's keep a database with five columns:

* Hash

* Sequence

* Family.insect

* Family.blast

* Species.insect

* Species.blast

And save it in disk

```{r reduced database}

clas.hash.tibble %>%
  select (Hash = representative,
          family.insect = family,
          species.insect = species) %>%
  left_join (all.hashes.blast)  -> clas.hash.tibble.to.write

write_csv(clas.hash.tibble.to.write, path = "hash.annotated.family.species.csv")

```

With this reference dataset, we were able to classify ~ 10% of the unique sequences to at least order level. 17k sequences did not have any match.

Let's chave a look at how many reads are classified

```{r combining }

ALL.ASVs %>% left_join(clas.hash.tibble.to.write, by = "Hash") -> All.ASVs.with.tax

All.ASVs.with.tax %>%
  mutate(Blast.Species = case_when(str_detect ( string = Blast.Species, pattern = "No |Not ") ~ "",
                               TRUE ~ Blast.Species) ) -> All.ASVs.with.tax

All.ASVs.with.tax %>% 
  filter(str_detect(sample, "K")) %>%
  distinct(sample)

All.ASVs.with.tax %>%
  group_by (Blast.Species, sample) %>%
  summarise (by.spp.blast = sum(nReads)) %>%
  arrange(desc(by.spp.blast)) -> summary.by.spp.blast

summary.by.spp.blast %>%
  filter(str_detect( string = Blast.Species, pattern = "No"))

All.ASVs.with.tax %>%
  group_by (species.insect) %>%
  summarise (by.spp.insect = sum(nReads)) %>%
  arrange(desc(by.spp.insect)) %>% 
  full_join(summary.by.spp.blast, by = c("species.insect" = "Blast.Species" ))

```

```{r}

All.ASVs.with.tax %>%
  group_by(rank) %>%
  summarise (n = sum(nReads)) %>%
  ggplot (aes(x = fct_reorder(rank, desc(n)), y = n)) +
  geom_col() +
  labs (x = "Rank")

#### Generate a table of the number of reads identified at least to family level

ALL.ASVs %>% left_join(all.hashes.blast) %>%
  group_by(Family) %>%
  summarise (n = sum(nReads)) %>%
  ggplot (aes(x = fct_reorder(Family, desc(n)), y = n)) +
  geom_col() +
  labs (x = "Family") +
  theme (axis.text.x = element_blank(),
         axis.ticks.x = element_blank())

All.ASVs.with.tax %>% 
  separate(sample, into = c("SampleShort", "rep"), sep = "\\.", remove = F) -> All.ASVs.with.tax

summary(All.ASVs.with.tax$SampleShort %in% OA_Samples$SampleShort)

All.ASVs.with.tax %>% filter (!SampleShort %in% OA_Samples$SampleShort)

All.ASVs.with.tax %>% filter (phylum != "" ) %>%
  mutate (family = case_when(family == "" ~ taxon,
                             TRUE         ~ family)) %>% 
  separate(sample, into = c("SampleShort", "rep"), sep = "\\.", remove = F) -> All.ASVs.with.tax

All.ASVs.with.tax %>% inner_join(OA_Samples, by = "SampleShort")  %>% 
  group_by(phylum, family, Site) %>%
  dplyr::summarise(nReads = sum(nReads)) %>%
  ggplot(aes(x= phylum, y = nReads)) +
  geom_col()

All.ASVs.with.tax %>% 
  group_by(sample, SampleShort, family) %>%
  summarise(nReads = sum(nReads))  %>%
  inner_join(OA_Samples, by = "SampleShort") -> ASVs.by.family

  summarise(nReads = sum(nReads))

```

Now we can use this dataset to elaborate our first exploratory analysis

## How many families are shared by each level of replication

```{r shared families}

ASVs.by.family %>% separate(sample, into = c("biol", "rep"), sep = "\\.", remove = F) %>%
  separate(biol, into = c("Site", "date"), remove = F) %>% 
  separate(date, into = "date", sep = -1) %>%
  separate (biol, into = "Site-date", sep = -1, remove = F) -> ASVs.by.family

## Now Start with site

ASVs.by.family %>% group_by(family) %>%
  summarise(n_sites = n_distinct (Site))

```

### 2020-09-24 Manuscript/What.lies.ahead.reviewed.tex
minor text fixes and editing.

@@
 Consequently, we know relatively little about the changes underway at real-world scales of biological communities, which typically have hundreds or thousands of interacting species.
-Here, we use monthly samples of environmental DNA to survey over 222 planktonic taxa along a gradient of temperature, salinity, dissolved oxygen, and carbonate chemistry in nearshore marine habitat.
+Here, we use monthly samples of environmental DNA to survey 222 planktonic taxa along a gradient of temperature, salinity, dissolved oxygen, and carbonate chemistry in nearshore marine habitat.
 The result is a high-resolution picture of changes in ecological communities using a technique replicable across a wide variety of ecosystems.
@@
 
-By treating amplification efficiency as consistent within a given taxon, we created an index of abundance for each taxon across space and time \cite{kelly2019understanding}, using pooled data from technical replicates and mean proportions across biological replicates.
+By treating amplification efficiency as consistent within a given taxon, we created an index of abundance for each taxon across space and time (``eDNA index'' \cite{kelly2019understanding}), using pooled data from technical replicates and mean proportions across biological replicates.
 We used this index of abundance in the multivariate community analysis, and used binary (presence/absence) data to capture species-level responses to environmental conditions.
-{\color{red}An implementation in R of this index is provided in the Supplementary information.}
+In the Supplementary information, we provide a function in R to calculate this index.
 
@@
 Independent of environmental parameters, we separately clustered samples by pairwise Bray-Curtis dissimilarities (k-means; N = 3) to identify groups of samples that were similar to one another with respect to biological community.
-The SIMPER procedure in \textit{vegan} revealed the taxa most contributing to between-cluster differences.
+The SIMPER procedure in \textit{vegan} revealed the taxa most strongly contributing to between-cluster differences.
 
@@
 Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}, and because salinity was correlated with temperature in our dataset, it was not an important predictor variable and we subsequently dropped it from our models.
-We calculated the probability of each community, given these predictors, using the r package \textit{nnet} \cite{nnet2002}.
+We calculated the probability of each community, given these predictors, using a multinomial logistic regression in the R package \textit{nnet} \cite{nnet2002}.
 
@@
 We performed the Wilcoxon test globally (total species richness) and on each phylum independently.
-% Reviewer suggests that the paragraoh below belongs to the discussion
-{\color{red} RYAN: Move paragraph below to discussion?}
-Taxa surveyed are a function of our metabarcoding PCR primers \cite{leray_new_2013} and reflect the current status of genetic databases, rather than a complete sampling of the planktonic community; we view these results as a cross-section of common taxa useful for understanding the biological effects of ocean conditions.
 
@@
 These samples represented biological and technical replicates from 86 unique sampling events.
-After bioinformatics quality-control \st{(See Methods)} the dataset included $\sim${}***45M sequences, from ***4849 unique amplicon sequence variants (ASVs).
+After bioinformatics quality-control the dataset included $\sim${}***45M sequences, from ***4849 unique amplicon sequence variants (ASVs).
 Of these, ***1364 ASVs (22.6M reads) could be annotated to a taxonomic level of Family or lower.
@@
 Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to geographic Area ***(Hood Canal vs.
-San Juan Island; F-model = 1.6184; p $<$ 0.01).
+San Juan Island; F = 1.6184; p $<$ 0.01).
 We therefore performed a constrained analysis of principal components (CAP) for samples within each Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig.
@@
 In the Hood Canal for example, the community linked with colder water and higher-pH (community **2 in Fig.
-2C) showed higher eDNA indexes of diatoms ***(\textit{Minutocellus polymorphus}), green algae  (\textit{Bathycoccus prasinos} and \textit{Micromonas pusilla}), and dinoflagellates like \textit{Karlodinium} sp than that on the low pH spectrum.
+2C) showed higher eDNA indexes of diatoms ***(\textit{Minutocellus polymorphus}), green algae  (\textit{Bathycoccus prasinos} and \textit{Micromonas pusilla}), and dinoflagellates like \textit{Karlodinium} sp relative to the lower-pH cluster at the same temperature.
 In that community (cluster 1) \textit{Emiliania huxleyi}, \textit{Ditylum brightwellii} and the copepod \textit{Paracalanus} sp.
 C are more prevalent.
-The community thriving in the high temperature range (community 3 in Fig.
+The community common in the high temperature range (cluster 3 in Fig.
 2C) shows high values of \textit{Phaeocystis globosa},\textit{Ditylum brightwellii} among other species.
-Community 3 (Fig.
+Cluster 3 (Fig.
 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls.
@@
 On the other hand, conditions that make cluster 2 the most likely community drop from 9\% to 1\% of the time between 2017 and 2095.
-
-% Under a common environmental condition, for example, in Hood Canal in summer, we expect such a planktonic community to go from occurring less than 1\% of the time in 2017 (0.001\% at T $=$ 15$\degree$C and pH $=$ 8.4; conditions more extreme than this occur ca.
-50\% of the time) to occurring 37.5\% of the time under an equivalently-common condition in 2095 (84\% at T $=$ 22$\degree$C and pH $=$ 8.05).
 
@@
 
-Diatoms (Bacillariophyta) show the steepest richness decline under future conditions (Figure 3 B.2); the probability of occurrence decreases markedly for diverse diatom taxa including both pennate (e.g., \textit{Nitzschia}) and centric (e.g., \textit{Coscinodiscus}, \textit{Stephanodiscus}) body forms.
+Diatoms (Bacillariophyta) show the steepest richness decline under future conditions (Figure 3 B.2); the probability of occurrence decreases markedly for diatom taxa including \textit{Coscinodiscus} and \textit{Stephanodiscus}, both centric body forms.
 These declines in diatom richness were more accentuated at lower pH values and higher temperatures.
-Other taxa likely to find less-suitable habitat in the future include dinoflagellates \textit{Karlodinium} sp.
-and Gymnodiniaceae sp.
+Other taxa likely to find less-suitable habitat in the future include the dinoflagellate Gymnodiniaceae sp.
 
@@
  Changes in environmental conditions and associated shifts in planktonic communities will likely reshape ecosystems and food webs, although some environmental processes may be conserved even as the particular taxa change.
-A switch from a diatom-dominated ecosystem to one in which dinoflagellate blooms extend in space and time could provoke cascade effects \cite{xiao2018warming} including fish mortality, anoxia \cite{anderson2002harmful}, and {\color{red} carbon sinking dynamics \st{alter current bentho-pelagic coupling processes} \cite{bienfang1984sinking}}.
+A switch from a diatom-dominated ecosystem to one in which dinoflagellate blooms extend in space and time could provoke cascade effects \cite{xiao2018warming} including fish mortality, anoxia \cite{anderson2002harmful}, and carbon sinking dynamics \cite{bienfang1984sinking}.
 Beyond the phylum-specific patterns, the increase in suitable habitat for harmful algae species will alone be an engine for ecosystem change \cite{trainer2019pelagic, wells2018harmful}.
@@
 
+The taxa surveyed here are a function of our metabarcoding PCR primers \cite{leray_new_2013} and reflect the current status of genetic databases, rather than a complete sampling of the planktonic community; we view these results as a cross-section of common taxa useful for understanding the biological effects of ocean conditions.
 Our observations are strong evidence of the kinds of changes likely in future marine communities, and they offer testable predictions about the magnitude and direction of effects on focal species.
-
-%%% Material y metodos
 
@@
 
- We are grateful to the UW Center for Environmental Genomics and Linda Park's lab at NOAA Fisheries for access to lab equipment and associated expertise and support.
+ We are grateful to the UW Center for Environmental Genomics and the Conservation Biology Molecular Genetics Laboratory at NOAA Fisheries (Montlake) for access to lab equipment and associated expertise and support.
 Thanks to Micah Horwith and the Washington State Department of Natural Resources for in-field support, to Terrie Klinger for feedback on an earlier draft, to Stephanie Moore for advice on phytoplankton generally, to Ana Ram\'{o}n-Laca for lab help, to Dan Drinan for field help, and to Alex Gagnon for access to carbonate-chemistry expertise and equipment.
- % Display the acknowledgments section
 

### 2020-09-25 Manuscript/What.lies.ahead.reviewed.tex
almost ready for submission

@@
 San Juan Island was less seasonally variable than the Hood Canal in every measured parameter (Figure 1C); the island is more directly affected by summer coastal upwelling as a function of bathymetry and circulation patterns \cite{murray2015inland}, and this appears to be the dominant influence on carbonate chemistry there.
-By contrast, photosynthesis and respiration likely drive much of the carbonate chemistry variation in the Hood Canal (See Supporting Information).
+By contrast, photosynthesis and respiration likely drive much of the carbonate chemistry variation in the Hood Canal (See Supplementary Information).
 
@@
 
-Code for all quality-screening and bioinformatics is available in the Supporting Information, implemented in Unix and R \cite{r_core_development_team_r:_2013}.
+Code for all quality-screening and bioinformatics is available in the Supplementary Information, implemented in Unix and R \cite{r_core_development_team_r:_2013}.
 Briefly, we used a Unix script that calls secondary programs for primer-trimming and preliminary quality-control \cite{martin2011cutadapt, Callahan:2016aa} we estimated the likely composition of each sample using DADA2.
@@
 We conservatively kept only taxa annotated at the level of taxonomic family, genus, or species, so we could reliably infer taxon natural history under the assuming that taxa within the same family shared broad natural-history characteristics.
-Using published literature and online databases, we placed every recovered taxon into a benthic/planktonic category and focused our analysis on the planktonic community (see Supporting Information).
+Using published literature and online databases, we placed every recovered taxon into a benthic/planktonic category and focused our analysis on the planktonic community (see Supplementary Information).
 
@@
 We used this index of abundance in the multivariate community analysis, and used binary (presence/absence) data to capture species-level responses to environmental conditions.
-In the Supplementary information, we provide a function in R to calculate this index.
+In the Supplementary Information, we provide a function in R to calculate this index.
 
@@
 The scenario labeled as 2095 is the set of parameters falling within the 95\% percentile in simulations for the years 2091-2095.
-See the R code in the supplementary information (lines 98-135).
+See the R code in the Supplementary Information (lines 98-135).
 
@@
 For each taxon, we fit these models using the Bayesian generalized linear mixed effects functions in R package rstanarm \cite{rstanarm} for R.
-Model selection using WAIC \cite{watanabe2010asymptotic} supported this as the preferred model over several similar ones (see Supporting Information for model comparison information and code) and helped to avoid model overfitting and maintain out-of-sample predictive power.
+Model selection using WAIC \cite{watanabe2010asymptotic} supported this as the preferred model over several similar ones (see Supplementary Information for model comparison information and code) and helped to avoid model overfitting and maintain out-of-sample predictive power.
 
@@
 \textbf{C} The observed environmental profiles of these two regions reflect a broad range of environmental conditions, with the Hood Canal resembling future conditions in temperate areas worldwide.
-Shown: Dissolved Inorganic Carbon (DIC, $\mu$moles/kg); Total Alkalinity (TA, $\mu$moles/kg); Temperature ($\degree$C) and Salinity (PSU).
-Individual sampling sites shown as vertical facets.}}
+Shown: Dissolved Inorganic Carbon (DIC, $\mu$moles/kg); Total Alkalinity (TA, $\mu$moles/kg); Temperature ($\degree$C) and Salinity (PSU).}}
 \label{fig:Fig1}
@@
 
-%% All changes due to new run of data are shown with *** before them for ease of search
 Despite geographic proximity and similar overall species composition (127 of the 222 planktonic taxa were found in both regions and accounted for 98\% of the sequences), the areas under study -- San Juan Island and the Hood Canal -- varied widely in pH, temperature, and other environmental parameters (Figure 1C), with a smooth gradient in conditions along the Hood Canal, and San Juan Island more closely resembling full marine conditions.
@@
 
-Metabarcoding analysis of eDNA samples generated more than ***50.8M sequences from 778 samples.
+Metabarcoding analysis of eDNA samples generated more than 50.8M sequences from 778 samples.
 These samples represented biological and technical replicates from 86 unique sampling events.
-After bioinformatics quality-control the dataset included $\sim${}***45M sequences, from ***4849 unique amplicon sequence variants (ASVs).
-Of these, ***1364 ASVs (22.6M reads) could be annotated to a taxonomic level of Family or lower.
+After bioinformatics quality-control the dataset included $\sim${}45M sequences, from 4849 unique amplicon sequence variants (ASVs).
+Of these, 1364 ASVs (22.6M reads) could be annotated to a taxonomic level of Family or lower.
 These $\sim$ 500 taxa from 43 phyla were split according to their natural history and habitat (benthic vs.
-planktonic; see Supporting Information).
+planktonic; see Supplementary Information).
 Because we expect planktonic taxa to vary with water mass \cite{kelly2018tides} and therefore with bottle-sampled carbonate chemistry, here we focus on only the planktonic taxa (N = 221).
@@
 
-Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to geographic Area ***(Hood Canal vs.
+Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to geographic Area (Hood Canal vs.
 San Juan Island; F = 1.6184; p $<$ 0.01).
 We therefore performed a constrained analysis of principal components (CAP) for samples within each Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig.
-2; for clarity, results for Hood Canal shown; full analysis in the Supporting Information).
+2; for clarity, results for Hood Canal shown; full analysis in the Supplementary Information).
 
@@
 These communities were distinguished by changes in the relative abundances of a wide variety of taxa.
-In the Hood Canal for example, the community linked with colder water and higher-pH (community **2 in Fig.
-2C) showed higher eDNA indexes of diatoms ***(\textit{Minutocellus polymorphus}), green algae  (\textit{Bathycoccus prasinos} and \textit{Micromonas pusilla}), and dinoflagellates like \textit{Karlodinium} sp relative to the lower-pH cluster at the same temperature.
+In the Hood Canal for example, the community linked with colder water and higher-pH (community 2 in Fig.
+2C) showed higher eDNA indexes of diatoms (\textit{Minutocellus polymorphus}), green algae  (\textit{Bathycoccus prasinos} and \textit{Micromonas pusilla}), and dinoflagellates like \textit{Karlodinium} sp relative to the lower-pH cluster at the same temperature.
 In that community (cluster 1) \textit{Emiliania huxleyi}, \textit{Ditylum brightwellii} and the copepod \textit{Paracalanus} sp.
@@
 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls.
-For example, we expect the Hood Canal in 2095 to have the conditions in which community 3 is the most likely community 66\% of the time, an increase of 11\% compared to 2017(see Supplemental Material).
+For example, we expect the Hood Canal in 2095 to have the conditions in which community 3 is the most likely community 66\% of the time, an increase of 11\% compared to 2017(see Supplementary Information).
 On the other hand, conditions that make cluster 2 the most likely community drop from 9\% to 1\% of the time between 2017 and 2095.
@@
 \textbf{C} Relative abundance (eDNA index; see \cite{kelly2019understanding}) of the taxa best distinguishing the three communities illustrated (SIMPER analysis).
-See Supporting Information for full analysis.}}
+See Supplementary Information for full analysis.}}
 \label{fig:Fig2}
@@
 Among others not shown in Fig 3, \textit{Chaetoceros} (diatom) and many hydrozoans (Cnidaria) likely increase in San Juan Island, and the potentially fish-killing heterokont flagellate \textit{Pseudochattonella} increases in both study areas.
-See Supporting Information for a complete list of taxon-specific projections.
+See Supplementary Information for a complete list of taxon-specific projections.
 
@@
 and the harmful algal bloom (HAB)-forming species \textit{Pseudo-nitzschia} spp.
-will see their habitat suitability remain constant or slightly increased, especially at low pH levels (see Supplemental Material).
+will see their habitat suitability remain constant or slightly increased, especially at low pH levels (see Supplementary Information).
 
@@
 Thanks to Micah Horwith and the Washington State Department of Natural Resources for in-field support, to Terrie Klinger for feedback on an earlier draft, to Stephanie Moore for advice on phytoplankton generally, to Ana Ram\'{o}n-Laca for lab help, to Dan Drinan for field help, and to Alex Gagnon for access to carbonate-chemistry expertise and equipment.
+The comments of two reviewers helped us improve the manuscripts and encourage us to streamline the analysis of our data.
 

### 2020-09-25 Scripts/Manuscript.Main.Analysis.Rmd
almost ready for submission

@@
 
-Because the standarization was done for each region separately, we have to convert them into standard values seapratedly as well.
+Because the standarization was done for each region separately, we have to convert them into standard values separately as well.
 
@@
 
-Make the ridges plot
+Make the box plots
 
@@
   ggplot(aes(x = `value`, y = `Month`, fill = Area)) +
-  geom_density_ridges_gradient() +
+  geom_boxplot() +
  # scale_fill_viridis(name = "Temp. [F]", option = "C") +
@@
 
-THe top will be the environmental gradients, the lower panels will have the map and the boxplot. Create a custom palette of colors for the two regions
+The top will have the map and the boxplots, the lower panel will be the environmental gradients. Create a custom palette of colors for the two regions
 
@@
 data.plankton %>% 
-  separate (event, into = c("Site", "Date"), sep = "_") %>% 
+  separate (event, into = c("Site", "Date"), sep = "_", remove = F) %>% 
   mutate(Area = case_when(Site %in% c("CP","LK", "FH") ~ "San Juan Island",
                           TRUE                         ~ "Hood Canal")) %>% 
-   unite (Site, Date, col = "event", sep = "_") %>% 
   cap.functions.only3things(Area) -> joined.dataset.for.CAP
@@
 
-joined.dataset.for.CAP %>% pull(CAP.significance)
 ```
@@
 
-  CAP.clusters = map2 (CAP.plot,n_clust,  function (.x, .y) {
-
-    final <- kmeans(.x$sites, .y, nstart = 15)
-
-    fviz_cluster(final, data = .x$sites, geom = "point", palette = "Set2")
-
-  } ),
+  # CAP.clusters = map2 (CAP.plot,n_clust,  function (.x, .y) {
+  # 
+  #   final <- kmeans(.x$sites, .y, nstart = 15)
+  # 
+  #   fviz_cluster(final, data = .x$sites, geom = "point", palette = "Set2")
+  # 
+  # } ),
 
@@
 Clusters.analysis.and.cap %>% 
-  mutate(full.cluster.info = map2(env, Cluster.df, ~ .x %>% unite(Site, Date, col = "event") %>%right_join(.y) ),
+  mutate(full.cluster.info = map2(env, Cluster.df, ~ .x %>% unite(Site, Date, col = "event") %>% right_join(.y) ),
          full.cluster.info = map2(full.cluster.info,CAP.plot, ~ .y$sites %>% 
@@
 
-  })) %>% pull(Hand.made.CAP) -> leftplot.Fig2
-
- leftplot.Fig2 %>% set_names(Clusters.analysis.and.cap$Area) %>% map(~ .x +  theme_light() + theme(legend.position = "none") ) -> left_top
+  })) %>% pull(Hand.made.CAP) -> left_top
+
+ left_top %>% set_names(Clusters.analysis.and.cap$Area) %>% map(~ .x +  theme_light() + theme(legend.position = "none") ) -> left_top
 
@@
          width = 14,height = 11, dpi = "retina") 
-
+plot
 ```
@@
 
-### Richness by phyla with time
- First prepare the grid - expand it to include the taxa
+### Richness by phyla across the pH and Temperature continuum
+
+THe idea of this panel is to account for the limitations of a simulation of future scenarios, and just look at the behaviour of the model in the continuum - we know that temperature and pH will change in a direction, and we can explore broad patterns in the raster grid.
+
+ First prepare the grid - expand it to include the taxa information.
 ```{r}
@@
   summarise(richness = sum(present)) %>% 
-#  left_join(richness.by.phylum) %>% 
   group_by(phylum) %>% 
@@
 
+Create a function to plot the richness across temperature and pH
 ```{r plot4 function}
@@
 
+Create a richness plot for each phylum, pull them and name them
 ```{r plot4 in play}
@@
 
-Following reviewer's suggestions, we are limiting the information displayed in the main Figures to two of most important phyla for coastal ecosystems dinamics: Diatoms and Dinoflagellates.
+Following reviewer's suggestions, we are limiting the information displayed in the main Figures to three important phyla for coastal ecosystems dynamics: Diatoms, Haptophytes and Dinoflagellates.
 ```{r}
@@
 
+You can find similar figures for all phyla found in the Folder `Figures/Supplemental`
+
 ```{r}
@@
   geom_vline(data = medians, aes(xintercept = grp.median, color = Year), linetype = "dashed") +
-  facet_wrap (~Area)
+  facet_wrap (~Area)+
+  ggtitle(phyl)
 
@@
 
-Now the Global - richness by point
+Now the Global - richness by point. If you want to see the change on any other phyla, type(`explore.phylum("phylum.name")`), you can see here plots for Dinoflagellates and diatoms
 

### 2020-09-26 Manuscript/Response.to.reviewers.tex
YES IT IS DONE

--- 2020-09-04 Manuscript/Response.to.reviewers.tex
@@
 In an effort to declutter Figure 3, and following suggestions from reviewer 1, we have focused the figure in those species showing a more dramatic change in suitability.
-The new version of the figure has two panels (A and B) and we refer to them specifically in the text
+The new version of the figure has three panels (A, B, C; one per phylum) displaying three aspects of the data (.1, .2 and .3) and we refer to them specifically in the text
 
@@
 
+{\color{red}RYAN: How about this:
+In sum, yes: these clusters are coherent groups of species and predictably occur in different environmental conditions.
+However, differences in relative proportions of species will also contribute to cluster dissimilarities.
+}
+
 We think the relationship between clusters and the environmental variables is important in a few aspects.
@@
 Second, these clusters (maybe communities is a more committed term that requires a longer dataset in which to confirm co-ocurrence of members) do indeed reflect changes in the relative importance of the species present.
-
-{\color{red} RAMON -- if we should use `cluster' instead of `community' we will need to change that in the main text. }
 
@@
 
-{\color{red} TODO - check }
+\textbf{Response:}
+
+ Figure 2C includes now taxa with the higher contribution to the dissimilarity between clusters, highlighting both positive and negative contributions.
 
@@
 (2019) (see comments on the Methods section as well).}
+
+\textbf{Response:}
 
@@
 
+\textbf{Response:}
+
 Done.
@@
 \item{Line 121: Refer to the relevant section of Figure 3 (A, B, or C)}
+
+\textbf{Response:}
 
@@
 
-{\color{red} TODO - check }
+\textbf{Response:}
+
+ We thank the reviewer for clarifying this point.
+We have now decluttered Figure 3; and highlight taxa that are now mentioned in the text: the Dinoflagellates Karlodinium, Gymnodinaceae and Alexandrium; the Diatoms Stephanodiscus, Nitzschia, Coscinodiscus and Chaetoceros; and the Haptophytes Emiliania and Chrysochromulina spp.
 
@@
 
-{\color{red} TODO - check }
+\textbf{Response:}
+
+Our effort has been on visualizing species with more pronounced change, not to highlight species that respond in contrasting ways in the San Juan Islands and the Hood Canal.
+Most taxa react in the same direction in both regions; with a few exceptions in which species are almost absent on one study area.
 
@@
 
-{\color{red} TODO - check }
+\textbf{Response:}
+
+Coscinodiscus sees its probability of presence reduced on both Areas.
+The same happens for D.
+brightwelli, although our approach for a more readable figure has seen this taxa drop into the Supplementary figures
 
 \item{Line 129: Write out Emiliania.
-The ?sp.?
-after Alexandrium should not be italicized.
-This sentence might read better if written: ?The coccolithophore Emilinania huxleyi and the dinoflagellate Alexandrium sp.
-both find more suitable habitat?
-(rather than ?each finds?)}
+The 'sp.' after Alexandrium should not be italicized.
+This sentence might read better if written: 'The coccolithophore Emilinania huxleyi and the dinoflagellate Alexandrium sp.
+both find more suitable habitat' (rather than 'each finds')}
 
-{\color{red} TODO - check }
+\textbf{Response:}
+
+ We have updated that sentence, which now reads: "The haptophytes \textit{Emiliania huxleyi} and \textit{Chrysochromulina} sp.
+the dinoflagellate \textit{Alexandrium} sp.
+all find more suitable habitat in both of our study areas." Thanks for the suggestion.
 
@@
 
-{\color{red} TODO - check }
+\textbf{Response:}
+
+We thank the reviewer for pointing this out: We have rethought this and applied Wilcoxon tests for both global and phylum-specific changes in diversity.
+As we are comparing pairs of data we prefer not to assume normal distributions of the datasets or the differences between them, Wilcoxon test was a better choice.
 
@@
 
-{\color{red} TODO - check }
+\textbf{Response:}
+
+ A reworded version of this paragraph is now at the end of the discussion section.
 
@@
 
-{\color{red} TODO - check }
+\textbf{Response:}
+
+ Agreed: We now cite the work of J. Zamon (2002) and S. Pasquaud and colleagues (2010) to support the role of Coscinodiscus in the food chain.
 
@@
 
-{\color{red} TODO - check }
+\textbf{Response:}
+
+  Done.
 
@@
 
-{\color{red} TODO - check }
+\textbf{Response:}
+
+ Agreed - it will definitely have an effect on the `fate` of the photosynthesized carbon- a quickly sunk carbon will join marine sediments instead of being recycled in the upper layers of the water masses.
+We now cite the work of Polimene and colleagues on the role of phytoplankton and carbon sequestration.
 
@@
 
-{\color{red} TODO - check }
+\textbf{Response:}
 
-\item{Line 208-214: This is a big aspect of why this study is valuable! I think it?s worth discussing this in the introduction.}
+ In their work, D.
+Anderson and colleagues cite a range of consequences of increased HAB blooms, including fish mortality and anoxia - and not eutrophication as we wrongly pointed out.
+Thanks for noticing.
+On the second argument - we saw potential consequences in two ways: by affecting recruiting of benthic species with pelagic larvae with reduced diatom diversity/biomass; and by changing the nature of sinking matter for filter feeders.
+We realize that bentho pelagic coupling is not the only source of carbon for benthic communities in the area of study so we have decided to remove this assertion from the discussion.
 
-{\color{red} TODO - check }
+\item{Line 208-214: This is a big aspect of why this study is valuable! I think it's worth discussing this in the introduction.}
+
+\textbf{Response:}
+
+ We thank the reviewer for this suggestion.
 
@@
 
-{\color{red} TODO - check }
+\textbf{Response:}
+
+ Done.
 
@@
 
-{\color{red} TODO - check }
+\textbf{Response:}
 
-\item{Line 273-282: It?s hard to discern based on the limited code and data provided exactly how the authors subsampled the different subregions of Puget Sound, since the paper they base their data on has overall values for the whole area.
+We used the eDNAindex (the proportion of reads assigned to each taxa per sample, scaled to the maximum proportion that taxa achieved).
+We provide the R implementation of this index in the Supplementary material, and have noted so in the text (line 128).
+
+\item{Line 273-282: It's hard to discern based on the limited code and data provided exactly how the authors subsampled the different subregions of Puget Sound, since the paper they base their data on has overall values for the whole area.
 I think this section could use a bit of clarification, and all the code and raw data should be given (the provided github link is broken).}
 
-{\color{red} TODO - check }
+\textbf{Response:}
+
+We apologize for the broken github code - We have clarified how we simulated the future projection in the Supplementary Information - Also see response to reviewer1 on line 122 - 137 of this letter.
 

### 2020-09-26 Manuscript/What.lies.ahead.reviewed.tex
YES IT IS DONE

@@
 
-\textbf{Ram\'{o}n Gallego \textsuperscript{$a,b,1$}, Emily Jacobs-Palmer \textsuperscript{$b$}, Kelly Cribari \textsuperscript{$b$}, Ryan P. Kelly}\textsuperscript{$b$}
+\textbf{Ram\'{o}n Gallego \textsuperscript{$a,b,1$}, Emily Jacobs-Palmer \textsuperscript{$a$}, Kelly Cribari \textsuperscript{$a$}, Ryan P. Kelly}\textsuperscript{$a$}
 \vspace{1em}
 
-\textbf{$^a$} {\small NRC Research Associateship Program,
+\textbf{$^a$} {\small School of Marine and Environmental Affairs. University of Washington. 3707 Brooklyn Ave NE Seattle, WA 98105}
+
+\textbf{$^b$} {\small Current affiliation: NRC Research Associateship Program,
 Northwest Fisheries Science Center,
@@
 National Oceanic and Atmospheric Administration. 2725 Montlake Boulevard East, Seattle, WA 98112}
-
-\textbf{$^b$} {\small School of Marine and Environmental Affairs. University of Washington. 3707 Brooklyn Ave NE Seattle, WA 98105}
 
@@
 These sites experience substantial variation in water chemistry and other environmental conditions despite geographic proximity (ca.
-300km; Fig.
-1).
+300km; Figure 1).
 We sampled eight sites monthly for approximately 1.5 years (March 2017 to August 2018), taking three 1L samples (biological replicates; ca.
@@
 Because we expect planktonic taxa to vary with water mass \cite{kelly2018tides} and therefore with bottle-sampled carbonate chemistry, here we focus on only the planktonic taxa (N = 221).
-These taxa showed a seasonal richness gradient between study areas, consistent with documented biodiversity clines in the area (Fig.
-1B.1) \cite{dethier_decadal_2008}.
+These taxa showed a seasonal richness gradient between study areas, consistent with documented biodiversity clines in the area (Figure 1B.1) \cite{dethier_decadal_2008}.
 
@@
 San Juan Island; F = 1.6184; p $<$ 0.01).
-We therefore performed a constrained analysis of principal components (CAP) for samples within each Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig.
-2; for clarity, results for Hood Canal shown; full analysis in the Supplementary Information).
+We therefore performed a constrained analysis of principal components (CAP) for samples within each Area, showing the differences among communities as a function of temperature, pH, and salinity (Figure 2; for clarity, results for Hood Canal shown; full analysis in the Supplementary Information).
 
- Each biological cluster (colored hulls, Fig.
-2A) occupied a unique area of environmental parameter space.
+ Each biological cluster (colored hulls, Figure 2A) occupied a unique area of environmental parameter space.
 Planktonic communities therefore varied predictably with water temperature, salinity, and pH, across a range of those parameters likely to be encountered in many near-term future-ocean scenarios \cite{murray2015inland}.
- Multinomial logistic regression yielded predictions of the most-likely community for any combination of environmental parameters (Fig. 2B).
+ Multinomial logistic regression yielded predictions of the most-likely community for any combination of environmental parameters (Figure 2B).
 
-These communities were distinguished by changes in the relative abundances of a wide variety of taxa.
-In the Hood Canal for example, the community linked with colder water and higher-pH (community 2 in Fig.
-2C) showed higher eDNA indexes of diatoms (\textit{Minutocellus polymorphus}), green algae  (\textit{Bathycoccus prasinos} and \textit{Micromonas pusilla}), and dinoflagellates like \textit{Karlodinium} sp relative to the lower-pH cluster at the same temperature.
-In that community (cluster 1) \textit{Emiliania huxleyi}, \textit{Ditylum brightwellii} and the copepod \textit{Paracalanus} sp.
+These clusters were distinguished by changes in the relative abundances of a wide variety of taxa.
+In the Hood Canal for example, the cluster linked with colder water and higher-pH (cluster 2 in Figure 2B) showed higher eDNA indexes of diatoms (\textit{Minutocellus polymorphus}), green algae  (\textit{Bathycoccus prasinos} and \textit{Micromonas pusilla}), and dinoflagellates like \textit{Karlodinium} sp.
+relative to the lower-pH cluster at the same temperature range.
+In that community (cluster 3) \textit{Emiliania huxleyi}, \textit{Ditylum brightwellii} and the copepod \textit{Paracalanus} sp.
 C are more prevalent.
-The community common in the high temperature range (cluster 3 in Fig.
-2C) shows high values of \textit{Phaeocystis globosa},\textit{Ditylum brightwellii} among other species.
-Cluster 3 (Fig.
-2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls.
-For example, we expect the Hood Canal in 2095 to have the conditions in which community 3 is the most likely community 66\% of the time, an increase of 11\% compared to 2017(see Supplementary Information).
-On the other hand, conditions that make cluster 2 the most likely community drop from 9\% to 1\% of the time between 2017 and 2095.
+The cluster common in the high temperature range (cluster 1 in Figure 2B) shows high values of \textit{Phaeocystis globosa},\textit{Emiliania huxleyi}, \textit{Ditylum brightwellii} among other species.
+Cluster 1 (Figure 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls.
+For example, we expect the Hood Canal in 2095 to have the conditions in which community 1 is the most likely community 67\% of the time, an increase of 11\% compared to 2017(see Supplementary Information).
+On the other hand, conditions that make cluster 2 the most likely community drop from 12\% to 2\% of the time between 2017 and 2095.
 
@@
 Salinity was not informative in our models, as it is highly correlated with temperature in our dataset and is moreover predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}.
-Model projections let us show the change in the probability of presence of each individual taxon for 2095 vs 2017 (Fig.
-3 A.1, B.1, C.1), estimate richness for larger taxonomic grounds as a whole for these two climates (Fig.
-3 A.2, B.2, C.2), and estimate richness within taxonomic groups across the pH-temperature continuum (Fig.
-3 A.3, B.3, C.3).
+Model projections let us show the change in the probability of presence of each individual taxon for 2095 vs 2017 (Figure 3 A.1, B.1, C.1), estimate richness for larger taxonomic grounds as a whole for these two climates (Figure 3 A.2, B.2, C.2), and estimate richness within taxonomic groups across the pH-temperature continuum (Figure 3 A.3, B.3, C.3).
 
@@
 all find more suitable habitat in both of our study areas.
-Among others not shown in Fig 3, \textit{Chaetoceros} (diatom) and many hydrozoans (Cnidaria) likely increase in San Juan Island, and the potentially fish-killing heterokont flagellate \textit{Pseudochattonella} increases in both study areas.
+Among others not shown in Figure 3, \textit{Chaetoceros} (diatom) and many hydrozoans (Cnidaria) likely increase in San Juan Island, and the potentially fish-killing heterokont flagellate \textit{Pseudochattonella} increases in both study areas.
 See Supplementary Information for a complete list of taxon-specific projections.
@@
 
-Temperate surface oceans worldwide average approximately 14$\degree$C and pH of 8.1 \cite{bopp2013multiple}, and will change substantially in this century \citep[mean $\Delta$ T 2.5$\degree$C, $\Delta$ pH -0.35 globally; RCP 8.5;][]{gattuso2015contrasting}.
+Temperate surface oceans worldwide average approximately 14$\degree$C and pH of 8.1 \cite{bopp2013multiple}, and will change substantially in this century \citep[mean $\Delta$T 2.5$\degree$C, $\Delta$pH -0.35 globally; RCP 8.5;][]{gattuso2015contrasting}.
 Here we document communities exposed to this same range of projected conditions in the present day, along an environmental gradient only ca.
@@
  We are grateful to the UW Center for Environmental Genomics and the Conservation Biology Molecular Genetics Laboratory at NOAA Fisheries (Montlake) for access to lab equipment and associated expertise and support.
-Thanks to Micah Horwith and the Washington State Department of Natural Resources for in-field support, to Terrie Klinger for feedback on an earlier draft, to Stephanie Moore for advice on phytoplankton generally, to Ana Ram\'{o}n-Laca for lab help, to Dan Drinan for field help, and to Alex Gagnon for access to carbonate-chemistry expertise and equipment.
+Thanks to Micah Horwith and the Washington State Department of Natural Resources for in-field support, to Terrie Klinger for feedback on an earlier draft, to Stephanie Moore for advice on phytoplankton generally, to Ana Ram\'{o}n-Laca for laboratory help, to Dan Drinan for field help, and to Alex Gagnon for access to carbonate-chemistry expertise and equipment.
 The comments of two reviewers helped us improve the manuscripts and encourage us to streamline the analysis of our data.

### 2020-09-26 Scripts/Manuscript.Main.Analysis.Rmd
YES IT IS DONE

@@
 
+    set.seed(1231)
+
     final <- kmeans(.x$sites, .y, nstart = 15)
@@
  left_top %>% set_names(Clusters.analysis.and.cap$Area) %>% map(~ .x +  theme_light() + theme(legend.position = "none") ) -> left_top
+ left_top
 
@@
 testquick %>% pull(plots) %>% set_names(testquick$Area) %>% map(~ .x +  theme(axis.title.y = element_text(angle = 0),legend.position = "none") )-> center_top
+
+center_top
 ```

### 2020-09-27 Manuscript/Response.to.reviewers.tex
revised response to reviewers very slightly. Nice work, Ramon.

@@
 
-We are thankful for the feedback, comments and suggestions Our manuscript RSPB-2019-2901  "Environmental DNA Metabarcoding Reveals Winners and Losers of Global Change in Coastal Waters" has received.
+We are thankful for the feedback, comments and suggestions on our manuscript RSPB-2019-2901  "Environmental DNA Metabarcoding Reveals Winners and Losers of Global Change in Coastal Waters".
 Find below our response to their suggestions, and how we implement them in our manuscript.
@@
 \item{2.      This is certainly not the first eDNA metabarcoding paper to track shifts in community composition across environmental gradients, yet there are almost no references to other studies that have used this approach, nor how eDNA metabarcoding is being applied more broadly as a monitoring tool.
-Pleas incorporate this literature into the Introduction.
+Please incorporate this literature into the Introduction.
 }
@@
 In an effort to declutter Figure 3, and following suggestions from reviewer 1, we have focused the figure in those species showing a more dramatic change in suitability.
-The new version of the figure has three panels (A, B, C; one per phylum) displaying three aspects of the data (.1, .2 and .3) and we refer to them specifically in the text
+The new version of the figure has three panels (A, B, C; one per phylum) displaying three aspects of the data (.1, .2 and .3) and we refer to them specifically in the text.
 
@@
 
-In sum, yes: these clusters are coherent groups of species and predictably occur in different environmental conditions.  We have revised the main text to be more clear in this regard.
-
-{\color{red}RYAN: How about this:
 In sum, yes: these clusters are coherent groups of species and predictably occur in different environmental conditions.
-However, differences in relative proportions of species will also contribute to cluster dissimilarities.
-}
+However, it is not that taxa occur exclusively in different clusters;  differences in relative proportions of species will also contribute to cluster dissimilarities.
+We see the clusters as different communities, in the sense that they are characteristic \textit{mixes} of taxa that occur under different conditions.
 
@@
  We thank the reviewer for clarifying this point.
-We have now decluttered Figure 3; and highlight taxa that are now mentioned in the text: the Dinoflagellates Karlodinium, Gymnodinaceae and Alexandrium; the Diatoms Stephanodiscus, Nitzschia, Coscinodiscus and Chaetoceros; and the Haptophytes Emiliania and Chrysochromulina spp.
+We have now decluttered Figure 3; and highlight taxa that are now mentioned in the text: the Dinoflagellates \textit{Karlodinium}, Gymnodinaceae sp.
+and \textit{Alexandrium}; the Diatoms \textit{Stephanodiscus}, \textit{Nitzschia}, \textit{Coscinodiscus} and \textit{Chaetoceros}; and the Haptophytes \textit{Emiliania huxlei} and \textit{Chrysochromulina} sp.
 
@@
 
-Coscinodiscus sees its probability of presence reduced on both Areas.
-The same happens for D.
-brightwelli, although our approach for a more readable figure has seen this taxa drop into the Supplementary figures
+\textit{Coscinodiscus} sees its probability of presence reduced in both Areas.
+The same happens for \textit{D.
+brightwelli}, although our approach for a more readable figure has seen this taxon drop into the Supplementary figures.
 
@@
 
- Agreed: We now cite the work of J. Zamon (2002) and S. Pasquaud and colleagues (2010) to support the role of Coscinodiscus in the food chain.
+ Agreed: We now cite the work of J. Zamon (2002) and S. Pasquaud and colleagues (2010) to support the role of \textit{Coscinodiscus} in the food chain.
 
-\item{Line 178: Spell out harmful algal bloom the first time it?s used.}
+\item{Line 178: Spell out harmful algal bloom the first time it's used.}
 
@@
 On the second argument - we saw potential consequences in two ways: by affecting recruiting of benthic species with pelagic larvae with reduced diatom diversity/biomass; and by changing the nature of sinking matter for filter feeders.
-We realize that bentho pelagic coupling is not the only source of carbon for benthic communities in the area of study so we have decided to remove this assertion from the discussion.
+We realize that bentho-pelagic coupling is not the only source of carbon for benthic communities in the area of study so we have decided to remove this assertion from the discussion.
 
@@
 
-{\color{red} Done }
+Done.
 
@@
 We used the eDNAindex (the proportion of reads assigned to each taxa per sample, scaled to the maximum proportion that taxa achieved).
-We provide the R implementation of this index in the Supplementary material, and have noted so in the text (line 128).
+We have now specified this and provide the R implementation of this index in the Supplementary material(line 128).
 
@@
 
-We apologize for the broken github code - We have clarified how we simulated the future projection in the Supplementary Information - Also see response to reviewer1 on line 122 - 137 of this letter.
+We apologize for the broken github code - We have clarified how we simulated the future projection in the Supplementary Information - Also see response to reviewer 1 on line 122 - 137 of this letter.
 

### 2020-09-28 Manuscript/CoverLetter/Cover_letter_ProcB_revision.tex
cover letter

\documentclass[11pt,letterpaper]{letter}

\usepackage{RPKLetter}
\usepackage{url}
\usepackage{changepage}
\usepackage{multicol}

\begin{document}

%	TO ADDRESS

\begin{letter}
{}
%\\
%	LETTER CONTENT

\opening{}

\noindent Dear Editors,
\vspace{0.05in}

Please accept the enclosed revised manuscript, \textit{Environmental DNA Metabarcoding Reveals Winners and Losers of Global Change in Coastal Waters}, for consideration in Proceedings of the Royal Society B. The original submission number was RSPB-2019-2901.

We have taken the opportunity to revise the manuscript significantly in response to reviewers' comments. In addition, of course, the COVID crisis delayed the revision somewhat.

In the piece, we report genetic data taken from a ca. 18-month time-series of water samples (N = 227), from along a spatial gradient in sea-surface temperature, pH, and other environmental parameters. After isolating the effects of these water conditions on eukaryotic communities, we derive models that describe species-specific responses to changes in environment. We then use these models to forecast community change in a future ocean scenario, finding large shifts in community composition that will likely result in fundamental changes to these ecosystems. We are very excited about this work.

A central problem in understanding the effects of global change -- and indeed, ecology generally -- is the usual tradeoff between depth and breadth: traditional approaches offer either an in-depth assessment of one or few species (e.g., laboratory experiments) or a shallow assessment of many species (e.g., meta-analyses and modeling).  Here, we use eDNA metabarcoding to overcome this challenge; our empirical dataset consists of hundreds of species sampled from along an environmental gradient that captures the conditions many nearshore habitats will experience in the coming decades.

Most strikingly, we see a decline in diatom richness and abundance and a rise in dinoflagellates across our study area, portending major shifts in community composition and the marine food web during this century.

We think this work represents a significant advance, using eDNA sequencing data to move beyond proofs-of-concept and into the realm of empirical ecology. By linking this emerging method to water chemistry and oceanographic projections, we offer a way of characterizing community-level changes that are otherwise impossible to capture. Moreover, the technique is broadly useful in the study of ecological change in the ocean; importantly, it complements data from model sites such as CO$_2$ vents by revealing patterns in non-model systems that are the rule in the rest of the world.

We think the manuscript has a wide appeal for ecologists, geneticists, marine food-web scholars, and scientists interested in the different ways we can forecast changes in marine ecosystems with climate change. For such a potentially wide audience, Proceedings B is a perfect fit and we want to see our work published here. We sincerely hope you share our view and we are looking forward to hearing from you,

\closing{Cheers,}

%----------------------------------------------------------------------------------------
\end{letter}
\end{document}

### 2020-11-04 Manuscript/Response.to.reviewers.tex
Final revision

@@
 \documentclass[11pt]{article}
-
+\usepackage{hyperref}
 \usepackage{textcomp}
@@
 Moreover, due to a technical error, the raw data and scripts were not available to reviewers in the first submission -- our sincere apologies.
-All data and scripts are now available in an open github repository (raw fastq files are in a google drive link, because of their size).
+All data and scripts are now available in an open github repository \url{https://github.com/ramongallego/eDNA.and.Ocean.Acidification.Gallego.et.al.2020} (raw fastq files are in a google drive \href{https://drive.google.com/drive/folders/1XtuDMdlnk9V2acNiPJo5T6O1Tp0KjaCz?usp=sharing}{link}, because of their size).
 We have strived for transparency and repeatability throughout.
@@
 \item{Figures: the colors red and blue are used both to distinguish time periods (2019/2095) and locality (San Juan/Hood Canal) in different figures.
-Figure 2 legend doesn?t say which color is which (presumably they are the same as in Figure 1).
+Figure 2 legend doesn't say which color is which (presumably they are the same as in Figure 1).
 Use consistent colors for categories across figures and use different colors for time period and locality, if possible.}

### 2020-11-04 Manuscript/What.lies.ahead.reviewed.tex
Final revision

@@
 
-\subsection{Background}
+\section*{Background}
 
@@
 Environmental DNA (eDNA) metabarcoding \cite{holman2019detection, kelly2016genetic} addresses this problem by amplifying a common gene region out of DNA present in a water sample; the technique can detect hundreds to thousands of taxa per sample, potentially with species-level identification.
-A growing body of evidence supports the efficacy of eDNA metabarcoding for monitoring biodiversity (see a review in \cite{ruppert2019past}), and this approach has been successfully used to detect community composition variation across environmental changes in aquatic  \cite{emilson2017dna}, estuarine \cite{chariton2015metabarcoding, lallias2015environmental}, and marine ecosystems \cite{berry2019marine, djurhuus2019eDNA}.
+A growing body of evidence supports the efficacy of eDNA metabarcoding for monitoring biodiversity (see a review in \cite{ruppert2019past}), and this approach has been successfully used to detect community composition variation across environmental changes in aquatic  \cite{emilson2017dna}, estuarine \cite{chariton2015metabarcoding, lallias2015environmental}, and marine ecosystems \cite{berry2019marine, djurhuus2020environmental}.
 
@@
 
-\section{Methods}
+\section*{Methods}
 \subsection*{Sampling}
@@
 
+All raw data used in this manuscript and the scripts that generate the final figures and analysis can be found at https://github.com/ramongallego/eDNA.and.Ocean.Acidification.Gallego.et.al.2020
+
 %%%%% RESULTS %%%%%
 
-\subsection*{Results}
+\section*{Results}
 \underline{Variation in Carbonate Chemistry and in Ecological Communities}
@@
 \includegraphics[width=1\linewidth]{	Figures_for_MS/Fig2}
-\caption{\footnotesize{Biological communities and their relationship with environmental variables in the summer months of the Hood Canal.
+\caption{\footnotesize{Biological communities and their relationship with environmental variables in the Hood Canal.
 \textbf{A} Constrained Analysis of Principal Coordinates (CAP) of Bray-Curtis dissimilarities among biological communities, as constrained by pH, temperature and salinity (arrows).
@@
 
-\subsection*{Discussion}
+\section*{Discussion}
 
@@
  Our results therefore suggest a possible change in relative dominance between diatoms and other phytoplankton species such as dinoflagellates, consistent with those seen at ecological regime shifts found elsewhere \cite{wasmund2017diatom, hinder2012changes}.
+Recent studies using mesocosm manipulations also suggest a foundational shift in community composition in that direction \cite{Hyun_2020, Crummett_2020}.
 Such a shift could affect ecosystems in many ways; even under the assumption that the surviving taxa would maintain the primary production levels, for example, the smaller cell-size of dinoflagellates and the differential sinking rates of the two groups would likely alter regional patterns of nutrient cycling and carbon sequestration \cite{xiao2018warming, anderson2002harmful, bienfang1984sinking, polimene2017biological}.
@@
 
-\section{Acknowledgments}
+\section*{Acknowledgments}
 

### 2020-11-04 Scripts/Manuscript.Main.Analysis.Rmd
Final revision

@@
 ggsave(Fig1, filename = here("Figures","Manuscript","Fig1.png"), width = 14, height = 12, dpi = "retina")
+ggsave(Fig1, filename = here("Figures","Manuscript","Fig1.pdf"), width = 14, height = 12, dpi = "retina")
 
@@
 ggsave(plot,filename = here("Figures","Manuscript",paste0("Figure2.", Sys.Date(), ".png")),
+         width = 14,height = 11, dpi = "retina") 
+ggsave(plot,filename = here("Figures","Manuscript",paste0("Figure2.", Sys.Date(), ".tiff")),
          width = 14,height = 11, dpi = "retina")
@@
    ggsave(plot , filename = here("Figures","Manuscript", paste0("Top_fig3_",Sys.Date(),".Other.sel.png")), units = "in", dpi = "retina", width = 20, height = 14)
-
+   ggsave(plot , filename = here("Figures","Manuscript", paste0("Top_fig3_",Sys.Date(),".Other.sel.tiff")), units = "in", dpi = "retina", width = 20, height = 14)
+ ggsave(plot , filename = here("Figures","Manuscript", paste0("Top_fig3_",Sys.Date(),".Other.sel.pdf")), units = "in", dpi = "retina", width = 20, height = 14)
 ```
